---
title: "Integration of donor only and with lab"
date: "2024-01-05"
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  library(rgl)
  library(patchwork)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(ggnewscale)
  library(tidyr)
  library(cowplot)
  library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  library(scmap)
  library(SingleCellExperiment)
  library(pheatmap)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(RColorBrewer)
  # library(MAST)
  library(Platypus)
  library(ggvenn)
  library('biomaRt')
  library(conflicted)
  library(SoupX)
  library(MetBrewer)
  library(knitr)
  library(readxl)
  library(SCpubr)
  library(ggh4x)
  library(htmlwidgets)
  library(slingshot)
  library(harmony)
  library(ggforce)
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
})
```


#  integration of all MSC


```{r}



sp_colors <- c('Early ring'='#D1EC9F',
                     'Late ring'='#78C679',
                     'Early trophozoite'='#FEEEAA',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                     'Gametocyte (developing)'='thistle',
                     'Gametocyte (female)'='#551A8B',
                     'Gametocyte (male)'='mediumpurple',
               'Unassigned'='red', 
               'Not assigned'='#D6CFC7',
               'Female (LE)'='#208cf7',
               'Female'='#551A8B',
               # 'Female (HE)'='#551A8B',
               'Male (LE)'='#a894d1',
               'Male'='mediumpurple',
               # 'Male (HE)'='mediumpurple',
               'Failed QC' = 'grey',
               'Gametocyte (committed)' = '#C399A2', 
               'Gametocyte (developing)' = '#66444C', 
               'Gametocyte (branching)' = '#66444C', 
               'Gametocyte (early female)' = '#749E89', 
               'Gametocyte (early male)' = '#7D87B2', 
               'Gametocyte (late female)' = '#4E6D58', 
               'Gametocyte (late male)' = '#41507B',
                   NULL = 'lightgrey'
)

sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   'Female'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male'= 'mediumpurple',#'#039911',
                   ## 'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Atlas' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle',
                   'early female' = '#749E89',
                   'late female'= '#4E6D58',
                   NULL = 'lightgrey'
)

cluster_cols <- brewer.pal(7, name = 'Set2') %>% set_names(as.character(c(0:6)))



stg_lvls = c("Early ring", "Late ring", "Early trophozoite","Late trophozoite" , "Early schizont", "Late schizont",  "Gametocyte (developing)", "Female (HE)", "Female", "Female (LE)",  "Gametocyte (female)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (male)", "Unassigned", "Failed QC", "Not assigned")

stg_refnd_lvls = c("Early ring", "Late ring", "Early trophozoite", "Late trophozoite", "Early schizont", "Late schizont", "Gametocyte (committed)", "Gametocyte (developing)", "Gametocyte (branching)", "Female (HE)", "Female", "Female (LE)", "Gametocyte (female)", "Gametocyte (early female)", "Gametocyte (late female)", "Gametocyte (male)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (early male)", "Gametocyte (late male)", "Unassigned", "Failed qc", "Not assigned", "Atlas")

stg_refnd_lvls2 = c('Early ring','Late ring','Early trophozoite','Late trophozoite','Early schizont','Late schizont', "Gametocyte (committed)","Gametocyte (developing)","Gametocyte (branching)", "Female (HE)", "Female", "Female (LE)", "Gametocyte (early female)","Gametocyte (late female)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (early male)","Gametocyte (late male)", "Unassigned", "Failed QC", "Not assigned") #%>% str_to_sentence(.)

# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
# donor_cols_m = met.brewer("Thomas")[c(6,3,1,2)]  %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
donor_cols_m = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))
donor_cols = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))
# donor_cols_cap = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% str_to_upper(set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ))))


```

```{r pressure, echo=FALSE}
# ggplot(cars, aes(x= dist, y =speed)) +
#   geom_point()
# graph2ppt(file = '/Users/jr35/Google Drive/My Drive/PHD_documents/test_exp2')
```

```{r}
##Strain cluster colours

strain_cols_n <-c("#910000", "#ff1a8d", "#007c71", "#0ea300", "#00deca", "#9a96c7", "#5d57a4", "#a8875b","#708db3", "#c39a00", "#ffd94d", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","thistle","#a40000" , "#00b7a7", "#e7f1f9", 'black', 'grey')
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_n) <- c('SC2', 'SC1','SC5',  'SC4', 'SC3',  'SC6', 'SC7', 'SC8', 'S3_S1', 'S1_p_i','S1_b', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative', 'Lab', 'Less5', NA)


# strain_cols <- c(met.brewer(name = 'Austria')[c(1,7, 4,5,6, 7, 8)], 'light gray')
strain_cols <- c(met.brewer(name = 'Thomas')[c(1:8)],'black', 'light gray')
# names(strain_cols) = c('Strain 0', 'Strain 1', 'Strain 2', 'Doublet', 'Negative')
names(strain_cols) = c('G1', 'G2', 'G3', 'G4','G5', 'G6','G7', 'G8', 'Doublet', 'Negative')
# strain_cols = c("Strain 2" = "#dd5129", "Strain 1"= "#0f7ba2", "Strain 0"="#43b284", "Doublet"="#fab255")
strain_cols =c(strain_cols,strain_cols_n)

strain_dset_lvls = c( 'SC1', 'SC2','SC3',  'SC4', 'SC5',  'SC6', 'SC7', 'SC8')

umap_theme <- theme(legend.position = 'right',
        axis.title=element_text(size=12,face="bold",hjust = 0),
        # plot.title = element_blank(),
        legend.title=element_text(size=14,face="bold"),
        legend.text=element_text(size=14),
        legend.spacing.y = unit(0.25, "cm"),
        axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed"))
        )  

```

```{r setup}
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali')
```


initialize variable for the sample names

```{r}
# sample_id <- c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" )
sample_id_lst <-as.list(c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" ) %>% 
                          str_sort(., numeric = T))
mt_filt <- c(5,1.2,1,0.45,2)
sample_name_nms<-sample_id_lst
optimum_k = c(5,2,2,8,2) %>% set_names(sample_name_nms)
```

```{r}
msc_qcd_anotd <- readRDS("data/processed/DB52e_star/msc_qcd_anotd.RDS") %>% .[1:4]
```


```{r}
##Read MCA V3 blood stage reference dataset
# EF_V3_ref <- readRDS("../sc_velo/data/raw/RoadtoUMAP/pseudotime/raw_combined_cut_0.4.RDS")
EF_V3_ref <- readRDS("data/raw/mca_v3_ref/raw_combined_cut_0.4_nodoublets_stagenew.RDS")
```

```{r}
##make refined metadata column that includes the gametocyte substages for finer resolution mapping
EF_V3_ref <- EF_V3_ref@meta.data %>%
  mutate(stage_fine = case_when(optionB_Stage == 'no assignment' ~ as.character(Stage),
                                TRUE ~ as.character(optionB_Stage))
  ) %>% 
  mutate(across(stage_fine, ~str_to_sentence(str_replace_all(., c('Gametocytes'='Gametocyte', 'comitted' = 'committed'))))) %>%
  mutate(across(stage_fine, ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
  mutate(across(stagenewx, ~droplevels(factor(str_to_sentence(.), levels = stg_refnd_lvls)))) %>% 
  dplyr::select(stage_fine, stagenewx) %>%
  AddMetaData(EF_V3_ref, .)
```


```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id) )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  dplyr::mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                        TRUE ~ as.character(.)))) 

```

Write out annotated MSC14 dataset in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad
msc_qcd_anotd_diet <- map(msc_qcd_anotd, ~DietSeurat(
  .x,
  counts = TRUE,
  data = T,
  scale.data = FALSE,
  features = NULL,
  assays = NULL,
  dimreducs = c('pca', 'umap', 'umap_test'),
  graphs = NULL
)
)

##Change factor metadata to character as these are not handled properly when converting to h5ad
msc_qcd_anotd_diet<- map(msc_qcd_anotd_diet, ~{
  .x@meta.data %>% mutate(across(where(is.factor), as.character)) %>% AddMetaData(.x, .)
})

imap(msc_qcd_anotd_diet, ~system(paste0("mkdir /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/", str_to_lower(str_remove_all(.y, "_.*")))))

## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top
imap(msc_qcd_anotd_diet, ~SeuratDisk::SaveH5Seurat(.x, filename =  paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.y, "_.*")),'/',str_to_lower(str_remove_all(.y, "_.*")),'_qcd_anotd_diet.h5Seurat'), overwrite = T))


imap(msc_qcd_anotd_diet, ~Convert(paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.y, "_.*")),'/',str_to_lower(str_remove_all(.y, "_.*")),'_qcd_anotd_diet.h5Seurat'), dest = "h5ad", overwrite = T))

```


Write out raw objects into .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad

imap(msc_qcd_anotd_diet, ~{
  obj <- CreateSeuratObject(counts = .x@assays$RNA@counts, meta.data = .x@meta.data)
  obj@meta.data <- obj@meta.data %>% mutate(across(where(is.factor), as.character))
  obj@reductions$pca <- CreateDimReducObject(embeddings = .x@reductions$pca@cell.embeddings, loadings = .x@reductions$pca@feature.loadings, projected = .x@reductions$pca@feature.loadings.projected, key = "PC_", assay = "RNA")
  obj@reductions$umap <- CreateDimReducObject(embeddings = .x@reductions$umap@cell.embeddings, loadings = .x@reductions$umap@feature.loadings, projected = .x@reductions$umap@feature.loadings.projected, key = "UMAP_", assay = "RNA")
  
  
  SeuratDisk::SaveH5Seurat(obj, filename =  paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.y, "_.*")),'/',str_to_lower(str_remove_all(.y, "_.*")),'_qcd_anotd_diet_raw.h5Seurat'), overwrite = T)
  Convert(paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.y, "_.*")),'/',str_to_lower(str_remove_all(.y, "_.*")),'_qcd_anotd_diet_raw.h5Seurat'), dest = "h5ad", overwrite = T)
})

```

Transfer MCA and MSC14 .h5ad to farm
```{r}

##Copy h5ad
map(names(msc_qcd_anotd_diet), ~{
  system(paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.x, "_.*")),'/',str_to_lower(str_remove_all(.x, "_.*")),'_qcd_anotd_diet.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.x, "_.*")),'/'))
  system(paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.x, "_.*")),'/',str_to_lower(str_remove_all(.x, "_.*")),'_qcd_anotd_diet_raw.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/',str_to_lower(str_remove_all(.x, "_.*")),'/'))
  
  })

```


Intergrate each sample with lab one by one
```{r}
# select features that are repeatedly variable across datasets for integration

# msc_y21_mca_fem_combi <- list(unlist(list(msc_list_fnorm, msc_list_mnorm), recursive = F), rep(list(MCA_ref_fem_d10b,MCA_ref_m_d10b), each =4)) %>%
# msc_y21_mca_fem_combi <- list(msc_list_fle[c(1,3,4)], rep(list(MCA_ref_fem_d10b),3), c(200,200,100), c(100,100,50)) %>%
std_seu_int <- function(donor_obj, lab_obj=list(), k_filt = 200, k_wt =100, nf =200, fnf =500){
    msc_y21_mca_fem_features <- SelectIntegrationFeatures(object.list = unlist(list(donor_obj, lab_obj), recursive = F), verbose = F, nfeatures = nf, fvf.nfeatures = fnf)
    
    msc_y21_mca_fem_anchors <- FindIntegrationAnchors(object.list = unlist(list(donor_obj, lab_obj), recursive = F), anchor.features = msc_y21_mca_fem_features, verbose = F, k.filter = k_filt)
    
    # this command creates an 'integrated' data assay
    msc_y21_mca_int <- IntegrateData(anchorset = msc_y21_mca_fem_anchors, verbose = F, k.weight =  k_wt)
    
    rm(msc_y21_mca_fem_anchors, msc_y21_mca_fem_features)
    
    return(msc_y21_mca_int)

  }
  # , "few cells")


```


### Std seurat normalization and clustering
```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

std_norm_seu_umap <- function(seu_int_obj){
  seu_int_obj %>%
  ScaleData(., features = rownames(.), verbose = FALSE)%>% 
  RunPCA(., features = VariableFeatures(object = .), verbose = FALSE) %>% 
  # RunUMAP(., reduction = "pca", dims = 1:6, seed.use = 123456, n.components = 3, min.dist = 0.3, n.neighbors = 30, spread = 1, local.connectivity = 1, a =60, b = 0.9, verbose = F) %>%
  RunUMAP(., reduction = "pca", dims = 1:6, seed.use = 123456, n.components = 3) %>%
  FindNeighbors(k.param = 30, verbose = F)
}
```


```{r}
## Making list of objects and renaming before integrating
msc_list<-msc_qcd_anotd
don_ids = str_remove(names(msc_list), "sc")
msc_list <- map2(msc_list, don_ids, ~RenameCells( .x, add.cell.id = .y))
```


```{r}
## Run integration
msc_combi <- std_seu_int(donor_obj=msc_list) 

```

```{r}
## Run integration
DefaultAssay(msc_combi) <- "integrated"
msc_combi <- std_norm_seu_umap(seu_int_obj=msc_combi) 

```

```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4])))

DimPlot(msc_combi, group.by = c('stageHL', 'orig.ident'),pt.size = 0.1) 
DimPlot(msc_combi, group.by = 'stageHL', pt.size = 0.1, reduction = 'pca', label = T) 
DimPlot(msc_combi,pt.size = 1, group.by = c('stageHL', 'orig.ident'), dims = c(1,3)) 
```


```{r}

DimPlot(msc_combi,pt.size = 1, group.by = 'orig.ident') + scale_color_manual(values = donor_cols_m)
DimPlot(msc_combi,pt.size = 1, group.by = 'orig.ident', dims = c(1,3)) + scale_color_manual(values = donor_cols_m)


```


```{r}

DimPlot(msc_combi,pt.size = 1, group.by = 'StageMCA') + scale_color_manual(values = sp_colors)
DimPlot(msc_combi,pt.size = 1, group.by = 'StageMCA', dims = c(1,3)) + scale_color_manual(values = sp_colors)


```

```{r}
FeaturePlot(msc_combi, pt.size = 1,features  = 'nFeature_RNA') 

DimPlot(msc_combi, group.by = 'orig.ident',pt.size = 1, reduction = 'pca') + scale_color_manual(values = donor_cols_m)
FeaturePlot(msc_combi, features  = 'nFeature_RNA', pt.size = 1,reduction = 'pca') 

```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
# DimPlot(msc_qcd_anotd_stgstrqc[[1]], reduction = "umap", group.by = 'stageHL', dims = c(1, 2),
#                 cols = sp_fld_colors ) 
   
```


```{r, results='asis'}

trunc_axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(2.5, "cm")
)

pmap(list(rep(c("umap","pca"),2), c(2,2,3,3)), ~{
DimPlot(msc_combi, reduction = ..1, group.by = 'stageHL', dims=c(1,..2), 
                cols = sp_fld_colors , pt.size = 0.1,  label = F, repel = T, label.size = 6 ,
                ) +
      theme(axis.text=element_blank(), 
            axis.title=element_blank(),
            plot.title = element_blank(),
            aspect.ratio = 1/1,
            ) + 
    guides(colour = F,
         x= trunc_axis,
         y= trunc_axis) +
  theme(axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
        axis.title = element_text(hjust = 0,size=14, face = 'bold')) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
    
})


```



```{r, results='asis', fig.width=6, fig.height=3}

msc_combi@meta.data %>% dplyr::count(orig.ident, stageHL) %>% mutate(across(orig.ident, ~str_remove(., '.mrna'))) %>% ggplot(aes(x=orig.ident, y = n, fill = stageHL)) + geom_col(position = position_dodge(width = 0.8))+ scale_fill_manual(values = sp_fld_colors)+
  theme_classic()+
      theme(axis.text=element_text(size = 18),
            axis.title=element_blank(),
            plot.title = element_blank(),
            legend.position = 'none'
)

```


```{r}
plot_ly(data = as.data.frame(msc_combi@reductions[["umap"]]@cell.embeddings),
         x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3,
         color= msc_combi$stageHL,
         # colors = c('blue', 'red'),
         opacity = 0.8,
         size = 1, 
         type = 'scatter3d'
 )
```


```{r}
saveRDS(msc_combi, "data/processed/DB52e_star/msc_combi.RDS")

```

Write out annotated MCA reference in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad
msc_combi_diet <- DietSeurat(
  msc_combi,
  counts = TRUE,
  data = T,
  scale.data = FALSE,
  features = NULL,
  assays = NULL,
  dimreducs = c('pca', 'umap', 'umap_test'),
  graphs = NULL
)

##Change factor metadata to character as these are not handled properly when converting to h5ad
msc_combi_diet@meta.data <- msc_combi_diet@meta.data %>% mutate(across(where(is.factor), as.character))
## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top

SaveH5Seurat(msc_combi_diet, filename =  '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet.h5Seurat', overwrite = T)

Convert("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet.h5Seurat", dest = "h5ad", overwrite = T)

```


Transfer MCA and MSC14 .h5ad to farm
```{r}

system('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/'
)

```


Write out annotated MCA reference in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad

msc_combi_diet_raw <- CreateSeuratObject(counts = msc_combi@assays$RNA@counts, meta.data = msc_combi@meta.data)

##Change factor metadata to character as these are not handled properly when converting to h5ad
msc_combi_diet_raw@meta.data <- msc_combi_diet_raw@meta.data %>% mutate(across(where(is.factor), as.character))
## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top

msc_combi_diet_raw@reductions$pca <- CreateDimReducObject(embeddings = msc_combi@reductions$pca@cell.embeddings, loadings = msc_combi@reductions$pca@feature.loadings, projected = msc_combi@reductions$pca@feature.loadings.projected, key = "PC_", assay = "RNA")

msc_combi_diet_raw@reductions$umap <- CreateDimReducObject(embeddings = msc_combi@reductions$umap@cell.embeddings, loadings = msc_combi@reductions$umap@feature.loadings, projected = msc_combi@reductions$umap@feature.loadings.projected, key = "UMAP_", assay = "RNA")

SaveH5Seurat(msc_combi_diet_raw, filename =  '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet_raw.h5Seurat', overwrite = T)

Convert("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet_raw.h5Seurat", dest = "h5ad", overwrite = T)

```

Transfer MCA and MSC14 .h5ad to farm
```{r}

system('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_combi_diet_raw.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/'
)

```
## MSC combi without asexuals

```{r}
## Making list of objects and renaming before integrating
msc_gam_list<-map(msc_qcd_anotd, ~subset(.x, !(subset = StageMCA %in% c("Early trophozoite", "Late ring"))))
don_ids = str_remove(names(msc_gam_list), "sc")
msc_gam_list <- map2(msc_gam_list, don_ids, ~RenameCells( .x, add.cell.id = .y))
```


```{r}
## Run integration
msc_gam_combi <- std_seu_int(donor_obj=msc_gam_list) 

```



```{r}
## Run integration
DefaultAssay(msc_gam_combi) <- "integrated"
msc_gam_combi <- std_norm_seu_umap(seu_int_obj=msc_gam_combi) 

```


```{r, fig.width= 10, fig.height=3.5}
## Add donor and stage labels appropriate for all samples
msc_gam_combi@meta.data[,"Donor"] <- factor(str_remove(msc_gam_combi@meta.data$orig.ident, ".mrna"), levels = don_ids)

# msc_gam_combi <- msc_gam_combi@meta.data %>% 
#   mutate(stageHL_ref = case_when(stageHL == "Gametocyte (female)" ~ "Female", 
#                                                            stageHL == "Gametocyte (male)" ~ "Male", 
#                                                            T ~ stageHL)) %>% 
#   select(stageHL_ref) %>% AddMetaData(msc_gam_combi, .)

```

```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4])))

DimPlot(msc_gam_combi, group.by = c('stageHL', 'orig.ident'),pt.size = 0.1) 
DimPlot(msc_gam_combi, group.by = 'stageHL', pt.size = 0.1, reduction = 'pca', label = T) 
DimPlot(msc_gam_combi,pt.size = 1, group.by = c('stageHL', 'orig.ident'), dims = c(1,3)) 
```

```{r}
 mgc_umap_mdta_df <- msc_gam_combi@reductions$umap@cell.embeddings[,c('UMAP_1', 'UMAP_2')] %>% as.data.frame() %>% rownames_to_column('cell_bc') %>% left_join(., msc_gam_combi@meta.data[,c('Donor', 'nCount_RNA', 'nFeature_RNA')] %>% as.data.frame() %>% rownames_to_column('cell_bc'), by = 'cell_bc') 
  
  don_f_plt <- mgc_umap_mdta_df %>% 
    mutate(samp = sample(n())) %>%
    arrange(samp) %>%
    ggplot(aes(x = UMAP_1, y =UMAP_2, color = Donor)) + 
    geom_point(size=.1) + 
    scale_color_manual(name = "Donor", values = donor_cols) +
    theme_classic()+
    umap_theme+
    theme(legend.position = "bottom", 
          plot.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=4.5), title.position = "top", nrow = 2),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
  
  cnts_plt <- mgc_umap_mdta_df %>% 
    mutate(samp = sample(n())) %>%
    arrange(samp) %>%
    ggplot(aes(x = UMAP_1, y =UMAP_2, color = nFeature_RNA)) + 
    geom_point(size=.2) + 
    # scale_color_distiller(name='Transcript count (X1e3)', direction = -1, palette = 'GnBu') + 
    scale_color_distiller(name='Gene count', direction = -1, palette = 'GnBu') + 
    theme_classic()+  
    umap_theme+
    theme(legend.position = "bottom", 
          plot.title = element_blank()) +
    guides(colour = guide_colorbar(title.position = "top", barwidth = unit(5, "cm")),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
  
don_f_plt
cnts_plt
```

```{r, fig.width= 12, fig.height=4}
##Put together the plots for main gam integration

gam_int_plt <- plot_grid(don_f_plt,
          cnts_plt,
          NULL,
          NULL,
          ncol = 4,
          rel_widths = c(.2,.2, .22, .38),
          labels = c('A', 'B', 'C', 'D'),
          label_size = 24,
          vjust= 1.5)

gam_int_plt

```


```{r, fig.width= 12, fig.height=7}
##save svg for sk21
save_plot('plots/mali2021_all/fig4/gam_int_plt.svg',gam_int_plt, base_width = 12, base_height = 4)
```

```{r}

DimPlot(msc_gam_combi,pt.size = 1, group.by = 'orig.ident') + scale_color_manual(values = donor_cols_m)
DimPlot(msc_gam_combi,pt.size = 1, group.by = 'orig.ident', dims = c(1,3)) + scale_color_manual(values = donor_cols_m)

DimPlot(msc_gam_combi,pt.size = 1, group.by = 'StageMCA') + scale_color_manual(values = sp_colors)
DimPlot(msc_gam_combi,pt.size = 1, group.by = 'StageMCA', dims = c(1,3)) + scale_color_manual(values = sp_colors)


```

```{r}

stages_umap <- DimPlot(msc_gam_combi,pt.size = 1, group.by = 'stageHL_ref') + 
  scale_color_manual(values = sp_fld_colors) +
  umap_theme+
  theme(legend.position = "bottom", 
        plot.title = element_blank(),
        legend.title = element_text(face = "bold", size = 14),
        legend.text = element_text( size = 14)) +
  guides(color = guide_legend(title = "Stage", override.aes = list(size = 3), title.position = "top", nrow = 2) ,
         x= trunc_axis,
         y= trunc_axis) +
  theme(axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
        axis.title = element_text(hjust = 0,size=14, face = 'bold')) +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)

rna_cnt <- FeaturePlot(msc_gam_combi,pt.size = 1, features = 'nCount_RNA') + 
  umap_theme+
  theme(legend.position = "bottom", 
        plot.title = element_blank(),
        legend.title = element_text(face = "bold"),
        legend.key.width = unit(1.2, "cm"),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text  = element_blank(),
        axis.title  = element_blank()) #+
  # guides(x= trunc_axis,
  #        y= trunc_axis) +
  # theme(axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
  #       axis.title = element_text(hjust = 0,size=14, face = 'bold')) +
  # scale_x_continuous(breaks = NULL) +
  # scale_y_continuous(breaks = NULL)
  
stages_umap
rna_cnt
```

```{r, fig.width= 10, fig.height=3.5}

umap_split_don <- DimPlot(msc_gam_combi,pt.size = 1, group.by = 'Donor', split.by = 'Donor') + scale_color_manual(values = donor_cols) + theme(legend.position = "none", plot.title = element_blank())

umap_split_don
```


```{r, fig.width= 12, fig.height=7.5, eval =F}
##Put together the plots for annotation

plot_grid(umap_split_don,
          plot_grid(stages_umap, 
                    rna_cnt,
                    NULL,
                    NULL,
                    ncol = 4,
                    rel_widths = c(.25,.25,.25,.25),
                    labels = c('B', 'C', 'D', 'E'),
                    label_size = 24,
                    vjust= 0.3
                    ),
          nrow = 2,
          rel_widths = c(.45,.55),
          labels = c('A', ''),
          label_size = 24,
          vjust= 1.5)


```

```{r}
saveRDS(msc_gam_combi, "data/processed/DB52e_star/msc_gam_combi.RDS")
```



Write out annotated MCA reference in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad

msc_gam_combi_diet_raw <- CreateSeuratObject(counts = msc_gam_combi@assays$RNA@counts, meta.data = msc_gam_combi@meta.data)

##Change factor metadata to character as these are not handled properly when converting to h5ad
msc_gam_combi_diet_raw@meta.data <- msc_gam_combi_diet_raw@meta.data %>% mutate(across(where(is.factor), as.character))
## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top

msc_gam_combi_diet_raw@reductions$pca <- CreateDimReducObject(embeddings = msc_gam_combi@reductions$pca@cell.embeddings, loadings = msc_gam_combi@reductions$pca@feature.loadings, projected = msc_gam_combi@reductions$pca@feature.loadings.projected, key = "PC_", assay = "RNA")

msc_gam_combi_diet_raw@reductions$umap <- CreateDimReducObject(embeddings = msc_gam_combi@reductions$umap@cell.embeddings, loadings = msc_gam_combi@reductions$umap@feature.loadings, projected = msc_gam_combi@reductions$umap@feature.loadings.projected, key = "UMAP_", assay = "RNA")

SaveH5Seurat(msc_gam_combi_diet_raw, filename =  '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_gam_combi_diet_raw.h5Seurat', overwrite = T)

Convert("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_gam_combi_diet_raw.h5Seurat", dest = "h5ad", overwrite = T)

```


Transfer MCA and MSC14 .h5ad to farm
```{r}

system('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc_gam_combi_diet_raw.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/'
)

```


####integrate field gams without LE plus mca gams - all msc objects included separately

```{r}
# MCA_V3_ref_gam <- EF_V3_ref[, EF_V3_ref@meta.data$stagenewx %in% c("gametocyte (developing)", "gametocyte (male)", "gametocyte (female)")]
MCA_V3_ref_gam <- EF_V3_ref[, EF_V3_ref@meta.data$Stagex %in% c("early stalk","late stalk", "branching", "early female", "late female", "early male", "late male")]

# msc_y21_gams <-map(msc_list, ~.x[, !.x@meta.data$stageHL %in% c('Early trophozoite', 'Late ring')])
msc_gam_list_noLE<-map(msc_qcd_anotd, ~subset(.x, !(subset = stageHL %in% c("Early trophozoite", "Late ring", "Female (LE)", "Male (LE)"))))



DimPlot(MCA_V3_ref_gam, group.by = 'orig.ident', dims = c(1,3))
map(msc_gam_list_noLE, ~DimPlot(., group.by = c('stageHL')))
```


```{r}
## Run integration
msc_gam_combi_noLE <- std_seu_int(donor_obj=msc_gam_list_noLE, lab_obj = MCA_V3_ref_gam) 

```

```{r}
## Run integration
DefaultAssay(msc_gam_combi_noLE) <- "integrated"
msc_gam_combi_noLE <- std_norm_seu_umap(seu_int_obj=msc_gam_combi_noLE) 

```


```{r}
## Function to Add metadata for visualization to object when integrating field with lab

add_mdata <- function(seu_obj) {
  seu_obj@meta.data %>% 
    mutate(stage_int = case_when(is.na(stageHL) ~ Stagex, T ~ StageMCA),
           dset_int = case_when(str_detect(orig.ident, "D") ~ Stagex, T ~ orig.ident),
           stageHL_fine_int = case_when(str_detect(orig.ident, "D") ~ Stagex, T ~ stageHL),
           stageHL_int = case_when(str_detect(orig.ident, "D") ~ stagenewx, T ~ stageHL),
           stagec_int = case_when(str_detect(orig.ident, "D") ~ "Lab", T ~ StageMCA)) %>% 
    select(stage_int, dset_int, stageHL_int, stagec_int, stageHL_fine_int) %>% 
    AddMetaData(seu_obj, .)
}


```


```{r}
## Add metadata for visualization to object

msc_gam_combi_noLE<- add_mdata(seu_obj= msc_gam_combi_noLE)
```

```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4])))

DimPlot(msc_gam_combi_noLE, group.by = c('stageHL_int', 'orig.ident'),pt.size = 0.1) 
DimPlot(msc_gam_combi_noLE, group.by = 'stageHL_int', pt.size = 0.1, reduction = 'pca', label = T, order=c("Lab","Female", "Gametocyte (female)","Gametocyte (male)","Male")) 
DimPlot(msc_gam_combi_noLE,pt.size = 1, group.by = c('stageHL_int', 'orig.ident'), dims = c(1,3)) 
```


```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
DimPlot(msc_gam_combi_noLE, group.by = "stageHL_int", pt.size = 0.1, reduction = 'umap', order=c("Female (LE)","Male (LE)","Female","Male", "Gametocyte (female)","Gametocyte (male)","Lab"), cols = sp_fld_colors) 
```

```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
DimPlot(msc_gam_combi_noLE, group.by = "stagec_int", pt.size = 0.1, reduction = 'pca', order=c("Female","Male", "Gametocyte (female)","Gametocyte (male)","Lab"), cols = sp_fld_colors) 
```


```{r, fig.width=8, fig.height=3.5}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

stg_order <- c("Gametocyte (female)", "Gametocyte (male)","Female", "Female (LE)", "early female", "late female",  "early male", "late male")

donor_order <- c(names(donor_cols_m), "merged_MCA", "D5", "D10a","D10b")
```


```{r, fig.width= 10, fig.height=5}
# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(names(msc_qcd_anotd[1:4]), '.mrna'))
DimPlot(msc_gam_combi_noLE, group.by = "orig.ident", pt.size = 0.1, reduction = 'pca', order=donor_order, cols = donor_cols_m) 
```


#### Slingshot


## Trajectory inference of integrated object
Adjust to get appropriate clustering so as to input into slingshot
```{r}
## Adjusting resolution to find the optimum for slingshot trajectory inference to avoid error  'values must be length 150, but FUN(X[[1]]) result is length 38689'

msc_gam_combi_noLE <- FindClusters(msc_gam_combi_noLE, resolution = 0.4, verbose = F)
  
```

```{r}
## Adjusting resolution to find the optimum for slingshot trajectory inference to avoid error  'values must be length 150, but FUN(X[[1]]) result is length 38689'
DimPlot(msc_gam_combi_noLE, reduction = "pca", group.by = "seurat_clusters", pt.size = 0.5, dims = c(1,2), label = T)
  
```

```{r}
msc_gam_combi_noLE@meta.data %>% dplyr::count(stage_int, seurat_clusters)
msc_gam_combi_noLE@meta.data %>% dplyr::count(seurat_clusters,stage_int )
```

```{r}
## Adjusting resolution to find the optimum for slingshot trajectory inference to avoid error  'values must be length 150, but FUN(X[[1]]) result is length 38689'
msc_gam_combi_noLE_sce <- as.SingleCellExperiment(msc_gam_combi_noLE)
  
```

Convert seurat object into single cell experiment filling in the slots appropriately
```{r}
##Raw counts - Preparing single cell experiment object for slingshot trajectory inference and tradeseq DE

#logcounts(msc14_sub_sce[[i]]) <- log2(counts(msc14_sub_sce[[i]]) + 1)
##This command gives error below when running the entire notebook or this whole chunk so run the command independently - coulb be and issue with rmd
##Error in x$.self$finalize() : attempt to apply non-function - need blank line after

sce_obj_creatn <- function(sce_obj, seu_obj){
  for(i in 1:length(sce_obj)) {
    sce_obj[[i]] <- SingleCellExperiment(assays = list(counts = as.matrix(seu_obj[[i]]@assays$RNA@counts)), colData = seu_obj[[i]]@meta.data)
    
    logcounts(sce_obj[[i]]) <- log2((counts(sce_obj[[i]]) + 1))
    normcounts(sce_obj[[i]]) <- as.matrix(seu_obj[[i]]@assays$RNA@data)
    rowData(sce_obj[[i]])$feature_symbol <- str_replace(rownames(sce_obj[[i]]), "-", "_")
    reducedDims(sce_obj[[i]]) <- SimpleList(PCA = seu_obj[[i]]@reductions[["pca"]]@cell.embeddings, UMAP = seu_obj[[i]]@reductions[["umap"]]@cell.embeddings)

  }
  return(sce_obj)
}

```


```{r}
msc_gam_combi_noLE_sce <- sce_obj_creatn(sce_obj = list(msc_gam_combi_noLE_sce), seu_obj = list(msc_gam_combi_noLE))
msc_gam_combi_noLE_sce <- msc_gam_combi_noLE_sce[[1]]
```

Run slingshot for trajectory inference
```{r}
##Function -slingshot

##Run slingshot to determine trajectories
# msc_y21_mca_fem_combi_sce$seurat_clusters <- as.character(msc_y21_mca_fem_combi_sce$seurat_clusters)
# msc_y21_mca_fem_combi_sce <- list(msc_y21_mca_fem_combi_sce, start.clus, end.clus) %>%
sling_start <- function(sce_obj, s_clstr, clstr ="seurat_clusters"){ 
  slingshot(sce_obj, 
               clusterLabels = clstr, 
               reducedDim = 'PCA', 
               start.clus = s_clstr,
               shrink =1, 
               stretch = 1,
               allow.break = F)

}

```

```{r}
# msc_gam_combi_noLE_sce <- sling_start(sce_obj = msc_gam_combi_noLE_sce, s_clstr = "early stalk", clstr = "stage_int")
msc_gam_combi_noLE_sce <- sling_start(sce_obj = msc_gam_combi_noLE_sce, s_clstr = "2", clstr = "seurat_clusters")
```

```{r}
##Prepare slingshot dataset (SDS) object 
msc_gam_combi_noLE_sce_sds <- SlingshotDataSet(msc_gam_combi_noLE_sce)

```


```{r}
plot(reducedDims(msc_gam_combi_noLE_sce)$PCA[,1:2], col = brewer.pal(7,"Set1")[as.factor(msc_gam_combi_noLE_sce$stage_int)], pch=16, asp = 0.7)
lines(msc_gam_combi_noLE_sce_sds, lwd=2, col='black',dims = 1:2)
```


```{r, results='asis'}

plot(reducedDims(msc_gam_combi_noLE_sce)$PCA[,1:2], col = brewer.pal(7,"Set1")[as.factor(msc_gam_combi_noLE_sce$seurat_clusters)], pch=16, asp = 0.7)
    lines(msc_gam_combi_noLE_sce_sds, lwd=2, col='black',dims = 1:2)
    
    plot(reducedDims(msc_gam_combi_noLE_sce)$PCA[,2:3], col = brewer.pal(7,"Set1")[as.factor(msc_gam_combi_noLE_sce$seurat_clusters)], pch=16, asp = 0.7)
    lines(msc_gam_combi_noLE_sce_sds, lwd=2, col='black',dims = 2:3)
    
    umap_meta_df <- reducedDims(msc_gam_combi_noLE_sce)$PCA[,c(1,2)] %>% as.data.frame() %>%rownames_to_column('cell_bc') %>% left_join(., msc_gam_combi_noLE_sce@colData[,c('stageHL', "slingPseudotime_1", "slingPseudotime_2", "orig.ident")] %>% as.data.frame() %>% rownames_to_column('cell_bc'), by = 'cell_bc')  #%>% 
  # mutate(across(strain_dset, ~factor(., levels = rev(strain_dset_lvls_don)))) %>% arrange(strain_dset
  
  plt <- umap_meta_df %>% 
    ggplot(aes(x = PC_1, y =PC_2, color = slingPseudotime_1)) + geom_point(size=0.5) + 
    scale_color_distiller(name='Pseudotime', direction = 1, palette = 'Blues') + 
    theme_classic()+
    theme(axis.text=element_blank(), 
          axis.title=element_blank(),
          legend.text = element_text(size = 22) ,
          legend.title = element_text(size = 24, face = "bold") )
  
  print(plt)
  
  plt <- umap_meta_df %>% 
    ggplot(aes(x = PC_1, y =PC_2, color = slingPseudotime_2)) + geom_point(size=0.5) + 
    scale_color_distiller(name='Pseudotime', direction = 1, palette = 'Greens') + 
    theme_classic()+
    theme(axis.text=element_blank(), 
          axis.title=element_blank(),
          legend.text = element_text(size = 22) ,
          legend.title = element_text(size = 24, face = "bold") )
  
  print(plt)
  
  plt = umap_meta_df %>% 
    ggplot(., aes(x = slingPseudotime_1, col=stageHL, fill=stageHL)) + 
    scale_color_manual(name = "Stage", values = sp_fld_colors)+ 
    scale_fill_manual(name = "Stage", values = sp_fld_colors)+
    geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
    # geom_freqpoly()+
    theme_classic()+
    theme(axis.text=element_text(size=26),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title=element_text(size=30,face="bold"), 
          legend.text = element_text(size = 26) ,
          legend.title = element_text(size = 28, face = "bold"),
          # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
          # legend.position="bottom",
          # legend.direction="vertical"
    ) +
    labs(x = 'Pseudotime', y='Counts') +
    guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"))
  
  print(plt)
  
  
  plt = umap_meta_df %>% 
    ggplot(., aes(x = slingPseudotime_2, col=stageHL, fill=stageHL)) + 
    scale_color_manual(name = "Stage", values = sp_fld_colors)+ 
    scale_fill_manual(name = "Stage", values = sp_fld_colors)+
    geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
    # geom_freqpoly()+
    theme_classic()+
    theme(axis.text=element_text(size=26),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title=element_text(size=30,face="bold"), 
          legend.text = element_text(size = 26) ,
          legend.title = element_text(size = 28, face = "bold"),
          # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
          # legend.position="bottom",
          # legend.direction="vertical"
    ) +
    labs(x = 'Pseudotime', y='Counts') +
    guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"))
  
  print(plt)
  
  plt = umap_meta_df %>% 
    ggplot(., aes(x = slingPseudotime_1, col=orig.ident, fill=orig.ident)) + 
    scale_color_manual(name = "Stage", values = donor_cols_m)+
    scale_fill_manual(name = "Stage", values = donor_cols_m)+
    geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
    # geom_freqpoly()+
    theme_classic()+
    theme(axis.text=element_text(size=26),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title=element_text(size=30,face="bold"), 
          legend.text = element_text(size = 26) ,
          legend.title = element_text(size = 28, face = "bold"),
          # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
          # legend.position="bottom",
          # legend.direction="vertical"
    ) +
    labs(x = 'Pseudotime', y='Counts') +
    guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"))
  
  print(plt)

  plt = umap_meta_df %>% 
    ggplot(., aes(x = slingPseudotime_2, col=orig.ident, fill=orig.ident)) + 
    scale_color_manual(name = "Stage", values = donor_cols_m)+
    scale_fill_manual(name = "Stage", values = donor_cols_m)+
    geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
    # geom_freqpoly()+
    theme_classic()+
    theme(axis.text=element_text(size=26),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title=element_text(size=30,face="bold"), 
          legend.text = element_text(size = 26) ,
          legend.title = element_text(size = 28, face = "bold"),
          # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
          # legend.position="bottom",
          # legend.direction="vertical"
    ) +
    labs(x = 'Pseudotime', y='Counts') +
    guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"))
  
  print(plt)


```



```{r, results='asis'}
msc_gam_combi_noLE_pt<- msc_gam_combi_noLE@meta.data  %>% 
  rownames_to_column('cell_bc') %>% 
  mutate(n_cb = str_remove(cell_bc, '_1')) %>% 
  left_join(., umap_meta_df %>% 
              mutate(n_cb = str_remove(cell_bc, '_1')), by=c('n_cb', 'orig.ident') ) %>% 
  select(cell_bc.x, slingPseudotime_1, slingPseudotime_2) %>% 
  column_to_rownames('cell_bc.x') %>% 
  AddMetaData(msc_gam_combi_noLE, .)

```


```{r, results='asis'}
FeaturePlot(msc_gam_combi_noLE_pt, features = "slingPseudotime_1", reduction = "pca")
FeaturePlot(msc_gam_combi_noLE_pt, features = "slingPseudotime_2", reduction = "pca")
```



### Visualize trajectory{.tabset}


```{r, results='asis'}

# pt_plts <- function(sce_obj, sds_obj, pt){
#     plot(reducedDims(sce_obj)$PCA[,1:2], col = brewer.pal(7,"Set1")[as.factor(sce_obj$seurat_clusters)], pch=16, asp = 0.7)
#     lines(sds_obj, lwd=2, col='black',dims = 1:2)
#     
#     plot(reducedDims(sce_obj)$PCA[,2:3], col = brewer.pal(7,"Set1")[as.factor(sce_obj$seurat_clusters)], pch=16, asp = 0.7)
#     lines(sds_obj, lwd=2, col='black',dims = 2:3)
#     
#     umap_meta_df <- reducedDims(sce_obj)$PCA[,c(1,3)] %>% as.data.frame() %>%rownames_to_column('cell_bc') %>% left_join(., sce_obj@colData[,c('stageHL', pt)] %>% as.data.frame() %>% rownames_to_column('cell_bc'), by = 'cell_bc')  #%>% 
#   # mutate(across(strain_dset, ~factor(., levels = rev(strain_dset_lvls_don)))) %>% arrange(strain_dset
#   
#   plt <- umap_meta_df %>% 
#     ggplot(aes(x = PC_1, y =PC_3, color = !!rlang::sym(pt))) + geom_point(size=2) + 
#     scale_color_distiller(name='Pseudotime', direction = 1, palette = 'Blues') + 
#     theme_classic()+
#     theme(axis.text=element_blank(), 
#           axis.title=element_blank(),
#           legend.text = element_text(size = 22) ,
#           legend.title = element_text(size = 24, face = "bold") )
#   
#   print(plt)
#   
#   plt = umap_meta_df %>% 
#     ggplot(., aes(x = !! rlang::sym(pt), col=stageHL, fill=stageHL)) + 
#     scale_color_manual(name = "Stage", values = sp_fld_colors)+ 
#     scale_fill_manual(name = "Stage", values = sp_fld_colors)+
#     geom_density(aes(y = ..count..*0.18), size=2, bw = 0.18, alpha=.2)+
#     # geom_freqpoly()+
#     theme_classic()+
#     theme(axis.text=element_text(size=26),
#           axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
#           axis.title=element_text(size=30,face="bold"), 
#           legend.text = element_text(size = 26) ,
#           legend.title = element_text(size = 28, face = "bold"),
#           # legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),#######
#           # legend.position="bottom",
#           # legend.direction="vertical"
#     ) +
#     labs(x = 'Pseudotime', y='Counts') +
#     guides(color = guide_legend(override.aes = list(size = 4), title.position = "top"))
#   
#   print(plt)
#   
#   }
# 

```

```{r, results='asis'}
# pt_plts(sce_obj = msc_gam_combi_noLE_sce, sds_obj = msc_gam_combi_noLE_sce_sds, pt = c("slingPseudotime_1", "slingPseudotime_2"))
```


Save integrated dataset with pseudotimes
```{r}
saveRDS(msc_gam_combi_noLE, "data/processed/DB52e_star/msc_gam_combi_noLE.RDS")
saveRDS(msc_gam_combi_noLE_sce, "data/processed/DB52e_star/msc_gam_combi_noLE_sce.RDS")
saveRDS(msc_gam_combi_noLE_sce_sds, "data/processed/DB52e_star/msc_gam_combi_noLE_sce_sds.RDS")
```

########################### END OF INTERGRATION FOR FIRST PART

