---
title: "msc_strain_decon_new"
date: '2023-08-10'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---


```{r, message=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(pheatmap)
  library(rmarkdown)
  library('vcfR')
  library(conflicted)
  library(RColorBrewer)
  library(MetBrewer)
  library(ComplexHeatmap)
  library(SeqArray)
  library(ggrepel)
  library(ggvenn)
  library(magick)
  library(svglite)
  library(plotly)
  library(gganimate)
  library(av)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("count", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflict_prefer("saveRDS", "base")
})

```


# MSC parasites strain deconvolution and VCF visualization 

## Variable declarations

```{r}
##good coours 
# "#6eb5ff", "#f4a261"
##Strain cluster colours
strain_cols <- c(met.brewer(name = 'Austria')[c(1,7, 4,5,6)], 'light gray')

# strain_cols_n <-c("#910000", "#f20000", "#007c71", "#00f2dd", "#c39a00", "#ffd94d", "#7570B3","#708db3", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","#a40000" , "#00b7a7", "#ffcd12", 'black', 'lavender', 'lightgray', 'grey')
# names(strain_cols_n) <- c('SC2', 'SC1','SC4',  'SC3', 'SC6',  'SC5', 'SC7', 'S2_S3', 'S3_S1', 'S1_p_i','S1_b', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative',  'Less5')
# "#aeaa61" "#c695c9"

strain_cols_n <-c("#910000", "#ff1a8d", "#007c71", "#0ea300", "#00deca", "#9a96c7", "#5d57a4", "#a8875b","#708db3", "#c39a00", "#ffd94d", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","thistle","#a40000" , "#00b7a7", "#e7f1f9", 'black', '#050301','grey')
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_n) <- c('SC2', 'SC1','SC5',  'SC4', 'SC3',  'SC6', 'SC7', 'SC8', 'SC9', 'SC10', 'SC11', 'SC12', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative', 'Lab', 'Less5', 'PoorQC', NA)

strain_cols_fh <-c("#910000", "#007c71", "#5d57a4", "lightgray")
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_fh) <- c('SC2', 'SC5', 'SC7', 'Other')

strain_cols_nl <-c("#910000", "#f20000", "#007c71", "#00f2dd", "#c39a00", "#ffd94d", "#7570B3","#708db3", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#thistle","#a40000" , "#00b7a7", "#ffcd12", 'black', 'lavender', 'lightgray', 'grey')
names(strain_cols_nl) <- c('SC1b', 'SC1a', 'SC2b',  'SC2a', 'SC3b',  'SC3a', 'S2_S1', 'S2_S3', 'S3_S1', 'S1_pa','S1_b', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'S1_3i','Negative',  'Less5')

sp_colors <- c('Early ring'='#D1EC9F',
                     'Late ring'='#78C679',
                     'Early trophozoite'='#FEEEAA',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                     'Gametocyte (developing)'='thistle',
                     'Gametocyte (female)'='#551A8B',
                     'Gametocyte (male)'='mediumpurple',
               'Unassigned'='red', 
               'Not assigned'='#D6CFC7',
               'Female (LE)'='#208cf7',
               'Female (HE)'='#551A8B',
               'Male (LE)'='#a894d1',
               'Male (HE)'='mediumpurple',
               'Failed QC' = 'grey',
               'Gametocyte (committed)' = '#C399A2', 
               'Gametocyte (developing)' = '#66444C', 
               'Gametocyte (branching)' = '#66444C', 
               'Gametocyte (early female)' = '#749E89', 
               'Gametocyte (early male)' = '#7D87B2', 
               'Gametocyte (late female)' = '#4E6D58', 
               'Gametocyte (late male)' = '#41507B',
                   NULL = 'lightgrey'
)

sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle',
                   'NA' = 'lightgrey',
                   'stgQC_fail' = 'lightgrey'
)

c_stg_shapes = c('Asexual' = 16 , 'Male' = 3,  'Female' = 6)
c_stg_cols= c('Asexual' = "#D7B7C7", 'Male' = "#C4CAA4", 'Female'="#6289A9", 'Developing' = "#C9F6E1", 'Unassigned'="lightgrey")

stg_refnd_lvls = c("Early ring", "Late ring", "Early trophozoite", "Late trophozoite", "Early schizont", "Late schizont", "Gametocyte (committed)", "Gametocyte (developing)", "Gametocyte (branching)", "Female (HE)", "Female (LE)", "Gametocyte (female)", "Gametocyte (early female)", "Gametocyte (late female)", "Gametocyte (male)", "Male (HE)", "Male (LE)", "Gametocyte (early male)", "Gametocyte (late male)", "Unassigned", "Failed qc", "Not assigned", "Atlas", "stgQC_fail")

trip = 'Mali'
algn_nm = 'minmap'
```

```{r setup}
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/')
```



## Genotype clustering for MSC

Read in clusters.tsv files with assignment of cells to strains and PCA plots metrics

Transfer clusters.tsv files with assignment of cells to strains and PCA plots metrics for the best K from farm to my local Mac
```{r}
##Copy souporcell files for the optimum clusters

##Sample ID
# sample_name_nms <- c("msc33", "msc48","msc49", "msc55", "msc60", "msc66") %>% sort()
sample_name_nms <- c("msc1", "msc3","msc13","msc14","msc1272") %>% str_sort(numeric = T)
algn_nms = c('minmap', 'hsat')

sample_name_nms_al <- rep(sample_name_nms, 2)
algn_nms_al <- rep(algn_nms, each= length(sample_name_nms))

sm_aln_nms_al <- paste(rep(sample_name_nms, 2), rep(algn_nms, each= length(sample_name_nms)), sep ='_')
# rep(algn_nms, each= length(sample_name_nms)


## number of optimum clusters to investigate per sample
# opt_clusters_nms <- list("msc33" = c(7, 8, 9),"msc48" = c(4,5,6),"msc49" = c(3,4,5),"msc55" = c(4,5,6),"msc60" = c(3,4,5), "msc66" = c(8,9,10))
opt_clusters_nms <- list("msc1" = c(3:9),"msc3" = c(2:3),"msc13" = c(2:6),"msc14" = c(4:10),"msc1272" = c(2:7))

opt_clusters_len <- map(opt_clusters_nms, length)

##Samples
sample_name <- map2(sample_name_nms, opt_clusters_len[sample_name_nms], ~rep(.x, each = .y))

opt_clusters <- unlist(flatten(opt_clusters_nms[sample_name_nms]))

sample_name <- unlist(flatten(sample_name))

## sorted donor_opt_cluster names
sorted_names <- paste0(sample_name, "_", opt_clusters)

## sorted factoring alignment
sorted_names_al <-paste0(sample_name, "_",rep(algn_nms,each=length(sample_name)),"_", opt_clusters)


iter = opt_clusters_len
```


```{r, eval=F}
##Copy souporcell files for the optimum clusters

for (i in 1:length(sample_name)) {
  for (a in 1:length(algn_nms)) {
    
    ## Create new directories
    system(
      paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', sample_name[i],'/soupc/',algn_nms[a],'/k', opt_clusters[i])
    )
    
    ##Transfers
    ##souporcell genotyping files
    system(
      paste0('rsync -av --include="cluster_genotypes.vcf" --include="clusters.tsv" --exclude="*" jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', sample_name[i],'/soupc/',algn_nms[a],'/k' , opt_clusters[i], '/* /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', sample_name[i],'/soupc/',algn_nms[a],'/k', opt_clusters[i])
      )
  }  
}


```

## Log likelihood Elbow plots visualisations for all MSCs
```{r, message=FALSE, eval=FALSE}
## copy from farm
map2(rep(sample_name_nms, 2), rep(algn_nms, each= length(sample_name_nms)), ~system(paste0('rsync -av jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', .x,'/soupc/',.y,'/clusters_log_likelihoods.txt /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', .x,'/soupc/',.y,'/')))

```


```{r, message=FALSE}
## Read in log likelihoods for estimation of number of clusters from elbow plots
clusters_log_likelihoods_lst = map2(rep(sample_name_nms, 2), rep(algn_nms, each= length(sample_name_nms)), ~read_table(paste0('data/processed/DB52e_star/', .x,'/soupc/',.y,'/clusters_log_likelihoods.txt'), col_names = c('No_of_clusters', 'td1','td2','td3', 'td4','log_likelihood')) %>% select(-(starts_with("td"))) %>% mutate(across(No_of_clusters, ~str_remove_all(., "out_dir_k|/clusters.err:best")))) %>% set_names(paste(rep(sample_name_nms, 2), rep(algn_nms, each= length(sample_name_nms)), sep ='_'))
```



```{r} 
map(clusters_log_likelihoods_lst, ~.x %>%
  mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
  ggplot(., aes(x = No_of_clusters, y = log_likelihood)) +
  geom_point() +
  geom_line(aes(group = 1)) +
  # facet_grid(rows = .~id)+
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        title = element_text(size=16,face="bold"),
        strip.text.x = element_text(size = 18, face = "bold")))
```



Transfer clusters.tsv files with assignment of cells to strains and PCA plots metrics for the best K from farm to my local Mac
```{r}
##Copy souporcell files for the optimum clusters

```


Read all the souporcell files for all the iterations. May be excessive but since we have the memory and cpu and we may need to use the files downstream then why not.
```{r, results='hide'}
# list.files("data/processed/DB52e_star/", pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)
list.files("data/processed/DB52e_star/", pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>% set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|soupc/|k|/clusters.tsv') %>% str_replace_all(., '/', '_'))) 
```

minimap2
```{r}
soupc_all <- list.files("data/processed/DB52e_star/", pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>% set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|soupc/mi.*k|/clusters.tsv') %>% str_replace(., '/', '_'))) %>% .[sorted_names]%>%
  map_df(read_delim, delim = "\t", col_types = 'cccddd', .id = "sample_k") 


```

hisat2
```{r}
soupc_all <- list.files("data/processed/DB52e_star/", pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>%
  set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|soupc/h.*k|/k|/clusters.tsv') %>% str_replace_all(., '/', '_'))) %>% .[sorted_names] %>%
  map_df(read_delim, delim = "\t", col_types = 'cccddd', .id = "sample_k") %>% mutate(across(sample_k, ~str_remove(., "_hsat")))


```

combi
```{r}
soupc_all_al <- list.files("data/processed/DB52e_star/", pattern = "clusters.tsv", full.names = T, recursive = T, include.dirs = T)%>% set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|soupc/|k|/clusters.tsv') %>% str_replace_all(., '/', '_')))  %>% .[sorted_names_al]%>%
  map_df(read_delim, delim = "\t", col_types = 'cccddd', .id = "sample_k") 


```


Select appropriate k  (number of strains) for each MSC sample
```{r}
# knee_k <- c()
```


Select only information from clusters.tsv for the best k as determined from the log likelihood plot for each MSC sample

```{r}
# soupo_kselect <- soupc_all %>% 
# filter(sample_k %in% knee_k) %>%
# dplyr::select(sample_k, barcode, status, assignment) %>% 
#   dplyr::mutate("Strain" = case_when(status == "doublet" ~ "Doublet",
#                                      status == "unassigned" ~ "Negative",
#                                      TRUE ~ paste0('Strain ', as.character(assignment)))
#                 ) 

##Change Strain to G0,G1,G2 for publication figures and enhanced readability
# soupo_kselect <- soupc_all %>% 
soupo_kselect_al <- soupc_all_al %>% 
  # filter(sample_k %in% knee_k) %>%
  dplyr::select(sample_k, barcode, status, assignment) %>% 
  dplyr::mutate("Strain" = case_when(status == "doublet" ~ "Doublet",
                                     status == "unassigned" ~ "Negative",
                                     assignment == "0" ~ "SC1",
                                     assignment == "1" ~ "SC2",
                                     assignment == "2" ~ "SC3",
                                     assignment == "3" ~ "SC4",
                                     assignment == "4" ~ "SC5",
                                     assignment == "5" ~ "SC6",
                                     assignment == "6" ~ "SC7",
                                     assignment == "7" ~ "SC8",
                                     assignment == "8" ~ "SC9",
                                     assignment == "9" ~ "SC10",
                                     assignment == "10" ~ "SC11",
                                     assignment == "11" ~ "SC12",
                                     assignment == "12" ~ "SC13",
                                     assignment == "13" ~ "SC14",
                                     assignment == "14" ~ "SC15",
                                     assignment == "15" ~ "SC16")
  ) 
```


```{r}
# Compare HISAT and MINIMAP
soupo_kselect_al_cor <- soupo_kselect_al %>%
  separate_wider_delim(cols = "sample_k", delim = "_", names = c("sample", "aligner", "k")) %>% 
  unite(col = "sample_k", c('sample','k'), remove = FALSE)  %>%
  select("sample_k","aligner","barcode","Strain") %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  .[sorted_names]
  
map(soupo_kselect_al_cor, ~pivot_wider(.x, names_from = aligner, values_from = Strain) %>% count(minmap, hsat) %>% pivot_wider(names_from = hsat, values_from = n) %>% column_to_rownames('minmap') %>% as.matrix() %>% ComplexHeatmap::pheatmap(., cluster_rows = F, cluster_cols = F, display_numbers = T))
  

```


Split the information for each sample into different labels and put into a list named appropriately
```{r}
soupo_kselect_list <- soupo_kselect_al %>%
  filter(str_detect(sample_k, algn_nm)) %>%
  mutate(across(sample_k, ~str_remove(.,paste0("_", algn_nm)))) %>%
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>% dplyr::select(Strain))


soupo_kselect_list <- soupo_kselect_list[as.character(sorted_names)]


##Get strain levels for samples with more than one strain and mutate Strain column to reflect this
# strain_levels <- map(soupo_kselect_list[2:8], ~.x %>% filter(str_detect(Strain, 'Strain')) %>% pull(Strain) %>% unique() %>% sort() %>% c(.,"Doublet","Negative"))
strain_levels <- map(soupo_kselect_list, ~.x %>% dplyr::filter(str_detect(Strain, '^SC')) %>% pull(Strain) %>% unique() %>% sort() %>% c(.,"Doublet","Negative"))

soupo_kselect_list <- map2(soupo_kselect_list,strain_levels, ~.x %>% mutate(across(Strain, ~factor(., levels = .y)))) 
```

```{r}
# soupo_kselect_list <- soupo_kselect_list[c(1:32, 33, 36, 39, 40, 42:55)]
# sortd_names_new <- names(soupo_kselect_list)
```

```{r, message=FALSE}
##Copy souporcell files for the optimum clusters investigation on igv - all except negative

soupo_kmerge <- map(sample_name_nms, ~{
soupo_kselect_list[str_detect(names(soupo_kselect_list), paste0(.x, "_"))] %>% 
  imap(., ~{
    nms = .y
    .x %>% rename_with(~paste0(.,'_', nms), everything()) %>% rownames_to_column('bcode')
        }) %>% 
    purrr::reduce(., left_join, by = 'bcode')
  }) %>% 
  set_names(sample_name_nms)

```

```{r}
##Transfer stage labels and other metadata from farm
# list.files("data/processed/DB52e_star/", pattern = "msc_filt_w_dblts_scmap_metadata.RDS", full.names = T, recursive = T, include.dirs = T) %>% 
#   set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|_5736.*|/msc_filt_w_dblts_scmap_metadata.RDS') %>% str_replace(., '/', '_'))) 

all_msc_anotd_mdata <- readRDS("data/processed/DB52e_star/all_msc_anotd_mdata.RDS")
msc_mdata <- map(all_msc_anotd_mdata, ~.x %>% rownames_to_column('bcode'))
# msc_mdata <- map2(all_msc_anotd_mdata, msc_qcd_anotd_mdata, ~.x  %>% select(StagePrelim) %>% rownames_to_column('bcode') %>% left_join(., .y %>% rownames_to_column('bcode'), by='bcode'))
```

```{r}
##Transfer stage labels and other metadata from farm
# mdata_all <- list.files("data/processed/DB52e_star/", pattern = "msc_filt_w_dblts_scmap_metadata.RDS", full.names = T, recursive = T, include.dirs = T) %>% 
#   set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|_5736.*|/msc_filt_w_dblts_scmap_metadata.RDS') %>% str_replace(., '/', '_'))) %>%
mdata_all <- msc_mdata %>% .[sample_name_nms] %>%
  map(. %>% 
        # readRDS() %>% 
        # dplyr::rename('stageHL' = stageHR_msc14) %>%
        mutate(#across(stageHL, ~str_replace_all(.,c('high count' = 'HE', 'low count' = 'LE'))),
               ##Collapse asexuals and gams and number of cells per coarse group & add the number per xluster as a column in metadata
               stage_afm = case_when(str_detect(StagePrelim, 'ring|troph') ~ 'Asexual', 
                                     str_detect(StagePrelim, '\\(male') ~ 'Male', 
                                     str_detect(StagePrelim, 'female') ~ 'Female', 
                                     str_detect(StagePrelim, 'developing') ~ 'Developing',
                                     T ~StagePrelim),
               stage_ag = case_when(str_detect(StagePrelim, 'ring|troph') ~ 'Asexual', 
                                    str_detect(StagePrelim, 'ale|developing') ~ 'Gametocyte',
                                     T ~StagePrelim)) 
      ) 

```

```{r}
##read in stage labels 

soupo_seu_kmerge <- list(soupo_kmerge[sample_name_nms], mdata_all[sample_name_nms]) %>%
  pmap(~{
    # ..1 %>% left_join(., ..2 %>% rownames_to_column('bcode'), by = 'bcode')  
    ..1 %>% left_join(., ..2 , by = 'bcode')  
  })

```


```{r, fig.width = 16, fig.height = 4, message=FALSE, warning=FALSE}
iter=2

ss_bar_plt <-list( rep(soupo_seu_kmerge, times = opt_clusters_len), rep(names(soupo_seu_kmerge), times = opt_clusters_len), opt_clusters_nms[names(soupo_seu_kmerge)] %>% flatten() %>% map(.,max))  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stage_afm, ~droplevels(factor(., levels=c('Asexual', 'Female', 'Male', 'Developing', 'Unassigned'))))) %>%
      dplyr::count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_afm, .drop = FALSE) 
  
  ##get stage_ag labels and counts to add to the Strain_max+stage_afm counts generated above
  stage_ag_cnt <- ..1 %>% 
    dplyr::select(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_ag, stage_afm) %>% 
    add_count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_ag, name = 'stage_ag_n') %>% 
    distinct(pick(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_afm), .keep_all = T) %>% 
    left_join(a_tbl, ., by =c(paste('Strain',..2,..3, sep='_'), 'stage_afm')) %>% 
    rename(c('Stage'= stage_afm, 'Counts' =n, 'Strain' = !!rlang::sym(paste('Strain',..2,..3, sep='_'))))  
  
  print(stage_ag_cnt)
  
    b_plt <- stage_ag_cnt %>%
      # ggplot(., aes(x = Stage, y = Counts, fill = Stage, color=stage_ag_n<30)) +
      ggplot(., aes(x = Stage, y = Counts, fill = Stage)) +
      geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
      geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 7, show.legend = F, fontface ='bold', colour='black')+
      scale_fill_manual(values = c_stg_cols) + #'Asexual' = "#FFE5B4", 'Male' = "#BFD7B5", 'Female'="#8ab6d6"
      # scale_color_manual(values = c('FALSE' = "white", 'TRUE' = "blue")) +
  geom_tile(aes(x = Stage, y = -10, fill = Stage, height = 14)) +
  facet_wrap( Strain ~ ., nrow = 1) +
      scale_y_continuous(expand = expansion(mult = c(0.01, 0.15))) +
      scale_x_discrete(expand = expansion(add = c(1, 1))) +
      theme_classic() +
  labs( y = 'Counts')+
      theme(
        axis.text.y  = element_text(size = 22),
        axis.title.y  = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position="bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=20,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
      )+ 
      guides(color = F)
    print(b_plt)
    
    return(b_plt)
})

##!!!!!!!NB!!!! SUNILS STAGES
```

```{r, fig.width = 16, fig.height = 4, message=FALSE, warning=FALSE}
iter=2

ss_bar_plt <-list( rep(soupo_seu_kmerge, times = opt_clusters_len), rep(names(soupo_seu_kmerge), times = opt_clusters_len), opt_clusters_nms[names(soupo_seu_kmerge)] %>% flatten() %>% map(.,max))  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stage_afm, ~factor(., levels=c('Asexual', 'Female', 'Male', 'Developing', 'Unassigned')))) %>%
      dplyr::count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_afm, .drop = FALSE) 
  
  ##get stage_ag labels and counts to add to the Strain_max+stage_afm counts generated above
  stage_ag_cnt <- ..1 %>% 
    dplyr::select(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_ag, stage_afm) %>% 
    add_count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_ag, name = 'stage_ag_n') %>% 
    distinct(pick(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_afm), .keep_all = T) %>% 
    left_join(a_tbl, ., by =c(paste('Strain',..2,..3, sep='_'), 'stage_afm')) %>% 
    rename(c('Stage'= stage_afm, 'Counts' =n, 'Strain' = !!rlang::sym(paste('Strain',..2,..3, sep='_'))))  
  
  print(stage_ag_cnt)
  
    b_plt <- stage_ag_cnt %>%
      # ggplot(., aes(x = Stage, y = Counts, fill = Stage, color=stage_ag_n<30)) +
      ggplot(., aes(x = Stage, y = Counts, fill = Stage)) +
      geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
      geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 7, show.legend = F, fontface ='bold', colour='black')+
      scale_fill_manual(values = c_stg_cols) + #'Asexual' = "#FFE5B4", 'Male' = "#BFD7B5", 'Female'="#8ab6d6"
      # scale_color_manual(values = c('FALSE' = "white", 'TRUE' = "blue")) +
  geom_tile(aes(x = Stage, y = -10, fill = Stage, height = 14)) +
  facet_wrap( Strain ~ ., nrow = 1) +
      scale_y_continuous(expand = expansion(mult = c(0.01, 0.15))) +
      scale_x_discrete(expand = expansion(add = c(1, 1))) +
      theme_classic() +
  labs( y = 'Counts')+
      theme(
        axis.text.y  = element_text(size = 22),
        axis.title.y  = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position="bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=20,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
      )+ 
      guides(color = F)
    print(b_plt)
    return(b_plt)
})

##!!!!!!!NB!!!! SUNILS STAGES
```


## msc3 strains = sc1(includes 3) & sc2
## msc13 strains = sc1 & sc2(incudes sc3)
## msc1 sc1 from msc1_4 splits into sc1 n sc3 in msc_5 and sc1_6 in msc_6



```{r, message=FALSE}
##Coalesce the genotypes and get a consensus genotype taking labels for highest resolution K
##VARIABLE
## !!!NB!!!!!When trialing out different souporcell Ks use Strain_qc_cutoff > 20 but once decided use Strain_qc_cutoff > 0 the optimum K 

# Strain_qc_cutoff=20
Strain_qc_cutoff=0

optimum_k = c(5,2,2,8,2) %>% set_names(sample_name_nms)

opt_narrow_clusters_nms <- list("msc1" = c(4:7),"msc3" = c(2:3),"msc13" = c(2:5),"msc14" = c(5:8),"msc1272" = c(2:3))



strn_c_mdata <- list(soupo_seu_kmerge, names(soupo_seu_kmerge), opt_clusters_nms[names(soupo_seu_kmerge)] %>% map(.,max), optimum_k, imap(opt_narrow_clusters_nms, ~paste0(.y, '_', .x))) %>%
  pmap(~{
    ..1 %>%
      unite(col = 'Strn_c', starts_with("Strain_msc"), sep='_', remove = F) %>% 
      add_count(Strn_c) %>% 
      mutate(
        'Strain_max_qc' = case_when( 
          !str_detect(Strn_c, 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste('Strain',..2,..3, sep='_')), 
          T ~ 'PoorQC'),
        'Strain_qc' = case_when(
          !str_detect(!!rlang::sym(paste('Strain',..2,..4, sep='_')), 'Negative|Doublet') & n>Strain_qc_cutoff ~ !!rlang::sym(paste('Strain',..2,..4, sep='_')), 
          T ~ 'PoorQC'),
        'Strain_DN' = !!rlang::sym(paste('Strain',..2,..4, sep='_')),
        'dbt_neg_all' = case_when(
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet") ~ "DoubletAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Negative") ~ "NegativeAll",
          if_all(c(paste0('Strain_', ..5)), ~ .== "Doublet" | .== "Negative") ~ "DbtNegMix", 
          T ~ 'LikelySinglet')
        )%>%
        dplyr::add_count(Strain_qc, stage_ag, name = 'cluster_ag_n') %>% 
        dplyr::add_count(Strain_qc, stage_afm, name = 'cluster_afm_n')
  })



```



```{r} 

map2(clusters_log_likelihoods_lst[str_detect(names(clusters_log_likelihoods_lst), "hsat")] , optimum_k, ~.x %>%
  mutate(opt_k = .y) %>% head
)
```


```{r} 
optimum_kv2 <- c(4,3,3,7,3) %>% set_names(sample_name_nms)
optimum_kv3 <- c(6,3,3,6,2) %>% set_names(sample_name_nms)


sc_ll_knee_p_all <- map2(clusters_log_likelihoods_lst[str_detect(names(clusters_log_likelihoods_lst), "hsat")] %>% .[1:4] , optimum_k[1:4], ~.x %>% mutate(opt_k = .y)) %>%
  bind_rows(., .id = 'donor') %>% 
  mutate(across(donor, ~factor(str_remove(., '_hsat'), levels = names(optimum_k[1:4])))) %>%
    mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
    ggplot(., aes(x = No_of_clusters, y = log_likelihood/100000)) + 
    geom_point() +
    geom_line(aes(group = 1)) + 
    geom_vline(aes(xintercept = optimum_k), data.frame(optimum_k[1:4]) %>% set_names("optimum_k") %>% rownames_to_column('donor') %>% mutate(across(donor, ~factor(., levels = sample_name_nms[1:4])))) +
    scale_x_discrete(breaks = function(x){x[c(F, T)]}) +
    facet_wrap(donor~., nrow = 1)+
    theme_classic() +
    theme(axis.title = element_text(size=20, face = 'bold'),
          axis.text.y = element_text(size=18),
          axis.text.x = element_text(size=16),
          strip.text =element_text(size=24,face="bold", colour = 'black'),
          panel.border=element_rect(colour="black",fill='transparent')) +
    labs(y= "Log likelihood ( X1e5)", x = "Number of strains")

sc_ll_knee_p_all
saveRDS(sc_ll_knee_p_all, "plots/fig_QC_s1/sc_ll_knee_p_all.RDS")
```


Plot log likelihood

```{r, fig.width=7.5, fig.height= 5.5}
  
sc_ll_knee_p_lst <- list(clusters_log_likelihoods_lst[str_detect(names(clusters_log_likelihoods_lst), "hsat")], optimum_k, optimum_kv2, optimum_kv3) %>%
  pmap(., ~{
      ..1 %>% 
  mutate(across(c('No_of_clusters'), ~factor(as.numeric(.)))) %>%
  ggplot(., aes(x = No_of_clusters, y = log_likelihood/100000)) + 
  geom_point() +
  scale_x_discrete(breaks = function(x){x[c(F, T)]}) +
  geom_vline(xintercept=..2)+
  geom_vline(xintercept=..3)+
  geom_vline(xintercept=..4)+
  geom_line(aes(group = 1)) + 
  theme_classic() +
  theme(text = element_text(size=35)) +
  labs(y= "Log likelihood (X1e5)", x = "Number of strains")
    })

sc_ll_knee_p_lst
saveRDS(sc_ll_knee_p_lst, "plots/fig_QC_s1/sc_ll_knee_p_lst.RDS")
```


```{r}
## visualize top clusters
list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain", "_", ..2, "_", ..3))) %>% arrange(desc(n))
})
```


```{r}
## Doublets confirmed removal round 2
## visualize top clusters
list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
  purrr::pmap(.,~{
  ..1 %>% filter(if_all(starts_with('Strain_msc'), ~ .== "Doublet" | .== "Negative")) %>% dplyr::count(Strn_c, dbt_neg_all, !!rlang::sym(paste0("Strain", "_", ..2, "_", ..3))) 
})
```

```{r}
## visualize top clusters
list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain", "_", ..2, "_", ..3))) %>% arrange(desc(n))
})
```

```{r}
## visualize top clusters
list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain", "_", ..2, "_", ..3))) %>% arrange(desc(n))
})
```


### PCA calculation
```{r}
soupo_PC_mtx <- soupc_all %>% 
  dplyr::select(-c(status, assignment, log_prob_singleton, log_prob_doublet)) %>% 
  dplyr::group_split(sample_k) %>%
  purrr::set_names(purrr::map_chr(., ~.x$sample_k[1])) %>%
  map(. %>% column_to_rownames(., "barcode") %>%
        dplyr::select(where(~!all(is.na(.)))) %>%
        dplyr::select(-sample_k) %>%
        mutate(across(where(is.numeric), ~ (. / rowSums(across(where(is.numeric)))))) %>%
        as.matrix() %>%
        t())



soupo_PC_mtx <- soupo_PC_mtx[as.character(sorted_names)]
```



```{r}
##Calculate PCA
soupo_PCA <- soupo_PC_mtx %>% 
  map(. %>% prcomp(.))
```


Calculate proportion of variance explained by PC1 and PC2
```{r}
# Assuming your data is stored in a matrix called 'data'
# Extract the standard deviations of the principal components
# pc_standard_deviations <- soupo_PCA[[9]]$sdev
# 
# #PC1 variance explained
# (pc_standard_deviations[1] ^ 2) / sum(pc_standard_deviations^2)
# 
# #PC2 variance explained
# (pc_standard_deviations[2] ^ 2) / sum(pc_standard_deviations^2)
```


```{r}
## Save PCA
saveRDS(soupo_PCA, "data/processed/DB52e_star/soupo_PCA_m21.RDS")
```



```{r}
## visualize top clusters
list(strn_c_mdata)  %>% 
  purrr::pmap(.,~{
  # ..1 %>% dplyr::count(Strain_qc, StageMCA) %>% arrange(desc(n))
  ..1 %>% dplyr::count(Strain_qc, Stage_V3) %>% arrange(desc(n))
  # ..1 %>% dplyr::count(stageHL) %>% arrange(desc(n))
  # ..1 %>% dplyr::count(stage_lab) %>% arrange(desc(n))
})
```


Add stage labels collapsing asexuals and gams and number of cells per coarse group
```{r}
##Add the number per xluster as a column in metadata

map(strn_c_mdata, . %>% 
      filter(cluster_ag_n < 20) %>%
      dplyr::count(Strain_qc, stage_ag, stage_afm) 
)
 
```

```{r}
##Add the number per xluster as a column in metadata
map(strn_c_mdata, . %>% 
      filter(cluster_ag_n > 20) %>%
      dplyr::count(Strain_qc, stage_ag, stage_afm) 
)

```

```{r, fig.width = 12, fig.height = 4, message=FALSE, warning=FALSE}


ss_bar_plt <-list( strn_c_mdata)  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stage_afm, ~factor(., levels=c('Asexual', 'Female', 'Male', 'Developing', 'Unassigned')))) %>%
      dplyr::count(Strain_qc, stage_afm, .drop = FALSE) 
  
  ##get stage_ag labels and counts to add to the Strain_qc+stage_afm counts generated above
  stage_ag_cnt <- ..1 %>% 
    dplyr::select(Strain_qc, stage_ag, stage_afm) %>% 
    add_count(Strain_qc, stage_ag, name = 'stage_ag_n') %>% 
    distinct(pick(Strain_qc, stage_afm), .keep_all = T) %>% 
    left_join(a_tbl, ., by =c('Strain_qc', 'stage_afm')) %>% 
    rename(c('Stage'= stage_afm, 'Counts' =n, 'Strain' = Strain_qc))  
  
  print(stage_ag_cnt)
  
    b_plt <- stage_ag_cnt %>%
      # ggplot(., aes(x = Stage, y = Counts, fill = Stage, color=stage_ag_n<30)) +
      ggplot(., aes(x = Stage, y = Counts, fill = Stage)) +
      geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
      geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 7, show.legend = F, fontface ='bold', colour='black')+
      scale_fill_manual(values = c_stg_cols) + #'Asexual' = "#FFE5B4", 'Male' = "#BFD7B5", 'Female'="#8ab6d6"
      # scale_color_manual(values = c('FALSE' = "white", 'TRUE' = "blue")) +
  geom_tile(aes(x = Stage, y = -10, fill = Stage, height = 14)) +
  facet_wrap( Strain ~ ., nrow = 1) +
      scale_y_continuous(expand = expansion(mult = c(0.01, 0.15))) +
      scale_x_discrete(expand = expansion(add = c(1, 1))) +
      theme_classic() +
  labs( y = 'Counts')+
      theme(
        axis.text.y  = element_text(size = 22),
        axis.title.y  = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position="bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=20,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
      )+ 
      guides(color = F)
    print(b_plt)
    return(b_plt)
})


```


```{r, fig.width = 12, fig.height = 4, message=FALSE, warning=FALSE}


ss_bar_plt <-list( strn_c_mdata)  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stage_afm, ~factor(., levels=c('Asexual', 'Female', 'Male', 'Developing', 'Unassigned')))) %>%
      dplyr::count(Strain_DN, stage_afm, .drop = FALSE) 
  
  ##get stage_ag labels and counts to add to the Strain_DN+stage_afm counts generated above
  stage_ag_cnt <- ..1 %>% 
    dplyr::select(Strain_DN, stage_ag, stage_afm) %>% 
    add_count(Strain_DN, stage_ag, name = 'stage_ag_n') %>% 
    distinct(pick(Strain_DN, stage_afm), .keep_all = T) %>% 
    left_join(a_tbl, ., by =c('Strain_DN', 'stage_afm')) %>% 
    rename(c('Stage'= stage_afm, 'Counts' =n, 'Strain' = Strain_DN))  
  
  print(stage_ag_cnt)
  
    b_plt <- stage_ag_cnt %>%
      # ggplot(., aes(x = Stage, y = Counts, fill = Stage, color=stage_ag_n<30)) +
      ggplot(., aes(x = Stage, y = Counts, fill = Stage)) +
      geom_col(position = position_dodge2()) + #width=0.4, position = position_dodge(width=0.5) 
      geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 7, show.legend = F, fontface ='bold', colour='black')+
      scale_fill_manual(values = c_stg_cols) + #'Asexual' = "#FFE5B4", 'Male' = "#BFD7B5", 'Female'="#8ab6d6"
      # scale_color_manual(values = c('FALSE' = "white", 'TRUE' = "blue")) +
  geom_tile(aes(x = Stage, y = -10, fill = Stage, height = 14)) +
  facet_wrap( Strain ~ ., nrow = 1) +
      scale_y_continuous(expand = expansion(mult = c(0.01, 0.15))) +
      scale_x_discrete(expand = expansion(add = c(1, 1))) +
      theme_classic() +
  labs( y = 'Counts')+
      theme(
        axis.text.y  = element_text(size = 22),
        axis.title.y  = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position="bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=20,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')
      )+ 
      guides(color = F)
    print(b_plt)
    return(b_plt)
})


```


```{r}

# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

# msc_m21_qcd_anotd_strn <- readRDS("data/processed/msc_m21_qcd_anotd_strn.RDS")
# 
# list(msc_m21_qcd_anotd_strn, strn_c_mdata) %>%
#   pmap(.,~{
#     ..1@meta.data %>% rownames_to_column('bcode') %>% full_join(., ..2, by='bcode') %>% dplyr::count( Strain_qc, strain,stage_ag.y) 
#   })

```


```{r}
## visualize top clusters
list(strn_c_mdata, sample_name_nms, optimum_k)  %>% 
  purrr::pmap(.,~{
  ..1 %>% dplyr::count(Strn_c, Strain_qc, !!rlang::sym(paste0("Strain", "_", ..2, "_", ..3))) %>% arrange(desc(n))
})
```

###msc3 = SC1+SC3
###ms13 = SC2+SC3
```{r}
##Add the number per xluster as a column in metadata

map(strn_c_mdata, . %>% 
      # filter(cluster_ag_n >= 100) %>%
      dplyr::count(Strain_DN, stage_ag, stage_afm) 
)

```

```{r}

##Save strain + new metadata as RDS
strn_c_mdata <- map(strn_c_mdata, 
    . %>% 
      mutate(across(stage_afm, ~ifelse(is.na(.), 'Low_n', .)),
             across(stage_ag, ~ifelse(is.na(.), 'Low_n', .))) %>%
      # add_count(Strain_qc, stage_afm, name = 'strain_afm_n')%>% 
      # add_count(Strain_qc, stage_ag, name = 'strain_ag_n')%>%
      add_count(Strain_DN, stage_afm, name = 'strain_afm_n')%>% 
      add_count(Strain_DN, stage_ag, name = 'strain_ag_n')%>%
      # mutate('strain' = case_when(strain_afm_n>=15 ~ Strain_qc))%>%
      # mutate('strain' = case_when(strain_afm_n>=15 ~ Strain_DN))%>%
      mutate('strain' = case_when(strain_afm_n>5 ~ Strain_DN))
    ) 


```

```{r}
##Add the number per xluster as a column in metadata

map(strn_c_mdata, . %>% 
      # filter(cluster_ag_n >= 100) %>%
      dplyr::count(strain,Strain_DN, stage_ag, stage_afm) 
)

```

```{r}
##Clusters with more than 100

map(strn_c_mdata, . %>% 
      filter(cluster_ag_n >= 100) %>%
      dplyr::count(Strain_qc, stage_ag) %>% 
      filter(Strain_qc != 'PoorQC') %>%
      unite(col = 'cluster',Strain_qc, stage_ag, sep = '_') %>%
      pull(cluster) %>%
      str_remove_all(., 'ametocyte|sexual')
) %>% imap(., ~paste(.x, .y, sep='_')) %>%
  saveRDS(., "data/processed/DB52e_star/ss_ag_gtype_clusters_100_m21.RDS")

```

```{r}
##Save strain + new metadata as RDS
map(strn_c_mdata, 
    . %>% 
      # dplyr::select("bcode", "Strn_c", starts_with("Strain"),"stage_afm", "stage_ag", "cluster_ag_n", "cluster_afm_n", "stageHL", "strain_afm_n", "strain")) %>%
      dplyr::select("bcode", "Strn_c", starts_with("Strain"),"stage_afm", "stage_ag", "cluster_ag_n", "cluster_afm_n", "strain_afm_n", "strain", "dbt_neg_all")) %>%
  saveRDS("data/processed/DB52e_star/strn_c_new_mdata_m21.RDS")


```


Write out strain_qcd/strain + stage_ag/stage_afm barcodes to farm for pseudobulk genotyping

```{r, eval=T}
## DOUBLETS WRITTEN INTO THE SAME FOLDER IN MSC_PP.rmd - SHOULD BE COMBINED INTO ONE NOTEBOOK
# sample_ids = map2(sample_name_nms, source_irods_nms, ~paste0(.x, '_', .y))

for (s in names(strn_c_mdata)) {
  for (str_q in c('strain')){
    # ss_qcd_ano <- map(strn_c_mdata[s], ~.x %>% dplyr::select(str_q) %>% mutate(across(c(str_q), ~droplevels(factor(.)))) %>% rownames_to_column('bcode')%>% rename('str'=str_q) %>% split(f = ~ str, drop = T))
    ss_qcd_ano <- map(strn_c_mdata[s], ~.x %>% dplyr::select(str_q, bcode) %>% mutate(across(c(str_q), ~droplevels(factor(.)))) %>% rename('str'=str_q) %>% split(f = ~ str, drop = T))
    names(ss_qcd_ano[[1]]) <- names(ss_qcd_ano[[1]]) %>% str_replace(., '\\.', '_')
    print(names(ss_qcd_ano[[1]]))
    
    system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/'))
    system(paste0('rm /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/*'))

    imap(ss_qcd_ano[[1]], ~write(.x$bcode, paste0('data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/',.y, '.tsv')))
     
    system(
      paste0('rsync -av --delete /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s], ' jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/')
    )
    
    ##Add doublets
    # system(
    #   paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/Doublet.tsv ', '/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/')
    # )
    
    
    for(stg_c in c('stage_afm', 'stage_ag')){
      ss_qcd_ano <- map(strn_c_mdata[s], ~.x %>% dplyr::select(str_q, stg_c, bcode) %>% mutate(across(c(str_q, stg_c), ~droplevels(factor(.)))) %>% rename('stg'=stg_c, 'str'=str_q) %>% split(f = ~ str+stg, drop = T))
      names(ss_qcd_ano[[1]]) <- names(ss_qcd_ano[[1]]) %>% str_replace(., '\\.', '_')
      print(names(ss_qcd_ano[[1]]))
      
      system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/'))
      system(paste0('rm /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/*'))

      imap(ss_qcd_ano[[1]], ~write(.x$bcode, paste0('data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/',.y, '.tsv')))
      
      system(
        paste0('rsync -av --delete /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s], ' jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/')
      )
      
       ##Add doublets
    #   system(
    #   paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/Doublet.tsv ', '/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/')
    # )
      
    }
  }
    system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip, '/data/processed/DB52e_star/',s,'/bcodes/',algn_nm, '/bcodes_no_dbt' ))
    
    strn_c_mdata[[s]] %>% filter(dbt_neg_all == "LikelySinglet") %>% pull(bcode) %>% write(., paste0('data/processed/DB52e_star/', s, '/bcodes/',algn_nm, '/bcodes_no_dbt/barcodes.tsv'))
  
    ##Add barcodes after removing of doublets concordant across ks
    system(
      paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm, '/bcodes_no_dbt/barcodes.tsv ', 'jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/',s,'/bcodes/',algn_nm, '/bcodes_no_dbt/')
    )
}
```
END OF STRAIN FIRST PASS ANALYSIS

#####START##!!! NEW NOTEBOOK AFTER ASSIGNMENT OF LE/HE - PUTTING CART BEFORE HORSE HERE - can be sorted by assigning LE HE stages at the preprocessing and qc pp script


```{r}
##Transfer stage labels and other metadata from farm
# list.files("data/processed/DB52e_star/", pattern = "msc_filt_w_dblts_scmap_metadata.RDS", full.names = T, recursive = T, include.dirs = T) %>% 
#   set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|_5736.*|/msc_filt_w_dblts_scmap_metadata.RDS') %>% str_replace(., '/', '_'))) 

msc_qcd_anotd_mdata <- readRDS("data/processed/DB52e_star/msc_qcd_anotd_mdata.RDS")
msc_mdata_qc <- map2(msc_qcd_anotd_mdata[sample_name_nms], strn_c_mdata[sample_name_nms], ~.x  %>% select(stageHL) %>% rownames_to_column('bcode') %>% left_join(., .y, by='bcode'))
```

```{r}
##Transfer stage labels and other metadata from farm
# mdata_all <- list.files("data/processed/DB52e_star/", pattern = "msc_filt_w_dblts_scmap_metadata.RDS", full.names = T, recursive = T, include.dirs = T) %>% 
#   set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|_5736.*|/msc_filt_w_dblts_scmap_metadata.RDS') %>% str_replace(., '/', '_'))) %>%
mdata_all_qc <- msc_mdata_qc %>% .[sample_name_nms] %>%
  map(. %>% 
        # readRDS() %>% 
        # dplyr::rename('stageHL' = stageHR_msc14) %>%
        mutate(across(stageHL, ~str_replace_all(.,c('high count' = 'HE', 'low count' = 'LE'))))
  )

```

```{r}
##read in stage labels 

# soupo_seu_kmerge <- list(soupo_kmerge[sample_name_nms], mdata_all[sample_name_nms]) %>%
#   pmap(~{
#     # ..1 %>% left_join(., ..2 %>% rownames_to_column('bcode'), by = 'bcode')  
#     ..1 %>% left_join(., ..2 , by = 'bcode')  
#   })

```


```{r, fig.width = 16, fig.height = 4, message=FALSE, warning=FALSE}
iter=2

ss_bar_plt <-list( rep(mdata_all_qc, times = opt_clusters_len), rep(names(mdata_all_qc), times = opt_clusters_len), opt_clusters_nms[names(mdata_all_qc)] %>% flatten() %>% map(.,max))  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stage_afm, ~droplevels(factor(., levels=c('Asexual', 'Female', 'Male', 'Developing', 'Unassigned'))))) %>%
      dplyr::count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stage_afm, .drop = FALSE) 
  
  ##get stage_ag labels and counts to add to the Strain_max+stage_afm counts generated above
 
    c_plt <- ..1 %>%
      mutate(across(stageHL, ~droplevels(factor(., levels = stg_refnd_lvls))))  %>%
      dplyr::count(!!rlang::sym(paste('Strain',..2,..3, sep='_')), stageHL, .drop = FALSE)  %>%
      rename(c('Counts' =n, 'Strain' = !!rlang::sym(paste('Strain',..2,..3, sep='_')))) %>%
      ggplot(., aes(x = stageHL, y = Counts/100, fill = stageHL)) +
 geom_col() +
  geom_tile(aes(x = stageHL, y = -(10/100), fill = stageHL, height = 7/100)) +
  scale_fill_manual(values =  sp_fld_colors, guide = guide_legend( direction = "horizontal"  ))+
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors)))+
  geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 4.8, show.legend = F, fontface = 'bold')+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.2))) +
  scale_x_discrete(expand = expansion(add = c(1.2, 0.8))) +
  facet_grid(.~Strain) +
  theme_classic() +
  labs( y = 'Counts (X100)')+
      theme(
        axis.text.y  = element_text(size = 15),
        axis.title.y  = element_text(size = 16, face = "bold"),
        # axis.title.x = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16, face = "bold"),
        legend.position="bottom",
        # plot.margin = unit(c(1,2,1,1),"cm")
        # legend.margin=margin()
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=18,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')#,
        # legend.margin=margin(t = 0, unit='cm')
      ) +
    guides(fill = guide_legend( nrow = 1, title.position = "top"))
    print(c_plt)
})

##!!!!!!!NB!!!! SUNILS STAGES
```


```{r}
list(soupo_PCA, rep(mdata_all_qc, times = opt_clusters_len), names(soupo_PCA)) %>%
  pmap(~{..1$rotation %>% 
     as.data.frame() %>%
     rownames_to_column('bcode') %>% 
      left_join(..2, by = 'bcode') %>%
     ggplot(., aes(x=PC1, y=PC3, color = Strain_qc, shape = stageHL)) + 
      scale_color_manual(values = strain_cols_n)+
     geom_point(size=2) +
    labs(title = ..3) +
      theme_classic()
      })

```


```{r}
## visualize top clusters
list(mdata_all_qc)  %>% 
  purrr::pmap(.,~{
  ..1 %>% 
      ggplot(aes(x= Strain_qc , y = nFeature_RNA, color = stageHL)) +
      geom_boxplot()
      
})
```


```{r}
# list(soupo_PCA, rep(strn_c_mdata, times = opt_clusters_len), names(soupo_PCA)) %>%
list(soupo_PCA, rep(mdata_all_qc, times = opt_clusters_len), names(soupo_PCA)) %>%
  pmap(~{..1$rotation %>% 
     as.data.frame() %>%
     rownames_to_column('bcode') %>% left_join(..2, by = 'bcode') %>%
      # ggplot(., aes(x=PC1, y=PC3, color = Stage_scmap_msc14)) +
     ggplot(., aes(x=PC1, y=PC3, color = stageHL)) + 
      scale_color_manual(values = sp_fld_colors) +
     geom_point(size=2)+
    labs(title = ..3) +
      theme_classic()
    })

```


```{r, fig.width = 12, fig.height = 4.5, message=FALSE, warning=FALSE}


stg_strn_distb_bar<-list( mdata_all_qc)  %>% 
  purrr::pmap(.,~{
  a_tbl <- ..1 %>%
      mutate(across(stageHL, ~droplevels(factor(., levels = stg_refnd_lvls))))  %>%
      dplyr::count(Strain_DN, stageHL, .drop = FALSE)  %>%
      rename(c('Stage' = stageHL, 'Counts' =n, 'Strain' = Strain_DN))
  # 
  a_tbl %>%
      print.data.frame()
  
    a_tbl %>%
      ggplot(., aes(x = Stage, y = Counts/100, fill = Stage)) +
 geom_col() +
  geom_tile(aes(x = Stage, y = -(10/100), fill = Stage, height = 7/100)) +
  scale_fill_manual(values =  sp_fld_colors, guide = guide_legend( direction = "horizontal"  ))+
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors)))+
  geom_text(aes(label=Counts), hjust=0.5, vjust= -0.3, position = position_dodge(width = .9), size = 4.8, show.legend = F, fontface = 'bold')+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.2))) +
  scale_x_discrete(expand = expansion(add = c(1.2, 0.8))) +
  facet_grid(.~Strain) +
  theme_classic() +
  labs( y = 'Counts (X100)')+
      theme(
        axis.text.y  = element_text(size = 15),
        axis.title.y  = element_text(size = 16, face = "bold"),
        # axis.title.x = element_text(size = 22, face = "bold"),
        axis.title.x = element_blank(),
        # axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16, face = "bold"),
        legend.position="bottom",
        # plot.margin = unit(c(1,2,1,1),"cm")
        # legend.margin=margin()
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=18,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent')#,
        # legend.margin=margin(t = 0, unit='cm')
      ) +
    guides(fill = guide_legend( nrow = 1, title.position = "top"))
    # guides(fill = guide_legend( nrow = 2))
    # b_plt
})

stg_strn_distb_bar

```


#####END##!!! NEW NOTEBOOK AFTER ASSIGNMENT OF LE/HE - PUTTING CART BEFORE HORSE HERE

```{r, results='asis'}
# , fig.width=9.5

list( soupo_PCA, strn_c_mdata)  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('####', ..3,'   \n')
    a=..1$rotation %>% 
      as.data.frame() %>%
      rownames_to_column('bcode') %>% 
      left_join(..2, by = 'bcode') %>%
      # filter(!!rlang::sym(..4) != 'Doublet' & !!rlang::sym(..4) != 'Negative') %>%
      group_by(!!rlang::sym(..4), stage_ag) %>% 
      add_count(name='strn_ag_n') %>%
      # filter(strn_ag_n > 20) %>%
      ungroup() %>%
      mutate(across(!!rlang::sym(..4), ~str_replace_all(.,c('G1'='SC6','G2'='SC3','G3'='SC4', 'G4'='SC1','G5'='SC8','G6'='SC2','G7'='SC5','G8'='SC7')  ))) %>%
      ggplot(., aes(x=PC1, y=PC2, color = !!rlang::sym(..4))) + 
      geom_point(size=2) +
      scale_color_manual(values = strain_cols_n) +
      labs(color = 'Strain')+
      theme_void()+
      theme(#axis.text=element_text(size=18), 
            axis.text=element_blank(), 
            #axis.title=element_text(size=18,face="bold"), 
            axis.title=element_blank(),
            legend.text = element_text(size=22), 
            legend.title = element_text(size=22,face="bold")) + 
      guides(color = guide_legend(override.aes = list(size = 3.5))) 
    
    print(a)
    cat('\n')
    cat("\n")
  })

```


Write the stage and strain annotations for tagging souporcell bam files to allow visualization in igv
```{r, eval=FALSE}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k7_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc', 'Strain_k3_asx_v3_b_cc', 'Strain_k4_asx_v3_b_cc', 'Strain_k6_asx_v3_b_cc', 'Strain_k3_asx_m14_b_cc', 'Strain_k4_asx_m14_b_cc', 'Strain_k6_asx_m14_b_cc', 'Strain_k3_asx_m14_b_mlg_cc', 'Strain_k3_asx_lnt_m14_b_cc', 'Strain_k3_asx_m14_b_v3_ubq_cc')

strain_bam_lbls <- str_to_lower(str_remove(strain_cc_nms, '_cc'))

for (i in strain_bam_lbls) {
  for(j in c('lr_iso', 'sr_illum')){
    system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/cln_subset_bam/', i, '_bam/' ,j))
  }
}

dedup_code <- read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/dedup.info.mapped.csv', delim = '\t', col_names = F)

list(strain_bam_lbls, strain_cc_nms)  %>% 
  purrr::pmap(.,~{
    
    strn_stg <- soupo_kselect_list[[4]] %>% 
      dplyr::select(!!rlang::sym(..2)) %>% 
      filter(!is.na(!!rlang::sym(..2))) %>% 
      rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  dplyr::select(stageHL) %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>% 
      dplyr::rename(c('stage' = stageHL) )
    
    ##write barcodes and annotation for short read illumina
    strn_stg %>%
      rename_at(..2, ~str_to_lower(str_remove(.,'_cc'))) %>%
      write_csv(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/cln_subset_bam/",..1,"_bam/sr_illum/",..1,"_labs.csv"))
    
    
    ##write barcodes and annotation for long read isoseq
    strn_stg %>% 
      mutate(across(bcode, ~str_remove(.,'-1'))) %>% 
      left_join(., dedup_code, by = c('bcode' = 'X5'), multiple = "all") %>% 
      dplyr::select(X1, stage, !!rlang::sym(..2)) %>% 
      dplyr::rename( 'bcode' = X1) %>%
      rename_at(..2, ~str_to_lower(str_remove(.,'_cc'))) %>%
      write_csv(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/cln_subset_bam/",..1,"_bam/lr_iso/",..1,"_labs.csv"))
    
  }
  )

```


```{r, eval=F}
##transfer cell/molecule annotation to farm for tagging

# #FARM#make directories in farm using a loop
# # get the directory path sections to create then paste to farm
# paste0(strain_bam_lbls,'_bam','/sr_illum')
# paste0(strain_bam_lbls,'_bam','/lr_iso')
# 
# ##copy loop to farm in relevant parent directory
# for i in "strain_k4_bam/sr_illum"      "strain_k6_bam/sr_illum"      "strain_k4_s2_bam/sr_illum"   "strain_k4_s3_bam/sr_illum"   "strain_k6_cln_bam/sr_illum" "strain_k3_asex_bam/sr_illum" "strain_k4_asex_bam/sr_illum" "strain_k5_asex_bam/sr_illum" "strain_k3_gam_bam/sr_illum"  "strain_k4_gam_bam/sr_illum" "strain_k5_gam_bam/sr_illum"; do mkdir -p $i; done


for (i in strain_bam_lbls) {
  for(j in c('lr_iso', 'sr_illum')){
  system(paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/cln_subset_bam/',i,'_bam/',j,'/',i,'_labs.csv  jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/msc14/5736STDY11771546_DB52e_star/cln_subset_bam/',i,'_bam/',j,'/'
  ))
  }
}

```


```{r}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list( soupo_kmerge, strain_cc_nms)  %>% 
  purrr::pmap(.,~{
  ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  dplyr::select(stageHL) %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>%
      dplyr::count(!!rlang::sym(..2), stageHL) %>%
      print.data.frame()
})
```
```{r}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list( soupo_kmerge, str_remove(strain_cc_nms, '_cc'))  %>% 
  purrr::pmap(.,~{
  ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  dplyr::select(stageHL) %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>%
      dplyr::count(!!rlang::sym(..2), stageHL) %>%
      print.data.frame()
})
```

```{r}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list( soupo_kmerge,  str_remove(strain_cc_nms, '_cc'))  %>% 
  purrr::pmap(.,~{
  ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  dplyr::select(stage_afm) %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>%
      dplyr::count(!!rlang::sym(..2), stage_afm) %>%
      print.data.frame()
})
```



```{r, fig.height=4, fig.width=12}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list( soupo_kmerge, strain_cc_nms[11])  %>% 
  purrr::pmap(.,~{
  plt_tb <- ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>%
    filter(!is.na(stageHL)) %>% 
    # filter(!is.na(stageHL) & !!rlang::sym(..2) != 'Doublet') %>% 
    ## 2 NEGATIVES FILTERED OUT
    filter(!is.na(!!rlang::sym(..2))) %>% 
      mutate(across(!!rlang::sym(..2), ~str_replace_all(., c('G1'='SC6','G2'='SC3','G3'='SC4', 'G4'='SC1','G5'='SC8','G6'='SC2','G7'='SC5','G8'='SC7') ))) 
      
      print(plt_tb %>% filter(is.na(!!rlang::sym(..2))))
      
    plt_tb %>%
      mutate(across(stageHL, ~droplevels(factor(str_replace(., ' ', '\n'), levels = str_replace(stg_refnd_lvls, ' ', '\n'))))) %>%
      ggplot(., aes(colour = !!rlang::sym(..2), y = nFeature_RNA, x = stageHL )) +
      scale_color_manual(values = strain_cols_n)+
      geom_boxplot(alpha = 0.5, position = position_dodge(width = 1)) +
        geom_point(size = 0.7, position = position_dodge(width = 1))+ 
      labs(color = 'Strain', y = 'Gene count', x = 'Stage') +
  theme_classic()+
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size = 22, face = 'bold'),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = 'bold')) 
              
})

```



```{r, fig.height=4, fig.width=12}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list(soupo_kmerge, str_remove(strain_cc_nms[4], '_cc')) %>%
  purrr::pmap(.,~{
  plt_tb <- ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>%
    filter(!is.na(stageHL)) %>% 
    # filter(!is.na(stageHL) & !!rlang::sym(..2) != 'Doublet') %>% 
    mutate(across(!!rlang::sym(..2), ~str_replace_all(., c('G1'='SC6','G2'='SC3','G3'='SC4', 'G4'='SC1','G5'='SC8','G6'='SC2','G7'='SC5','G8'='SC7') ))) 
      
      # print(table(is.na(plt_tb$stageHL)))
      
    plt_tb %>%
      mutate(across(stageHL, ~droplevels(factor(str_replace(., ' ', '\n'), levels = str_replace(stg_refnd_lvls, ' ', '\n'))))) %>%
      mutate(across(sim_lab1, as.numeric)) %>%
      ggplot(., aes(colour = !!rlang::sym(..2), y = nFeature_RNA, x = stageHL )) +
      scale_color_manual(values = strain_cols_n)+
      geom_boxplot(alpha = 0.5, position = position_dodge(width = 1)) +
        geom_point(size = 0.7, position = position_dodge(width = 1))+ 
      labs(color = 'Strain', y = 'Gene count', x = 'Stage') +
  theme_classic()+
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size = 22, face = 'bold'),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = 'bold')) 
              
})
```


```{r, fig.height=10, fig.width=16}
# strain_cc_nms <- c('Strain_k4_cc', 'Strain_k6_cc', 'Strain_k4_s2_cc', 'Strain_k4_s3_cc', 'Strain_k6_cln_cc', 'Strain_k3_asex_cc', 'Strain_k4_asex_cc', 'Strain_k5_asex_cc', 'Strain_k3_gam_cc', 'Strain_k4_gam_cc', 'Strain_k5_gam_cc')

list( soupo_kmerge, strain_cc_nms[11])  %>% 
  purrr::pmap(.,~{
  ..1 %>% rownames_to_column('bcode') %>% 
      left_join(., msc14[[1]]@meta.data %>% 
                  rownames_to_column('bcode'), 
                by = 'bcode') %>% 
  dplyr::select(starts_with('prob'), 'stageHL', ..2) %>%
  pivot_longer(cols = starts_with('prob'), names_to = 'Machine learning predictions', values_to = 'probabilities') %>% 
  mutate(across(`Machine learning predictions`, ~str_replace_all(., c('prob_' = '', '\\.$'=')', '_\\.'= ' (', '\\.'=' ', '_'=' '))))%>% 
  mutate(across(`Machine learning predictions`, ~str_to_sentence(str_replace_all(., c('Gametocytes'='Gametocyte', 'comitted' = 'committed'))))) %>% 
  mutate(across(`Machine learning predictions`, ~droplevels(factor(., stg_refnd_lvls))))%>% 
  ggplot(., aes(y = probabilities, x = `Machine learning predictions`, fill = `Machine learning predictions`)) +
  geom_boxplot() +
  geom_point(aes(group=`Machine learning predictions`), size = 0.7, position = position_jitterdodge())+
  # geom_tile(aes(x = `Machine learning predictions`, y = -0.05, fill = `Machine learning predictions`, height = 0.06)) +
  # scale_fill_manual(values =  sp_colors, name = "MCA stage", guide = guide_legend( direction = "horizontal", 
                                                                                   #title.position = "top" 
  # ))+
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors)))+
      theme(axis.text.x=element_text(angle=45))+
  facet_grid(rows = vars(!!rlang::sym(..2)), cols =  vars(stageHL))
              
})
```

!!!!!NEED TO DECIDE ON WHERE BEST TO PUT THIS - THERES A SIMILAR CODE IN THE msc_annotation script FROM THE ORIGINAL 1ST ITERATION SOUPORCELL ASSIGNMENTS. - THIS IS THE 2nd ITERATION
!!!!!NEED NEW SCRIPT DOWNSTREAM OF THIS AND msc_annotation.rmd for DEEP STRAIN INVESTIGATION
#### Strain barcodes
Writing out barcodes for the strain information for resouporcell deu and plotting
