---
title: "MSC stage annotation using scmap and comparison with logistic regression machine learning"
date: "2023-07-05"
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(sctransform)
  library(rgl)
  library(patchwork)
  library(ggrepel)
  library(ggpubr)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(ggnewscale)
  library(tidyr)
  library(cowplot)
  library(scDblFinder)
  library(SeuratData)
  library(SeuratDisk)
  library(scmap)
  library(SingleCellExperiment)
  library(pheatmap)
  library(plotly)
  library(Hmisc)
  library(rmarkdown)
  library(RColorBrewer)
  # library(MAST)
  library(Platypus)
  library(ggvenn)
  library('biomaRt')
  library(conflicted)
  library(SoupX)
  library(MetBrewer)
  library(knitr)
  library(readxl)
  library(SCpubr)
  library(ggh4x)
  library(htmlwidgets)
  library(slingshot)
  library(harmony)
  library(ggforce)
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
})
```


# MSC annotation

## Variable declarations, read in input files and exploratory visualizations


```{r}



sp_colors <- c('Early ring'='#D1EC9F',
                     'Late ring'='#78C679',
                     'Early trophozoite'='#FEEEAA',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                     'Gametocyte (developing)'='thistle',
                     'Gametocyte (female)'='#551A8B',
                     'Gametocyte (male)'='mediumpurple',
               'Unassigned'='red', 
               'Not assigned'='#D6CFC7',
               'Female (LE)'='#208cf7',
               'Female'='#551A8B',
               # 'Female (HE)'='#551A8B',
               'Male (LE)'='#a894d1',
               'Male'='mediumpurple',
               # 'Male (HE)'='mediumpurple',
               'Failed QC' = 'grey',
               'Gametocyte (committed)' = '#C399A2', 
               'Gametocyte (developing)' = '#66444C', 
               'Gametocyte (branching)' = '#66444C', 
               'Gametocyte (early female)' = '#749E89', 
               'Gametocyte (early male)' = '#7D87B2', 
               'Gametocyte (late female)' = '#4E6D58', 
               'Gametocyte (late male)' = '#41507B',
                   NULL = 'lightgrey'
)

sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   'Female'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male'= 'mediumpurple',#'#039911',
                   ## 'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Atlas' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle',
                   'early female' = '#749E89',
                   'late female'= '#4E6D58',
                   NULL = 'lightgrey'
)

cluster_cols <- brewer.pal(7, name = 'Set2') %>% set_names(as.character(c(0:6)))



stg_lvls = c("Early ring", "Late ring", "Early trophozoite","Late trophozoite" , "Early schizont", "Late schizont",  "Gametocyte (developing)", "Female (HE)", "Female", "Female (LE)",  "Gametocyte (female)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (male)", "Unassigned", "Failed QC", "Not assigned")

stg_refnd_lvls = c("Early ring", "Late ring", "Early trophozoite", "Late trophozoite", "Early schizont", "Late schizont", "Gametocyte (committed)", "Gametocyte (developing)", "Gametocyte (branching)", "Female (HE)", "Female", "Female (LE)", "Gametocyte (female)", "Gametocyte (early female)", "Gametocyte (late female)", "Gametocyte (male)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (early male)", "Gametocyte (late male)", "Unassigned", "Failed qc", "Not assigned", "Atlas")

stg_refnd_lvls2 = c('Early ring','Late ring','Early trophozoite','Late trophozoite','Early schizont','Late schizont', "Gametocyte (committed)","Gametocyte (developing)","Gametocyte (branching)", "Female (HE)", "Female", "Female (LE)", "Gametocyte (early female)","Gametocyte (late female)", "Male (HE)", "Male", "Male (LE)", "Gametocyte (early male)","Gametocyte (late male)", "Unassigned", "Failed QC", "Not assigned") #%>% str_to_sentence(.)

# donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
# donor_cols_m = met.brewer("Thomas")[c(6,3,1,2)]  %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
donor_cols_m = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))
donor_cols = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))
# donor_cols_cap = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% str_to_upper(set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ))))


```

```{r pressure, echo=FALSE}
# ggplot(cars, aes(x= dist, y =speed)) +
#   geom_point()
# graph2ppt(file = '/Users/jr35/Google Drive/My Drive/PHD_documents/test_exp2')
```

```{r}
##Strain cluster colours

strain_cols_n <-c("#910000", "#ff1a8d", "#007c71", "#0ea300", "#00deca", "#9a96c7", "#5d57a4", "#a8875b","#708db3", "#c39a00", "#ffd94d", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","thistle","#a40000" , "#00b7a7", "#e7f1f9", 'black', 'grey')
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_n) <- c('SC2', 'SC1','SC5',  'SC4', 'SC3',  'SC6', 'SC7', 'SC8', 'S3_S1', 'S1_p_i','S1_b', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative', 'Lab', 'Less5', NA)


# strain_cols <- c(met.brewer(name = 'Austria')[c(1,7, 4,5,6, 7, 8)], 'light gray')
strain_cols <- c(met.brewer(name = 'Thomas')[c(1:8)],'black', 'light gray')
# names(strain_cols) = c('Strain 0', 'Strain 1', 'Strain 2', 'Doublet', 'Negative')
names(strain_cols) = c('G1', 'G2', 'G3', 'G4','G5', 'G6','G7', 'G8', 'Doublet', 'Negative')
# strain_cols = c("Strain 2" = "#dd5129", "Strain 1"= "#0f7ba2", "Strain 0"="#43b284", "Doublet"="#fab255")
strain_cols =c(strain_cols,strain_cols_n)

strain_dset_lvls = c( 'SC1', 'SC2','SC3',  'SC4', 'SC5',  'SC6', 'SC7', 'SC8')

```

```{r setup}
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali')
```


initialize variable for the sample names

```{r}
# sample_id <- c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" )
sample_id_lst <-as.list(c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" ) %>% 
                          str_sort(., numeric = T))
mt_filt <- c(5,1.2,1,0.45,2)
sample_name_nms<-sample_id_lst
optimum_k = c(5,2,2,8,2) %>% set_names(sample_name_nms)
```

```{r}
##Read quality controled msc objects with doublets and poor quality cell filtered out
msc_qcd <- readRDS('data/processed/DB52e_star/all_msc_filt_w_dblts.RDS')
```

```{r}
## !!! ONLY KEEP MSC14 IN THE LIST - WILL NEED TO BE UNDONE IN FUTURE

# msc_qcd <- msc_qcd[4]
# pc_list <- pc_list[4]
# mt_filt <- mt_filt[4]
# sample_id_lst <- sample_id_lst[4]
# soupo_kselect_list_pp <- soupo_kselect_list_pp[4]
```

```{r , echo=FALSE}
# FeaturePlot(msc_qcd[[1]], features = "PF3D7-0707900", dims = c(1,2), pt.size = 1.5, label = T, label.size = 5, repel = T) 
# graph2ppt(file = '/Users/jr35/Google Drive/My Drive/PHD_documents/test_exp2')
```

```{r}
##Read MCA V3 blood stage reference dataset
# EF_V3_ref <- readRDS("../sc_velo/data/raw/RoadtoUMAP/pseudotime/raw_combined_cut_0.4.RDS")
EF_V3_ref <- readRDS("data/raw/mca_v3_ref/raw_combined_cut_0.4_nodoublets_stagenew.RDS")
```

```{r}
##make refined metadata column that includes the gametocyte substages for finer resolution mapping
EF_V3_ref <- EF_V3_ref@meta.data %>%
  mutate(stage_fine = case_when(optionB_Stage == 'no assignment' ~ as.character(Stage),
                                TRUE ~ as.character(optionB_Stage))
  ) %>% 
  mutate(across(stage_fine, ~str_to_sentence(str_replace_all(., c('Gametocytes'='Gametocyte', 'comitted' = 'committed'))))) %>%
  mutate(across(stage_fine, ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
  mutate(across(stagenewx, ~droplevels(factor(str_to_sentence(.), levels = stg_refnd_lvls)))) %>% 
  dplyr::select(stage_fine, stagenewx) %>%
  AddMetaData(EF_V3_ref, .)
```


```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id) )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  dplyr::mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                        TRUE ~ as.character(.)))) 

```

```{r, fig.width = 32, fig.height = 7}
stat_box_data <- function(y) {
  return( 
    data.frame(
      y=quantile(y,probs=1)*1.4,  #may need to modify this depending on your data
      label = paste(
        'n:', length(y), '\n',
        'mu:', round(mean(y), 1)
      )
    )
  )
}
```

### Visualize reference datasets
#### MCA V3 
```{r}
##Visualize reference datasets
DimPlot(EF_V3_ref, reduction = "umap", label = T, group.by = 'stagenewx', pt.size = 2, dims = c(2,3))
```

7G8 VS NF54 distribution in UMAP
```{r}

DimPlot(EF_V3_ref, dims = c(2,3))
```

## scmap annotation

### Preprocessing references

#### MCA V3 reference dataset
```{r, message=FALSE, warning=FALSE}
##Convert to reference to sce
EF_V3_ref.sce <- as.SingleCellExperiment(EF_V3_ref)

##Place counts appropriately into single cell experiment object
SingleCellExperiment::logcounts(EF_V3_ref.sce) <- log2(SingleCellExperiment::counts(EF_V3_ref.sce) + 1)
rowData(EF_V3_ref.sce)$feature_symbol <- rownames(EF_V3_ref.sce)
EF_V3_ref.sce <- EF_V3_ref.sce[!duplicated(rownames(EF_V3_ref.sce)), ]

##scmap feature selection
EF_V3_ref.sce <- selectFeatures(EF_V3_ref.sce, suppress_plot = T)

##Cell indexing
set.seed(155)
EF_V3_ref.sce <- indexCell(EF_V3_ref.sce)
```


### Preprocessing MSC query datasets

```{r}

#Convert to SCE and subset
msc_qcd_sce <- msc_qcd %>%
  map(. %>% as.SingleCellExperiment(.))
```

```{r}

logcounts(msc_qcd_sce[[1]]) <- log2(SingleCellExperiment::counts(msc_qcd_sce[[1]]) + 1)
rowData(msc_qcd_sce[[1]])$feature_symbol <- rownames(msc_qcd_sce[[1]])
msc_qcd_sce[[1]] <- msc_qcd_sce[[1]][!duplicated(rownames(msc_qcd_sce[[1]])), ]
```

```{r}
##!!NB commenting out to retain only MSC14 - will need to revert in future
logcounts(msc_qcd_sce[[2]]) <- log2(SingleCellExperiment::counts(msc_qcd_sce[[2]]) + 1)
rowData(msc_qcd_sce[[2]])$feature_symbol <- rownames(msc_qcd_sce[[2]])
msc_qcd_sce[[2]] <- msc_qcd_sce[[2]][!duplicated(rownames(msc_qcd_sce[[2]])), ]


```

```{r}
##!!NB commenting out to retain only MSC14 - will need to revert in future
logcounts(msc_qcd_sce[[3]]) <- log2(SingleCellExperiment::counts(msc_qcd_sce[[3]]) + 1)
rowData(msc_qcd_sce[[3]])$feature_symbol <- rownames(msc_qcd_sce[[3]])
msc_qcd_sce[[3]] <- msc_qcd_sce[[3]][!duplicated(rownames(msc_qcd_sce[[3]])), ]
```

```{r}
##!!NB commenting out to retain only MSC14 - will need to revert in future
logcounts(msc_qcd_sce[[4]]) <- log2(SingleCellExperiment::counts(msc_qcd_sce[[4]]) + 1)
rowData(msc_qcd_sce[[4]])$feature_symbol <- rownames(msc_qcd_sce[[4]])
msc_qcd_sce[[4]] <- msc_qcd_sce[[4]][!duplicated(rownames(msc_qcd_sce[[4]])), ]
```

```{r, eval=T}
##!!NB commenting out to retain only MSC14 - will need to revert in future
logcounts(msc_qcd_sce[[5]]) <- log2(SingleCellExperiment::counts(msc_qcd_sce[[5]]) + 1)
rowData(msc_qcd_sce[[5]])$feature_symbol <- rownames(msc_qcd_sce[[5]])
msc_qcd_sce[[5]] <- msc_qcd_sce[[5]][!duplicated(rownames(msc_qcd_sce[[5]])), ]
```

### Stage label assignment (projection)
transferring labels from both reference datasets
```{r}
##Projection
msc_qcd_jc <- msc_qcd_sce %>%
  map(. %>% scmapCell(.,list(
    Elias_V3 = metadata(EF_V3_ref.sce)$scmap_cell_index
  )
  )
  )
```

Assigning stage labels only when the top 3 most closest reference cells to a query cell all have a similar stage label and a cosine similarity greater than 0.4
```{r}
##Assigning stage labels using thresholds
msc_qcd_lbls <- msc_qcd_jc %>%
  map(. %>% scmapCell2Cluster(.,list(
    as.character(SummarizedExperiment::colData(EF_V3_ref.sce)$stagenewx)
  ),w = 3, threshold = 0.4)
  )
```

Obtaining the reference cell that is closest to each query cell in neighbourhood space regardless of cosine similarity. For those where the top 10 most closest cells constitute several stages, we also obtain the second most similar stage.


```{r, eval=T}
## This chunk takes long to process so we avoid evaluating and read the saved output objects from file in the next chunk
##Ensure that none of the objects in the list are not empty ie all samples should have at least one assigned more than one stage
# sc_cells <- list(msc_qcd_jc) %>%
#   pmap(~ ..1$Elias_V3$cells %>% 
#          as.data.frame()  %>%
#          mutate(across(everything(), ~as.character(SummarizedExperiment::colData(EF_V3_ref.sce)$stagenewx)[.])) %>% 
#          dplyr::select(where(~n_distinct(.) > 1))
#   )
# 
# scmap_cells_sim <- list(msc_qcd_jc[unlist(map(sc_cells, ~ncol(.) > 0))], sc_cells[unlist(map(sc_cells, ~ncol(.) > 0))]) %>%
#   pmap(~ ..1$Elias_V3$similarities %>% 
#          as.data.frame()  %>%
#          dplyr::select(colnames(..2)) %>%
#          mutate(order_c = row_number()) %>%
#          mutate(across(everything(), ~as.character(.))) %>%
#          bind_cols(..2)  %>% 
#          pivot_longer(cols = -c(order_c),names_to = 'cell_ids', values_to = 'lab_or_sim') %>% 
#          mutate(across(cell_ids, ~str_remove(., '...[0-9]*$'))) %>% 
#          mutate(state = case_when(str_detect(lab_or_sim, '[0-9]') ~ 'sim', T ~ 'stage'))  %>% 
#          pivot_wider(names_from = 'state', values_from = 'lab_or_sim' ) %>% 
#          group_by(cell_ids, stage) %>% 
#          slice_min(order_c) %>% 
#          arrange(cell_ids) %>% 
#          ungroup(stage)  %>% 
#          add_count() %>% 
#          slice_min(order_c, n=2) %>% 
#          arrange(order_c)%>% 
#          mutate(assn = c('lab1', 'lab2')) %>%
#          pivot_wider(names_from = assn, values_from = c('order_c', 'stage', 'sim'))%>%
#          column_to_rownames("cell_ids") %>% 
#          dplyr::select(-n)
#   )
# 
# ##Add the empty element into the position with list lacking
# scmap_cells_sim_nw <- ifelse(T %in% unlist(map(sc_cells, ~ncol(.) == 0)), 
#                              list(append(scmap_cells_sim, list(NULL), which(unlist(map(sc_cells, ~ncol(.) == 0))) -1)), 
#                              list(scmap_cells_sim)) %>%
#   unlist(., recursive = F)
# 
# 
# ##Get stages for cell that have similar stage assignment by scmap
# sc_cells_hom <- list(msc_qcd_jc) %>%
#   pmap(~ ..1$Elias_V3$cells %>% 
#          as.data.frame()  %>%
#          mutate(across(everything(), ~as.character(SummarizedExperiment::colData(EF_V3_ref.sce)$stagenewx)[.])) %>% 
#          dplyr::select(where(~n_distinct(.) == 1)) %>%
#          slice_head(., n = 1)
#   )
# 
# ##Get similaraties for cells that have similar stage assignment by scmap and row bind
# scmap_cs <- list(msc_qcd_jc, sc_cells_hom, scmap_cells_sim_nw) %>%
#   pmap(~ ..1$Elias_V3$similarities %>% 
#          as.data.frame()  %>%
#          dplyr::select(colnames(..2)) %>%
#          slice_head(., n = 1) %>%
#          mutate(across(everything(), ~as.character(.))) %>%
#          bind_rows(.,..2) %>% 
#          t() %>% 
#          as.data.frame() %>% 
#          setNames(c('sim_lab1','stage_lab1')) %>%
#          bind_rows(., ..3)
#   )

##Get the top V3 cell and top 3 similarity scores that is most similar to each MSC14 cell 
sc_topv3_cells_sims_all <- list(msc_qcd_jc) %>%
    pmap(~{
      ..1$Elias_V3$similarities %>% 
        .[1:3,] %>% 
        t() %>% 
        data.frame() %>% 
        rename_all(., ~str_replace(., 'X', 'sim_lab')) %>% 
        rownames_to_column('bcode') %>%
        left_join(., ..1$Elias_V3$cells %>% 
                    .[1:3,] %>% t() %>% 
                    data.frame() %>% 
                    rename_all(., ~str_replace(., 'X', 'v3_ref_cells')) %>% 
                    rownames_to_column('bcode'),
                  by = 'bcode') %>% 
        mutate(across(starts_with('v3_ref_cells'), ~colnames(EF_V3_ref.sce)[.]))
    })

saveRDS(sc_topv3_cells_sims_all, 'data/processed/DB52e_star/sc_topv3_cells_sims_all.RDS')
```

```{r}
##Add the empty element into the position with list lacking
# scmap_cells_sim_nw <- readRDS('data/processed/MSC14/5736STDY11771546_DB52e_star/scmap_cells_sim_nw.RDS')

##Get similarities for cells that have similar stage assignment by scmap and row bind
# scmap_cs <- readRDS('data/processed/MSC14/5736STDY11771546_DB52e_star/scmap_cs.RDS')

##Get similarities for cells that have similar stage assignment by scmap and row bind
sc_topv3_cells_sims_all <- readRDS('data/processed/DB52e_star/sc_topv3_cells_sims_all.RDS')

```

Adding annotations to MSC query datasets
```{r}
## Adding annotations to query cells
stg_anots <- EF_V3_ref@meta.data[,c('stagenewx'), drop=F]

## scmap merge the top labels and similarity scores
top_stg_sim <- map(sc_topv3_cells_sims_all,
~left_join(.x, stg_anots %>% rownames_to_column('v3_ref_cells1') %>% rename('stage_lab1' = stagenewx), by = 'v3_ref_cells1') %>%
                       left_join(stg_anots %>% rownames_to_column('v3_ref_cells2') %>% rename('stage_lab2' = stagenewx), by = 'v3_ref_cells2') %>%
                       left_join(stg_anots %>% rownames_to_column('v3_ref_cells3') %>% rename('stage_lab3' = stagenewx), by = 'v3_ref_cells3') %>% 
  column_to_rownames('bcode')
)


# msc_qcd_anotd <- list(msc_qcd_lbls, msc_qcd_jc, msc_qcd, scmap_cs, sc_topv3_cell_id) %>%
msc_qcd_anotd <- list(msc_qcd_lbls, msc_qcd_jc, msc_qcd, top_stg_sim) %>%
  pmap(~ ..1$scmap_cluster_labs %>% 
         data.frame()  %>% 
         mutate(cell_bc = colnames(..2$Elias_V3$cells))%>%
         column_to_rownames("cell_bc") %>%
         dplyr::rename("Stage_V3" = Elias_V3)%>% 
         mutate(across(Stage_V3, ~droplevels(factor(str_to_sentence(.), levels = stg_refnd_lvls)))) %>%
         AddMetaData(..3, .) %>%
         AddMetaData(., ..4) #%>%
         # AddMetaData(., ..5)
       )

```


Glimpse the MSC datasets
```{r}
# rm(msc_qcd)
msc_qcd_anotd
```

```{r}
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(Stage_V3, stage_lab1, stage_lab2, stage_lab3) )
```

### Visualization of scmap annotation

#### MCA V3 labels split by strain {.tabset}

```{r, results='asis', fig.width = 10, fig.height = 4}
# cat('## Visualize field parasites {.tabset}   \n')
sample_id_lst<-sample_id_lst
list(msc_qcd_anotd, sample_id_lst)  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p <- DimPlot(..1, reduction = "umap", split.by = 'Strain_DN', group.by = 'Stage_V3', pt.size = 1.1)  + 
      theme(axis.text=element_text(size=18), 
            axis.title=element_text(size=20,face="bold"), 
            # legend.text = element_blank(), 
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )
    print(p)
    cat('\n  ')
    cat("\n")
  })

```


### sMap cosine similarities visualization
scmap cosine similarities for the top 2 assigned stage labels among the top 10 stage labels  


### Further MSC14 annotation adjustments given visualization above 
For those query cells where the top 3 most similar cells constitute several stages we assign the top stage based as long as it has a cosine similarity greater than 0.4. We notice that a majority of unassigned cells have both Early trophozoite and Late ring as the among top 3 similar cells and this leads to them not getting assigned stages as the top 3 are not homogeneous as is the requirement.


```{r}
##Modify the stage labels for MSC14
# msc_qcd_anotd[[4]] <- msc_qcd_anotd[[4]]@meta.data %>%
msc_qcd_anotd <- pmap(list(msc_qcd_anotd, c(rep(0.4,4), 0.4), c(rep('Late ring|Early trophozoite|developing', 4),'ring|trophozoite|developing|schizont')), ~{
# msc_qcd_anotd <- pmap(list(msc_qcd_anotd, rep(0.4,4), rep('Late ring|Early trophozoite|developing', 4)), ~{
  ..1@meta.data %>%
  mutate(across(Stage_V3, as.character)) %>%
  # mutate(Stage_V3 = Stage_V3) %>%
  mutate(StageMCA = case_when(
    Stage_V3 == 'Unassigned' & sim_lab1 > ..2 & str_detect(stage_lab1, ..3) & str_detect(stage_lab2, ..3) & str_detect(stage_lab3, ..3) ~ stage_lab1,
    Stage_V3 == 'Unassigned' & sim_lab1 > ..2 & str_detect(stage_lab1, '\\(male|developing') & str_detect(stage_lab2, '\\(male|developing') & str_detect(stage_lab3, '\\(male|developing') ~ stage_lab1,
    Stage_V3 == 'Unassigned' & sim_lab1 > ..2 & str_detect(stage_lab1, 'female|developing') & str_detect(stage_lab2, 'female|developing') & str_detect(stage_lab3, 'female|developing') ~ stage_lab1,
    sim_lab1 <= ..2 ~ 'Unassigned',
    TRUE ~ Stage_V3
  ) ,
  ) %>%
  dplyr::select(Stage_V3, StageMCA) %>%
  # mutate(across(starts_with('Stage_V'), ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
  AddMetaData(.x, .)
})

```


```{r}
map(msc_qcd_anotd, ~.@meta.data %>% filter(sim_lab1 <0.39) %>% dplyr::select(Stage_V3,stage_lab1,stage_lab2, stage_lab3, StageMCA, sim_lab1, sim_lab2, sim_lab3) )
```

```{r}
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(Stage_V3,stage_lab1,stage_lab2, stage_lab3, StageMCA) )
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(StageMCA) )
```

```{r}
##Modify the stage labels for MSC14
# msc_qcd_anotd[[4]] <- msc_qcd_anotd[[4]]@meta.data %>% 
# msc_qcd_anotd[[1]] <- msc_qcd_anotd[[1]]@meta.data %>% 
#   mutate(across(StageMCA, as.character)) %>%
#   mutate(Stage_V3 = StageMCA) %>%
#   mutate(StageMCA = case_when(#StageMCA == 'Unassigned' & sim_lab1 > 0.4 & str_detect(stage_lab1, 'Late ring|Early trophozoite') & str_detect(stage_lab2, 'Late ring|Early trophozoite') ~ stage_lab1,
#                               StageMCA == 'Unassigned' & sim_lab1 > 0.4 & str_detect(stage_lab2, 'Late ring|Early trophozoite|Gametocyte (developing)') & str_detect(stage_lab2, 'Late ring|Early trophozoite|Gametocyte (developing)') ~ stage_lab1,
#                               StageMCA == 'Unassigned' & sim_lab1 > 0.4 & str_detect(stage_lab1, '\\(male|developing') & str_detect(stage_lab2, '\\(male|developing') ~ stage_lab1,
#                               StageMCA == 'Unassigned' & sim_lab1 > 0.4 & str_detect(stage_lab1, 'female|developing') & str_detect(stage_lab2, 'female|developing') ~ stage_lab1,
#                               TRUE ~ StageMCA
#   ) ,
#   ) %>% 
#   dplyr::select(Stage_V3, StageMCA) %>%
#   mutate(across(starts_with('Stage_V'), ~droplevels(factor(., levels = stg_refnd_lvls)))) %>%
#   AddMetaData(msc_qcd_anotd[[1]], .)

```


Assign stage 2
```{r}
##Modify the stage labels for msc14
# msc_qcd_anotd[[4]] <- msc_qcd_anotd[[4]]@meta.data %>% 
# msc_qcd_anotd[[1]] <- msc_qcd_anotd[[1]]@meta.data %>%
#   mutate(StageMCA=case_when(StageMCA == 'Gametocyte (developing)' & sim_lab1 > 0.4 & str_detect(stage_lab2, 'Late ring|Early trophozoite') ~ stage_lab2,TRUE ~ StageMCA)) %>%
#   AddMetaData(msc_qcd_anotd[[1]], .)
# AddMetaData(msc_qcd_anotd[[4]], .)
```


```{r}
# msc_qcd_anotd[[1]]@meta.data %>% dplyr::count(Stage_V3,stage_lab1,stage_lab2, StageMCA, StageMCA)
```


```{r}
msc_qcd_anotd
```

```{r, fig.width = 13, fig.height = 5.5}
## Plotting top two scmap stages among the top 10 with their similarities for msc14

# map(msc_qcd_anotd[c(1,3:5)], ~{
map(msc_qcd_anotd[c(1,3,4)], ~{
  tryCatch({
.x@meta.data %>%
  # filter(Stage_V3 == 'Unassigned') %>% 
  dplyr::select(matches('s.*lab.'), StageMCA) %>%
  rownames_to_column('cell_bc')%>%
  pivot_longer(-c('cell_bc', 'StageMCA'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>%
  filter(!is.na(sim)) %>%
  mutate(across(sim, as.numeric)) %>%
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  ggplot(., aes(y = sim, x = stage, color = Label)) +
  geom_boxplot()+
  geom_point(aes(group=Label), size = 0.7, position = position_jitterdodge(jitter.width = 0.2))+
  facet_grid(.~StageMCA) +
  geom_hline(yintercept = 0.45)+
  labs(y = 'scmap similarity scores', ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4),
        axis.title=element_text(size=20,face="bold"),
        axis.text = element_text(size=16),
        strip.text =element_text(size=18,face="bold"),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=16))
  }, error = function(e) {
      message(as.character('no unassigned cells')) # exclude this line to return NULL quietly
      return(NULL)
    })
})
```






#### Assess and remove the MSC14 unassigned cells
Unassigned cells at this stage seem to be of poor quality not eliminated by QC
```{r}
# msc_qcd_anotd[[4]]@meta.data %>% filter(StageMCA != 'Unassigned') %>% dplyr::select(sim_lab1, stage_lab1, order_c_lab2, stage_lab2, sim_lab2)
# msc_qcd_anotd[[1]]@meta.data %>% filter(StageMCA != 'Unassigned') %>% dplyr::select(sim_lab1, stage_lab1, order_c_lab2, stage_lab2, sim_lab2)
```

Cells not assigned stage labels cells to be removed - one seems to be schizont
```{r}
# msc_qcd_anotd[[1]]@meta.data %>% filter(StageMCA == 'Unassigned') %>% dplyr::select(sim_lab1, stage_lab1, order_c_lab2, stage_lab2, sim_lab2)
```


### Visualize unassigned in all annotated MSC

#### UMAP & PCA coloured by stage {.tabset}
```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
list(msc_qcd_anotd, sample_id_lst)  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p_umap <- DimPlot(., reduction = "umap", cells.highlight = .@meta.data %>% filter(StageMCA == 'Unassigned') %>% row.names(), sizes.highlight = 1)  + 
      theme(axis.text=element_text(size=18), 
            axis.title=element_text(size=20,face="bold"), 
            # legend.text = element_blank(), 
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )
    print(p_umap)
    cat("\n")
    
    p_umap <- DimPlot(., reduction = "umap", cells.highlight = .@meta.data %>% filter(StageMCA == 'Gametocyte (developing)') %>% row.names(), sizes.highlight = 1)  + 
      theme(axis.text=element_text(size=18), 
            axis.title=element_text(size=20,face="bold"), 
            # legend.text = element_blank(), 
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )
    print(p_umap)
    cat("\n")
    
    p_pca <- DimPlot(., reduction = "pca",  cells.highlight = .@meta.data %>% filter(StageMCA == 'Unassigned') %>% row.names(), sizes.highlight = 1)  + 
      theme(axis.text=element_text(size=18), 
            axis.title=element_text(size=20,face="bold"), 
            # legend.text = element_blank(), 
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )
    
    print(p_pca)
    cat('\n')
    cat("\n")
    cat("\n")
  })

```
#### View and remove cells that we couldn't confidently annotate with scmap

```{r}
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(StagePrelim, StageMCA) )
```


Umap with stage labels in legend

```{r, fig.width = 7, fig.height = 5}
# Assign the MCA cluster of the projected cell to the nearest cell in the reference dataset
## NB - some of the query cells share one top cell in the reference

for (i in 1:length(msc_qcd_anotd)){
  # make a vector of ""NA" - not assigned with length of MCA
  mca_mapings_v <- rep("NA", ncol(EF_V3_ref))
  
  # Get only unique top cells in the reference since some will be shared by several msc cells
  uniq_labs <- msc_qcd_jc[[i]]$Elias_V3$cells[1,][!duplicated(msc_qcd_jc[[i]]$Elias_V3$cells[1,])]
  
  # Ensure unique reference cells are also in the query cells since we will be mapping 1-to-1
  uniq_labs <- uniq_labs[as.character(colnames(msc_qcd_anotd[[i]]))]
  uniq_labs <- uniq_labs[!is.na(uniq_labs)]
  print(length(uniq_labs))
  
  # Assign the values of interest from the lab to the relevant top reference cells in the lab leaving replacing "NA"
  mca_mapings_v[uniq_labs] <- as.character(msc_qcd_anotd[[i]][, names(uniq_labs)]$old_seu_cluster)
  print(table(mca_mapings_v))
  
  # Assign this to a new column in the reference metadata for plotting
  EF_V3_ref@meta.data[, paste0("stage_", names(msc_qcd_anotd[i]))] <-  mca_mapings_v
}

```


```{r, fig.width = 7, fig.height = 5}
trunc_axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(2.2, "cm")
)

labs_order <- c('0', '1','2',  '3', '4', '5','6',  'Atlas', 'NA')
names(labs_order) <- labs_order

## Counts of QC Pass
don_n <- map(msc_qcd_anotd, ~dim(.x)[2])

umap_theme <- theme(legend.position = 'right',
        axis.title=element_text(size=12,face="bold",hjust = 0),
        # plot.title = element_blank(),
        legend.title=element_text(size=14,face="bold"),
        legend.text=element_text(size=14),
        legend.spacing.y = unit(0.25, "cm"),
        axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed"))
        )  


# Idents(EF_V3_ref) <- 'MSC14_V3'
my21_mca_umap <- map2(names(msc_qcd_anotd), don_n, ~DimPlot(EF_V3_ref, reduction = "umap",  group.by = paste0("stage_",.x), pt.size = 1, label = T, label.size = 5, repel = T, dims = c(2,3), 
        order = labs_order) +
  scale_color_manual(breaks=labs_order, values = c(cluster_cols, "NA" = "lightgrey")[labs_order], labels = str_wrap(labs_order, 10))+ 
    geom_text(x=0.5, y=-2, label=paste0('N=', .y), size = 5) + 
    umap_theme+
    theme(plot.title = element_blank())+
  guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
)

##Save image for manuscript fig 4
# saveRDS(m14_scmap_mca_umap, 'plots/fig4/m14_scmap_mca_umap')

# my21_mca_umap


```



```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14
scm_barpt_unass <- list(msc_qcd_anotd[1:4]) %>%
pmap(.,~{
  
# stage_fct_labs <- ..1@meta.data %>% select(-c(stage_lab3,sim_lab3))%>% dplyr::group_by(StageMCA, seurat_clusters) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(StageMCA, seurat_clusters), as.character), across(seurat_clusters, ~paste0("Cluster ",., ':\n',StageMCA))) %>% deframe()

  scm_b_tbl <- ..1@meta.data %>% 
  mutate(across(stage_lab1, ~droplevels(factor(., stg_refnd_lvls)))) 

stg_vec <- levels(scm_b_tbl$stage_lab1)
  
  scm_b_tbl %>% 
  ggplot(aes(x = stage_lab1, y = sim_lab1)) + 
  geom_boxplot()+
  geom_point() +
  new_scale_fill()+
  geom_tile(data = data.frame(x = stg_vec), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
                    # guide = 'none'
  ) +
  # scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  # facet_grid(.~StageMCA, labeller = labeller(StageMCA = stage_fct_labs)) +
  facet_wrap(.~old_seu_cluster, nrow = 1) + 
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')
})
scm_barpt_unass
```



```{r, fig.width = 12, fig.height = 4}
## Prepare table for bar plot
md_don_an <- map(msc_qcd_anotd[1:4], ~.@meta.data #%>% 
                       # mutate(across(c(sim_lab1, sim_lab2), as.numeric))
                     ) %>% 
  bind_rows(., .id = 'donor') %>%
  mutate(across(donor, ~factor(., levels = as.character(sample_id_lst))),
         across(stage_lab1, droplevels))

dd <- md_don_an %>% select(donor, old_seu_cluster)  %>%  distinct() %>% arrange(donor, old_seu_cluster) %>%left_join(., data.frame(cluster_cols) %>% rownames_to_column("old_seu_cluster")) 

strip_cols_gg4hx <- dd%>% select(-c(old_seu_cluster)) %>% pivot_wider(names_from = donor, values_from = cluster_cols, values_fn = list) %>% as.list() %>% flatten() 
strip_cols_gg4hx <- as.character(c(names(strip_cols_gg4hx), unlist(strip_cols_gg4hx))) 


strip_cols_gg4hx[str_detect(strip_cols_gg4hx, "MS*")] <- "white"

```



```{r, fig.width = 12, fig.height = 4}

stg_vec <- levels(md_don_an$stage_lab1)

don_stg_sims_bar <- md_don_an %>% 
  ggplot(aes(x = stage_lab1, y = sim_lab1)) + 
  # scale_color_manual(values =  sp_colors, guide = "none")+
  geom_sina(size = 0.01) +
  geom_boxplot(outlier.shape = NA)+
  geom_tile(aes(x = stage_lab1, y = 0.34, fill = stage_lab1, height = 0.05)) +
  # scale_fill_manual(values =  sp_colors, name = "MCA stage",
  #                   guide = guide_legend(direction = "horizontal", title.position = "top", nrow = 1)
  #                   # guide = 'none'
  # ) +
  scale_fill_manual(values =  sp_colors, guide= "none") +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors))) +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x", strip = strip_nested(background_x = elem_list_rect(fill =  strip_cols_gg4hx))) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')


##Plotting top two scmap stages among the top 10 with their similarities for msc14
stg_vec <- levels(md_don_an$stage_lab1)

 md_don_an %>% 
  ggplot(aes(x = stage_lab2, y = sim_lab2, color = stage_lab2)) + 
  scale_color_manual(values =  sp_colors, guide = "none")+
  geom_sina(size = 0.01) +
  geom_boxplot(outlier.shape = NA)+
  geom_tile(aes(x = stage_lab1, y = 0.34, fill = stage_lab1, height = 0.05)) +
  scale_fill_manual(values =  sp_colors, guide= "none") +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors))) +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x", strip = strip_nested(background_x = elem_list_rect(fill =  strip_cols_gg4hx))) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')

don_stg_sims_bar
```


#### Gold standard gam and asexual markers from Lisette Meerstein-Kessel et al {.tabset}


```{r, message=FALSE}
kes_gam_m <- read_xlsx('data/raw/published_lit/kessel_markers_2018/41598_2017_18840_MOESM2_ESM.xlsx', sheet = 'Gametocytes-Positive set') %>% mutate(across(gene_ID, ~str_replace(., '_', '-')))
kes_asx_m <- read_xlsx('data/raw/published_lit/kessel_markers_2018/41598_2017_18840_MOESM2_ESM.xlsx', sheet = 'Asexual Stages-Negative Set') %>% mutate(across(gene_ID, ~str_replace(., '_', '-')))

lora_batille_egam_asex <- read_xlsx('data/raw/published_lit/lora_batille_2020/de_genes_file_s1.xlsx', sheet = 'fig5_hmap') %>% mutate(across(Gene_id, ~str_replace(., '_', '-')))
```


```{r, results='asis', message=FALSE, warning=FALSE}
##!!!NOTE - Stage legend plotted, cropped and then turned off - manual modification
# dot_pts <- list()

mk_gns <- data.frame("gene_id" = c(kes_gam_m[1:3,]$gene_ID,kes_asx_m[1:3,]$gene_ID,"PF3D7-0935900",  "PF3D7-0406200", "PF3D7-1031000", "PF3D7-0905300") ) %>% left_join(Pf3D7_genes_plasmoDB[, c("gene_id", "Gene")], by = "gene_id")

mark_dot_plts<- msc_qcd_anotd[1:4]  %>% 
  purrr::imap(.,~{
    # create tabset for each group 
    cat('###### ',.y,'   \n')
    cat(' \n Average gametocyte markers expression \n')
    # cat('\n')
    seu_obj<- .x
    nm<- .y
    print(seu_obj)
    dot_pts <- map(list(mk_gns$gene_id), 
                   ~do_DotPlot(seu_obj, features = .x, group.by = 'old_seu_cluster') +
                     scale_x_discrete(labels= mk_gns$Gene) +
                     scale_fill_gradientn(colours = brewer.pal(n = 9, name = "Blues"), na.value = 'grey50', name = "Avg. Expression", guide = guide_colorbar(title.position = "top", title.hjust = 0.5, order = 2, frame.colour ="grey50", ticks.colour = "white"))+
                     # new_scale_fill() +
                     # geom_tile(data = data.frame(Stage = factor(levels(seu_obj@meta.data$stageHL), levels = levels(seu_obj@meta.data$stageHL))), inherit.aes = F, aes(x=0, y=Stage, fill =Stage)) +
                     labs(title = nm) +
                     theme(axis.text.y = element_text(size = 16),
                           axis.text.x = element_text(size = 16),
                           # axis.ticks.x = element_blank(),
                           legend.text = element_text(size = 14), 
                           legend.title = element_text(size = 14, face = 'bold'),
                           legend.position = "bottom", 
                           legend.box = "horizontal",
                           legend.spacing.x = unit(0.02, "cm"),
                           legend.key.width = unit(0.4, "cm"),
                           legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),
                           plot.title = element_text(hjust = 0.5)
                     ) +
                     # scale_fill_manual(values = sp_fld_colors, name = "Stage", guide = guide_legend(nrow = 2,  title.position = "top", order = 1)) + 
                     ##!!NOTE - turning off stage label after screen shotting - switch back on if needed
                     # guides(fill=F, size = guide_legend(title = "% Expressed", title.position = "top")) +
                     coord_flip()
    )
    print(dot_pts)
    
    
    return(dot_pts)
    
    cat("\n")
    cat("\n")
  })

# mark_dot_plts


```


```{r, results='asis', message=FALSE, warning=FALSE}
##!!!NOTE - Stage legend plotted, cropped and then turned off - manual modification
# dot_pts <- list()

mark_dot_plts2_4<- msc_qcd_anotd[2:4]  %>% 
  purrr::imap(.,~{
    # create tabset for each group 
    cat('###### ',.y,'   \n')
    cat(' \n Average gametocyte markers expression \n')
    # cat('\n')
    seu_obj<- .x
    nm<- .y
    print(seu_obj)
    dot_pts <- map(list(mk_gns$gene_id), 
                   ~do_DotPlot(seu_obj, features = .x, group.by = 'old_seu_cluster') +
                     scale_x_discrete(labels= mk_gns$Gene) +
                     scale_fill_gradientn(colours = brewer.pal(n = 9, name = "Blues"), na.value = 'grey50', name = "Avg. Expression", guide = guide_colorbar(title.position = "top", title.hjust = 0.5, order = 2, frame.colour ="grey50", ticks.colour = "white"))+
                     # new_scale_fill() +
                     # geom_tile(data = data.frame(Stage = factor(levels(seu_obj@meta.data$stageHL), levels = levels(seu_obj@meta.data$stageHL))), inherit.aes = F, aes(x=0, y=Stage, fill =Stage)) +
                     labs(title = nm)+
                     theme(axis.text.y = element_blank(),
                           axis.text.x = element_text(size = 16),
                           # axis.ticks.x = element_blank(),
                           legend.text = element_text(size = 14), 
                           legend.title = element_text(size = 14, face = 'bold'),
                           legend.position = "bottom", 
                           legend.box = "horizontal",
                           legend.spacing.x = unit(0.02, "cm"),
                           legend.key.width = unit(0.4, "cm"),
                           legend.box.margin = margin(t = 0, r = 4.1, b = 0, l = 0, unit = "cm"),
                           plot.title = element_text(hjust = 0.5)
                     ) +
                     # scale_fill_manual(values = sp_fld_colors, name = "Stage", guide = guide_legend(nrow = 2,  title.position = "top", order = 1)) + 
                     ##!!NOTE - turning off stage label after screen shotting - switch back on if needed
                     # guides(fill=F, size = guide_legend(title = "% Expressed", title.position = "top")) +
                     coord_flip()
    )
    print(dot_pts)
    
    
    return(dot_pts)
    
    cat("\n")
    cat("\n")
  })

# mark_dot_plts


```

```{r}
map(msc_qcd_anotd, ~.@meta.data %>% filter(StageMCA == "Unassigned") %>% dplyr::count(StagePrelim, StageMCA) )
```

```{r}
##Remove cells that aren't asigned labels by scmap
msc_qcd_anotd <- map(msc_qcd_anotd, ~subset(.x, subset = StageMCA != 'Unassigned'))
# msc_qcd_anotd[3:4] <- map(msc_qcd_anotd[3:4], ~subset(.x, subset = StageMCA != 'Gametocyte (developing)'))

# msc_qcd_anotd[[4]] <- subset(msc_qcd_anotd[[4]], subset = StageMCA != 'Gametocyte (developing)')
```


```{r}
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(StagePrelim, StageMCA) )
```
### Visualize all annotated MSC

#### UMAP & PCA coloured by old seurat clusters before QC {.tabset}
The old seurat clusters metadata are maintained to trace identity with the clusters before QC and which clusters had the most poor QC cells
```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
# list(msc_qcd_anotd, sample_id_lst)  %>% 
#   purrr::pmap(.,~{
#     # create tabset for each group 
#     cat('##### ',..2,'   \n')
#     p_umap <- DimPlot(., reduction = "umap",  group.by  = 'old_seu_cluster'#, pt.size = 1, label=T, cells.highlight = .@meta.data %>% filter(seurat_clusters == '2') %>% row.names(), sizes.highlight = 1
#     )  + 
#       theme(axis.text=element_text(size=18), 
#             axis.title=element_text(size=20,face="bold"), 
#             # legend.text = element_blank(), 
#             plot.title = element_blank(),
#             strip.text.x = element_text(size = 24, face = "bold") )
#     print(p_umap)
#     cat("\n")
#   })

```


```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14
scm_barpt_s <- list(msc_qcd_anotd[1:4]) %>%
pmap(.,~{
  
stage_fct_labs <- ..1@meta.data %>% select(-c(stage_lab3,sim_lab3))%>% dplyr::group_by(StageMCA, old_seu_cluster) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(StageMCA, old_seu_cluster), as.character), across(old_seu_cluster, ~paste0("Cluster ",., ':\n',StageMCA))) %>% deframe()



scm_b_tbl <- ..1@meta.data %>% 
  select(-c(stage_lab3,sim_lab3))%>% 
  dplyr::select(matches('s.*lab.'), StageMCA) %>% 
  rownames_to_column('cell_bc')%>%
  pivot_longer(-c('cell_bc', 'StageMCA'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>% 
  filter(!is.na(sim)) %>%
  mutate(across(sim, as.numeric)) %>% 
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls)))) 

stg_vec <- levels(scm_b_tbl$stage)

scm_b_tbl %>%
  ggplot(., aes(y = sim, x = stage, color = Label)) +
  geom_boxplot(aes(fill = Label))+
  scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 2)
  ) +
  scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
  geom_point(aes(group=Label), size = 0.05, position = position_jitterdodge(jitter.width = 0.2))+
  new_scale_fill()+
  geom_tile(data = data.frame(x = stg_vec), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
                    # guide = 'none'
  )+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  facet_grid(.~StageMCA, labeller = labeller(StageMCA = stage_fct_labs)) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')
})
scm_barpt_s
```


### Annotate female/male low-expression and high-expression clusters using plots above
Use the original clusters for msc14 to allow for concordance in the paper

```{r}
# msc_qcd_anotd[1] %>%
#   map(~.@meta.data  %>% 
#         dplyr::count(seurat_clusters, old_seu_cluster) 
#   )
```

```{r}
umap_stg_mca <- map(msc_qcd_anotd[1:4], ~DimPlot(.x, reduction = "umap", group.by = "StageMCA")+
      scale_color_manual(values = sp_colors) +
      theme(axis.text=element_blank(), 
            axis.title=element_blank(),
            axis.ticks=element_blank(),
            legend.position = "None",
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )
)
    
umap_stg_mca
```


#### Violin plots
```{r, fig.height = 4, fig.width = 10}
# options(repr.plot.width=32, repr.plot.height=7)
fct_theme <- theme(text=element_text(size=14),
        axis.text=element_text(size=20), 
            axis.title=element_text(size=20,face="bold"), 
            legend.text = element_text(size=20), 
            legend.title = element_text(size=20,face="bold"),
        strip.text =element_text(size=22,face="plain", colour = 'black'))

msc_qcd_mdata <- map(msc_qcd_anotd[1:4], ~.@meta.data) %>% 
  bind_rows(., .id = 'sample') %>% 
  mutate(across(c(sample), ~factor(., levels = unlist(sample_id_lst)))) 
  

viol_gn_cnt <- msc_qcd_mdata %>% 
  ggplot(aes(x = seurat_clusters, color = StageMCA, y = nFeature_RNA)) +
  scale_color_manual(values = sp_colors)+
  theme_classic() +
  geom_violin() +
  ggforce::geom_sina(size = 0.01)+
  facet_wrap(.~sample, ncol = 4) +
  fct_theme+
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = 'Cluster',override.aes = list(size = 4), nrow = 1))

viol_gn_cnt
```


#### UMAP & PCA coloured by new seurat clusters after QC and stage {.tabset}
The old seurat clusters metadata are maintained to trace identity with the clusters before QC and which clusters had the most poor QC cells
```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
std_theme <- theme(axis.text=element_text(size=18), 
            axis.title=element_text(size=20,face="bold"), 
            # legend.text = element_blank(), 
            plot.title = element_blank(),
            strip.text.x = element_text(size = 24, face = "bold") )

list(msc_qcd_anotd, sample_id_lst)  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p_umap <- DimPlot(., reduction = "umap",  group.by  = 'old_seu_cluster', pt.size = 1, label=T)  + 
      std_theme
    print(p_umap)
    cat("\n")
    
    p_umap <- DimPlot(., reduction = "pca",  group.by  = 'old_seu_cluster', pt.size = 1, label=T)  + 
      std_theme
    print(p_umap)
    
    p_umap <- DimPlot(., reduction = "pca",  group.by  = 'StageMCA', pt.size = 1, label=T)  + 
      std_theme
    print(p_umap)
    
    
    cat('\n')
    p_umap <- DimPlot(., reduction = "pca", dims=c(1,2),  group.by  = 'seurat_clusters', pt.size = 1, label=T)  + 
      std_theme
    print(p_umap)
    
    cat('\n')
    
    cat("\n")
  })

```

#### Gene count in the different seurat_clusters

Violin plots with UMAP of seurat clusters as reference
```{r, fig.width = 14, fig.height = 5}
msc_qcd_anotd[1:4] %>%
  map(. %>% VlnPlot(., features = c( "nFeature_RNA"), group.by = 'old_seu_cluster')) %>% 
  cowplot::plot_grid(plotlist = ., labels = sample_id_lst, ncol = 3) 
```

```{r}
map(msc_qcd_anotd, ~.@meta.data %>% dplyr::count(old_seu_cluster, seurat_clusters) )
```

Violin plots with UMAP of seurat clusters as reference
```{r, fig.width = 14, fig.height = 5}
msc_qcd_anotd[1:4] %>%
  map(~.@meta.data %>% ggplot(aes(y=nFeature_RNA, x=StageMCA, color = old_seu_cluster)) + geom_boxplot()+ geom_point(position=position_jitterdodge(), size = .2) + geom_hline(yintercept = 600)) %>% 
  cowplot::plot_grid(plotlist = ., labels = sample_id_lst, ncol = 2)

```


```{r}
##coding for run with percent.mt cut-off
# f_high = list(c(""), c(2),c(2),c(2))
# f_low =  list(c(""),c(3), c(0),c(1))
# 
# m_high = list(c(""), c(0), c(1),c(""))
# m_low =  list(c(""), c(1), c(3),c(""))

# f_high = list(c("1","2"), c("4"),c("3"),c("2"))
# f_low =  list(c("3"),     c("3"), c("0"),c("1"))
# f_high = f_high[2:3]
# f_low =  f_low[2:3]
f_high = list(c(""), c("2"), c("3"),c("2"))
f_low =  list(c(""), c("1"), c("0"),c("1"))


# m_high = list(c("0"), c("2","0"), c("1"),c("3"))
# m_low =  list(c(""),  c("1"),      c("2"),c("4"))
# m_high = m_high[1:3]
# m_low =  m_low[1:3]
m_high = list(c(""), c(""), c("1"),c("3"))
m_low =  list(c(""), c(""), c("2"),c("4"))

# cluster_2_use<- c(rep('seurat_clusters',4))
cluster_2_use<- c(rep('old_seu_cluster',4))

strains_anno_fHL = list(msc_qcd_anotd[1:4], f_low, f_high, m_low, m_high, cluster_2_use) %>%
# strains_anno_fHL = list(msc_qcd_anotd[1:3], f_low, f_high, m_low, m_high, cluster_2_use) %>%
  pmap(~..1 %>% 
         .@meta.data %>% 
         mutate(stageHL = case_when(str_detect(StageMCA, '\\(female') & !!rlang::sym(..6) %in% ..2 ~ 'Female (LE)',
                                    str_detect(StageMCA, '\\(female')  & !!rlang::sym(..6) %in%  ..3 ~'Female',
                                    # StageMCA == 'Gametocyte (developing)' & !!rlang::sym(..6) %in% ..3 ~ 'Female (LE)',
                                    str_detect(StageMCA, '\\(male')  & !!rlang::sym(..6) %in% ..4 ~ 'Male (LE)',
                                    # str_detect(StageMCA, '\\(male')  & seurat_clusters %in% ..4 ~ 'Male (LE)',
                                    str_detect(StageMCA, '\\(male')  & !!rlang::sym(..6) %in%  ..5 ~'Male',
                                    T~as.character(StageMCA)) ) %>%
         dplyr::select(stageHL) %>%
         mutate(across(stageHL, ~droplevels(factor(., levels = stg_refnd_lvls)))))

msc_qcd_anotd[1:4] = list(strains_anno_fHL, msc_qcd_anotd[1:4]) %>%
# msc_qcd_anotd[1:3] = list(strains_anno_fHL, msc_qcd_anotd[1:3]) %>%
  pmap(~ ..1 %>% 
         dplyr::select('stageHL') %>%
         AddMetaData(..2, .))
```

```{r}
msc_qcd_anotd[1:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        dplyr::count(StageMCA, stageHL,seurat_clusters) 
  )
```

Check out the weird cells with stage not corresponding to cluster
```{r}
map(msc_qcd_anotd[1:4], ~DimPlot(.x, reduction = "umap", cells.highlight = .x@meta.data %>% filter(str_detect(stageHL,'Gam')) %>% rownames()))

```


```{r}
msc_qcd_anotd[3:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        filter(stageHL == 'Gametocyte (female)' | stageHL == 'Gametocyte (male)' | stageHL == 'Gametocyte (developing)') %>%
        dplyr::add_count(StageMCA, stageHL,seurat_clusters, stage_lab2, stage_lab3) %>%
        dplyr::select(StageMCA, stageHL,seurat_clusters, stage_lab2, stage_lab3, sim_lab1, n) 
  )

```
```{r}
map(msc_qcd_anotd[3:4], ~DimPlot(.x, reduction = "umap", label =T))
map(msc_qcd_anotd[3:4], ~DimPlot(.x, reduction = "umap", group.by = "StageMCA"))

```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
gn_nms = c("REX1", "Pfs16", "P25", "DHC")#"ETRAMP5",
gns=c("PF3D7-0935900",  "PF3D7-0406200", "PF3D7-1031000", "PF3D7-0905300")
    
makr_fp <- list(msc_qcd_anotd[1:4])  %>% 
  purrr::pmap(.,~{
    obj=..1
    map2(gns,gn_nms, ~FeaturePlot(obj, reduction = "umap",  features  = .x) + 
           labs(title = .y)+ 
           theme(legend.position = "none", 
                 axis.title = element_blank(),
                 axis.text = element_blank(),
                 axis.ticks = element_blank())) %>%
      plot_grid(plotlist = ., ncol=2)
  })

makr_fp

```

```{r}
msc_qcd_anotd[4] <- msc_qcd_anotd[4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.x@meta.data  %>% 
        mutate(across(stageHL, ~case_when(. == "Gametocyte (developing)" & old_seu_cluster %in% c("0", "3") ~ "Early trophozoite", T~ stageHL)),
               across(StageMCA, ~case_when(. == "Gametocyte (developing)" & old_seu_cluster %in% c("0", "3") ~ "Early trophozoite", T~ StageMCA))) %>%
        select(stageHL, StageMCA) %>% 
        AddMetaData(.x, .) 
  )


```

```{r}
msc_qcd_anotd[3:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        filter(stageHL == 'Gametocyte (female)' | stageHL == 'Gametocyte (male)' | stageHL == 'Gametocyte (developing)') %>%
        dplyr::add_count(StageMCA, stageHL,old_seu_cluster, stage_lab2, stage_lab3) %>%
        dplyr::select(StageMCA, stageHL,seurat_clusters, stage_lab2, stage_lab3, sim_lab1, n) 
  )

```


```{r}
msc_qcd_anotd[1:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        dplyr::count(old_seu_cluster, StageMCA, stageHL) 
  )

```

```{r}
msc_qcd_anotd[1:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        dplyr::count(old_seu_cluster, seurat_clusters,StageMCA) 
  )

```


```{r}
##ACGTACAGTGCTGCAC-1 CCAAGCGCACAGTCAT-1 TCTGTCGTCCACGTAA-1 TGCATCCCATGCCGAC-1
##Remove cells that aren't asigned labels by scmap
# msc_qcd_anotd[[4]] <- subset(msc_qcd_anotd[[4]], subset = StageMCA != 'Unassigned')
# msc_qcd_anotd[1:3] <- subset(msc_qcd_anotd[[1]], !(subset = stageHL == 'Gametocyte (female)' & seurat_clusters == '4'))

map(msc_qcd_anotd[3:4], ~dim(subset(.x, (subset = stageHL == 'Gametocyte (female)' | stageHL == 'Gametocyte (developing)'))))
map(msc_qcd_anotd[1], ~dim(subset(.x, (subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (female)')))))
map(msc_qcd_anotd[1], ~dim(subset(.x, (subset = old_seu_cluster == '2' & (StageMCA == 'Gametocyte (male)' | StageMCA == 'Gametocyte (female)')))))
map(msc_qcd_anotd[2], ~dim(subset(.x, (subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (male)')))))
```


```{r}
##ACGTACAGTGCTGCAC-1 CCAAGCGCACAGTCAT-1 TCTGTCGTCCACGTAA-1 TGCATCCCATGCCGAC-1
##Remove cells that aren't asigned labels by scmap
# msc_qcd_anotd[[4]] <- subset(msc_qcd_anotd[[4]], subset = StageMCA != 'Unassigned')
# msc_qcd_anotd[1:3] <- subset(msc_qcd_anotd[[1]], !(subset = stageHL == 'Gametocyte (female)' & seurat_clusters == '4'))

msc_qcd_anotd[3:4] <- map(msc_qcd_anotd[3:4], ~subset(.x, !(subset = stageHL == 'Gametocyte (female)' | stageHL == 'Gametocyte (developing)')))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (female)'))))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '2' & (StageMCA == 'Gametocyte (male)' | StageMCA == 'Gametocyte (female)'))))
msc_qcd_anotd[2] <- map(msc_qcd_anotd[2], ~subset(.x, !(subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (male)'))))
```

```{r}
##ACGTACAGTGCTGCAC-1 CCAAGCGCACAGTCAT-1 TCTGTCGTCCACGTAA-1 TGCATCCCATGCCGAC-1
##Remove cells that aren't asigned labels by scmap
# msc_qcd_anotd[[4]] <- subset(msc_qcd_anotd[[4]], subset = StageMCA != 'Unassigned')
# msc_qcd_anotd[1:3] <- subset(msc_qcd_anotd[[1]], !(subset = stageHL == 'Gametocyte (female)' & seurat_clusters == '4'))

msc_qcd_anotd[3:4] <- map(msc_qcd_anotd[3:4], ~subset(.x, !(subset = stageHL == 'Gametocyte (female)' | stageHL == 'Gametocyte (developing)')))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '3' & (StageMCA == 'Gametocyte (male)' | stageHL == 'Gametocyte (female)'))))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (female)'))))
msc_qcd_anotd[2] <- map(msc_qcd_anotd[2], ~subset(.x, !(subset = old_seu_cluster == '2' & (StageMCA == 'Gametocyte (male)'))))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '2' & (StageMCA == 'Gametocyte (male)'))))
msc_qcd_anotd[1] <- map(msc_qcd_anotd[1], ~subset(.x, !(subset = old_seu_cluster == '2' & (StageMCA == 'Gametocyte (female)'))))
msc_qcd_anotd[2] <- map(msc_qcd_anotd[2], ~subset(.x, !(subset = old_seu_cluster == '1' & (StageMCA == 'Gametocyte (male)'))))
```

```{r}
msc_qcd_anotd[1:4] %>%
# msc_qcd_anotd[1:3] %>%
  map(~.@meta.data  %>% 
        dplyr::count(old_seu_cluster, StageMCA, stageHL) 
  )

```


```{r, fig.width = 12, fig.height = 4}
## Prepare table for bar plot
md_don_an <- map(msc_qcd_anotd[1:4], ~.@meta.data #%>% 
                       # mutate(across(c(sim_lab1, sim_lab2), as.numeric))
                     ) %>% 
  bind_rows(., .id = 'donor') %>%
  mutate(across(donor, ~factor(., levels = as.character(sample_id_lst))),
         across(stage_lab1, droplevels))

dd <- md_don_an %>% select(donor, old_seu_cluster)  %>%  distinct() %>% arrange(donor, old_seu_cluster) %>%left_join(., data.frame(cluster_cols) %>% rownames_to_column("old_seu_cluster")) 

strip_cols_gg4hx <- dd%>% select(-c(old_seu_cluster)) %>% pivot_wider(names_from = donor, values_from = cluster_cols, values_fn = list) %>% as.list() %>% flatten() 
strip_cols_gg4hx <- as.character(c(names(strip_cols_gg4hx), unlist(strip_cols_gg4hx))) 


strip_cols_gg4hx[str_detect(strip_cols_gg4hx, "MS*")] <- "white"

```


```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14
stg_vec <- levels(md_don_an$stage_lab1)


stage_fct_labs <- msc_qcd_anotd[[1]]@meta.data  %>% dplyr::group_by(stageHL, seurat_clusters) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(stageHL, seurat_clusters), as.character), across(seurat_clusters, ~paste0("Cluster ",., ':\n',stageHL))) %>% deframe()

don_stg_sims_bar_cln <- md_don_an %>% 
  ggplot(aes(x = stage_lab1, y = sim_lab1)) + 
  scale_color_manual(values =  sp_colors, guide = "none")+
  geom_sina(size = 0.01) +
  geom_boxplot(outlier.shape = NA)+
  geom_tile(aes(x = stage_lab1, y = 0.34, fill = stage_lab1, height = 0.05)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", nrow = 1)
                    # guide = 'none'
  ) +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors))) +
  # ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x") +
  ggh4x::facet_nested_wrap(vars(donor, old_seu_cluster),  nrow = 1, scales = "free_x", strip = strip_nested(background_x = elem_list_rect(fill =  strip_cols_gg4hx))) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')

don_stg_sims_bar_cln
```
```{r}
msc_qcd_anotd[1:4] %>%
  map(~.@meta.data  %>% 
        dplyr::count(stageHL) 
  )
```


```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14
scm_barptHL <- list(msc_qcd_anotd[1:4]) %>%
pmap(.,~{
  
stage_fct_labs <- ..1@meta.data %>% select(-c(stage_lab3,sim_lab3))%>% dplyr::group_by(stageHL, seurat_clusters) %>% 
  dplyr::group_keys() %>% 
  dplyr::mutate(across(c(stageHL, seurat_clusters), as.character),
                "stageHL2" = str_replace_all(stageHL, c("Gametocyte \\(female\\)" = "Female", "Gametocyte \\(male\\)" = "Male", "Early trophozoite" = "Early\ntrophozoite", " \\(" = "\n(")),
                across(seurat_clusters, ~paste0("Cl ",., ':\n',stageHL2))) %>% 
  select(stageHL, seurat_clusters) %>%
  deframe()

print(stage_fct_labs)

scm_b_tbl <- ..1@meta.data %>% 
  select(-c(stage_lab3,sim_lab3))%>% 
  dplyr::select(matches('s.*lab.'), stageHL) %>% 
  rownames_to_column('cell_bc')%>%
  pivot_longer(-c('cell_bc', 'stageHL'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>% 
  filter(!is.na(sim)) %>%
  mutate(across(sim, as.numeric)) %>% 
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls))))

stg_vec <- levels(scm_b_tbl$stage)

scm_b_tbl %>%
  ggplot(., aes(y = sim, x = stage, color = Label)) +
  geom_boxplot(aes(fill = Label))+
  scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 2)
  ) +
  scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
  geom_point(aes(group=Label), size = 0.05, position = position_jitterdodge(jitter.width = 0.2))+
  new_scale_fill()+
  geom_tile(data = data.frame(x = stg_vec), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
                    # guide = 'none'
  )+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  facet_grid(.~stageHL, labeller = labeller(stageHL = stage_fct_labs)) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')
})
scm_barptHL
```


#### Renormalize and cluster annotated cells
```{r}

# msc_qcd_anotd[4] <- msc_qcd_anotd[4] %>%
msc_qcd_anotd <- msc_qcd_anotd %>%
  map(. %>%
        Seurat::NormalizeData(normalization.method = "LogNormalize", scale.factor = 10000, verbose = FALSE) %>%
        FindVariableFeatures(selection.method = "vst", nfeatures = 700, verbose = FALSE) %>% 
        ScaleData(., features = rownames(.), verbose = FALSE)%>% 
        RunPCA(., features = VariableFeatures(object = .), verbose = FALSE)
  )
```

```{r, message=FALSE, warning=FALSE}
pc_list <-list(6L,4L,5L, 5L, 12L)
# pc_list <-pc_list
seed_lst <- list(043,540,540, 270, 540)
##555 105 829

##!!NOTE - TURN OFF SO AS NOT TO RECLUSTER AGAIN - MAY BE NECESSARY TO RECLUSTER
# msc_qcd_anotd[4] <- list(msc_qcd_anotd[4], pc_list[4], res_list[4]) %>%
msc_qcd_anotd <- list(msc_qcd_anotd, pc_list, seed_lst) %>%
  pmap(~RunUMAP(..1, dims = 1:..2, n.components = 3L, verbose=F, seed.use = ..3) %>%
         FindNeighbors(., dims = 1:..2, verbose = FALSE)
  )



```

```{r, message=FALSE, warning=FALSE}
# res_list <- c(0.18,0.12,0.12,0.12,0.12)
res_list <- c(0.1,0.1,0.2,0.2,0.2)

res_list <- res_list
##555 105

##!!NOTE - TURN OFF SO AS NOT TO RECLUSTER AGAIN - MAY BE NECESSARY TO RECLUSTER
# msc_qcd_anotd[4] <- list(msc_qcd_anotd[4], res_list[4]) %>%
msc_qcd_anotd <- list(msc_qcd_anotd, res_list) %>%
  pmap(~FindClusters(..1, resolution = ..2, verbose=T)
  )

```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
anot_umaps <- list(rep(msc_qcd_anotd[1:4], 2), rep(sample_id_lst[1:4], 2), c(rep(2,4), rep(3,4)))  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p_umap <- DimPlot(., reduction = "umap", group.by = 'stageHL', dims = c(1, ..3),
                cols = sp_fld_colors , pt.size = 1,  label = F, repel = T, label.size = 6 ,
                )  + 
      labs(title = ..2)+
      umap_theme+
  guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)

    return(p_umap)
    cat("\n")
  })

anot_umaps

```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
map(msc_qcd_anotd, ~DimPlot(.x, group.by = "old_seu_cluster" ))
```


```{r, results='asis'}

## numbers 
## Counts of QC Pass
don_n_cln <- map(msc_qcd_anotd, ~dim(.x)[2])

# cat('## Visualize field parasites {.tabset}   \n')
anot_umaps_st <- list(rep(msc_qcd_anotd[1:4], 1), rep(sample_id_lst[1:4], 1), c(rep(2,4)), c(-5, 6, -1, -2), c(-5, -5, -10,-8), don_n_cln[1:4])  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p_umap <- DimPlot(., reduction = "umap", group.by = 'StageMCA', dims = c(1, ..3),
                cols = sp_colors , pt.size = 1,  label = F, repel = T, label.size = 6 ,
                )  + 
      labs(title = ..2)+
      geom_text(x=..4, y=..5, label=paste0('N=', ..6), size = 5) + 
      umap_theme +
  guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
    
p_umap$data <- p_umap$data %>% rownames_to_column('bcode') %>% left_join(., ..1@meta.data[,'old_seu_cluster', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')

p_umap<- LabelClusters(p_umap, id = "old_seu_cluster", size = 5, repel = T)


    return(p_umap)
    cat("\n")
  })

anot_umaps_st

```

```{r, results='asis'}

## numbers 
## Counts of QC Pass

# cat('## Visualize field parasites {.tabset}   \n')
list(rep(msc_qcd_anotd[1:4], 2), rep(sample_id_lst[1:4], 2), c(rep(2,4), rep(3,4)))  %>% 
  purrr::pmap(.,~{
    # create tabset for each group 
    cat('##### ',..2,'   \n')
    p_umap <- DimPlot(., reduction = "umap", group.by = 'StageMCA', dims = c(1, ..3),
                cols = sp_colors , pt.size = 1,  label = F, repel = T, label.size = 6 ,
                )  + 
      labs(title = ..2)+
      umap_theme +
  guides(color = guide_legend(title = 'Stage',override.aes = list(size = 4), ncol = 1, byrow = TRUE),
         x= trunc_axis,
         y= trunc_axis)+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)
    
p_umap$data <- p_umap$data %>% rownames_to_column('bcode') %>% left_join(., ..1@meta.data[,'old_seu_cluster', drop = F] %>% rownames_to_column('bcode'), by='bcode') %>% column_to_rownames('bcode')

p_umap<- LabelClusters(p_umap, id = "old_seu_cluster", size = 5, repel = T)


    return(p_umap)
    cat("\n")
  })



```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
list(msc_qcd_anotd)  %>% 
  purrr::pmap(.,~{
    FeaturePlot(., reduction = "umap",  features  = 'nFeature_RNA')
  })

```





#### Violin plots
```{r, fig.height = 4, fig.width = 10}
# options(repr.plot.width=32, repr.plot.height=7)
fct_theme <- theme(text=element_text(size=14),
        axis.text=element_text(size=14), 
            axis.title=element_text(size=16,face="bold"), 
            legend.text = element_text(size=16), 
            legend.title = element_text(size=16,face="bold"),
        strip.text =element_text(size=18,face="plain", colour = 'black'))

msc_qcd_mdata <- map(msc_qcd_anotd[1:4], ~.@meta.data %>% mutate(across(c(sim_lab1, sim_lab2), as.numeric))) %>% 
  bind_rows(., .id = 'sample') %>% 
  mutate(across(c(sample), ~factor(., levels = unlist(sample_id_lst)))) 
  

viol_gn_cnt_cln <- msc_qcd_mdata %>% 
  ggplot(aes(x = old_seu_cluster, color = StageMCA, y = nFeature_RNA/1000)) +
  scale_color_manual(values = sp_colors)+
  theme_classic() +
  geom_boxplot(outlier.shape = NA) +
  ggforce::geom_sina(size = 0.01)+
  labs(y = "Gene count (X 1e3)", x = "Cluster")+
  facet_wrap(.~sample, ncol = 4, scales = "free_x") +
  fct_theme+
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = 'Cluster',override.aes = list(size = 4), nrow = 1))

viol_gn_cnt_cln
```



#### gene depth and cell count plot
```{r, fig.height = 4, fig.width = 10}
# options(repr.plot.width=32, repr.plot.height=7)
msc_qcd_mdata %>% 
  group_by(sample, stageHL, Strain_DN) %>%
  add_count(name = "ss_n") %>%
  mutate(mean_gnC = mean(nFeature_RNA)) %>% 
  select(sample, Strain_DN, mean_gnC, ss_n)

```

```{r, fig.height = 4, fig.width = 16}
# options(repr.plot.width=32, repr.plot.height=7)
msc_qcd_mdata %>% 
  filter(sample != "MSC1272") %>%
  filter(!str_detect(stageHL, "LE|ring|troph")) %>%
  group_by(sample, stageHL, Strain_DN) %>%
  add_count(name = "ss_n") %>%
  mutate(mean_gnC = mean(nFeature_RNA)) %>% 
  select(sample, Strain_DN, mean_gnC, ss_n) %>%
  ggplot(aes(x = stageHL, y = mean_gnC, color = Strain_DN)) +
  geom_point(position=position_dodge(width=0.8), size = 4)+
  geom_point(aes(y = ss_n*4), shape = 3, position=position_dodge(width=0.8), size = 4)+
  scale_y_continuous(
    name = "First Axis",
    sec.axis = sec_axis( trans=~.*4, name="Count(/2)")
  ) +
  facet_wrap(sample ~ ., nrow = 1, scales = "free") +
  theme_bw()

```


```{r, fig.height = 4, fig.width = 16}
# options(repr.plot.width=32, repr.plot.height=7)
msc_qcd_mdata %>% 
  filter(sample != "MSC1272") %>%
  filter(!str_detect(stageHL, "LE|ring|troph")) %>%
  group_by(sample, stageHL, Strain_DN) %>%
  # summarise("sum_gn" = sum(nFeature_RNA)) %>%
  mutate(sum_gn = sum(nCount_RNA)) %>% 
  # add_count(name = "ss_n") %>%
  mutate(mean_gnC = mean(nCount_RNA)) %>% 
  select(sample, Strain_DN, mean_gnC, sum_gn) %>%
  ggplot(aes(x = stageHL, y = sum_gn, color = Strain_DN)) +
  geom_point(position=position_dodge(width=0.8), size = 4)+
  facet_wrap(sample ~ ., nrow = 1, scales = "free_x") +
  theme_bw()

```

Replace MSC14 here with the old published msc14

```{r}
##READ IN MSC14
# msc_qcd_anotd_stgstrqc <- readRDS('data/processed/msc14/5736STDY11771546_DB52e_star/msc_qcd_anotd_stgstrqc.RDS') 

##Change the Stage_V3 in msc14 to match the same labelling in the other objects
# msc_qcd_anotd_stgstrqc[[1]]@meta.data$StageMCA <- msc_qcd_anotd_stgstrqc[[1]]@meta.data$Stage_V3
```



```{r}
##integration of MSC14 and V3 reference
# msc_qcd_anotd[1:4] <- append(msc_qcd_anotd[1:3], msc_qcd_anotd_stgstrqc)%>% 
#   set_names(sample_id_lst[1:4])

```



```{r, fig.width= 4, fig.height=4}
msc_qcd_anotd[1:4] %>%
  imap(.,~VlnPlot(., features = c( "nFeature_RNA"), group.by = 'stageHL', cols = sp_fld_colors) + labs(title = .y)+ theme(legend.position = 'none') ) #%>% 
  #cowplot::plot_grid(plotlist = ., labels = sample_id_lst, ncol = 3) 
```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
gn_nms = c("REX1", "Pfs16", "P25", "DHC")#"ETRAMP5",
gns=c("PF3D7-0935900",  "PF3D7-0406200", "PF3D7-1031000", "PF3D7-0905300")
    
list(msc_qcd_anotd[1:4])  %>% 
  purrr::pmap(.,~{
    FeaturePlot(..1, reduction = "umap",  features  = gns, ncol = 2)
  })

```

```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
gn_nms = c("REX1", "Pfs16", "P25", "DHC")#"ETRAMP5",
gns=c("PF3D7-0935900",  "PF3D7-0406200", "PF3D7-1031000", "PF3D7-0905300")
    
makr_fp <- list(msc_qcd_anotd[1:4])  %>% 
  purrr::pmap(.,~{
    obj=..1
    map2(gns,gn_nms, ~FeaturePlot(obj, reduction = "umap",  features  = .x) + 
           labs(title = .y)+ 
           theme(legend.position = "none", 
                 axis.title = element_blank(),
                 axis.text = element_blank(),
                 axis.ticks = element_blank())) %>%
      plot_grid(plotlist = ., ncol=2)
  })

makr_fp

```

#### Violin plots
```{r, fig.height = 4, fig.width = 16}
# options(repr.plot.width=32, repr.plot.height=7)

msc_qcd_mdata <- map(msc_qcd_anotd, ~.@meta.data %>% mutate(across(c(sim_lab1, sim_lab2), as.numeric))) %>% 
  bind_rows(., .id = 'sample') %>% 
  mutate(across(sample, ~factor(., levels = unlist(sample_id_lst)))) 
  

msc_qcd_mdata %>% 
  ggplot(aes(x = stageHL, color = sample, y = nFeature_RNA)) +
  scale_color_manual(values = donor_cols)+
  theme_classic() +
  geom_violin() +
  ggforce::geom_sina(size = 0.01)+
  theme(text=element_text(size=14))

```



Put together all the necessary plots
```{r, fig.width= 12, fig.height=16, eval =F}
##Put together the plots for annotation
stg_lgnd <- cowplot::get_legend(
  scm_barpt_s[[4]]+
    guides(fill=guide_legend(title.position = "top", nrow = 2)))

plot_grid(plot_grid(plotlist = map(my21_mca_umap[1:4], ~.+ theme(legend.position = "none")), ncol = 4),
          plot_grid(plotlist = umap_stg_mca[1:4], ncol = 4),
          viol_gn_cnt +theme(legend.position = "none"),
          plot_grid(plotlist = map(scm_barptHL[1:2], ~.+ theme(legend.position = "none")), ncol = 2, rel_widths = c(.6, .4)),
          plot_grid(plotlist = map(scm_barptHL[3:4], ~.+ theme(legend.position = "none")), ncol = 2, rel_widths = c(.4, .6)),
          stg_lgnd,
          nrow = 6,
            rel_heights = c(.2,.175,.225,.175,.175,.05),
            labels = c('A', 'B', 'C', 'D', 'E'),
            label_size = 24,
            vjust= 0.3)
```


Put together all the necessary plots
```{r, fig.width= 12, fig.height=16, eval =F}
##Put together the plots for annotation
stg_lgnd <- cowplot::get_legend(
  scm_barpt_s[[4]]+
    guides(fill=guide_legend(title.position = "top", nrow = 2)))

plot_grid(plot_grid(plotlist = map(my21_mca_umap[1:4], ~.+ theme(legend.position = "none")), ncol = 4),
          don_stg_sims_bar,
          plot_grid(plotlist = map(anot_umaps[1:4], ~.+ theme(legend.position = "none")), ncol = 4),
          viol_gn_cnt +theme(legend.position = "none"),
          stg_lgnd,
          nrow = 6,
            rel_heights = c(.2,.175,.225,.175,.175,.05),
            labels = c('A', 'B', 'C', 'D', 'E'),
            label_size = 24,
            vjust= 0.3)
```



Put together all the necessary plots
```{r, fig.width= 12, fig.height=16, eval =F}
##Put together the plots for annotation
stg_lgnd <- cowplot::get_legend(
  viol_gn_cnt_cln+
    guides(color=guide_legend(title.position = "top", nrow = 1)))

dp_lgnd <- cowplot::get_legend(
  mark_dot_plts[[1]][[1]])

plot_grid(plot_grid(plotlist = map(my21_mca_umap[1:4], ~.+ theme(legend.position = "none")), ncol = 4),
          don_stg_sims_bar,
          plot_grid(plotlist = map(unlist(c(mark_dot_plts[1],mark_dot_plts2_4), recursive = F), ~ .+ theme(legend.position = "none")), ncol = 4, rel_widths = c(.34,.22,.22,.22)),
          dp_lgnd,
          plot_grid(plotlist = map(anot_umaps_st[1:4], ~.+ theme(legend.position = "none", axis.title = element_blank(), axis.text = element_blank())), ncol = 4),
          viol_gn_cnt_cln +theme(legend.position = "none"),
          stg_lgnd,
          nrow = 7,
            rel_heights = c(.2,.185,.18,.05,.195,.185,.05),
            labels = c('A', 'B', 'C', '', 'D', 'E', ''),
            label_size = 24,
            vjust= 0.3)


```


```{r, fig.width= 12, fig.height=16}
##Put together the plots for annotation
# rrca()
```

```{r, fig.width= 12, fig.height=16, eval =F}
##Put together the plots for annotation
msc_ttl <- map(sample_name_nms[1:4], ~.x %>% ggdraw() + 
  draw_label(.,  fontface = 'bold', size =20,  hjust =  0.5) +
  theme(
    # plot.margin = margin(3, 3, 3, 3)
  )
)

stg_lgnd <- cowplot::get_legend(
  viol_gn_cnt_cln+
    guides(color=guide_legend(title.position = "top", nrow = 1)))

dp_lgnd <- cowplot::get_legend(
  mark_dot_plts[[1]][[1]])

plot_grid(plot_grid(plotlist = msc_ttl, nrow = 1),
          plot_grid(plotlist = map(my21_mca_umap[1:4], ~.+ theme(legend.position = "none")), ncol = 4, scale = 0.95),
          don_stg_sims_bar,
          plot_grid(plotlist = map(unlist(c(mark_dot_plts[1],mark_dot_plts2_4), recursive = F), ~ .+ theme(legend.position = "none")), ncol = 4, rel_widths = c(.34,.22,.22,.22)),
          dp_lgnd,
          plot_grid(plotlist = map(anot_umaps_st[1:4], ~.+ theme(legend.position = "none", axis.text = element_blank())), ncol = 4),
          viol_gn_cnt_cln +theme(legend.position = "none"),
          stg_lgnd,
          nrow = 8,
            rel_heights = c(.02,.19,.185,.18,.05,.185,.185,.05),
            labels = c('','A', 'B', 'C', '', 'D', 'E', ''),
            label_size = 24,
            vjust= 0.3)


```


```{r}
# for (i in 1:length(msc_qcd_anotd)){
#   msc_qcd_anotd[[i]]@meta.data["strain_qcd"] <- msc_qcd_anotd[[i]]@meta.data$Strain_DN
# }
# msc_qcd_anotd[1:4] <- list(msc_qcd_anotd[1:4], unified_strain_nms, c("n_strains", "Strain_2nd", "Strain_2nd", "n_strains2")) %>%
#   pmap(., ~AddMetaData(..1, ..2 %>% select(..3) %>% rename("Strain_DN" = ..3)))
```


```{r}
##Transfer stage labels and other metadata from farm
# mdata_all <- list.files("data/processed/DB52e_star/", pattern = "msc_filt_w_dblts_scmap_metadata.RDS", full.names = T, recursive = T, include.dirs = T) %>% 
#   set_names(nm = (str_remove_all(.,'data/processed/DB52e_star//|_5736.*|/msc_filt_w_dblts_scmap_metadata.RDS') %>% str_replace(., '/', '_'))) %>%
msc_qcd_anotd <- map(msc_qcd_anotd, ~{
    .x@meta.data %>%
    rownames_to_column('bcode') %>%
    mutate(stage_afm = case_when(str_detect(StageMCA, 'ring|troph') ~ 'Asexual', 
                                     str_detect(StageMCA, '\\(male') ~ 'Male', 
                                     str_detect(StageMCA, 'female') ~ 'Female',
                                     T ~StageMCA),
            stage_ag = case_when(str_detect(StageMCA, 'ring|troph') ~ 'Asexual', 
                                    str_detect(StageMCA, 'ale') ~ 'Gametocyte',
                                     T ~StageMCA)) %>%
      add_count(Strain_DN, stage_afm, name = 'strain_afm_n')%>% 
      add_count(Strain_DN, stage_ag, name = 'strain_ag_n') %>%
    mutate("strain_qcd" = case_when(strain_ag_n >= 30 ~ Strain_DN)) %>%
    column_to_rownames('bcode') %>%
    select(stage_afm, stage_ag, strain_afm_n, strain_ag_n, strain_qcd) %>%
    AddMetaData(.x, .)
      })

msc_qcd_anotd[1:4] <- map(msc_qcd_anotd[1:4], ~{
    .x@meta.data %>%
    mutate(stageHL_ref = case_when(stageHL == "Gametocyte (female)" ~ "Female", 
                                   stageHL == "Gametocyte (male)" ~ "Male", 
                                   T ~ stageHL)) %>%
    select(stageHL_ref) %>%
    AddMetaData(.x, .)
      })

```


```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
# msc_qcd_anotd<-readRDS("data/processed/DB52e_star/msc_qcd_anotd.RDS")
saveRDS(msc_qcd_anotd, "data/processed/DB52e_star/msc_qcd_anotd.RDS")
saveRDS(map(msc_qcd_anotd, ~.x@meta.data), "data/processed/DB52e_star/msc_qcd_anotd_mdata.RDS")

```


Write out strain_qcd/strain + stage_ag/stage_afm barcodes to farm for pseudobulk genotyping

```{r, eval=T}
## DOUBLETS WRITTEN INTO THE SAME FOLDER IN MSC_PP.rmd - SHOULD BE COMBINED INTO ONE NOTEBOOK
# sample_ids = map2(sample_name_nms, source_irods_nms, ~paste0(.x, '_', .y))
algn_nm = "minmap"
trip = "Mali"

for (s in names(msc_qcd_anotd[1:4])) {
  for (str_q in c('strain_qcd')){
    # ss_qcd_ano <- map(msc_qcd_anotd[s], ~.x@meta.data %>% dplyr::select(str_q) %>% mutate(across(c(str_q), ~droplevels(factor(.)))) %>% rownames_to_column('bcode')%>% rename('str'=str_q) %>% split(f = ~ str, drop = T))
    ss_qcd_ano <- map(msc_qcd_anotd[s], ~.x@meta.data %>% rownames_to_column('bcode') %>% dplyr::select(str_q, bcode) %>% mutate(across(c(str_q), ~droplevels(factor(.)))) %>% rename('str'=str_q) %>% split(f = ~ str, drop = T))
    names(ss_qcd_ano[[1]]) <- names(ss_qcd_ano[[1]]) %>% str_replace(., '\\.', '_')
    print(names(ss_qcd_ano[[1]]))
    
    system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/'))
    system(paste0('rm /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/*'))

    imap(ss_qcd_ano[[1]], ~write(.x$bcode, paste0('data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/',.y, '.tsv')))
     
    ##Add doublets
    system(
      paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/Doublet.tsv ', '/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s],'/')
    )
    
    system(
      paste0('rsync -av --delete /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',str_q,'_k',optimum_k[s], ' jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/')
    )
    
    
    
    
    for(stg_c in c('stage_afm', 'stage_ag')){
      ss_qcd_ano <- map(msc_qcd_anotd[s], ~.x@meta.data %>% rownames_to_column('bcode') %>% dplyr::select(str_q, stg_c, bcode) %>% mutate(across(c(str_q, stg_c), ~droplevels(factor(.)))) %>% rename('stg'=stg_c, 'str'=str_q) %>% split(f = ~ str+stg, drop = T))
      names(ss_qcd_ano[[1]]) <- names(ss_qcd_ano[[1]]) %>% str_replace(., '\\.', '_')
      print(names(ss_qcd_ano[[1]]))
      
      system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/'))
      system(paste0('rm /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/*'))

      imap(ss_qcd_ano[[1]], ~write(.x$bcode, paste0('data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/',.y, '.tsv')))
      
      ##Add doublets consistent across different ks - and removed in QC step 'msc_QC_pp_all.Rmd'
      system(
      paste0('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/Doublet.tsv ', '/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s],'/')
    )
      
      system(
        paste0('rsync -av --delete /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/',stg_c,'_',str_q,'_k',optimum_k[s], ' jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', str_to_lower(s),'/bcodes/',algn_nm,'/')
      )
      
       
      
    }
  }
  
}
```


```{r}
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay

# msc_m21_qcd_anotd_strn<- readRDS("data/processed/msc_m21_qcd_anotd_strn.RDS")

##!!! Need to comment this out to only account off strain qcd object
msc_m21_qcd_anotd_strn <- msc_qcd_anotd[1:4]

# msc_m21_qcd_anotd_strn[[4]]@meta.data$StageMCA <- msc_m21_qcd_anotd_strn[[4]]@meta.data$Stage_V3
```


#### Cluster verification using Rafa irzary method - https://github.com/igrabski/sc-SHC
Male LE lumped together with asexuals and Female (LE) and (HE) both split into 2
```{r, eval=FALSE}
library(scSHC)
clusters <- scSHC(msc_qcd_anotd[[4]]@assays$RNA@counts, num_features = 500, num_PCs = 7)
AddMetaData(msc_qcd_anotd[[4]], data.frame(clusters[[1]]) %>% rename('scSHC_clusters' = 1))@meta.data %>% dplyr::count(scSHC_clusters, seurat_clusters, stageHL)

##Test our annotated clusters - takes long
new_clusters <- testClusters(msc_qcd_anotd[[4]]@assays$RNA@counts, as.character(msc_qcd_anotd[[4]]@meta.data$seurat_clusters))
new_clusters <- testClusters(msc_qcd_anotd[[4]]@assays$RNA@counts, as.character(msc_qcd_anotd[[4]]@meta.data$stageHL))
```

### sMap cosine similarities visualization
scmap cosine similarities for the top 2 assigned stage labels among the top 10 stage labels  

```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14

stage_fct_labs <- msc_qcd_anotd[[4]]@meta.data  %>% dplyr::group_by(stageHL, seurat_clusters) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(stageHL, seurat_clusters), as.character), across(seurat_clusters, ~paste0("Cluster ",., ':\n',stageHL))) %>% deframe()

scm_barpt <- msc_qcd_anotd[[4]]@meta.data %>% 
  dplyr::select(matches('s.*lab.'), stageHL) %>% 
  rownames_to_column('cell_bc')%>%
  pivot_longer(-c('cell_bc', 'stageHL'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>% 
  filter(!is.na(sim)) %>%
  mutate(across(sim, as.numeric)) %>% 
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls)))) %>%
  ggplot(., aes(y = sim, x = stage, color = Label)) +
  geom_boxplot(aes(fill = Label))+
  scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 2)
  ) +
  scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
  geom_point(aes(group=Label), size = 0.5, position = position_jitterdodge(jitter.width = 0.2))+
  new_scale_fill()+
  geom_tile(data = data.frame(x = c( 'Late ring', 'Early trophozoite', 'Gametocyte (developing)', 'Gametocyte (female)', 'Gametocyte (male)')), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
                    # guide = 'none'
  )+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  facet_grid(.~stageHL, labeller = labeller(stageHL = stage_fct_labs)) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')

scm_barpt
```

```{r, fig.width = 12, fig.height = 4}

##Plotting top two scmap stages among the top 10 with their similarities for msc14

stage_fct_labs <- msc_qcd_anotd[[4]]@meta.data  %>% dplyr::group_by(stageHL, seurat_clusters) %>% dplyr::group_keys() %>% dplyr::mutate(across(c(stageHL, seurat_clusters), as.character), across(seurat_clusters, ~paste0("Cluster ",., ':\n',stageHL))) %>% deframe()

scm_barpt <- msc_qcd_anotd[[4]]@meta.data %>% 
  dplyr::select(matches('s.*lab.'), stageHL) %>% 
  rownames_to_column('cell_bc')%>%
  pivot_longer(-c('cell_bc', 'stageHL'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>% 
  filter(!is.na(sim)) %>%
  mutate(across(sim, as.numeric)) %>% 
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls)))) %>%
  ggplot(., aes(y = sim, x = stage, color = Label)) +
  geom_boxplot(aes(fill = Label))+
  scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 2)
  )+
  scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
  geom_point(aes(group=Label), size = 0.5, position = position_jitterdodge(jitter.width = 0.2))+
  new_scale_fill()+
  geom_tile(data = data.frame(x = c( 'Late ring', 'Early trophozoite', 'Gametocyte (developing)', 'Gametocyte (female)', 'Gametocyte (male)')), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage",
                    guide = guide_legend(direction = "horizontal", title.position = "top", ncol = 3)
                    # guide = 'none'
  )+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  facet_grid(.~stageHL, labeller = labeller(stageHL = stage_fct_labs)) +
  geom_hline(yintercept = 0.4)+
  labs(y = 'scmap \nsimilarity scores', x = 'MCA stage') +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')

scm_barpt
```


Plotting only the stages where there are 2 different stage labels in the top 10 stage labels

```{r, fig.width = 12, fig.height = 4}
msc_qcd_anotd[[4]]@meta.data %>% 
  filter(!is.na(sim_lab2))%>% 
  dplyr::select(matches('s.*lab.'), stageHL) %>% 
  rownames_to_column('cell_bc') %>%
  pivot_longer(-c('cell_bc', 'stageHL'),
               names_to = c( ".value", "Label"),
               names_pattern = "(s.*)_(l.*)")%>% 
  mutate(across(stage, ~droplevels(factor(., stg_refnd_lvls))))%>% 
  mutate(across(sim, as.numeric)) %>% 
  mutate(across(Label, ~str_replace(., 'lab','Label '))) %>%
  ggplot(., aes(y = sim, x = stage,  color = Label)) +
  geom_boxplot(aes(fill = Label)) +
  scale_fill_manual(values =  c('Label 1' = '#007FFF', 'Label 2' = '#FFFF9F'), name = "Label")+
  scale_color_manual(values =  c('Label 1' = 'black', 'Label 2' = 'black'), name = "Label")+ 
  geom_point(aes(group=Label), size = 0.7, position = position_jitterdodge(jitter.width = 0.2))+
  new_scale_fill()+
  geom_tile(data = data.frame(x = c( 'Late ring', 'Early trophozoite', 'Gametocyte (developing)', 'Gametocyte (female)', 'Gametocyte (male)')), inherit.aes = F, aes(x = x, y = 0.37, fill = x, height = 0.03)) +
  scale_fill_manual(values =  sp_colors, name = "MCA stage")+
  facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors))) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4), 
        axis.title=element_text(size=20,face="bold"), 
        axis.text = element_text(size=18), 
        strip.text =element_text(size=18,face="bold"),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=16))
```

Tabulate the top 2 stages from scmap for those with 2 stages
```{r}
# scmap_cells_sim_nw[[1]] %>% dplyr::count(stage_lab1,stage_lab2)
```


## Logistic regression machine learning (LR-ML) label transfer - on farm

Write out annotated MSC14 dataset in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad
msc14_qcd_anotd_diet <- DietSeurat(
  msc_qcd_anotd[[4]],
  counts = TRUE,
  data = T,
  scale.data = FALSE,
  features = NULL,
  assays = NULL,
  dimreducs = c('pca', 'umap', 'umap_test'),
  graphs = NULL
)

##Change factor metadata to character as these are not handled properly when converting to h5ad
msc14_qcd_anotd_diet@meta.data <- msc14_qcd_anotd_diet@meta.data %>% mutate(across(where(is.factor), as.character))

## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top
SeuratDisk::SaveH5Seurat(msc14_qcd_anotd_diet, filename =  '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/msc14_qcd_anotd_diet.h5Seurat', overwrite = T)

Convert("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/msc14_qcd_anotd_diet.h5Seurat", dest = "h5ad", overwrite = T)

```

Write out annotated MCA reference in .h5ad 
```{r, message=FALSE, warning=FALSE}
##Save msc14 h5ad
EF_V3_ref_diet <- DietSeurat(
  EF_V3_ref,
  counts = TRUE,
  data = T,
  scale.data = FALSE,
  features = NULL,
  assays = NULL,
  dimreducs = c('pca', 'umap', 'umap_test'),
  graphs = NULL
)

##Change factor metadata to character as these are not handled properly when converting to h5ad
EF_V3_ref_diet@meta.data <- EF_V3_ref_diet@meta.data %>% mutate(across(where(is.factor), as.character))
## !!!NB - NEEDS SORTING IN FUTURE UPGRADING TO SEURAT 5 RESULTED IN ERROR AS DESCRIBED https://github.com/mojaveazure/seurat-disk/issues/147#top

SaveH5Seurat(EF_V3_ref_diet, filename =  '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/EF_V3_ref_diet.h5Seurat', overwrite = T)

Convert("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/EF_V3_ref_diet.h5Seurat", dest = "h5ad", overwrite = T)

```

Transfer MCA and MSC14 .h5ad to farm
```{r}

##Copy h5ad
system('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/msc14_qcd_anotd_diet.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/'
)

system('rsync -av /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/EF_V3_ref_diet.h5ad jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/'
)

```

Read in MSC14 logistic regression labels
```{r, message=FALSE}

##Copy logistic regression predictions from farm
system('rsync -av jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/ml_lr_predictions.csv /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/msc14/5736STDY11771546_DB52e_star/'
)

ml_lr_predictions <- read_csv('data/processed/msc14/5736STDY11771546_DB52e_star/ml_lr_predictions.csv') %>% column_to_rownames('...1') %>% dplyr::select(classifier_best_prediction, starts_with('prob')) %>% rename_all(., ~str_replace(.,' ', '_'))

```

### LR-ML probabilities per cell for each stage
Barplot 



```{r, fig.width = 14, fig.height = 4, message=FALSE, warning=FALSE}


msc_qcd_anotd_m14_scm_lrml <- AddMetaData(msc_qcd_anotd[[4]], ml_lr_predictions)

lrml_barpt <-msc_qcd_anotd_m14_scm_lrml@meta.data %>% 
  dplyr::select(c(starts_with('prob'), 'stageHL')) %>%
  pivot_longer(cols = starts_with('prob'), names_to = 'Machine learning predictions', values_to = 'probabilities') %>% 
  mutate(across(`Machine learning predictions`, ~str_replace_all(., c('prob_' = '', '\\.$'=')', '_\\.'= ' (', '\\.'=' ', '_'=' '))))%>% 
  mutate(across(`Machine learning predictions`, ~str_to_sentence(str_replace_all(., c('Gametocytes'='Gametocyte', 'comitted' = 'committed'))))) %>% 
  mutate(across(`Machine learning predictions`, ~droplevels(factor(., stg_refnd_lvls))))%>% 
  ggplot(., aes(y = probabilities, x = `Machine learning predictions`, fill = `Machine learning predictions`)) +
  geom_boxplot() +
  geom_point(aes(group=`Machine learning predictions`), size = 0.5, position = position_jitterdodge())+
  geom_tile(aes(x = `Machine learning predictions`, y = -0.05, fill = `Machine learning predictions`, height = 0.04)) +
  scale_fill_manual(values =  sp_colors, name = "MCA\nstage", guide = guide_legend( direction = "horizontal", 
                                                                                   #title.position = "top" 
  ))+
  # facet_grid2(.~stageHL, strip = strip_themed(background_x = elem_list_rect(fill = sp_fld_colors)))+
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  facet_grid(.~stageHL, labeller = labeller(stageHL = stage_fct_labs)) +
  theme_classic() +
  labs(x = 'MCA stage', y = 'Probabilities')+
  theme(axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4), 
        axis.title=element_text(size=20,face="bold"), 
        axis.text = element_text(size=16), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=18,face="bold", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15.5),
        legend.position = 'bottom',
        legend.margin=margin(t = 0, unit='cm'))


lrml_barpt
```


## Compare scmap and LR-ML stage annotations
Tabulate
```{r}
msc_qcd_anotd_m14_scm_lrml@meta.data %>% dplyr::count(stageHL, classifier_best_prediction)
```



scmap cosine similarities VS top LR-ML probabilities faceted by LR-ML labels on columns and scmap labels on rows
```{r, fig.width = 14, fig.height = 14, eval=FALSE}
msc_qcd_anotd_m14_scm_lrml@meta.data %>% 
  dplyr::select(c(starts_with('prob'), 'stageHL', 'stage_lab1', 'sim_lab1', 'classifier_best_prediction'))%>%
  rownames_to_column('cell_bc') %>%
  pivot_longer(cols = starts_with('prob'), names_to = 'ML_stages', values_to = 'probabilities') %>% 
  mutate(across(ML_stages, ~str_remove_all(., 'pr.*_')),
         across(ML_stages, ~str_replace(., '\\.', ' ')),
         across(ML_stages, ~str_replace(., '\\.', '\\(')),
         across(ML_stages, ~str_replace(., '\\.', '\\)')),
         across(sim_lab1, as.numeric))%>%
  ggplot(., aes(y = probabilities, x = sim_lab1)) +
  geom_point() +
  facet_grid(stage_lab1~ML_stages) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4),
        axis.title=element_text(size=18,face="bold"), 
        axis.text = element_text(size=16), 
        strip.text =element_text(size=14,face="bold"))

```


### Visualize MSC14
#### UMAPs of labels

```{r, fig.width = 14, fig.height = 4}
print(plot_grid(plotlist = list(DimPlot(msc_qcd_anotd_m14_scm_lrml, group.by = 'stage_lab1'),
                                DimPlot(msc_qcd_anotd_m14_scm_lrml, group.by = 'stageHL'),
                                DimPlot(msc_qcd_anotd_m14_scm_lrml, group.by = 'classifier_best_prediction')),
                nrow = 1
))

```

#### Heatmaps scmap VS LR-ML

```{r, fig.width = 10, fig.height = 10}
## function for plotting pheatmap for number of cells assigned labels from Clara's and Roser's datasets
phmap_lab_transfr <- function(dset_lab, col_labs, row_labs){
  anno_row = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>% 
    pivot_wider(values_from = n, names_from = {{ col_labs }}) %>% 
    column_to_rownames(names(.)[1]) %>% 
    mutate(LogReg_ML = "LogReg_ML") %>% 
    dplyr::select(LogReg_ML)
  anno_col = dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    t() %>%
    data.frame() %>%
    mutate(scmap = "scmap") %>%
    dplyr::select(scmap)
  dset_lab %>% dplyr::count({{ col_labs }}, {{ row_labs }}) %>%
    pivot_wider(values_from = n, names_from = {{ col_labs }}) %>%
    column_to_rownames(names(.)[1]) %>%
    pheatmap::pheatmap(display_numbers = T,treeheight_row=0, treeheight_col=0, annotation_row = anno_row, annotation_col = anno_col, 
                       legend = F, annotation_legend = F, cluster_rows =F, cluster_cols = F, fontsize = 26)
  
}


phmap_lab_transfr(msc_qcd_anotd_m14_scm_lrml@meta.data  %>%mutate(across(classifier_best_prediction, ~case_when(is.na(.) ~ 'missing', TRUE ~ .))), stageHL, classifier_best_prediction)
```



### MSCs UMAP coloured by stageHL {.tabset}

```{r, results='asis'}
purrr::imap(msc_qcd_anotd[1:4],  ~ {
    # create tabset for each group
    cat('####', .y, '   \n')
    
    p = DimPlot(.x, reduction = "umap", group.by = 'stageHL', 
                cols = sp_fld_colors , pt.size = 2,  label = T, repel = T, label.size = 7 ) +
      # labs(title = 'Gene expression clusters') +
      theme_void()+
      theme(#axis.text=element_text(size=18), 
            #axis.title=element_text(size=18,face="bold"), 
            axis.text=element_blank(), 
            axis.title=element_blank(),
            # title = element_text(size=18,face="bold"),
            title = element_blank(),
            aspect.ratio = 1/1,
            # legend.text = element_text(size=15),
            # legend.position = "right",
            # legend. = element_blank()#,
            # legend.position = "right"
            ) + 
      # guides(colour = guide_legend(nrow = 7, override.aes = list(size = 5))) +
      guides(colour = F)
    
    print(p)
    
    cat('\n')
    cat("\n")
  })

```












