---
title: "MSC strain deconvolution"
date: '2022-07-11'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---


```{r, message=FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(pheatmap)
  library(rmarkdown)
  library('vcfR')
  library(conflicted)
  library(RColorBrewer)
  library(MetBrewer)
  library(ComplexHeatmap)
  library(SeqArray)
  library(ggrepel)
  library(ggvenn)
  library(magick)
  library(VariantAnnotation)
  library(gridExtra)
  library(Matrix)
  library(UpSetR)
  library(SNPRelate)
  library(openxlsx)
  
  conflict_prefer("select", "dplyr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
})
```


# MSC parasites strain deconvolution and VCF visualization 

## Variable declarations

```{r}
##Strain cluster colours
strain_cols <- c(met.brewer(name = 'Austria')[c(1,7, 4,5,6)], 'light gray')

strain_cols_o_i <-c("#910000", "#f20000", "#007c71", "#00f2dd", "#c39a00", "#ffd94d", "#7570B3","#708db3", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","#a40000" , "#00b7a7", "#ffcd12", 'black', 'lavender', 'lightgray', 'grey')
names(strain_cols_o_i) <- c('G1_o', 'G1_i','G2_o',  'G2_i', 'G3_o',  'G3_i', 'G2_G1', 'G2_G3', 'G3_G1', 'G1_p_i','G1_b', 'G2_b', 'G3_b', 'G1', 'G2', 'G3', 'Doublet', 'G1_3i','Negative',  'Less5')
strain_cols_n <-c("#910000", "#ff1a8d", "#007c71", "#0ea300", "#00deca", "#9a96c7", "#5d57a4", "#a8875b","#708db3", "#c39a00", "#ffd94d", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","thistle","#a40000" , "#00b7a7", "#e7f1f9", 'black', 'grey')
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_n) <- c('SC2', 'SC1','SC5',  'SC4', 'SC3',  'SC6', 'SC7', 'SC8', 'S3_S1', 'S1_p_i','S1_b', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative', 'Lab', 'Less5', NA)

sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Atlas' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle'
)


donor_cols_m = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("msc1", "msc3","msc13","msc14"), '.mrna'))
donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(c("msc1", "msc3","msc13","msc14"))

stg_refnd_lvls = c('Early ring','Late ring','Early trophozoite','Late trophozoite','Early schizont','Late schizont', "Gametocyte (committed)","Gametocyte (developing)","Gametocyte (branching)", "Female (low count)", "Female (high count)" , "Gametocyte (early female)","Gametocyte (late female)","Male (low count)", "Male (high count)", "Gametocyte (early male)","Gametocyte (late male)", "Unassigned", "Failed QC", "Not assigned") %>% str_to_sentence(.)

trip="Mali"
```

```{r setup}
knitr::opts_knit$set(root.dir = paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/'))
```



```{r}

# system(paste0("hmmIBD -i ibd_analysis/msc14.strain_qcd_cln_gt.tsv  -o ibd_analysis/tr -m 100000"))

##Sample names
# sample_name_nms <- c("msc33", "msc48","msc49", "msc55", "msc60", "msc66") %>% sort()
sample_name_nms <- c("msc1", "msc3","msc13","msc14") %>% str_sort(numeric = T)
sample_name <- c("msc1", "msc3","msc13","msc14") %>% str_sort(numeric = T)


##Sample irods IDs
# source_irods_nms <- list("5736STDY13782211","5736STDY13782212","5736STDY13782213","5736STDY13782214","5736STDY13782215", "5736STDY13437393") %>% set_names(sample_name_nms)

# source_irods_nms <- list("5736STDY11771535","5736STDY11771536","5736STDY11771544") %>% set_names(sample_name_nms)


# sample_ids <- paste0(sample_name_nms, '_', source_irods_nms) %>% set_names(sample_name_nms)
# sample_ids <- paste0(sample_name_nms, '_', source_irods_nms) %>% set_names(sample_name_nms)
# optimum_k = c(8,6, 4,5,5,10) %>% set_names(sample_name_nms)
optimum_k = c(5,2,2,8) %>% set_names(sample_name_nms)

## sorted donor_opt_cluster names
sorted_names <- paste0(sample_name, "_", optimum_k)

grps <- c('stage_afm_strain_qcd', 'stage_ag_strain_qcd', 'strain_qcd')

trip = 'Mali'

##Alignment incorporation
algn_nms = c('minmap', 'hsat')
algn_nm = 'minmap'


sample_name_nms_al <- rep(sample_name_nms, 2)
algn_nms_al <- rep(algn_nms, each= length(sample_name_nms))

sm_aln_nms_al <- paste(rep(sample_name_nms, 2), rep(algn_nms, each= length(sample_name_nms)), sep ='_')

donor_cols_m_lcase = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("msc1", "msc3", "msc13", "msc14" ), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("msc1", "msc3", "msc13", "msc14" )))
donor_cols_lcase = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("msc1", "msc3", "msc13", "msc14" )))

donor_cols_m = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" ), '.mrna'))
# donor_cols = met.brewer("Egypt")[c(2,4,1,3)] %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))
donor_cols = c(met.brewer("Thomas")[c(1)],c('#F39EC8','#99ceff', '#b9e589')) %>% set_names(paste0(c("MSC1", "MSC3", "MSC13", "MSC14" )))


```


```{r, eval=T}
##Copy souporcell files for the optimum clusters


# for (k in c(6,8)) {
# for (smpl in c('stage_afm_strain_qcd', 'stage_ag_strain_qcd', 'strain_qcd')) {
for (id in sample_name_nms) {
    for (smpl in grps) {
    # for (smpl in c('stage_afm_strain')) {
      for (a in 1:length(algn_nms)) {
        # for (p in c(1,2)) {
        for (p in c(1)) {
          # for (al in c("multiallelic", "biallelic")) {
          system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', id,'/pbulk_gtypes/',smpl,'_k', optimum_k[id], '_',algn_nms[a],'/p', p)
            )
          system(paste0('rsync -av --include=\'*.vcf\' --include=\'vartrix*\' --include=\'vartrix*\\/SC*\' --include=\'vartrix*\\/D*\'  --include=\'vartrix*\\/N*\' --include=\'vartrix*\\/P*\' --exclude=\'*\' jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', id,'/pbulk_gtypes/',smpl,'_k', optimum_k[id], '_',algn_nms[a],'/p', p, '/ /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/',id,'/pbulk_gtypes/',smpl,'_k', optimum_k[id], '_',algn_nms[a],'/p', p)
            )
        }
      }
    }
}

```


VCF to GDS conversion
```{r, eval=T}
##Read in the genotypes and annotate

gt_ss_afm_p1_bi <-pmap(list(sample_name_nms, optimum_k), ~{
  sm <- ..1
  gp <- ..2
  map(grps, ~seqVCF2GDS(paste0("/Users/jr35/Google\ Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sm ,"/pbulk_gtypes/", .x,"_k", gp,"_", algn_nm,"/p1/fb_biallelic.vcf"), paste0("/Users/jr35/Google\ Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sm ,"/pbulk_gtypes/", .x,"_k", gp,"_", algn_nm,"/p1/fb_biallelic.gds"),  storage.option="ZIP_RA")
)
})

```

Prepare table for highly informative SNPs table for making bam files in farm by adding gene postions and formating appropriately for the format required by farm scripts

```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id)
  )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                 TRUE ~ as.character(.))))
'/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/msc14/pbulk_gtypes/stage_'
```



```{r, results="hide"}
##Merge SNP information above with annotations from plasmoDB to get gene positions etc for plotting
# dp  %>% filter_if(any(!is.na()))
gene_n_pos <- Pf3D7_genes_plasmoDB %>% 
  mutate(
    chr = str_extract(`Genomic Location (Gene)`, '.*v3'), 
    gn_start = str_remove_all(`Genomic Location (Gene)`, '.*v3:|,|\\.\\..*'), 
    gn_end = str_remove_all(`Genomic Location (Gene)`, '.*\\.\\.|,|\\(.*')
  ) %>%
  mutate(across(c(gn_start, gn_end), ~as.numeric(.))) %>% 
  dplyr::select(gene_id, chr, gn_start, gn_end, Gene) 

# write_csv(strain_top_snps, '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali2/data/processed/DB52e_star/strain_top_snps.csv')

```



filter pf7 variants
```{r}

pf7_waf <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/pf7_variants/AF-W-freqs.csv")

pf7_waf %>% 
  filter(is_invariant == F &  is_biallelic == T & minor_af > 0.3 & major_af < 0.7) %>% dim

pf7_waf_all <- data.table::fread("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/waf/Pf3D7_v3_g.txt")

pf7_waf_all_snp_qc <- pf7_waf_all %>% dplyr::select(V1, V2,V3,V4, paste0('V', c(11:17))) %>% rename(c('chrom' = V1, 'pos'=V2, 'ref'=V3, 'alt'=V4, 'filter'=V11,'type'=V12, 'VQSLOD'=V13,   'CDS'=V14, 'AC'=V15, 'AN'=V16, 'AF'=V17)) %>%  mutate(pf_all = 'yes')

```

filter pf7 variants
```{r}

pf7_mali <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/pf7_variants/Mali-freqs.csv")

pf7_mali_ibd_whitelst <- pf7_mali %>% 
  filter(is_invariant == F &  is_biallelic == T & minor_af > 0.1)

# pf7_waf <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_variants/AF-W-freqs.csv")
# 
# pf7_waf_ibd_whitelst <- pf7_waf %>% 
#   filter(is_invariant == F &  is_biallelic == T & minor_af > 0.1)

```


filter pf7 all variants
```{r}

pf7_mali_all <- data.table::fread("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/pf7_variants/all_variants.txt")

pf7_mali_all_snp_qc <- pf7_mali_all %>% dplyr::select(V1, V2, V10, V11, V12,  V13, V14, V15, V16) %>% rename(c('chrom' = V1, 'pos'=V2, 'filter'=V10, 'type'=V11, 'VQSLOD'=V12, 'CDS'=V13, 'AC'=V14, 'AN'=V15, 'AF'=V16)) %>%  mutate(pf_all = 'yes')


pf7_mali_only <- data.table::fread("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/pf7_variants/Pf3D7_v3_g_bisplit.txt", na.strings='.')
pf7_mali_only_snp_qc <- pf7_mali_only %>% dplyr::select(V1, V2,V3,V4, paste0('V', c(11:17))) %>% rename(c('chrom' = V1, 'pos'=V2, 'ref'=V3, 'alt'=V4, 'filter'=V11,'type'=V12, 'VQSLOD'=V13,   'CDS'=V14, 'AC'=V15, 'AN'=V16, 'AF'=V17)) %>%  mutate(pf_all = 'yes')

pf7_mali_only_nmajor001 <- data.table::fread("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/pf7_data/pf7_variants/Pf3D7_v3_g_pass_bisplit_snp_001.txt", na.strings='.')
pf7_mali_only_nmajor001_snp_qc <- pf7_mali_only_nmajor001 %>% dplyr::select(V1, V2,V3,V4, paste0('V', c(11:17))) %>% rename(c('chrom' = V1, 'pos'=V2, 'ref'=V3, 'alt'=V4, 'filter'=V11,'type'=V12, 'VQSLOD'=V13,   'CDS'=V14, 'AC'=V15, 'AN'=V16, 'AF'=V17)) %>%  mutate(pf_all = 'yes')

```


```{r}
##Write pf7 snps for sekou
pf7_mali_only_snp_qc %>% filter(type == "SNP" & filter == "PASS") %>% write_csv(., "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/collaborators/sekou_files/pf7_mali_only_snp_qc_pass.csv")
```

###########k8_raw
```{r}
## Function to read in the genotypes and annotate

gds_rd <- function(sample_id, grp, opt_k){  
  
  p1_bi <- seqOpen(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sample_id ,"/pbulk_gtypes/", grp, "_k", opt_k,"_", algn_nm,"/p1/fb_biallelic.gds"), allow.duplicate=T)
  print(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sample_id ,"/pbulk_gtypes/", grp, "_k", opt_k,"_", algn_nm,"/p1/fb_biallelic.gds"))
  
  print(head(seqGetData(p1_bi, "allele")))
  # print(head(seqGetData(p1_bi, "$dosage")))
  # print(head(seqGetData(p1_bi, "annotation/format/QR")))

  gtypes_p1_bi <- base::rbind(seqGetData(p1_bi, "allele"), seqGetData(p1_bi, "annotation/qual"),seqGetData(p1_bi, "chromosome"), seqGetData(p1_bi, "position"), seqGetData(p1_bi, "annotation/info/DP"), seqGetData(p1_bi, "annotation/filter"), seqGetData(p1_bi, "$dosage"), seqGetData(p1_bi, "annotation/format/DP"), seqGetData(p1_bi, "annotation/format/QR"), seqGetData(p1_bi, "annotation/format/QA")) %>% t()
  
  seqClose(p1_bi)
  print(gtypes_p1_bi[1:5,1:5])
  
  strn_nms <- seqVCF_SampID(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sample_id ,"/pbulk_gtypes/", grp, "_k", opt_k,"_", algn_nm,"/p1/fb_biallelic.vcf")) %>% str_replace_all(c("PoorQC"="P", "Low_n"="Lo", "Female"="F", "Male" = "M", "Asexual" = "A", "Gametocyte" = "G"))
  
  colnames(gtypes_p1_bi) <- c('alleles', 'quality', 'chrom', 'pos', 'depth', 'filt', strn_nms, paste0(strn_nms, rep(c('_DP','_RefN', '_AltN'), each=length(strn_nms))))

  return(gtypes_p1_bi)
  
}
```


###########k8_raw stage_afm_strain
```{r}
##Read in the genotypes and annotate
# gt_ss_afm_p1_bi <- gds_rd(sample_id = "msc48_5736STDY13782212", grp = "stage_afm_strain_k6_minimp")

gt_ss_afm_p1_bi <-pmap(list(sample_name_nms, optimum_k), ~{
  sm <- ..1
  k <- ..2
  map(grps, ~gds_rd(sample_id = sm, grp = .x, opt_k=k)) %>% set_names(paste0(..1 ,'.', grps))
  
})
```

###########k8_raw strain
```{r}
# ##k8 raw strain stage labels
# k6r_s_vcf_ids <- c("SC7", "SC6", "SC5", "SC3", "SC2", "SC1", "Doublet")
# 
# ##Read in the genotypes and annotate
# gt_s_k6r_p1_bi <- gds_rd(gds_path = "strain_qcd_k8_minimp/p1/fb_biallelic.gds", strn_nms = k6r_s_vcf_ids)
# dim(gt_s_k6r_p1_bi)
```

```{r}
##soupo k6
#seqClose(fb_p1_bi)

##Replace the clean cells genotype with the raw cells genotype
# gtypes_fb_p1_bi <-gt_ss_k6r_p1_bi
```

```{r}
##remove doublets as well
# gt_fb_vtx_p1_cx_bi <- gtypes_fb_p1_cx_bi %>% 
#   as.data.frame() %>% 
#   dplyr::select(!contains('Doublet')) %>%
#   dplyr::select(!contains('G1_i_F')) %>%
#   mutate(across(c(pos, depth, quality, matches('^DP.*|^Ref.*|^Alt.*'), matches('.*sNum$|.*cProp$')), as.numeric)) %>%
#   filter(quality>75 & depth>40)  #%>%
#   #filter(!(quality>200 & depth>30)
```

###k6 Clean
```{r}
# list.files("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/pbulk_gtypes/stg_strn_k8_minimp_qcd/p1/vartrix_biallelic/", pattern = "S.*.mtx", full.names = T, recursive = T, include.dirs = T) 
```


Vartrix matrix to get number of molecules supporting either reference or alternate allele - all

```{r}
##Function to read vartrix output and annotate
vtx_rd <- function(sample_id, grp, opt_k ){
  
  ##Read in the alternate allele matrix for all the stg/strn combinations and combine into a list - also shortening the names of the matrices (list items)
  alt_mtx <- list.files(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sample_id ,"/pbulk_gtypes/", grp, "_k", opt_k, "_", algn_nm,"/p1/vartrix_biallelic/"), pattern = "^[D|SC|G|P|N].*.mtx", full.names = T, recursive = T, include.dirs = T) %>% 
    set_names(nm = (str_remove_all(.,paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/", sample_id ,"/pbulk_gtypes/", grp, "_k", opt_k, "_", algn_nm,"/p1/vartrix_biallelic/", '/|.mtx')) %>% str_replace_all(c("PoorQC"="P", "Low_n"="Lo", "Female"="F", "Male" = "M", "Asexual" = "A", "Gametocyte" = "G", "_a_" = "_A_"))  %>%  tools::file_path_sans_ext())) %>%
    map(readMM) 
  
  ##Count, for each SNP, the total alternate allele count over all cells and proportion of cell that have alternate allele - Using base R because it is faster
  alt_mtx_stats = lapply(alt_mtx, function(x) base::cbind(x, rowSums(x) , rowSums(x != 0)/ncol(x))[,c(ncol(x)+1, ncol(x)+2)])
  
  print(c("1", sample_id, grp))
  
  ##Name the matrix columns with the name of the stg/strn and suffix of snpNum (total snp count over all cell) and celProp (proportion of cell that have alternate allele)
  nmd_alt_mt<- lapply(names(alt_mtx_stats), function(x) {
       colnames(alt_mtx_stats[[x]]) <- paste0(x, c('_sNum', '_cProp'))
       as.matrix(alt_mtx_stats[[x]])
   })
  
  print(c("2", sample_id, grp))
  
  nmd_alt_mt <- do.call(base::cbind, nmd_alt_mt) 
  print(c("finito", sample_id, grp))
  return(nmd_alt_mt)
  
}

```


```{r}
##Function to read vartrix output and annotate
# a=list.files(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/msc14/pbulk_gtypes/stage_afm_strain_qcd_k8_minmap/p1/vartrix_biallelic/"), pattern = "^[D|SC|G|P|N].*.mtx", full.names = T, recursive = T, include.dirs = T) %>% 
#     set_names(nm = (str_remove_all(.,paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/",trip,"/data/processed/DB52e_star/msc14/pbulk_gtypes/stage_afm_strain_qcd_k8_minmap/p1/vartrix_biallelic/", '/|.mtx')) %>% str_replace_all(c("PoorQC"="P", "Low_n"="Lo", "Female"="F", "Male" = "M", "Asexual" = "A", "Gametocyte" = "G", "_a_" = "_A_"))  %>%  tools::file_path_sans_ext())) %>%
#     map(readMM) 
#   
#   ##Count, for each SNP, the total alternate allele count over all cells and proportion of cell that have alternate allele - Using base R because it is faster
#   alt_mtx_stats = lapply(a, function(x) base::cbind(x, rowSums(x) , rowSums(x != 0)/ncol(x))[,c(ncol(x)+1, ncol(x)+2)])
#   
#   ##Name the matrix columns with the name of the stg/strn and suffix of snpNum (total snp count over all cell) and celProp (proportion of cell that have alternate allele)
#   nmd_alt_mt<- lapply(names(alt_mtx_stats), function(x) {
#        colnames(alt_mtx_stats[[x]]) <- paste0(x, c('_sNum', '_cProp'))
#        as.matrix(alt_mtx_stats[[x]])
#    })
#   
#   nmd_alt_mt <- do.call(base::cbind, nmd_alt_mt) 
#   
# }

```


###########k8_raw stage_afm_strain
```{r}
##Read in the genotypes and annotate
# gt_ss_afm_p1_bi <- gds_rd(sample_id = "msc48_5736STDY13782212", grp = "stage_afm_strain_k6_minimp")

ss_afm_vtx_mtx <-pmap(list(sample_name_nms, optimum_k), ~{
  sm <- ..1
  k <- ..2
  map(grps, ~vtx_rd(sample_id = sm, grp = .x, opt_k=k)) %>% set_names(paste0(..1 ,'.', grps))
})
```

###########k8_raw -TO BE DELETED IF RAW NOT PREFERED

```{r}
##Function to read stage afm + strain (ss) k8 raw (r) vartrix output and annotate
# ss_afm_vtx_mtx <- vtx_rd(sample_id = "msc48_5736STDY13782212", grp = "stage_afm_strain_k6_minimp")

```

```{r}
##Function to read stage ag + strain (ss) k8 raw (r) vartrix output and annotate
# ss_ag_k6r_vtx_mtx <- vtx_rd(vtx_path = 'stage_ag_strain_qcd_k8_minimp/p1/vartrix_biallelic/')

```

```{r}
##Function to read strain (s) k8 raw vartrix output and annotate
# s_k6r_vtx_mtx <- vtx_rd(vtx_path = 'strain_qcd_k8_minimp/p1/vartrix_biallelic/')

```


Combine vartix snp count per cell with freebayes calls
```{r}
##remove doublets as well and low quality SNPS - maybe not include doublets in variant calling with freebayes. Adds no value? or could be used to demonstrate high number of non-distinct and rule out small clusters as doublets and be part of QC
# fb_vt_bind <- function(fb_gtype , vtx_mtx ) {
#   gt_fb_vtx_p1_bi <- base::cbind(fb_gtype, vtx_mtx) %>% 
#     as.data.frame() %>% 
#     mutate(across(c(pos, depth, quality, matches('.*_DP$|.*_RefN$|.*_AltN$|.*sNum$|.*cProp$')), as.numeric)) %>%
#     filter(quality>75 & depth>40)  
#   
#   print(dim(gt_fb_vtx_p1_bi))
#   
#   ##remove  mitochondrial calls
#   gt_fb_vtx_p1_bi <- gt_fb_vtx_p1_bi %>%
#     filter(!str_detect(chrom, 'Pf_M'))
#   
#   print(dim(gt_fb_vtx_p1_bi))
#   
#   return(gt_fb_vtx_p1_bi)
#   
# }

# Same as above but option to remove some groups  

##remove doublets as well and low quality SNPS - maybe not include doublets in variant calling with freebayes. Adds no value? or could be used to demonstrate high number of non-distinct and rule out small clusters as doublets and be part of QC
fb_vt_bind <- function(fb_gtype , vtx_mtx, rm_grps="None" ) {
  gt_fb_vtx_p1_bi <- base::cbind(fb_gtype, vtx_mtx) %>% 
    as.data.frame() %>% 
    mutate(across(c(pos, depth, quality, matches('.*_DP$|.*_RefN$|.*_AltN$|.*sNum$|.*cProp$')), as.numeric)) %>%
    dplyr::select(-c(matches(rm_grps))) %>%
    filter(if_any(starts_with('SC'), ~!is.na(.))) %>%
    # filter(quality>75 & depth>40)  %>%
    mutate(filt_reason1 = case_when(quality<=75 | depth<=40 ~ 'Low quality',
                                    str_detect(chrom, 'Pf_M') ~ 'Mt'))
  
  print(dim(gt_fb_vtx_p1_bi))
  print(table(gt_fb_vtx_p1_bi$filt_reason1, useNA = 'ifany'))
  
  ##remove  mitochondrial calls
  # gt_fb_vtx_p1_bi <- gt_fb_vtx_p1_bi %>%
    # filter(!str_detect(chrom, 'Pf_M'))%>%
    # mutate(filt_reason1 = case_when(str_detect(chrom, 'Pf_M') ~ 'Mt'))
  
  print(dim(gt_fb_vtx_p1_bi))
  
  return(gt_fb_vtx_p1_bi)
}
```



```{r}
##strain stage k8 raw freebayes vcf + vatrix high quality no mitochondrial DNA
# ss_k6r_fb_vtx_p1bi <- fb_vt_bind(fb_gtype = gt_ss_k6r_p1_bi, vtx_mtx = ss_k6r_vtx_mtx, rm_grps = paste(c("SC8_F", "SC8_A", "SC7_A", "SC1_M", "SC1_F", "SC4_F", "SC4_M", "SC5_A.*"), collapse = '.*|'))
```

```{r}
##strain stage k8 raw freebayes vcf + vatrix high quality no mitochondrial DNA
ss_afm_fb_vtx_p1bi <- pmap(list(unlist(gt_ss_afm_p1_bi, recursive = F), unlist(ss_afm_vtx_mtx, recursive = F)), ~fb_vt_bind(fb_gtype = ..1, vtx_mtx = ..2))
```

```{r}
##strain stage k8 raw freebayes vcf + vatrix high quality no mitochondrial DNA
# ss_ag_k6r_fb_vtx_p1bi <- fb_vt_bind(fb_gtype = gt_ss_ag_k6r_p1_bi, vtx_mtx = ss_ag_k6r_vtx_mtx)
```

```{r}
##strain k8 raw freebayes vcf + vatrix high quality no mitochondrial DNA
# s_k6r_fb_vtx_p1bi <- fb_vt_bind(fb_gtype = gt_s_k6r_p1_bi, vtx_mtx = s_k6r_vtx_mtx)
```


VCF with minimum alternate allele fraction being 0.2 and ploidy 1 - NOT USED
```{r}
##Filter to retain those genotypes where the sum of quality for either alternate or reference alleles is > 50
# 
# gt_fb_vtx_p1_bi_qc_q <- gt_fb_vtx_p1_bi_qc %>%
#   mutate(G1_i_A_qc = case_when(G1_i_A_RefN >50 |G1_i_A_AltN >50 ~ G1_i_A_qc),
#          # G1_i_F_qc = case_when(G1_i_F_RefN >50 |G1_i_F_AltN >50 ~ G1_i_F_qc),
#          G1_i_F_qc = case_when(G1_i_F_RefN >50 |G1_i_F_AltN >50 ~ G1_i_F_qc),
#          G1_o_A_qc = case_when(G1_o_A_RefN >50 |G1_o_A_AltN >50 ~ G1_o_A_qc),
#          G1_o_F_qc = case_when(G1_o_F_RefN >50 |G1_o_F_AltN >50 ~ G1_o_F_qc),
#          G1_o_M_qc = case_when(G1_o_M_RefN >50 |G1_o_M_AltN >50 ~ G1_o_M_qc),
#          G2_i_A_qc = case_when(G2_i_A_RefN >50 |G2_i_A_AltN >50 ~ G2_i_A_qc),
#          G2_i_F_qc = case_when(G2_i_F_RefN >50 |G2_i_F_AltN >50 ~ G2_i_F_qc),
#          G2_i_M_qc = case_when(G2_i_M_RefN >50 |G2_i_M_AltN >50 ~ G2_i_M_qc),
#          G2_o_F_qc = case_when(G2_o_F_RefN >50 |G2_o_F_AltN >50 ~ G2_o_F_qc),
#          G2_o_M_qc = case_when(G2_o_M_RefN >50 |G2_o_M_AltN >50 ~ G2_o_M_qc),
#          G3_i_A_qc = case_when(G3_i_A_RefN >50 |G3_i_A_AltN >50 ~ G3_i_A_qc),
#          G3_o_F_qc = case_when(G3_o_F_RefN >50 |G3_o_F_AltN >50 ~ G3_o_F_qc),
#          G3_o_M_qc = case_when(G3_o_M_RefN >50 |G3_o_M_AltN >50 ~ G3_o_M_qc),
#          Doublet_A_qc = case_when(Doublet_A_RefN >50 |Doublet_A_AltN >50 ~ Doublet_A_qc)#,
#          # Doublet_A_qc = case_when(RefN_Doublet_A - AltN_Doublet_A >20 ~ Doublet_A)
#          ) #%>%
#   #filter(!if_all(ends_with('_qc'), ~ is.na(.)))

```

## SNPS from the pf7 paper without filtering

```{r}
##soupo k6 with pf7 all variants including poor quality

gt_pf7_merge <- function(gt_tbl){  
  gt_pf7_mali <- pf7_mali_only_snp_qc %>% 
  left_join(gt_tbl,., by = c("chrom", "pos"), multiple = "all")
  
  grpd_tbl <- gt_pf7_mali %>% group_by(chrom,pos) %>% mutate(across(type, ~factor(., levels = c('SNP', 'INDEL', 'OVERLAP')))) %>% arrange(type, .by_group = T) %>% arrange(dplyr::desc(AF), .by_group = T) 
    
  grpd_tbl %>% dplyr::select(c('chrom', 'pos','alleles','ref', 'alt',  'type', 'AF', 'AC')) %>% dim() %>% print()
  
  return(gt_pf7_mali)
}

```

```{r}
##strain stage k8 raw genotypes merge with pf Mali

ss_afm_pf7_mali <- map(ss_afm_fb_vtx_p1bi, ~gt_pf7_merge(gt_tbl=.x))

```

```{r}
##strain stage k8 raw genotypes merge with pf Mali

# ss_ag_k6r_pf7_mali <- gt_pf7_merge(gt_tbl=ss_ag_k6r_fb_vtx_p1bi, gt_tbl_nm='ss_ag_k6r_pf7_mali')
```

```{r}
##strain stage k8 raw genotypes merge with pf Mali

# s_k6r_pf7_mali <- gt_pf7_merge(gt_tbl=s_k6r_fb_vtx_p1bi, gt_tbl_nm='s_k6r_pf7_mali')
```


```{r}
##Visualize the distribution of SNPs, INDELs and overlaps within the split multi-alleles

ss_afm_pf7_mali[[1]] %>%  group_by(chrom,pos) %>% mutate(across(type, ~factor(., levels = c('SNP', 'INDEL', 'OVERLAP')))) %>% arrange(type, .by_group = T) %>% arrange(dplyr::desc(AF), .by_group = T) #%>% View()

# s_k6r_pf7_mali %>% group_by(chrom,pos) %>% mutate(across(type, ~factor(., levels = c('SNP', 'INDEL', 'OVERLAP')))) %>% arrange(type, .by_group = T) %>% arrange(dplyr::desc(AF), .by_group = T) %>% distinct(., pick(chrom, pos), .keep_all = T) %>% dplyr::select(c('chrom', 'pos','alleles','ref', 'alt',  'type', 'AF', 'AC')) %>% View()
```

```{r}
##Select only the SNP record from the split alleles and if more than one SNP in position then keep the most frequent alternate allele - distinct keeps the first row. No big difference when we dont first sort by type and just choose the most frequent alternate allele - SNP still top except for around 6 where indel swaps with SNP
## In vcf drop non-mali alleles, split after pass filter then select SNPs from this and maybe filter by allele frequency if too many SNPs
# merge with SNPs only after spliting

pf7_snps_qc_multitop <- function(gt_pf7_mali) {
  
  pf7_mali_nqc_snptop <- gt_pf7_mali %>% group_by(chrom,pos) %>% mutate(across(type, ~factor(., levels = c('SNP', 'INDEL', 'OVERLAP')))) %>% arrange(type, .by_group = T) %>% arrange(dplyr::desc(AF), .by_group = T) %>% distinct(., pick(chrom, pos), .keep_all = T) %>% ungroup()
  
}


```

```{r}
##strain k8 raw genotypes merge with pf7M (pf7Mali) retaining only the top SNPs from the Mali dataset where there are complex alleles with overlaps with indel etc

ss_afm_pf7M_snptop <- map(ss_afm_pf7_mali, ~pf7_snps_qc_multitop(gt_pf7_mali=.x))
```

```{r}
##strain k8 raw genotypes merge with pf7M (pf7Mali) retaining only the top SNPs from the Mali dataset where there are complex alleles with overlaps with indel etc

# ss_ag_k6r_pf7M_snptop <- pf7_snps_qc_multitop(gt_pf7_mali=ss_ag_k6r_pf7_mali)
```

```{r}
##strain k8 raw genotypes merge with pf7M (pf7Mali) retaining only the top SNPs from the Mali dataset where there are complex alleles with overlaps with indel etc

# s_k6r_pf7M_snptop <- pf7_snps_qc_multitop(gt_pf7_mali=s_k6r_pf7_mali)
```


```{r}
##Jacob filtered SNPs - some missing in MSC14
m14_pf7_prep_viz <- function(m14_gt, pf7_base = pf7_mali) {
  
  m14_pf7_all_jcb <- pf7_mali %>% 
    mutate(pf7_jcb = 'yes') %>%
    left_join(m14_gt,., by = c("chrom" = "chromosome", "pos" = "position"))  %>% 
    dplyr::select(c(chrom, pos, depth, type, filter,VQSLOD,CDS,AC,AN,AF, pf_all, quality, ends_with('_qc'), pf7_jcb,filt_reason1))%>% 
    mutate(p_dset = case_when(pf_all == 'yes' & pf7_jcb == 'yes' ~ 'm14_pAll_pJcb', 
                              pf_all == 'yes' & is.na(pf7_jcb) ~ 'm14_pAll', 
                              is.na(pf_all) & pf7_jcb == 'yes' ~ 'm14_pJcb', 
                              is.na(pf_all) & is.na(pf7_jcb)  ~ 'm14')) 
  
  m14_pf7_all_jcb %>% dplyr::count(p_dset, type, filter) %>%
    print()
  
  plt<- m14_pf7_all_jcb  %>% dplyr::count(p_dset, type, filter) %>% 
    filter(n>1) %>%
    mutate(across(filter, ~str_replace_all(., c(';'= '\n','Hyp'= '\nHyp')))) %>% 
    ggplot(., aes(x =p_dset, y=n, fill = type)) + 
    geom_col(position = position_dodge(width = 1)) + 
    facet_grid(.~filter) + 
    theme(axis.text.x = element_text(angle=90), text = element_text(size = 14)) 
    
  print(plt)
  
  return(m14_pf7_all_jcb)
  
}
```


```{r, fig.height=5, fig.width=10}
##strain stage k8 raw genotypes merge with pf Mali

ss_afm_pf7M_jcb <- map(ss_afm_pf7M_snptop, ~m14_pf7_prep_viz(m14_gt=.x))
```

```{r, fig.height=5, fig.width=10}
##strain stage k8 raw genotypes merge with pf Mali

# ss_ag_k6r_pf7M_jcb <- m14_pf7_prep_viz(m14_gt=ss_ag_k6r_pf7M_snptop)
```

```{r, fig.height=5, fig.width=10}
##strain stage k8 raw genotypes merge with pf Mali

# s_k6r_pf7M_jcb <- m14_pf7_prep_viz(m14_gt=s_k6r_pf7M_snptop)
```



```{r}
kas_togo_pf <- readxl::read_xlsx("data/raw/published_lit/kassegne_togo_pf_wgs/Table 1.XLSX", sheet = 'Table 1', skip = 1) %>% 
  rename(c('chrom'='CHROMOSOME', 'pos'='POS')) %>% 
  mutate(kas_togo = 'yes')
```


```{r, fig.height=5, fig.width=10}
ss_afm_pf7M_jcb[[3]] %>% 
  left_join(., kas_togo_pf, by = c('chrom', 'pos'))  %>% dplyr::count(pf_all, kas_togo)

# ss_ag_k6r_pf7M_jcb %>% 
#   left_join(., kas_togo_pf, by = c('chrom', 'pos')) %>% filter(is.na(filt_reason1)) %>% dplyr::count(pf_all, kas_togo)
# 
# s_k6r_pf7M_jcb %>% 
#   left_join(., kas_togo_pf, by = c('chrom', 'pos')) %>% filter(is.na(filt_reason1)) %>% dplyr::count(pf_all, kas_togo)
```


```{r, fig.height=5, fig.width=10}
regions_hyp <- read_delim("../Pf3D7_genomes_gtfs_indexs/pf7_data/regions-20130225.onebased.txt", col_names = F)
regions_hyp <- regions_hyp %>% mutate(chrom = as.numeric(str_remove_all(X1, 'Pf3D7_|_v3')))
```

mark snps not in pf7 dataset
```{r}
##remove doublets as well and low quality SNPS - maybe not include doublets in variant calling with freebayes. Adds no value? or could be used to demonstrate high number of non-distinct and rule out small clusters as doublets and be part of QC
pf7_hlght <- function(m14_gt ) {
  pf7_h <-  m14_gt %>%
    mutate(across(filt_reason1, ~case_when(is.na(.) & (filter != 'PASS' | is.na(filter)) ~ 'Absent in Pf7', 
                                           is.na(.) ~ 'Pass',
                                           T ~ as.character(.)))
           ) 
  
  pf7_h %>% dplyr::count(filt_reason1)%>% print()
  
  return(pf7_h)
}
```


```{r}
ss_afm_pf7M_pf7qc <- map(ss_afm_pf7M_snptop, ~pf7_hlght(m14_gt=.x))
```

```{r}
# ss_ag_k6r_pf7M_pf7qc <- pf7_hlght(m14_gt=ss_ag_k6r_pf7M_snptop)
```

```{r}
# s_k6r_pf7M_pf7qc <- pf7_hlght(m14_gt=s_k6r_pf7M_snptop)
```


```{r, fig.height=4, fig.width=15}
##Function for cleaner plot of SNPs falling in the different filter categories for each stage +/ strain cluster

broadqc_lvls <- c('Pass', 'Absent in Pf7', 'Mt', 'Low quality')
broadqc_cols <- met.brewer('Isfahan2')[c(2:4,1)] %>% set_names(broadqc_lvls)


cln_plt_snp_filt_categ_pf7qc <- function(annot_snp_qc, filt_c = 'filt_reason1', xlb = 'Strain', qc_cols = broadqc_cols,  qc_lvls = broadqc_lvls, filt_rm=''){
  
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
annot_snp_qc %>% 
  filter(!(!!rlang::sym(filt_c) %in% filt_rm)) %>%
  mutate(across(!!rlang::sym(filt_c), ~droplevels(factor(., levels = qc_lvls)))) %>%
  # dplyr::count(groups, !!rlang::sym(filt_c), .drop=F) %>%
    dplyr::count(!!rlang::sym(filt_c), .drop=F) %>%
  # group_by(groups) %>% mutate(perc = round((n/sum(n))*100, digits = 1)) %>%
    mutate(perc = round((n/sum(n))*100, digits = 1)) %>%
  ggplot(aes(x=!!rlang::sym(filt_c),fill=!!rlang::sym(filt_c), y = n/1000)) + 
  geom_col(position = position_dodge(width = 1)) + 
  geom_text(aes(label=paste0(n, "(", perc,")"), color = !!rlang::sym(filt_c)), vjust = 0.5, hjust = -0.1, position = position_dodge(width = 1), size = 6.5, angle = 90, show.legend = F, fontface= 'bold') + 
  scale_color_manual(values = qc_cols) +
  scale_fill_manual(values = qc_cols) +
  scale_y_continuous(expand = expansion(mult =c(0, 0.8))) + 
  # facet_wrap(.~groups, nrow = 1,scales = "free_x", drop = T) +
  labs(x = xlb, y = 'Number of SNPs (X 1e3)', fill = 'SNP category') +
  theme_classic()+
  theme(axis.text.y = element_text(size = 14),
        axis.title.x = element_blank(),
        strip.text.x.top = element_text(size = 15, face='bold'),
        axis.text.x = element_blank(),
        axis.title = element_text(size = 18, face = 'bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18, face = 'bold'),
        legend.position="bottom",
        panel.border=element_rect(colour="grey",fill='transparent'),#,
        # plot.margin = unit(c(1,2,1,1),"cm")
        # legend.margin=margin()
      ) +
    # coord_fixed()+
    # coord_flip() +
    guides(fill = guide_legend( nrow = 1, title.position = "top"))
}
```


```{r}
##!!!CHEATING - SORT OUT
ss_afm_broadqc_plt <- map(ss_afm_pf7M_pf7qc, ~cln_plt_snp_filt_categ_pf7qc(annot_snp_qc = .x, filt_c = 'filt_reason1', xlb = 'Strain', qc_cols = broadqc_cols,  qc_lvls = broadqc_lvls, filt_rm=''))

ss_afm_broadqc_plt

```

```{r}
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
# ss_ag_broadqc_plt <- cln_plt_snp_filt_categ_pf7qc(annot_snp_qc = ss_ag_k6r_pf7M_pf7qc, filt_c = 'filt_reason1', xlb = 'Strain', qc_cols = broadqc_cols,  qc_lvls = broadqc_lvls, filt_rm='')
# ss_ag_broadqc_plt
# 
# s_k6r_broadqc_plt <- cln_plt_snp_filt_categ_pf7qc(annot_snp_qc = s_k6r_pf7M_pf7qc, filt_c = 'filt_reason1', xlb = 'Strain', qc_cols = broadqc_cols,  qc_lvls = broadqc_lvls, filt_rm='')
# s_k6r_broadqc_plt

```

##Assess filtered SNPs

```{r}
##!!!! NOTE :Adding S to [G|D] [G|D|S] so as to detect the SC labeles in the k6 raw
##!!!! NOTE :Using instead of gt_fb_vtx_p1_bi_qc_pf7 since we want to only look at SNPs in pf 7 Mali
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
# snp_qc_tbl <- gt_fb_vtx_p1_bi_qc_pf7 %>% rename_with( ~str_replace(., '$','_raw'), matches("^[G|D|S].*_[A-Z]$")) %>% dplyr::select(chrom, pos,

snps_qc_filts <- function(m14_gt){
  
  snp_qc_tbl <- m14_gt %>% filter(filt_reason1 == 'Pass') %>% rename_with( ~str_replace(., '$','_raw'), matches("^(G|D|SC|P_|N).*(_A|_F|_M|_G|_Lo|_G|blet|tive|Unassigned|fail|[0-9])$")) %>% dplyr::select(chrom, pos, matches('.*_raw$|.*_qc$|.*_sNum$|.*_cProp$|.*_DP$|.*_RefN$|.*_AltN$'), type, filter,VQSLOD,CDS,AC,AN,AF, alleles)   %>% 
      pivot_longer(.,cols = matches('^(G|D|SC|P_|N).*'), 
                   names_pattern = "(^[G|D|SC|P_|N].*[A|M|F|G|Lo|t|tive|Unassigned|stgQC_fail|(0-9)])_(alt_sNum|alt_cProp|ref_sNum|ref_cProp|raw|qc|DP|RefN|AltN)", 
                   names_to = c( "groups" ,".value")) %>%
      mutate(filt_reason = case_when(filter != 'PASS' | is.na(filter) ~ 'Absent in Pf7', ##is.na(type) ~ 'Absent in pf7',### !!NB - Marking those missing in the pf7 dataset first. May need to be put last below if we want to include those absent if pf7 as some of these have been seen to be of good quality
                                     alt_sNum < 8 & ref_sNum < 8 ~  "Low UMI support",
                                     alt_sNum >= 15 & ref_sNum >= 15 ~  "Non distinct",
                                     abs(alt_sNum - ref_sNum) < 8 ~ 'Noisy' ,
                                     raw == '1' & alt_sNum > ref_sNum ~ 'Ref inverted' ,
                                     raw == '0' & ref_sNum > alt_sNum ~ 'Alt inverted',
                                     is.na(raw)&alt_sNum!=0&ref_sNum!=0  ~ "No fbayes call",
                                     str_detect(filter, 'Hyper') ~ 'Hypervariable',
                                     # is.na(type) ~ 'Absent in pf7',
                                     T ~ 'Clean'
                                     # !is.na(qc) ~ "Clean"
                                     )
             ) %>%
    mutate(qc2 = case_when(filt_reason == 'Clean' ~ raw))
  
  return(snp_qc_tbl)

  }
```

```{r}
ss_afm_pf7M_qc <- map(ss_afm_pf7M_pf7qc, ~snps_qc_filts(m14_gt=.x))

## Make snp_qc directory for storing qc files
imap(ss_afm_pf7M_qc, ~system(paste0('mkdir /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_remove(.y, '\\..*'),'/pbulk_gtypes/',str_remove(.y, 'm.*\\.'),'_k', optimum_k[str_remove(.y, '\\..*')],'_', algn_nm,'/p1/snp_qc')))

imap(ss_afm_pf7M_qc, ~write_csv(.x, paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_remove(.y, '\\..*'),'/pbulk_gtypes/',str_remove(.y, 'm.*\\.'),'_k', optimum_k[str_remove(.y, '\\..*')],'_', algn_nm,'/p1/snp_qc/ss_afm_pf7M_qc.csv')))


```


```{r}
# ss_ag_k6r_pf7M_qc <- snps_qc_filts(m14_gt=ss_ag_k6r_pf7M_pf7qc)
# 
# ss_ag_k6r_pf7M_qc %>% write_csv(., '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/msc48/pbulk_gtypes/stage_ag_strain_qcd_k8_minimp/p1/snp_qc/ss_ag_k6r_pf7M_qc.csv')

```

```{r}
# s_k6r_pf7M_qc <- snps_qc_filts(m14_gt=s_k6r_pf7M_pf7qc)
# 
# s_k6r_pf7M_qc %>% write_csv(., '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/msc48/pbulk_gtypes/strain_qcd_k8_minimp/p1/snp_qc/s_k6r_pf7M_qc.csv')

```

Minimal QC (lessQC [lqc]) retaining the "Noisy", "Non distinct", 'Ref inverted' and 'Alt inverted' categories
```{r}
##!!!! NOTE :Same script as above just commenting out the lines to remove "Noisy", "Non distinct", 'Ref inverted' and 'Alt inverted' categories - Need to find a clever way to do this to avoid repetition
##!!!! NOTE :Adding S to [G|D] [G|D|S] so as to detect the SC labels in the k6 raw
##!!!! NOTE :Using instead of gt_fb_vtx_p1_bi_qc_pf7 since we want to only look at SNPs in pf 7 Mali
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
# snp_qc_tbl <- gt_fb_vtx_p1_bi_qc_pf7 %>% rename_with( ~str_replace(., '$','_raw'), matches("^[G|D|S].*_[A-Z]$")) %>% dplyr::select(chrom, pos,

snps_qc_filts_lqc <- function(m14_gt){
    
    snp_qc_tbl <- m14_gt %>% filter(filt_reason1 == 'Pass') %>% rename_with( ~str_replace(., '$','_raw'), matches("^(G|D|SC|P_|N).*(_A|_F|_M|_G|_Lo|_G|blet|tive|Unassigned|fail|[0-9])$")) %>% dplyr::select(chrom, pos, matches('.*_raw$|.*_qc$|.*_sNum$|.*_cProp$|.*_DP$|.*_RefN$|.*_AltN$'), type, filter,VQSLOD,CDS,AC,AN,AF, alleles)   %>% 
        pivot_longer(.,cols = matches('^(G|D|SC|P_|N).*'), 
                     # names_pattern = "(^[G|D|S].*[A|M|F|blet])_([alt|ref|raw|qc|DP|RefN|AltN].*)",
                     # names_pattern = "(^[G|D|S].*[A|M|F|t])_(alt_sNum|alt_cProp|ref_sNum|ref_cProp|raw|qc|DP|RefN|AltN)", 
                     names_pattern = "(^[G|D|SC|P_|N].*[A|M|F|G|Lo|t|tive|Unassigned|stgQC_fail|(0-9)])_(alt_sNum|alt_cProp|ref_sNum|ref_cProp|raw|qc|DP|RefN|AltN)", 
                     names_to = c( "groups" ,".value")) %>%
        mutate(filt_reason = case_when(filter != 'PASS' | is.na(filter) ~ 'Absent in Pf7', ##is.na(type) ~ 'Absent in pf7',### !!NB - Marking those missing in the pf7 dataset first. May need to be put last below if we want to include those absent if pf7 as some of these have been seen to be of good quality
                                       # alt_sNum < 5 & ref_sNum < 5 ~  "Low UMI support",
                                       # alt_sNum >= 15 & ref_sNum >= 15 ~  "Non distinct",
                                       # abs(alt_sNum - ref_sNum) < 8 ~ 'Noisy' ,
                                       # raw == '1' & alt_sNum > ref_sNum ~ 'Ref inverted' ,
                                       # raw == '0' & ref_sNum > alt_sNum ~ 'Alt inverted',
                                       # is.na(raw)&alt_sNum!=0&ref_sNum!=0  ~ "No fbayes call",
                                       # str_detect(filter, 'Hyper') ~ 'Hypervariable',
                                       # is.na(type) ~ 'Absent in pf7',
                                       T ~ 'Clean'
                                       # !is.na(qc) ~ "Clean"
        )
        ) %>%
        mutate(qc2 = case_when(filt_reason == 'Clean' ~ raw))
    
    return(snp_qc_tbl)
    
}
```


```{r}
# s_lqc_k6r_pf7M_qc <- snps_qc_filts_lqc(m14_gt=s_k6r_pf7M_pf7qc)
ss_afm_pf7M_lqc <- map(ss_afm_pf7M_pf7qc, ~snps_qc_filts_lqc(m14_gt = .x))
# 
# s_lqc_k6r_pf7M_qc %>% write_csv(., '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/',trip,'/data/processed/msc48/pbulk_gtypes/strain_qcd_k8_minimp/p1/snp_qc/s_lqc_k6r_pf7M_qc.csv')

```

```{r}
# ss_afm_pf7M_qc <- map(ss_afm_pf7M_pf7qc, ~snps_qc_filts(m14_gt=.x))
```

```{r, fig.height=4, fig.width=20}
##Function to plot snps per the number of SNPs falling in the different filter categories for each stage + strain cluster
plt_snp_filt_categ <- function(annot_snp_qc, xlb = 'Strain'){
  
  annot_snp_qc %>% dplyr::count(groups, filt_reason) %>% 
  ggplot(aes(x=groups,fill=filt_reason, y = n)) + 
  geom_col(position = position_dodge(width = 1)) + 
  geom_text(aes(label=n, color = filt_reason), vjust = 0.5, hjust = -0.1, position = position_dodge(width = 1), size = 5, label.padding = unit(0.1, "lines"), angle = 90, show.legend = F) + 
  scale_y_continuous(expand = expansion(mult =c(0, 0.25))) + 
  labs(x = xlb, y = 'Number of SNPs', fill = 'SNP category') +
  theme_classic()+
  theme(axis.text.x = element_text(size = 16, angle = 90),
        axis.title = element_text(size = 18, face = 'bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18, face = 'bold')) 
  
}


```


```{r, fig.height=4, fig.width=20}
##Function to plot snps per the number of SNPs falling in the different filter categories for each stage + strain cluster
map(ss_afm_pf7M_qc, ~plt_snp_filt_categ(annot_snp_qc = .x))

```


```{r, fig.height=4, fig.width=20}
##Function to plot snps per the number of SNPs falling in the different filter categories for each stage + strain cluster
# plt_snp_filt_categ(annot_snp_qc = ss_ag_k6r_pf7M_qc, xlb = 'Strain stage combinations')

```


```{r, fig.height=4, fig.width=20}
##Function to plot snps per the number of SNPs falling in the different filter categories for each stage + strain cluster
# plt_snp_filt_categ(annot_snp_qc = s_k6r_pf7M_qc)
# plt_snp_filt_categ(annot_snp_qc = s_lqc_k6r_pf7M_qc)
```

Count those absent in pf7

```{r, fig.height=5, fig.width=14}
##SNPs in pf7
# ss_ag_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% distinct(pick(chrom,pos)) %>% dim
# s_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% distinct(pick(chrom,pos)) %>% dim
# 
# ##SNPs where there's less than 8 UMI
# ss_ag_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% distinct(chrom,pos) %>% dim
# s_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% distinct(chrom,pos) %>% dim
# 
# ##Genotypes where there's less than 8 UMI
# ss_ag_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dim
# s_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dim
# 
# ##Counts of genotypes where there's less than 8 UMI across groups (range medium)
# ss_ag_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dplyr::count(groups, filt_reason)
# s_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dplyr::count(groups, filt_reason)
# 
# ##Summary of genotypes where there's less than 8 UMI across groups (range medium)
# ss_ag_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dplyr::count(groups, filt_reason) %>% pull(n) %>% summary
# s_k6r_pf7M_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason == 'Low UMI support') %>% dplyr::count(groups, filt_reason) %>% pull(n) %>% summary

```

```{r, fig.height=4, fig.width=15}
##Function for cleaner plot of SNPs falling in the different filter categories for each stage +/ strain cluster

snp_qc_lvls <- c('Clean','Low UMI support', 'Non distinct', 'Noisy', 'Ref inverted', 'Alt inverted')
snp_qc_cols <- c(met.brewer('NewKingdom')[c(2,3,1,4,5)],met.brewer('Navajo')[c(2)])  %>% set_names(snp_qc_lvls)


cln_plt_snp_filt_categ <- function(annot_snp_qc, xlb = 'Strain'){
  
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
annot_snp_qc %>% 
  # filter(!filt_reason %in% c('Absent in Pf7', 'Low UMI support')) %>% 
  mutate(across(filt_reason, ~droplevels(factor(., levels = snp_qc_lvls)))) %>%
  dplyr::count(groups, filt_reason, .drop=F) %>%
    mutate(n_imp = as.numeric(case_when(filt_reason == 'Low UMI support' ~ NA, T ~ n)))%>%
  group_by(groups) %>% mutate(perc = round((n_imp/sum(n_imp, na.rm = T))*100, digits = 1)) %>%
  ggplot(aes(x=filt_reason,fill=filt_reason, y = n/1000)) + 
  geom_col(position = position_dodge(width = 1)) + 
  geom_text(aes(label=ifelse(is.na(n_imp), "", paste0(n_imp, "(", perc,")")), color = filt_reason), vjust = 0.5, hjust = -0.1, position = position_dodge(width = 1), size = 6.5, angle = 90, show.legend = F, fontface= 'bold') + 
  scale_color_manual(values = snp_qc_cols) +
  scale_fill_manual(values = snp_qc_cols) +
  scale_y_continuous(expand = expansion(mult =c(0, 0.05))) + 
  facet_wrap(.~groups, nrow = 1,scales = "free_x", drop = T) +
  labs(x = xlb, y = 'Number of SNPs (X 1e3)', fill = 'SNP category') +
  theme_classic()+
  theme(axis.text.y = element_text(size = 14),
        axis.title.x = element_blank(),
        strip.text.x.top = element_text(size = 15, face='bold'),
        axis.text.x = element_blank(),
        axis.title = element_text(size = 18, face = 'bold'),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18, face = 'bold'),
        legend.position="bottom",
        panel.border=element_rect(colour="grey",fill='transparent'),#,
        # plot.margin = unit(c(1,2,1,1),"cm")
        # legend.margin=margin()
      ) +
    # coord_fixed()+
    # coord_flip() +
    guides(fill = guide_legend( nrow = 1, title.position = "top"))
}
```

```{r, fig.height=6, fig.width=24}
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
##!!!CHEATING - SORT OUT
ss_afm_k6r_qc_plt <- map(ss_afm_pf7M_qc, ~cln_plt_snp_filt_categ(annot_snp_qc = .x))
ss_afm_k6r_qc_plt

```


```{r, fig.height=6, fig.width=24}
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
##!!!CHEATING - SORT OUT
ss_afm_k6r_lqc_plt <- map(ss_afm_pf7M_lqc, ~cln_plt_snp_filt_categ(annot_snp_qc = .x))
ss_afm_k6r_lqc_plt

```


```{r, fig.width= 12, fig.height=15.4}
##Put together the plots for annotation
sag_bqc <- ss_afm_broadqc_plt[str_detect(names(ss_afm_broadqc_plt), "\\.stage_ag_strain_qcd")]
s_bqc <- ss_afm_broadqc_plt[str_detect(names(ss_afm_broadqc_plt), "\\.strain_qcd")]
sag_nqc <- ss_afm_k6r_qc_plt[str_detect(names(ss_afm_k6r_qc_plt), "\\.stage_ag_strain_qcd")]
s_nqc <- ss_afm_k6r_qc_plt[str_detect(names(ss_afm_k6r_qc_plt), "\\.strain_qcd")]


pmap(list(sag_bqc, s_bqc, sag_nqc, s_nqc), ~{
        plot_grid(plot_grid(..1 +
            theme(plot.margin = margin(t=1.1, b=0.3, l=0.1,r=0.3, unit = "cm")),
                    ..2 +
            theme(plot.margin = margin(t=1.1, b=0.3,l=0.1,r=0.3, unit = "cm")), nrow = 1,
              labels = c('A', 'B'),
              label_size = 24,
              vjust= 1) ,
          ..3 +
            theme(plot.margin = margin(t=1.1, b=0.3, unit = "cm")),
          ..4 +
              theme(plot.margin = margin(t=1.1, b=0.3, unit = "cm")),
              nrow = 3,
              labels = c('', 'C', 'D'),
              label_size = 24,
              vjust= 1)
      }
)

``` 



```{r, fig.height=5, fig.width=14}
##Visualize and count the different classes of SNPs removed in QC
snp_filt_bias_viz <- function(annot_snp_qc, filt_cat){
  
  snps_filt_bias <- annot_snp_qc %>% dplyr::distinct(chrom, pos, filt_reason, groups)  %>% filter(filt_reason != 'Absent in Pf7') %>% group_by(chrom,pos) %>% filter(!all(filt_reason == 'Low UMI support')) 
  
  snps_filt_bias_fc <- snps_filt_bias %>% filter(any(filt_reason == filt_cat)) %>% group_by(chrom, pos) %>% ungroup() 
  # snps_filt_bias_fc %>%  pivot_wider(names_from = groups, values_from = filt_reason) %>% View()
  snps_filt_bias_fc %>%  pivot_wider(names_from = groups, values_from = filt_reason) %>% dim() %>% print()
  snps_filt_bias_fc %>% filter(filt_reason == filt_cat) %>% dplyr::count(groups) %>% print()
  snps_filt_bias_fc %>%  dplyr::count(filt_reason, groups) %>% print()
  snps_filt_bias %>% filter(any(filt_reason == filt_cat)) %>% filter(any(filt_reason == 'Clean'))  %>% group_by(chrom, pos) %>% filter(sum(filt_reason == filt_cat) > 2) %>% ungroup() %>%  pivot_wider(names_from = groups, values_from = filt_reason) %>% dim %>% print()

}
```


```{r, fig.height=5, fig.width=14}
##Visualize and count the different classes of SNPs removed in QC
for (fc in c('Non distinct', 'Noisy', 'Alt inverted', 'Ref inverted')){
  map(ss_afm_pf7M_qc, ~snp_filt_bias_viz(annot_snp_qc = .x, filt_cat = fc))
}
```


```{r, fig.height=5, fig.width=14}
##Visualize and count SNPs where some stage strain groups support 'Non distinct' or 'Noisy' and other groups support 'Clean'
ss_afm_pf7M_qc[["msc14.stage_ag_strain_qcd"]] %>% dplyr::distinct(chrom, pos, filt_reason, groups) %>% filter(filt_reason != 'Absent in Pf7') %>% filter(filt_reason != 'Absent in Pf7') %>% group_by(chrom,pos) %>% filter(!all(filt_reason == 'Low UMI support')) %>% filter(any(filt_reason == 'Non distinct' | filt_reason == 'Noisy')) %>% filter(any(filt_reason == 'Clean'))  %>% group_by(chrom, pos) %>% ungroup %>% dplyr::count(filt_reason, groups) 
#%>% View


```


```{r}
##Viz and Remove doublets and get informative SNPs
snp_rm_dblt_infomtv <- function(annot_snp_qc){
  annot_snp_qc %>% filter(str_detect(groups, "Doublet")) %>% distinct(., pick(chrom,pos), .keep_all = T) %>% mutate(id =paste0(chrom,'_',pos)) %>% pull(id) %>% length() %>% print()

  annot_snp_qc_nd <- annot_snp_qc %>% filter(!str_detect(groups, "Doublet"))%>% filter(!is.na(qc2))
  
  annot_snp_qc_nd %>% group_by(chrom, pos) %>% filter(n_distinct(qc2) == 1 ) %>% distinct(., pick(chrom,pos), .keep_all = T) %>% dim() %>% print()
  
  annot_snp_qc_infomtv <- annot_snp_qc_nd  %>% group_by(chrom, pos) %>% filter(n_distinct(qc2) ==2 ) %>% ungroup
  
  annot_snp_qc_infomtv %>% distinct(., pick(chrom,pos), .keep_all = T) %>% dim() %>% print()


return(list(annot_snp_qc_nd, annot_snp_qc_infomtv))

}
```

```{r, fig.height=5, fig.width=14}
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
ss_afm_pf7M_qc_infomtv <- map(ss_afm_pf7M_qc, ~snp_rm_dblt_infomtv(annot_snp_qc = .x))
# ss_ag_k6r_pf7M_qc_infomtv <- snp_rm_dblt_infomtv(annot_snp_qc = ss_ag_k6r_pf7M_qc) 
```

```{r, fig.height=5, fig.width=14}
##By retaining only SNPs with absolute difference between alt and ref >10 we also factor in depth per allele per cluster
ss_afm_pf7M_lqc_infomtv <- map(ss_afm_pf7M_lqc, ~snp_rm_dblt_infomtv(annot_snp_qc = .x))
# ss_ag_k6r_pf7M_qc_infomtv <- snp_rm_dblt_infomtv(annot_snp_qc = ss_ag_k6r_pf7M_qc) 
```

```{r}
pf_chr_size = read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-52_Pfalciparum3D7_Genome.size', col_names = c('chrom', 'size')) %>% mutate(chr = str_remove_all(chrom,'Pf3D7_0|Pf3D7_|_v3')) 
```

IBS megacells from Mi

#------------------------ 2. compute IBS for megaCells -----------------------------------------#####

#------------------------------------ #
# Compute ibs for pseudo 'mega' cells #
#------------------------------------ #


```{r}

gtypes_chr_cln<- function(snp_qc_fnl ){
  gtypes_chr_pos_pos <- snp_qc_fnl %>% dplyr::select(chrom, pos, groups, qc2, alleles) %>% pivot_wider(names_from = groups, values_from = qc2) %>% unite(., col="id", c(chrom, pos), sep = "_", remove = F) %>% mutate(chr = str_remove_all(chrom, 'Pf3D7_0|Pf3D7_|_v3'))
}
```



```{r}

ss_afm_pf7M_cln_gt<- map(ss_afm_pf7M_qc_infomtv, ~gtypes_chr_cln(snp_qc_fnl = .x[[1]]))
```

```{r}
## !!!! NOTE - Remove SC8 in MSC14 as shown to be poor quality
ss_afm_pf7M_cln_gt[c("msc14.stage_afm_strain_qcd", "msc14.stage_ag_strain_qcd", "msc14.strain_qcd")] <- ss_afm_pf7M_cln_gt[c("msc14.stage_afm_strain_qcd", "msc14.stage_ag_strain_qcd", "msc14.strain_qcd")] %>% map(., ~dplyr::select(.x, -c(contains("SC8")))) 
```


```{r}
ss_afm_pf7M_cln_gt_lqc <- map(ss_afm_pf7M_lqc_infomtv, ~gtypes_chr_cln(snp_qc_fnl = .x[[1]]))
```


```{r}
## !!!! NOTE - Remove SC8 in MSC14 as shown to be poor quality
ss_afm_pf7M_cln_gt_lqc[c("msc14.stage_afm_strain_qcd", "msc14.stage_ag_strain_qcd", "msc14.strain_qcd")] <- ss_afm_pf7M_cln_gt_lqc[c("msc14.stage_afm_strain_qcd", "msc14.stage_ag_strain_qcd", "msc14.strain_qcd")] %>% map(., ~dplyr::select(.x, -c(contains("SC8")))) 
```


```{r}

##Function to save tables in excel as sheets

createWorksheet <- function(x, y){
  wb <- createWorkbook()
  for(i in seq_along(x)){
    addWorksheet(wb, names(x)[i])
    writeData(wb, sheet = names(x)[i], x[[i]])
    saveWorkbook(wb, file = y, overwrite = TRUE)
  }
}

```



Export the SNPs with QC metrics and only clean SNPs for stage&strain combinations and just strain
```{r, eval=T}
##!!!! NOTE - MANUSCRIPT SNPS table

## subset SNP tables to write and rename lists/sheets 
### QC
snps_export_qc <- ss_afm_pf7M_qc[str_detect(names(ss_afm_pf7M_qc), "stage_ag|\\.strain_qcd")] %>% map(., ~rename_with(., ~paste0('Pf7_',.), c("filter","type","VQSLOD","CDS","AC","AN","AF")))
names(snps_export_qc) <- paste0(10:(length(snps_export_qc)+9), ".", str_replace_all(names(snps_export_qc), c("\\.stage_ag_strain_qcd" = " stage_strain QC", "\\.strain_qcd" = " strain QC", "msc" = "MSC")))

### Clean
snps_export_cln <- ss_afm_pf7M_cln_gt[str_detect(names(ss_afm_pf7M_cln_gt), "stage_ag|\\.strain_qcd")]  %>% map(., ~dplyr::select(.x, -c(chr))) 
names(snps_export_cln) <- paste0(2:(length(snps_export_cln)+1), ".", str_replace_all(names(snps_export_cln), c("\\.stage_ag_strain_qcd" = " stage_strain clean", "\\.strain_qcd" = " strain clean", "msc" = "MSC")))

createWorksheet(c(snps_export_cln, snps_export_qc), paste0("../",trip,"/v3_science_manuscript/tables/data S9 - Strain & stage-strain SNPs including QC categories.xlsx"))
```

```{r, fig.height=3.8, fig.width=4.2}
##Heatmap function

gtypes_chr_cln_ibs_hm <- function(gtypes_chr_pos){
  
  # ##Remove SC8 and SC1_F
  # gtypes_chr_pos <- snp_qc_tbl %>% filter(!(groups %in% c("SC8_F", "SC1_F")))%>% dplyr::select(chrom, pos, groups, qc2, alleles) %>% pivot_wider(names_from = groups, values_from = qc2) %>% unite(., col="id", c(chrom, pos), sep = "_", remove = F) %>% mutate(chr = str_remove_all(chrom, 'Pf3D7_0|Pf3D7_|_v3'))
  
  gtMat = gtypes_chr_pos %>% dplyr::select(-c(alleles,chrom, pos, chr))  %>% column_to_rownames("id") %>% mutate(across(everything(), ~as.numeric(str_replace(., '1', '2'))))%>% as.matrix() 
  
  
  gdsFile = tempfile()
    
  snpgdsCreateGeno(gdsFile, genmat = gtMat, sample.id = colnames(gtMat), 
                     snp.id = rownames(gtMat), snp.chromosome = gtypes_chr_pos$chr, 
                     snp.position = gtypes_chr_pos$pos, snp.allele = as.character(gtypes_chr_pos$alleles), snpfirstdim = TRUE)  
  
  gds = snpgdsOpen(gdsFile)
  
  gdsSub = snpgdsLDpruning(gds, ld.threshold = 0.2)
  
  ibs = snpgdsIBS(gds, num.thread = 4)
  
  rownames(ibs$ibs) = colnames(ibs$ibs) = colnames(gtMat)
    
    
  colFun = suppressWarnings(brewer.pal(100, 'Greens'))
  colFun = circlize::colorRamp2(seq(0.5, 1, length.out = length(colFun)), 
                                  colFun)
  # celltype_col = brewer.pal(8,'Dark2')[as.numeric(as.factor(sapply(strsplit(colnames(ibs$ibs),split=':'),'[',3)))]
  # names(celltype_col) = as.factor(sapply(strsplit(colnames(ibs$ibs),split=':'),'[',3))
  
  # hm = Heatmap(ibs$ibs, col = colFun, show_row_names = TRUE, 
  #                show_column_names = TRUE, show_row_dend = T, 
  #                show_column_dend = T, row_title_rot = 0, cluster_rows = F, cluster_columns = F)
  
  # colnames(a) <- stg_strn_ordr %>% str_replace_all(.,  c('G'='SC', '1_i'='1','1_o'='2', '2_i_A'='3_A', '2_i_M' = '4_M', '2_i_F' = '4_F', '2_o'='5','3_i'='6','3_o'='7') )
  # rownames(a) <- stg_strn_ordr %>% str_replace_all(.,  c('G'='SC', '1_i'='1','1_o'='2', '2_i_A'='3_A', '2_i_M' = '4_M', '2_i_F' = '4_F', '2_o'='5','3_i'='6','3_o'='7') )
    
  hm = Heatmap(ibs$ibs, name='IBS', col = colFun, show_row_names = TRUE, 
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = T, cluster_columns = T)
  
  mat<-ibs$ibs

  hm2<-Heatmap(mat, name='IBS', col = colFun, show_row_names = TRUE, 
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = T, cluster_columns = T, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), border='black',   heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
                  layer_fun = function(j, i, x, y, width, height, fill) {
                      v = pindex(mat, i, j)
                      l = v >0.98
                      grid.rect(x[l], y[l], width = width, height = height, gp = gpar(col = "#a4c2f4", lwd = 2, fill = "transparent"))
})
  ##Make rectangle heatmap https://jokergoo.github.io/2021/07/22/make-triangle-heatmap/
  
  od =  hclust(dist(mat))$order
  hm2_od = mat[od, od]
  
  hm3<- Heatmap(hm2_od, rect_gp = gpar(type = "none"), name='IBS', col = colFun, show_row_names = TRUE, row_names_side = c("right"),
                 # show_column_names = TRUE, 
                show_row_dend = T, 
                 show_column_dend = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
              heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
              border='black',
              cell_fun = function(j, i, x, y, w, h, fill) {
        if(as.numeric(x) <= 1 - as.numeric(y) + 1e-6) {
            grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
        }
    })
  
  print(hm)
  print(hm2)
  print(hm3)
}
```

```{r, fig.height=3.8, fig.width=4.2}
##Heatmap function

gtypes_chr_cln_ibs_mtx <- function(gtypes_chr_pos){
  
  # ##Remove SC8 and SC1_F
  # gtypes_chr_pos <- snp_qc_tbl %>% filter(!(groups %in% c("SC8_F", "SC1_F")))%>% dplyr::select(chrom, pos, groups, qc2, alleles) %>% pivot_wider(names_from = groups, values_from = qc2) %>% unite(., col="id", c(chrom, pos), sep = "_", remove = F) %>% mutate(chr = str_remove_all(chrom, 'Pf3D7_0|Pf3D7_|_v3'))
  
  gtMat = gtypes_chr_pos %>% dplyr::select(-c(alleles,chrom, pos, chr))  %>% column_to_rownames("id") %>% mutate(across(everything(), ~as.numeric(str_replace(., '1', '2'))))%>% as.matrix() 
  gtMat = gtMat[,sort(colnames(gtMat))]
  
  
  gdsFile = tempfile()
    
  snpgdsCreateGeno(gdsFile, genmat = gtMat, sample.id = colnames(gtMat), 
                     snp.id = rownames(gtMat), snp.chromosome = gtypes_chr_pos$chr, 
                     snp.position = gtypes_chr_pos$pos, snp.allele = as.character(gtypes_chr_pos$alleles), snpfirstdim = TRUE)  
  
  gds = snpgdsOpen(gdsFile)
  
  gdsSub = snpgdsLDpruning(gds, ld.threshold = 0.2)
  
  ibs = snpgdsIBS(gds, num.thread = 4)
  
  rownames(ibs$ibs) = colnames(ibs$ibs) = colnames(gtMat)
    
  mat<-ibs$ibs
   
}
```

```{r, fig.height=3.8, fig.width=4.2}
##Heatmap function

gtypes_ibs_heatmap <- function(mtx, dendo = F){
  
  colFun = suppressWarnings(brewer.pal(100, 'Greens'))
  colFun = circlize::colorRamp2(seq(0.5, 1, length.out = length(colFun)), 
                                  colFun)
   print(str(mtx))
   
  fa = str_extract(colnames(mtx), 'msc.*')
  print(str(fa))
  
  # c_d = cluster_between_groups(mtx, fa)
  # dendog = if(dendo == T) c_d else F
  dendog = if(dendo == T) cluster_between_groups(mtx, fa) else F
  print(str(fa))
  print(str(dendog))
  
  # fa = str_extract(colnames(mtx), 'msc.*')
  # # fa_col = c("a" = 2, "b" = 3, "c" = 4)
  # dend1 = cluster_between_groups(mtx, fa)
  

  hm = Heatmap(mtx, name='IBS', col = colFun, show_row_names = TRUE, 
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = dendog, cluster_columns = dendog)
  
  
  hm2<-Heatmap(mtx, name='IBS', col = colFun, show_row_names = TRUE, 
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = dendog, cluster_columns = dendog, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), border='black',   heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
                  layer_fun = function(j, i, x, y, width, height, fill) {
                      v = pindex(mtx, i, j)
                      l = v >0.98
                      grid.rect(x[l], y[l], width = width, height = height, gp = gpar(col = "#a4c2f4", lwd = 2, fill = "transparent"))
})
  
  print(hm)
  print(hm2)
  # print(hm3)
}
```

```{r, fig.height=4.1, fig.width=4.5}

ss_afm_ibs_mtx<- map(ss_afm_pf7M_cln_gt, ~gtypes_chr_cln_ibs_mtx(gtypes_chr_pos=.x))
ss_afm_ibs_hmap<- map(ss_afm_ibs_mtx[1:8], ~gtypes_ibs_heatmap(mtx=.x))
```


```{r, fig.height=4.1, fig.width=4.5}
## All genotypes done separately since when done together the error for DUPLICATE colnames appears
ss_cln_gt_tbl_afm <- ss_afm_pf7M_cln_gt[c(seq(1,12,3))] %>%
    imap(~ rename_with(.x, function(x) paste(.y, x, sep = "_"), -c(id,chrom, pos, alleles, chr))) %>%
    purrr::reduce(left_join, by = c('id','chrom', 'pos', 'alleles', 'chr')) %>% 
  rename_with(~str_remove(., '\\.st.*ain'), starts_with(c("msc")) )

ss_cln_gt_tbl_afm<-ss_cln_gt_tbl_afm[,sort(colnames(ss_cln_gt_tbl_afm))]

ss_cln_gt_tbl_ag <- ss_afm_pf7M_cln_gt[c(seq(2,12,3))] %>%
    imap(~ rename_with(.x, function(x) paste(.y, x, sep = "_"), -c(id,chrom, pos, alleles, chr))) %>%
    purrr::reduce(left_join, by = c('id','chrom', 'pos', 'alleles', 'chr')) %>% 
  rename_with(~str_remove(., '\\.st.*ain'), starts_with(c("msc")) )

ss_cln_gt_tbl_ag<-ss_cln_gt_tbl_ag[,sort(colnames(ss_cln_gt_tbl_ag))]

ss_cln_gt_tbl_s <- ss_afm_pf7M_cln_gt[c(seq(3,12,3))] %>%
    imap(~ rename_with(.x, function(x) paste(.y, x, sep = "_"), -c(id,chrom, pos, alleles, chr))) %>%
    purrr::reduce(left_join, by = c('id','chrom', 'pos', 'alleles', 'chr')) %>% 
  rename_with(~str_remove(., '\\.st.*ain'), starts_with(c("msc")) )

ss_cln_gt_tbl_s<-ss_cln_gt_tbl_s[,sort(colnames(ss_cln_gt_tbl_s))]

## 
ss_cln_all_mtx <- list(ss_cln_gt_tbl_afm, ss_cln_gt_tbl_ag, ss_cln_gt_tbl_s) %>% set_names(c("all.stage_afm_strain_qcd", "all.stage_afm_strain_qcd", "all.strain_qcd"))
```


```{r, fig.height=4.1, fig.width=4.5}
## gtype for clusters with more than 100
# strn_c_new_mdata <- readRDS("data/processed/DB52e_star/strn_c_new_mdata_m21.RDS")
# # strn_c_new_mdata <- readRDS("data/processed/DB52e_star/strn_c_new_mdata_m21.RDS")
# 
# clstr_200 <- map(strn_c_new_mdata, . %>% 
#       filter(cluster_ag_n >= 15) %>%
#       dplyr::count(Strain_qc, stage_ag) %>% 
#       filter(stage_ag == 'Gametocyte') %>%
#       filter(Strain_qc != 'PoorQC') %>%
#       unite(col = 'cluster',Strain_qc, stage_ag, sep = '_') %>%
#       pull(cluster) %>%
#       str_remove_all(., 'ametocyte|sexual')
# ) %>% imap(., ~paste(.x, .y, sep='_')) %>% unlist() %>% as.vector()
# 
# clstr_200_no55 <- map(strn_c_new_mdata, . %>% 
#       filter(cluster_ag_n >= 15) %>%
#       dplyr::count(Strain_qc, stage_ag) %>% 
#       filter(stage_ag == 'Gametocyte') %>%
#       filter(Strain_qc != 'PoorQC') %>%
#       unite(col = 'cluster',Strain_qc, stage_ag, sep = '_') %>%
#       pull(cluster) %>%
#       str_remove_all(., 'ametocyte|sexual')
# ) %>% .[names(.) != 'msc55'] %>% imap(., ~paste(.x, .y, sep='_')) %>% unlist() %>% as.vector()
# 
# ss_cln_gt_tbl_100 <- ss_cln_gt_tbl[c("id","chrom" ,"pos","alleles","chr",clstr_200_no55)]

```

```{r}
## Read in msc annotations 

msc_qcd_anotd_stgstrqc <- readRDS('../Mali/data/processed/DB52e_star/msc_qcd_anotd.RDS')
```

```{r}
##Clusters with more than 100

# clustr_numbers <- map(strn_c_new_mdata, . %>% 
#         filter(cluster_ag_n >= 20) %>%
#         dplyr::count(Strain_qc, stage_ag) %>%  
#       filter(stage_ag == 'Gametocyte') %>%
#         mutate(across(stage_ag, ~str_remove_all(., 'ametocyte|sexual'))) %>%
#         filter(Strain_qc != 'PoorQC') %>%
#         unite(col = 'cluster',Strain_qc, stage_ag, sep = '_'))%>% 
#     .[names(.) != 'msc55'] %>%
#     bind_rows(., .id='sample') %>%
#     unite(col = 'sample_cluster',cluster, sample, sep = '_')

clustr_numbers <- map(msc_qcd_anotd_stgstrqc, ~.x@meta.data %>% 
        filter(strain_ag_n >= 20) %>%
        dplyr::count(strain_qcd, stage_ag) %>%  
      filter(stage_ag == 'Gametocyte') %>%
        mutate(across(stage_ag, ~str_remove_all(., 'ametocyte|sexual'))) %>%
        filter(strain_qcd != 'PoorQC') %>%
        unite(col = 'cluster',strain_qcd, stage_ag, sep = '_'))%>% 
    # .[names(.) != 'msc55'] %>%
    bind_rows(., .id='sample') %>%
    unite(col = 'sample_cluster',cluster, sample, sep = '_')

```

```{r}
## Read in msc annotations 
# clustr_numbers_plusm14 <- msc_qcd_anotd_stgstrqc@meta.data %>% dplyr::count(stage_ag, strain_qcd) %>% filter(stage_ag == "Gametocyte" & !is.na(strain_qcd)) %>% mutate('sample_cluster'= paste(strain_qcd, str_remove(stage_ag, "ametocyte"), "msc14", sep = '_')) %>% select(sample_cluster, n) %>% bind_rows(., clustr_numbers)
```



```{r, fig.height=15, fig.width=20}

# gt_all_ibs_mtx<- gtypes_chr_cln_ibs_mtx(gtypes_chr_pos=ss_cln_gt_tbl)
# gt_all_ibs_hmap<- gtypes_ibs_heatmap(mtx=gt_all_ibs_mtx)

gt_all_ibs_mtx<- map(ss_cln_all_mtx, ~gtypes_chr_cln_ibs_mtx(gtypes_chr_pos=.x))
gt_all_ibs_hmap<- map(gt_all_ibs_mtx, ~gtypes_ibs_heatmap(mtx=.x))


# gt_all_ibs_100_mtx<- gtypes_chr_cln_ibs_mtx(gtypes_chr_pos=ss_cln_gt_tbl_100)
# gt_all_ibs_100_hmap<- gtypes_ibs_heatmap(mtx=gt_all_ibs_100_mtx, dendo = T)

```


```{r}
## Read in msc annotations 
clustr_numbers_afm <- map(msc_qcd_anotd_stgstrqc[1:4], ~.x@meta.data %>% 
        filter(strain_ag_n >= 30 & !is.na(strain_qcd)) %>%
        distinct(strain_qcd,stage_afm,strain_afm_n) %>% 
        mutate(across(stage_afm, ~str_remove_all(., 'emale|ale|sexual|ametocyte'))) %>%
        unite(col = 'cluster', strain_qcd, stage_afm, sep = '_'))  %>% 
    bind_rows(., .id='sample') %>%
  mutate(across(sample, ~str_remove_all(., 'sc'))) %>%
    unite(col = 'sample_cluster', sample, cluster, sep = '_')%>%
  mutate(across(sample_cluster, ~str_replace(., "MSC", "m")))

rownames(clustr_numbers_afm) <- NULL

clustr_numbers_ag <- map(msc_qcd_anotd_stgstrqc[1:4], ~.x@meta.data %>% 
        filter(strain_ag_n >= 30 & !is.na(strain_qcd)) %>%
        distinct(strain_qcd,stage_ag,strain_ag_n) %>% 
        mutate(across(stage_ag, ~str_remove_all(., 'emale|ale|sexual|ametocyte'))) %>%
        unite(col = 'cluster', strain_qcd, stage_ag, sep = '_'))  %>% 
    bind_rows(., .id='sample') %>%
  mutate(across(sample, ~str_remove_all(., 'sc'))) %>%
    unite(col = 'sample_cluster', sample, cluster, sep = '_')%>%
  mutate(across(sample_cluster, ~str_replace(., "MSC", "m")))

rownames(clustr_numbers_ag) <- NULL

clustr_numbers_s <- map(msc_qcd_anotd_stgstrqc[1:4], ~.x@meta.data %>% 
        filter(strain_ag_n >= 30 & !is.na(strain_qcd)) %>%
          add_count(strain_qcd, name = "strain_n") %>% 
        distinct(strain_qcd, strain_n) %>% 
        unite(col = 'cluster', strain_qcd, sep = '_'))  %>% 
    bind_rows(., .id='sample') %>%
  mutate(across(sample, ~str_remove_all(., 'sc')))%>% 
    unite(col = 'sample_cluster', sample, cluster, sep = '_')%>%
  mutate(across(sample_cluster, ~str_replace(., "MSC", "m")))

## This will replace the above once the donor labels msc are capitalized
clustr_numbers_s_cap <- map(msc_qcd_anotd_stgstrqc[1:4], ~.x@meta.data %>% 
        filter(strain_ag_n >= 30 & !is.na(strain_qcd)) %>%
          add_count(strain_qcd, name = "strain_n") %>% 
        distinct(strain_qcd, strain_n) %>% 
        unite(col = 'cluster', strain_qcd, sep = '_'))  %>% 
    bind_rows(., .id='sample') %>%
  mutate(across(sample, ~str_remove_all(., 'sc')))%>% 
    unite(col = 'sample_cluster', sample, cluster, sep = '_')#%>%
  #mutate(across(sample_cluster, ~str_replace(., "MSC", "m")))

rownames(clustr_numbers_ag) <- NULL
```



```{r, fig.height=8, fig.width=10.5}
##Merge strain numbers to IBD results
ibs_don_hm <- function(ibsmtx, clustr_numbers, strn_cat, start_no = 0.5, test_nm = "IBS"){
colFun = suppressWarnings(brewer.pal(100, 'Greens'))
colFun = circlize::colorRamp2(seq(start_no, 1, length.out = length(colFun)), 
                                  colFun)
  
sibd_pht_mtx <- ibsmtx
print(colnames(sibd_pht_mtx))
sibd_pht_mtx <- sibd_pht_mtx[,row.names(sibd_pht_mtx)]

# fa = str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]')
fa = str_remove(colnames(sibd_pht_mtx), '_qcd.*')
c_d = cluster_within_group(sibd_pht_mtx, fa)

# fr = str_extract(rownames(sibd_pht_mtx), 'ms.*[0-9]')
fr = str_remove(rownames(sibd_pht_mtx), '_qcd.*')
print(fr)
c_r = cluster_within_group(sibd_pht_mtx, fr)

# nms = unique(str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]'))
nms = unique(str_remove(colnames(sibd_pht_mtx), '_qcd.*'))
# sample_c_cols <- met.brewer("Thomas")[1:length(nms)] %>% set_names(nms)
sample_c_cols <- donor_cols_lcase

# sibd_pht_mtx_cln <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('G_msc' = 'm')))) %>% rename_with(~str_replace_all(., c('G_msc' = 'm')), starts_with('SC')) %>% column_to_rownames('sid') %>% as.matrix()
sibd_pht_mtx_cln <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('msc' = 'm', '_qcd'='')))) %>% rename_with(~str_replace_all(., c('msc' = 'm', '_qcd'='')), starts_with('msc')) %>% 
  left_join(., clustr_numbers, by = c('sid'='sample_cluster')) %>% mutate(across(sid, ~paste0(., ' [', !!rlang::sym(strn_cat), ']'))) %>% select(-c(!!rlang::sym(strn_cat))) %>%
  column_to_rownames('sid') %>% as.matrix()

  
sibd_pht<- Heatmap(sibd_pht_mtx_cln, name=test_nm, 
                   col = colFun, 
                   show_row_names = TRUE, 
                   row_names_side = c("left"), 
                   show_column_names = TRUE,  
                   show_row_dend = T,  
                   show_column_dend = F, 
                   row_title_rot = 0, 
                   cluster_rows = c_r, 
                   clustering_method_rows = "complete",
                   cluster_columns = c_d, 
                   column_names_gp = gpar(fontsize = 15), 
                   row_names_gp = gpar(fontsize = 15), 
                   heatmap_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15)),
                   bottom_annotation = HeatmapAnnotation(Donor = fa, 
                                                         col = list(Donor = sample_c_cols), 
                                                         annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), 
                                                         annotation_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15))),
                   column_split = length(nms), 
                   border = TRUE, 
                   left_annotation=rowAnnotation(Donor = fa, col = list(Donor = sample_c_cols), annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), show_legend = F), row_split = length(nms), 
                   column_title = NULL,
                   column_title_gp = gpar(fontsize = 16), 
                   row_title = NULL,
                   row_title_gp = gpar(fontsize = 16)
              # border='black',row_gap=unit(.015, "npc"), column_gap=unit(.015, "npc"), 
              ) %>%
  draw(., merge_legend = TRUE)

sibd_pht
}
```


```{r, fig.height=8, fig.width=10}

hms<- pmap(list(gt_all_ibs_mtx, list(clustr_numbers_afm, clustr_numbers_ag, clustr_numbers_s), c("strain_afm_n", "strain_ag_n", "strain_n")), ~ibs_don_hm(ibsmtx=..1, clustr_numbers=..2, strn_cat=..3))

```


```{r, fig.height=4.5, fig.width=6}

hms<- pmap(list(gt_all_ibs_mtx[3], list(clustr_numbers_s), c("strain_n")), ~ibs_don_hm(ibsmtx=..1, clustr_numbers=..2, strn_cat=..3))

```


```{r, fig.height=4.5, fig.width=6}
## Capitalize the donor labels - will need to sort out later for entire notebook
cap_gt_all_ibs_mtx <- gt_all_ibs_mtx[[3]] %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('msc' = 'MSC', '_qcd'='')))) %>% rename_with(~str_replace_all(., c('msc' = 'MSC', '_qcd'='')), starts_with('msc')) %>% column_to_rownames('sid') %>% as.matrix()

```

```{r, fig.height=4, fig.width=5}
##Merge strain numbers to IBD results

colFun = suppressWarnings(brewer.pal(100, 'Greens'))
colFun = circlize::colorRamp2(seq(0.5, 1, length.out = length(colFun)), 
                                  colFun)
  
sibd_pht_mtx <- cap_gt_all_ibs_mtx
print(colnames(sibd_pht_mtx))
sibd_pht_mtx <- sibd_pht_mtx[,row.names(sibd_pht_mtx)]

# fa = str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]')
fa = str_remove(colnames(sibd_pht_mtx), '_S.*')
c_d = cluster_within_group(sibd_pht_mtx, fa)

# fr = str_extract(rownames(sibd_pht_mtx), 'ms.*[0-9]')
fr = str_remove(rownames(sibd_pht_mtx), '_S.*')
print(fr)
c_r = cluster_within_group(sibd_pht_mtx, fr)

# nms = unique(str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]'))
nms = unique(str_remove(colnames(sibd_pht_mtx), '_S.*'))
# sample_c_cols <- met.brewer("Thomas")[1:length(nms)] %>% set_names(nms)
sample_c_cols <- donor_cols

row_labs <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(r_labs =str_remove(sid, "MS.*_")) %>% left_join(., clustr_numbers_s_cap, by = c('sid'='sample_cluster')) %>% mutate(across(r_labs, ~paste0(., ' [', !!rlang::sym("strain_n"), ']'))) %>% select(sid, r_labs) %>% deframe

col_labs <- str_remove(colnames(sibd_pht_mtx), "MS.*_") %>% set_names(colnames(sibd_pht_mtx))

  
sibd_pht<- Heatmap(sibd_pht_mtx, name="IBS", 
                   col = colFun, 
                   show_row_names = TRUE, 
                   row_labels = row_labs,
                   column_labels = col_labs,
                   row_names_side = c("left"), 
                   show_column_names = TRUE,  
                   show_row_dend = F,  
                   show_column_dend = F, 
                   row_title_rot = 0, 
                   cluster_rows = c_r, 
                   clustering_method_rows = "complete",
                   cluster_columns = c_d, 
                   column_names_gp = gpar(fontsize = 15), 
                   row_names_gp = gpar(fontsize = 15), 
                   heatmap_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15)),
                   bottom_annotation = HeatmapAnnotation(Donor = fa, 
                                                         col = list(Donor = sample_c_cols), 
                                                         annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), 
                                                         annotation_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15))),
                   column_split = length(nms), 
                   border = TRUE, 
                   left_annotation=rowAnnotation(Donor = fa, col = list(Donor = sample_c_cols), annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), show_legend = F), row_split = length(nms), 
                   column_title = NULL,
                   column_title_gp = gpar(fontsize = 16), 
                   row_title = NULL,
                   row_title_gp = gpar(fontsize = 16)
              # border='black',row_gap=unit(.015, "npc"), column_gap=unit(.015, "npc"), 
              ) %>%
  draw(., merge_legend = TRUE)

sibd_pht

```

```{r, fig.width= 12, fig.height=7}
##save svg for sk21
save_plot('plots/mali2021_all/fig4/sibd_pht.svg', plot_grid(ggdraw(sibd_pht)), base_width = 4, base_height = 5)
```


```{r, fig.height=8, fig.width=10.5, eval=FALSE}
# ##Merge strain numbers to IBD results - needs IBD!!!!!
# colFun = suppressWarnings(brewer.pal(100, 'Greens'))
# colFun = circlize::colorRamp2(seq(0.5, 1, length.out = length(colFun)), 
#                                   colFun)
#   
# sibd_pht_mtx <- gt_all_ibs_mtx
# 
# sibd_pht_mtx <- sibd_pht_mtx[,row.names(sibd_pht_mtx)]
# 
# # fa = str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]')
# fa = str_remove(colnames(sibd_pht_mtx), '_qcd.*')
# c_d = cluster_within_group(sibd_pht_mtx, fa)
# 
# # fr = str_extract(rownames(sibd_pht_mtx), 'ms.*[0-9]')
# fr = str_remove(rownames(sibd_pht_mtx), '_qcd.*')
# print(fr)
# c_r = cluster_within_group(sibd_pht_mtx, fr)
# 
# # nms = unique(str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]'))
# nms = unique(str_remove(colnames(sibd_pht_mtx), '_qcd.*'))
# # sample_c_cols <- met.brewer("Thomas")[1:length(nms)] %>% set_names(nms)
# sample_c_cols <- donor_cols
# 
# # sibd_pht_mtx_cln <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('G_msc' = 'm')))) %>% rename_with(~str_replace_all(., c('G_msc' = 'm')), starts_with('SC')) %>% column_to_rownames('sid') %>% as.matrix()
# sibd_pht_mtx_cln <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('msc' = 'm', '_qcd'='')))) %>% rename_with(~str_replace_all(., c('msc' = 'm', '_qcd'='')), starts_with('msc')) %>% 
#   left_join(., clustr_numbers, by = c('sid'='sample_cluster')) %>% mutate(across(sid, ~paste0(., ' [', strain_afm_n, ']'))) %>% select(-c(strain_afm_n)) %>%
#   column_to_rownames('sid') %>% as.matrix()
# 
#   
# sibd_pht<- Heatmap(sibd_pht_mtx_cln, name='IBS', 
#                    col = colFun, 
#                    show_row_names = TRUE, 
#                    row_names_side = c("left"), 
#                    show_column_names = TRUE,  
#                    show_row_dend = T,  
#                    show_column_dend = F, 
#                    row_title_rot = 0, 
#                    cluster_rows = c_r, 
#                    clustering_method_rows = "complete",
#                    cluster_columns = c_d, 
#                    column_names_gp = gpar(fontsize = 15), 
#                    row_names_gp = gpar(fontsize = 15), 
#                    heatmap_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15)),
#                    bottom_annotation = HeatmapAnnotation(Donor = fa, 
#                                                          col = list(Donor = sample_c_cols), 
#                                                          annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), 
#                                                          annotation_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15))),
#                    column_split = length(nms), 
#                    border = TRUE, 
#                    left_annotation=rowAnnotation(Donor = fa, col = list(Donor = sample_c_cols), annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), show_legend = F), row_split = length(nms), 
#                    column_title = NULL,
#                    column_title_gp = gpar(fontsize = 16), 
#                    row_title = NULL,
#                    row_title_gp = gpar(fontsize = 16)
#               # border='black',row_gap=unit(.015, "npc"), column_gap=unit(.015, "npc"), 
#               ) %>%
#   draw(., merge_legend = TRUE)
# 
# sibd_pht

# Legend(labels = month.name[1:10], legend_gp = gpar(fill = 1:10), 
#     title = "foo", nrow = 3)
```



Read in MSC14 genotypes
```{r, fig.height=5, fig.width=6}
# m14_ss_ag_k8r_pf7M_cln_gt <- read_tsv("/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_ag_k8r_pf7M_cln_gt.tsv") %>% rename_with(~paste0(., "_msc14"), starts_with("SC"))
```


```{r}
##!!NOTE - gfv_p1_bi_strn_pf7_c is created somewhere downstream so this is likely to give an error

# ss_cln_gt_tbl_100 %>% dplyr::select(-c(id, chrom, alleles))%>% dplyr::select(chr, pos, any_of(sort(colnames(.)))) %>% filter(!if_all(starts_with('SC'), ~is.na(.))) %>% mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% write_tsv(., paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21.tsv"))
# 
# ##merge m14 with other donors
# ss_cln_gt_100_plus_m14 <-m14_ss_ag_k8r_pf7M_cln_gt %>% mutate(across(-pos, as.character)) %>% mutate(across(starts_with('SC'), ~na_if(.,"-1"))) %>% select(-c(contains("_A_"))) %>% full_join(.,ss_cln_gt_tbl_100, by=c('chr', 'pos')) %>% mutate(id= paste0(chr, "_", pos))
# 
# ss_cln_gt_100_plus_m14 %>% dplyr::select(-c(id, chrom, alleles))%>% dplyr::select(chr, pos, any_of(sort(colnames(.)))) %>% filter(!if_all(starts_with('SC'), ~is.na(.))) %>% mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% arrange(chr, pos) %>% write_tsv(., paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21_plus_m14.tsv"))

```


```{r}
##Run IBD 
wdir<- '/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/sware/hmmIBD-master'

system(paste0(wdir,'/hmmIBD -i ',wdir,'/ss_cln_gt_tbl_100_m21.tsv -o ',wdir,'/ss_cln_gt_tbl_100_m21 -m 100000'))

```


```{r}
wdir<- '/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master'

ss_cln_gt_tbl_100_hmmibd <- read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21.hmm.txt')
ss_cln_gt_tbl_100_hmmibd_f <- read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21.hmm_fract.txt')

ss_cln_gt_tbl_100_plus_m14_hmmibd <- read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21_plus_m14.hmm.txt')
ss_cln_gt_tbl_100_plus_m14_hmmibd_f <- read_delim('/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/ss_cln_gt_tbl_100_m21_plus_m14.hmm_fract.txt')


```




```{r, fig.height=5, fig.width=6}
colFun = suppressWarnings(brewer.pal(100, 'Blues'))
colFun = circlize::colorRamp2(seq(0, 1, length.out = length(colFun)), colFun)
  
cols_rand <- c("#fb9a99", "#cab2d6", "#b2df8a", "#e31a1c", "#ffff99", "#33a02c", "#ff7f00", "#b15928", "#fdbf6f")
```



```{r, fig.height=8, fig.width=10.5, eval=FALSE}
##Merge strain numbers to IBD results
# ss_n_ibd_tbl100 = ss_cln_gt_tbl_100_hmmibd_f %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., clustr_numbers, by = c('sample1'='sample_cluster')) %>% left_join(., clustr_numbers, by = c('sample2'='sample_cluster'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1, '\n[', n_1, ']'), id2 = paste0(sample2, '\n[', n_2, ']'))  
ss_n_ibd_tbl100 = ss_cln_gt_tbl_100_plus_m14_hmmibd_f %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., clustr_numbers_plusm14, by = c('sample1'='sample_cluster')) %>% left_join(., clustr_numbers_plusm14, by = c('sample2'='sample_cluster'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1), id2 = paste0(sample2))  

##Get the IBD and the number of sites for each pairwise comparison to annotate heatmap
s_ibdf_nsites_anno <- ss_n_ibd_tbl100 %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., ss_n_ibd_tbl100) %>% mutate(ibdf_nsites = paste0(round(fract_sites_IBD,2), '\n(',N_informative_sites,')'))%>% dplyr::select(starts_with('id'), ibdf_nsites) %>% pivot_wider(names_from = id2, values_from = ibdf_nsites) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., '1')))%>% column_to_rownames('id1')


##Create mirrored matrix and plot
sibd_pht_mtx <- ss_n_ibd_tbl100 %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., ss_n_ibd_tbl100) %>% dplyr::select(starts_with('id'), fract_sites_IBD) %>%  pivot_wider(names_from = id2, values_from = fract_sites_IBD) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., 1)))%>% column_to_rownames('id1')  
  

sibd_pht_mtx <- sibd_pht_mtx[,row.names(sibd_pht_mtx)]

fa = str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]')
c_d = cluster_within_group(sibd_pht_mtx, fa)

fr = str_extract(rownames(sibd_pht_mtx), 'ms.*[0-9]')
print(fr)
c_r = cluster_within_group(sibd_pht_mtx, fr)

nms = unique(str_extract(colnames(sibd_pht_mtx), 'ms.*[0-9]'))
# sample_c_cols <- met.brewer("Thomas")[1:length(nms)] %>% set_names(nms)
sample_c_cols <- cols_rand[1:length(nms)] %>% set_names(nms)

sibd_pht_mtx_cln <- sibd_pht_mtx %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('G_msc' = 'm')))) %>% rename_with(~str_replace_all(., c('G_msc' = 'm')), starts_with('SC')) %>% column_to_rownames('sid') %>% as.matrix()

  
sibd_pht<- Heatmap(sibd_pht_mtx_cln, name='IBD', 
                   col = colFun, 
                   show_row_names = TRUE, 
                   row_names_side = c("left"), 
                   show_column_names = TRUE,  
                   show_row_dend = T,  
                   show_column_dend = F, 
                   row_title_rot = 0, 
                   cluster_rows = c_r, 
                   cluster_columns = c_d, 
                   column_names_gp = gpar(fontsize = 15), 
                   row_names_gp = gpar(fontsize = 15), 
                   heatmap_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15)),
                   bottom_annotation = HeatmapAnnotation(Donor = fa, 
                                                         col = list(Donor = sample_c_cols), 
                                                         annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), 
                                                         annotation_legend_param = list(title_gp = gpar(fontsize = 15, fontface = "bold"), labels_gp = gpar(fontsize = 15))),
                   column_split = length(nms), 
                   border = TRUE, 
                   left_annotation=rowAnnotation(Donor = fa, col = list(Donor = sample_c_cols), annotation_name_gp = gpar(fontsize = 15, fontface = 'bold'), show_legend = F), row_split = length(nms), 
                   column_title = NULL,
                   column_title_gp = gpar(fontsize = 16), 
                   row_title = NULL,
                   row_title_gp = gpar(fontsize = 16)
              # border='black',row_gap=unit(.015, "npc"), column_gap=unit(.015, "npc"), 
              ) %>%
  draw(., merge_legend = TRUE)

sibd_pht

# 
# hm1<- Heatmap(sibd_pht_mtx_cln, rect_gp = gpar(type = "none"), name='IBD', col = colFun, show_row_names = TRUE, row_names_side = c("right"),
#                  # show_column_names = TRUE, 
#                 show_row_dend = T, cluster_rows = c_r, cluster_columns = c_d,
#                  show_column_dend = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
#               heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
#     top_annotation = HeatmapAnnotation(Sample = fa, col = list(Sample = sample_c_cols)),column_split = length(nms), border = TRUE, right_annotation=rowAnnotation(Sample = fa, col = list(Sample = sample_c_cols)), row_split = length(nms),
#               cell_fun = function(j, i, x, y, w, h, fill) {
#         if(as.numeric(x) <= 1 - as.numeric(y) + 1e-6) {
#             grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
#         }
#     })
#   
# hm1



# od_a = c_r %>% labels
# sibd_pht_mtx_t = sibd_pht_mtx[od_a, od_a]
# 
# sibd_pht_mtx_t_cln <- sibd_pht_mtx_t %>% as.data.frame()%>% rownames_to_column('sid') %>% mutate(across(sid, ~str_replace_all(., c('G_msc' = 'm')))) %>% rename_with(~str_replace_all(., c('G_msc' = 'm')), starts_with('SC')) %>% column_to_rownames('sid') %>% as.matrix()
# 
# nmslab = str_extract(od_a, 'ms.*[0-9]')
# nms = unique(nmslab)
# sample_c_cols <- met.brewer("Thomas")[1:length(nms)] %>% set_names(nms)
#   
# hm3<- Heatmap(sibd_pht_mtx_t_cln,rect_gp = gpar(type = "none"),  name='IBS', col = colFun, show_row_names = TRUE, row_names_side = c("right"),
#                  # show_column_names = TRUE, 
#                 show_row_dend = T, cluster_rows = F, cluster_columns = F,
#                  show_column_dend = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
#               heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
#     top_annotation = HeatmapAnnotation(Sample = nmslab, col = list(Sample = sample_c_cols)), border = TRUE, right_annotation=rowAnnotation(Sample = nmslab, col = list(Sample = sample_c_cols)), row_split = factor(nmslab, levels = nms),column_split = factor(nmslab, levels = nms), 
#               cell_fun = function(j, i, x, y, w, h, fill) {
#         if(as.numeric(x) <= 1 - as.numeric(y) + 1e-6) {
#             grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
#         }
#     })
# hm3
Legend(labels = month.name[1:10], legend_gp = gpar(fill = 1:10), 
    title = "foo", nrow = 3)
```


```{r, fig.height=5, fig.width=6, eval=FALSE}
##Merge strain numbers to IBD results
s_strn_n_ibd_tbl = s_k6r_pf7M_cln_gt_hmmibd_f %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., strn_qcd_n, by = c('sample1'='strain_qcd')) %>% left_join(., strn_qcd_n, by = c('sample2'='strain_qcd'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1, '\n[', n_1, ']'), id2 = paste0(sample2, '\n[', n_2, ']'))  

##Get the IBD and the number of sites for each pairwise comparison to annotate heatmap
s_combi_ibdf_nsites_anno <- s_strn_n_ibd_tbl %>% 
  rename('id2' = id1,  'id1' = id2) %>% 
  bind_rows(., s_lqc_strn_n_ibd_tbl) %>% 
  mutate(ibdf_nsites = paste0(round(fract_sites_IBD,2), '\n(',N_informative_sites,')'))%>% 
  dplyr::select(starts_with('id'), ibdf_nsites) %>% 
  pivot_wider(names_from = id2, values_from = ibdf_nsites) %>% 
  arrange(id1) %>%
  mutate(across(starts_with('SC'), ~replace_na(., '1')))%>%
  column_to_rownames('id1')


##Create mirrored matrix and plot
##!!! TIP - Refer to this if you want more complex arrangements https://github.com/jokergoo/ComplexHeatmap/issues/712 - eg different colours for the different half matrices
s_combi_ibd_pht <- s_strn_n_ibd_tbl %>% 
  rename('id2' = id1,  'id1' = id2) %>% 
  bind_rows(., s_lqc_strn_n_ibd_tbl) %>% 
  dplyr::select(starts_with('id'), fract_sites_IBD) %>%  
  pivot_wider(names_from = id2, values_from = fract_sites_IBD) %>% 
  arrange(id1) %>% 
  mutate(across(starts_with('SC'), ~replace_na(., 1)))%>%
  column_to_rownames('id1') %>% 
  as.matrix()%>%
  Heatmap(., name='IBD', col = colFun, show_row_names = TRUE, row_names_side = c("left"),
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = F, cluster_columns = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
              heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),border='black',
              cell_fun = function(j, i, x, y, width, height, fill) {
        grid.text(sprintf("%s", s_combi_ibdf_nsites_anno[i, j]), x, y, gp = gpar(fontsize = 13))
})

s_combi_ibd_pht
  
```

#########€#€######€€€€€€€

```{r}
# m = matrix(rnorm(120), nc = 12)
# colnames(m) = 1:12
# fa = rep(c("a", "b", "c"), times = c(2, 4, 6))
# # fa_col = c("a" = 2, "b" = 3, "c" = 4)
# dend1 = cluster_between_groups(m, fa)
# 
# Heatmap(m, cluster_columns = dend1, 
#     row_title = "cluster_between_groups")
# 
# m
# fa
# fa_col
```


```{r, fig.height=4.1, fig.width=4.5}

ss_afm_hmap_ibs<- map(ss_afm_pf7M_cln_gt, ~gtypes_chr_cln_ibs_hm(gtypes_chr_pos=.x))
```

```{r, fig.height=4.1, fig.width=4.5}
## Remove SC8
ss_afm_pf7M_cln_gt_m14_noSC8 <- ss_afm_pf7M_cln_gt[c("msc14.stage_ag_strain_qcd", "msc14.strain_qcd")] %>% map(~.x %>% select(!contains("SC8")))
ss_afm_hmap_ibs_m14_noSC8 <- map(ss_afm_pf7M_cln_gt_m14_noSC8, ~gtypes_chr_cln_ibs_hm(gtypes_chr_pos=.x))

ss_afm_hmap_ibs_m14_noSC8
```


###########k8_raw -TO BE DELETED IF k8 RAW NOT PREFERED


```{r}
##Function to creat pairwise comparison groups which are the used for plotting chromosome paints
strn_pw_comps <- function(gtypes_chr_pos, strn_regex){
  
  snp_pf7_afm <- gtypes_chr_pos %>% dplyr::select(id, chr, pos, starts_with(strn_regex)) #%>%  dplyr::select(!matches('^SC[0-9]_.$')) 
  
  col_pairs_strn <- snp_pf7_afm %>% dplyr::select(starts_with(strn_regex)) %>% colnames() %>% sort() %>% combn(., 2, simplify = F)
  
  strn_afm_comps <- map_dfc(col_pairs_strn, function(pair) {
    snp_pf7_afm %>%
      transmute(!!paste0(pair[1], "_vs_", pair[2]) := case_when(!is.na(.data[[pair[1]]]) & !is.na(.data[[pair[2]]]) & .data[[pair[1]]] == .data[[pair[2]]] ~ TRUE, 
                                                                !is.na(.data[[pair[1]]]) & !is.na(.data[[pair[2]]]) & .data[[pair[1]]] != .data[[pair[2]]] ~ FALSE))
  })
  
  strn_afm_comps <- base::cbind(snp_pf7_afm %>% dplyr::select(chr, pos), 
                                         strn_afm_comps)  %>%
    filter(!if_all(starts_with(strn_regex), ~is.na(.)))
  
  return(strn_afm_comps)
}
```


```{r}
##stage strain pairwise comparison groups
ss_afm_pf7M_afm_comps <- map(ss_afm_pf7M_cln_gt, ~strn_pw_comps(gtypes_chr_pos = .x, strn_regex = c('SC','N')))

# ss_100_plus_m14_comps<- strn_pw_comps(gtypes_chr_pos = ss_cln_gt_100_plus_m14, strn_regex = c('SC','N'))

```

```{r}
##Function to create binned windows across genome and compare pairwise strains and generate the plotting details for chromosome painting
b_strn_c_wndw_fn <- function(strn_c_tbl, wndw = 100000) {
  strn_c_tbl %>% 
    group_by(chr) %>% 
    filter(!is.na(chr)) %>%
    mutate(pos_range = cut(pos, breaks = c(-Inf,seq(0, max(pos), by = wndw),Inf))) %>% 
    pivot_longer(., cols = starts_with(c('SC','N')), names_to = 'comps', values_to = 'sim') %>% 
    dplyr::count(comps, chr, pos_range, sim) %>% 
    group_by(chr, comps, pos_range) %>% 
    mutate(overlap_TF = ifelse(any(is.na(sim)), sum(!is.na(sim)), n_distinct(sim)))%>%
    ungroup() %>% 
    filter(!(is.na(sim) & overlap_TF>0))  %>% 
    mutate(rltd = case_when(overlap_TF == 0 ~ 'No SNP', overlap_TF == 2 ~ 'Both', overlap_TF==1& sim==TRUE ~ 'Same', overlap_TF==1&sim==FALSE ~ 'Different')) %>% 
    group_by(chr, comps, pos_range, rltd) %>% 
    arrange(dplyr::desc(sim), .by_group=T) %>%
    summarise(overlap_snps_No = paste(n, collapse = ","), 
              overlap_snps_prop = log10(n[1]/n[2]))  %>% 
    mutate(across(overlap_snps_No, ~ifelse(rltd == 'No SNP', NA, .))) %>%
    ungroup()  %>%
    mutate(normalized_snps_prop = case_when(rltd == "Same" ~ max(overlap_snps_prop, na.rm = TRUE), 
                                            rltd == "Different" ~ min(overlap_snps_prop, na.rm = TRUE), 
                                            # rltd == 'No SNP' ~ NA, 
                                            T ~ overlap_snps_prop)) %>%
    mutate(start = as.numeric(str_extract(pos_range, "\\d+.*(?=,)")), end = as.numeric(str_extract(pos_range, "(?<=,)\\d+.*(?=])"))) %>% 
    left_join(.,pf_chr_size, by =c('chr')) %>%
    mutate(across(chr, ~factor(., levels = c(1:14)))) 
}
```


```{r}
##stage strain pairwise comparison chromosome binning of SNPs and creation of windows for plotting
ss_afm_wndws <- map(ss_afm_pf7M_afm_comps, ~b_strn_c_wndw_fn(strn_c_tbl = .x))

# ss_ag_k6r_wndws <- b_strn_c_wndw_fn(strn_c_tbl = ss_ag_k6r_pf7M_afm_comps)
# s_k6r_wndws <- b_strn_c_wndw_fn(strn_c_tbl = s_k6r_pf7M_afm_comps)
# s_lqc_k6r_wndws <- b_strn_c_wndw_fn(strn_c_tbl = s_lqc_k6r_pf7M_comps)
```

```{r, fig.height=4.1, fig.width=4.5}
# ss_100_plus_m14_wndws<- b_strn_c_wndw_fn(strn_c_tbl=ss_100_plus_m14_comps)

# ss_ag_hmap_ibs<-gtypes_chr_cln_ibs_hm(gtypes_chr_pos=ss_ag_k6r_pf7M_cln_gt)
```

```{r}
## Function for chromosome painting

chrom_paintr_smth_gradient <- function(gtypes_comps_tbl = b_strn, ttl, expn=100000, x_font_sz = 13) {

      # gs <-gtypes_comps_tbl %>% mutate(across(start, ~(./expn)),across(end, ~(./expn)), across(size, ~(./expn))) %>%
      gs <-gtypes_comps_tbl %>% mutate(across(start, ~(./expn)),across(end, ~(./expn)), across(size, ~(./expn))) %>%
      ggplot()+
        geom_segment(
    aes(x = 0-2000/expn, xend = size+2000/expn, y = chr, yend = chr),
    linewidth = 9, lineend = "round", colour = "black"
  ) + 
      geom_segment(
    aes(x = 0, xend = size, y = chr, yend = chr),
    linewidth = 8, lineend = "round", colour = "white"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  chr, yend =  chr,
      colour = normalized_snps_prop#,
      # colour = rltd
    ),
    linewidth = 8
  ) +
  # scale_color_manual(values = c('Same'='black', 'Different'='#ffd94d', 'Both'='lightgrey', 'No SNP'='white'))+
  scale_color_gradient2(low="black" , mid = "#F5F5F5", high ="#ffd94d" ,
                        limits = c(min(gtypes_comps_tbl$normalized_snps_prop, na.rm = TRUE), max(gtypes_comps_tbl$normalized_snps_prop, na.rm = TRUE)),
                        breaks = seq(min(gtypes_comps_tbl$normalized_snps_prop, na.rm = TRUE), max(gtypes_comps_tbl$normalized_snps_prop, na.rm = TRUE), length.out = 30),
                        oob = scales::squish,
                        na.value = "white",
                        midpoint = 0) +
  # geom_label_repel(aes(label=new_column, x=(start + end)/2, y = comps), size=3,label.padding = unit(0.01, "lines"),alpha = 1,label.size = 0.01, color = alpha('black', .2))+
  # facet_wrap(comps~., ncol = 5) ++
      labs(title = ttl) +
  theme_classic() +
  theme(strip.text = element_text(size = 12, face = 'bold'),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size =  x_font_sz)) 
    
    pp <- gtypes_comps_tbl %>% 
      separate(.,overlap_snps_No, into=(c('same2', 'diff2')), sep=',' ) %>% 
      pivot_longer(cols = c('same2','diff2'), names_to = 'shared2', values_to = 'shared_n') %>% 
      filter(!is.na(shared_n)) %>%
      mutate('Shared' = case_when(rltd == 'Both' & shared2 == 'same2' ~ 'Same', 
                                  rltd == 'Both' & shared2 == 'diff2' ~ 'Different', 
                                  rltd == 'Different' & shared2 == 'same2' ~ 'Different', 
                                  T ~ rltd)) %>%
      mutate(across(shared_n, as.numeric)) %>%
      group_by(Shared) %>% 
      summarise('snpsNo' = sum(shared_n))%>%
      arrange(dplyr::desc(snpsNo), .by_group = T) %>%
    ggplot(., aes(x="", y=snpsNo, fill=Shared))+
    geom_bar(width = 1, stat = "identity") + 
    coord_polar("y", start=0)+
    geom_text(aes(y = snpsNo/2 + c(0, cumsum(snpsNo)[-length(snpsNo)]), 
                  label = snpsNo), size=5) +
    theme_void()

    ggdraw(gs + theme_half_open(12)) +
      draw_plot(pp, .5, .11, .36, .36) 
}


```


strains split by stage Without colour gradients
```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp_smth <- map(ss_afm_wndws, ~{
  smp_tbl <- .x
  smp_tbl %>% split(., ~factor(comps)) %>% imap(.,possibly(~chrom_paintr_smth_gradient(gtypes_comps_tbl =.x, ttl=.y), 'No overlapping SNPs between the 2 groups'))
})

ss_afm_chromp_smth[1:3]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp_smth[4:6]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp_smth[7:9]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp_smth[10:12]
```

strains split by stage Without colour gradients
```{r}
# strain stage (afm) chromosome painting
# ss_100_pm14_chromp <- ss_100_plus_m14_wndws %>% split(., ~factor(comps)) %>% imap(.,possibly(~chrom_paintr_smth_gradient(gtypes_comps_tbl =.x, ttl=.y), 'No overlapping SNPs between the 2 groups'))
# 
# ss_100_pm14_chromp
```


```{r}
# strain stage (afm) chromosome painting
# names(ss_100_pm14_chromp)
# ss_100_pm14_chromp[["SC1_G_msc60_vs_SC3_G_msc49"]]
# ss_100_pm14_chromp[["SC1_G_msc48_vs_SC2_G_msc49"]]
# ss_100_pm14_chromp[["SC2_G_msc49_vs_SC5_G_msc48"]]
```


```{r}
# strain stage (afm) chromosome painting
# ss_100_pm14_chromp
```

strains low QC (lqc) split by stage Without colour gradients

```{r, eval=FALSE}
# strain stage (afm) chromosome painting
s_lqc_k6r_chromp <- s_lqc_k6r_wndws %>% split(., ~factor(comps)) %>% imap(.,possibly(~chrom_paintr_smth_gradient(gtypes_comps_tbl =.x, ttl=.y), 'No overlapping SNPs between the 2 groups'))

s_lqc_k6r_chromp
```

```{r}
##SC8_F comparison plots
# ss_afm_chromp[names(ss_afm_chromp)[grep('SC8_F', names(ss_afm_chromp))]]
```


Coalescing genotypically similar groups into coarse asexual and sexual stages
```{r}
##Coalescing asexual, male, female groups that appear similar from the IBS heatmap into asexual, gametocytes ## MAY NEED TO REGENOTYE later?  ??Remove SC8_F and SC1_F??
# ss_ag_k6r_pf7M_cln_gt <- ss_k6r_pf7M_cln_gt %>% mutate(SC1A=SC1_A, SC1G=coalesce(SC1_F, SC1_M), SC2A=SC2_A, SC2G = coalesce(SC2_F, SC2_M), SC3A=SC3_A, SC4G = coalesce(SC4_M, SC4_F),SC5A=SC5_A, SC5G = coalesce(SC5_M, SC5_F), SC6A=SC6_A, SC6G = coalesce(SC6_F, SC6_M),SC7A=SC7_A, SC7G = coalesce(SC7_F, SC7_M), SC8A=SC8_A, SC8F=SC8_F) %>%  dplyr::select(!matches('^SC[0-9]_.$'))

##Same as above but removing the groups already removed with 20 or less parasites
# ss_ag_k6r_pf7M_cln_gt <- ss_k6r_pf7M_cln_gt %>% mutate(SC1A=SC1_A, SC2A=SC2_A, SC2G = coalesce(SC2_F, SC2_M), SC3A=SC3_A, SC5G = coalesce(SC5_M, SC5_F), SC6A=SC6_A, SC6G = coalesce(SC6_F, SC6_M), SC7G = coalesce(SC7_F, SC7_M)) %>%  dplyr::select(!matches('^SC[0-9]_.$'))
                                                      
```

##Simpler chromosome plots(no continous gradient) for SS_AG and S

```{r}
## Function for chromosome painting

chrom_paintr <- function(gtypes_comps_tbl = b_strn_c, ttl, expn=100000, x_font_sz = 13) {

      gs <-gtypes_comps_tbl %>% mutate(across(start, ~(./expn)),across(end, ~(./expn)), across(size, ~(./expn))) %>%
      ggplot()+
      geom_segment(
    aes(x = 0-(2000/expn), xend = size+(2000/expn), y = chr, yend = chr),
    linewidth = 3.5, lineend = "round", colour = "black"
  ) + 
      geom_segment(
    aes(x = 0, xend = size, y = chr, yend = chr),
    linewidth = 2.8, lineend = "round", colour = "white"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  chr, yend =  chr,
      colour = rltd
    ),
    linewidth = 2.8
  ) +
  # scale_color_manual(values = c('Same'= '#D4AF37', 'Different'= '#2C3E50','Both'='lightgrey', 'No SNP'='white'))+      
  scale_color_manual(values = c('Same'= '#DACB7D', 'Different'= '#2C3E50','Both'='#A9B7C5', 'No SNP'='white'))+ ##prev colours "black" ,"#ffd94d" 
  # facet_wrap(comps~., ncol = 5) +
      labs(title = ttl) +
      labs(colour = "Allele \ncomparison", x = 'Position (bases)', y ="Chromosome") +
  theme_classic() +
  theme(#strip.text = element_text(size = 9, face = 'bold'),
        axis.text.y = element_text(size =  13),
        axis.text.x = element_text(size =  x_font_sz),
        # axis.title = element_text(size = 9),
        axis.title = element_blank(),
        # legend.title = element_text(size = 9, face = "bold"),
        # legend.text = element_text(size = 9),
        title = element_text(size = 12, face = "bold"),
        legend.position = "none") 
    
    pie_bar_tbl <- gtypes_comps_tbl %>% 
      separate(.,overlap_snps_No, into=(c('same2', 'diff2')), sep=',' ) %>% 
      pivot_longer(cols = c('same2','diff2'), names_to = 'shared2', values_to = 'shared_n') %>% 
      filter(!is.na(shared_n)) %>%
      mutate('Shared' = case_when(rltd == 'Both' & shared2 == 'same2' ~ 'Same', 
                                  rltd == 'Both' & shared2 == 'diff2' ~ 'Different', 
                                  rltd == 'Different' & shared2 == 'same2' ~ 'Different', 
                                  T ~ rltd)) %>%
      mutate(across(shared_n, as.numeric))
    
    pp <- pie_bar_tbl %>% 
      group_by(Shared) %>% 
      summarise('snpsNo' = sum(shared_n))%>%
    arrange(dplyr::desc(snpsNo), .by_group = T) %>%
    #mutate(lab.ypos = cumsum(prop) - 0.5*prop) %>%
    ggplot(., aes(x="", y=snpsNo, fill=Shared))+
    geom_bar(width = 1, stat = "identity") + 
    coord_polar("y", start=0)+
    # geom_text(aes(y = snpsNo/2 + c(0, cumsum(snpsNo)[-length(snpsNo)]), 
    #               label = snpsNo), size=3, colour = "#90c7ff") +
      geom_label(aes(y = snpsNo/2 + c(0, cumsum(snpsNo)[-length(snpsNo)]), 
                  label = snpsNo), size=5, label.padding = unit(0.07, "lines"), label.size = 0, color = 'black', fill = 'white') +
      scale_fill_manual(values = c('Same'= '#DACB7D', 'Different'= '#2C3E50'))+ #'#D4AF37'
    # labs(title = .y)+
    theme_void()+
      theme(legend.position = "none")
  #theme_minimal_hgrid(11)

    
    ggdraw(gs ) +
                draw_plot(pp, x=.5, y=.1, width = .50, height= .50)

  }

```


strains split by stage Without colour gradients
```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp <- map(ss_afm_wndws, ~{
  smp_tbl <- .x
  smp_tbl %>% split(., ~factor(comps)) %>% imap(.,possibly(~chrom_paintr(gtypes_comps_tbl =.x, ttl=.y), 'No overlapping SNPs between the 2 groups'))
})

```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp[1:3]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp[4:6]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp[7:9]
```

```{r}
# strain stage (afm) chromosome painting
ss_afm_chromp[10:12]
```

```{r}

plot_grid(plotlist = ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC1_A_vs_SC2_A", "SC1_A_vs_SC2_A", "SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G", "SC6_A_vs_SC7_G", "SC6_G_vs_SC7_G")], ncol = 3)
# plot_grid(plotlist = plts_strn_ss[c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 2)
```


```{r}

# plot_grid(plotlist = ss_k6r_ag_wndws[c("SC1_A_vs_SC2A", "SC1_A_vs_SC2_G", "SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G", "SC6_A_vs_SC7_G", "SC6_G_vs_SC7_G")], ncol = 3)
plot_grid(plotlist = ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 1)
```

```{r, fig.height=6, fig.width=12}
plot_grid(plotlist = ss_afm_chromp[["msc14.strain_qcd"]], ncol = 7)
```


strains split by stage Without colour gradients
```{r}
# strain stage (afm) chromosome painting
s_chromp_xlab_mod <- map(ss_afm_wndws[str_detect(names(ss_afm_wndws), "\\.strain_qcd")], ~{
  smp_tbl <- .x
  smp_tbl %>% split(., ~factor(comps)) %>% imap(.,possibly(~chrom_paintr(gtypes_comps_tbl =.x, ttl=.y, x_font_sz = 0), 'No overlapping SNPs between the 2 groups'))
})

```


```{r}


b_ovrl <-map(ss_afm_pf7M_afm_comps[str_detect(names(ss_afm_pf7M_afm_comps), "\\.strain_qcd")], ~.x %>% 
    group_by(chr) %>% 
    filter(!is.na(chr)) %>%
    mutate(pos_range = cut(pos, breaks = c(-Inf,seq(0, max(pos), by = 100000),Inf)))  %>% 
    dplyr::group_by(chr, pos_range) %>%
    summarise(SNPs_sum = n()) %>%
    ungroup() %>%
    left_join(.,pf_chr_size, by =c('chr')) %>%
    mutate(across(chr, ~factor(., levels = c(1:14)))) %>%
    mutate(start = as.numeric(str_extract(pos_range, "\\d+.*(?=,)")), end = as.numeric(str_extract(pos_range, "(?<=,)\\d+.*(?=])"))) %>%
    mutate(across(c(end), ~case_when(is.na(.) ~ size, T ~ .)))
)

```


```{r, fig.width=14, fig.height=6.5}

plt_ovrl <- b_ovrl %>%
  map(~{ .x %>% ggplot() + 
      geom_segment(
    aes(x = 0-2000, xend = size+2000, y = chr, yend = chr),
    linewidth = 7.5, lineend = "round", colour = "black"
  ) + 
      geom_segment(
    aes(x = 0, xend = size, y = chr, yend = chr),
    linewidth = 6.5, lineend = "round", colour = "white"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  chr, yend =  chr
    ),
      colour = "#C9E0F1",
    linewidth = 6.5
  ) +
  geom_label(aes(label=SNPs_sum, x=(start + end)/2, y = chr), size=5.4, label.padding = unit(0.05, "lines"), fontface = 'bold', label.size = 0, color = alpha('black', .9))+
  # facet_wrap(comps~., ncol = 5) +
      labs(y = 'Chromosome', x = 'Position (bases)') +
  theme_classic() +
  theme(strip.text = element_text(size = 12, face = 'bold'),
        axis.text = element_text(size = 18),
        axis.title = element_text(size = 20, face = 'bold'),
        title = element_text(size = 15, face = "bold"))
})

plt_ovrl


```


######Moved IBD codes from above to below here - may create complications
hmmIBD START

```{r}
##!!NOTE - gfv_p1_bi_strn_pf7_c is created somewhere downstream so this is likely to give an error

system(paste0('mkdir data/processed/ibd/'))
system(paste0('rm data/processed/ibd/*'))

imap(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")], ~{
  .x %>% 
    dplyr::select(-c(id, chrom, alleles))%>% 
    dplyr::select(chr, pos, any_of(sort(colnames(.)))) %>% 
    filter(!if_all(starts_with('SC'), ~is.na(.))) %>% 
    mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% 
    write_tsv(., paste0("data/processed/ibd/", .y, ".tsv"))
})

imap(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")], ~{
  .x %>% 
    dplyr::select(-c(id, chrom, alleles))%>% 
    dplyr::select(chr, pos, any_of(sort(colnames(.)))) %>% 
    filter(!if_all(starts_with('SC'), ~is.na(.))) %>% 
    mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% 
    write_tsv(., paste0("data/processed/ibd/", .y, "_lqc.tsv"))
})

## Write out the combined snps across all samples for ibd
imap(ss_cln_all_mtx[3], ~{
  .x %>% 
    dplyr::select(-c(id, chrom, alleles))%>% 
    dplyr::select(chr, pos, any_of(sort(colnames(.)))) %>% 
    filter(!if_all(starts_with('msc14_qcd_SC'), ~is.na(.))) %>% 
    mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% 
    write_tsv(., paste0("data/processed/ibd/", .y, ".tsv"))
})

# ##!!!NOTE - Remove SC8 first - use this to include SC8 dplyr::select(chr, pos, paste0('SC', 1:8))
# s_k6r_pf7M_cln_gt %>% dplyr::select(-c(id, chrom, alleles))%>% dplyr::select(chr, pos, any_of(sort(colnames(.))))  %>% filter(!if_all(starts_with('SC'), ~is.na(.))) %>% mutate(across(everything(), ~as.numeric(replace_na(., '-1')))) %>% write_tsv(., "/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master/s_k6r_pf7M_cln_gt.tsv")

```


```{r}
##Run IBD 
##!!!!!!NOTE - sort out the output file -o argument - failure to save in the same directory. Goes to working dir

# imap(ss_afm_pf7M_cln_gt, ~system(paste0('hmmIBD -i data/processed/DB52e_star/', str_remove(.y, "\\.st.*"), '/ibd/', .y, '.tsv -o data/processed/DB52e_star/', str_remove(.y, "\\.st.*"), '/ibd/', .y, ' -m 100000')))

imap(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")], ~system(paste0('hmmIBD -i data/processed/ibd/', .y, '.tsv -o data/processed/ibd/', .y, ' -m 100000')))
imap(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")], ~system(paste0('hmmIBD -i data/processed/ibd/', .y, '_lqc.tsv -o data/processed/ibd/', .y, '_lqc -m 100000')))

## IBD of genotypes combined snps across all samples
imap(ss_cln_all_mtx[3], ~system(paste0('hmmIBD -i data/processed/ibd/', .y, '.tsv -o data/processed/ibd/', .y, ' -m 100000')))

# system('/Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/sware/hmmIBD-master/hmmIBD -i /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/sware/hmmIBD-master/ss_afm_pf7M_cln_gt.tsv -o ss_afm_pf7M_cln_gt -m 100000')
  
```


```{r}
# wdir<- '/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master'

ss_afm_pf7M_cln_gt_hmmibd <- map(names(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")]), ~read_delim(paste0('data/processed/ibd/', .x, '.hmm.txt'))) %>% set_names(names(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")]))
ss_afm_pf7M_cln_gt_hmmibd_f <- map(names(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")]), ~read_delim(paste0('data/processed/ibd/', .x, '.hmm_fract.txt'))) %>% set_names(names(ss_afm_pf7M_cln_gt[!str_detect(names(ss_afm_pf7M_cln_gt), "\\.stage_afm_strain_qcd")]))

ss_afm_pf7M_cln_gt_hmmibd_lqc <- map(names(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")]), ~read_delim(paste0('data/processed/ibd/', .x, '_lqc.hmm.txt'))) %>% set_names(names(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")]))
ss_afm_pf7M_cln_gt_hmmibd_f_lqc <- map(names(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")]), ~read_delim(paste0('data/processed/ibd/', .x, '_lqc.hmm_fract.txt'))) %>% set_names(names(ss_afm_pf7M_cln_gt_lqc[!str_detect(names(ss_afm_pf7M_cln_gt_lqc), "\\.stage_afm_strain_qcd")]))

## IBD of genotypes combined snps across all samples
all_donors_hmmibd <- map(names(ss_cln_all_mtx[3]), ~read_delim(paste0('data/processed/ibd/', .x, '.hmm.txt'))) %>% set_names(names(ss_cln_all_mtx[3]))
all_donors_hmmibd_f <- map(names(ss_cln_all_mtx[3]), ~read_delim(paste0('data/processed/ibd/', .x, '.hmm_fract.txt'))) %>% set_names(names(ss_cln_all_mtx[3]))


```


```{r}
# ## Remove SC8 from the msc14 results
# ss_afm_pf7M_cln_gt_hmmibd_m14 <- ss_afm_pf7M_cln_gt_hmmibd[[8]] %>% filter(sample1 != "SC8" & sample2 != "SC8" )
# ss_afm_pf7M_cln_gt_hmmibd_f_m14 <- ss_afm_pf7M_cln_gt_hmmibd_f[[8]] %>% filter(sample1 != "SC8" & sample2 != "SC8" )
# 
# ss_afm_pf7M_cln_gt_hmmibd_lqc_m14 <- ss_afm_pf7M_cln_gt_hmmibd_lqc[[8]] %>% filter(sample1 != "SC8" & sample2 != "SC8" )
# ss_afm_pf7M_cln_gt_hmmibd_f_lqc_m14 <- ss_afm_pf7M_cln_gt_hmmibd_f_lqc[[8]] %>% filter(sample1 != "SC8" & sample2 != "SC8" )

```


Export the SNPs with QC metrics and only clean SNPs for stage&strain combinations and just strain
```{r, eval=T}
##!!!! NOTE -MANUSCRIPT IBD table

## subset IBD tables to write and rename lists/sheets 
### IBD results
ibd_export <- ss_afm_pf7M_cln_gt_hmmibd_f[str_detect(names(ss_afm_pf7M_cln_gt_hmmibd_f), "\\.strain_qcd")] %>% map(., ~rename_with(., ~str_replace(.,'sample', 'Strain'), contains("sample")))
names(ibd_export) <- paste0(2:(length(ibd_export)+1), ".", str_replace_all(names(ibd_export), c("\\.strain_qcd" = " IBD", "msc" = "MSC")))

### IBD summary
ibd_f_export <- ss_afm_pf7M_cln_gt_hmmibd[str_detect(names(ss_afm_pf7M_cln_gt_hmmibd), "\\.strain_qcd")] %>% map(., ~rename_with(., ~str_replace(.,'sample', 'Strain'), contains("sample")))
names(ibd_f_export) <- paste0(6:(length(ibd_f_export)+5), ".", str_replace_all(names(ibd_f_export), c("\\.strain_qcd" = " IBD sites summary", "msc" = "MSC")))


createWorksheet(c(ibd_export, ibd_f_export), paste0("../",trip,"/v3_science_manuscript/tables/data S10 - Strain pairwise IBD.xlsx"))
```


```{r}
## Read in msc annotations 

# msc14_qcd_anotd <- readRDS('data/processed/msc14/msc_qcd_anotd.RDS') %>% .[[1]]

```


```{r, fig.height=5, fig.width=6}
colFun = suppressWarnings(brewer.pal(100, 'Blues'))
colFun = circlize::colorRamp2(seq(0, 1, length.out = length(colFun)), colFun)
  
##Strain relatedness counts
strn_qcd_n <- msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% dplyr::count(strain_qcd) %>% filter(!is.na(strain_qcd))

##Merge strain numbers to IBD results
s_lqc_strn_n_ibd_tbl = ss_afm_pf7M_cln_gt_hmmibd_f_lqc[["msc14.strain_qcd"]] %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., strn_qcd_n, by = c('sample1'='strain_qcd')) %>% left_join(., strn_qcd_n, by = c('sample2'='strain_qcd'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1, '\n[', n_1, ']'), id2 = paste0(sample2, '\n[', n_2, ']'))

##Get the IBD and the number of sites for each pairwise comparison to annotate heatmap
s_lqc_ibdf_nsites_anno <- s_lqc_strn_n_ibd_tbl %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., s_lqc_strn_n_ibd_tbl)  %>% mutate(ibdf_nsites = paste0(round(fract_sites_IBD,2), '\n(',N_informative_sites,')'))%>% dplyr::select(starts_with('id'), ibdf_nsites) %>% pivot_wider(names_from = id2, values_from = ibdf_nsites) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., '1')))%>% column_to_rownames('id1')

##Create mirrored matrix and plot
##!!! TIP - Refer to this if you want more complex arrangements https://github.com/jokergoo/ComplexHeatmap/issues/712 
s_lqc_sibd_pht_mtx <- s_lqc_strn_n_ibd_tbl %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., s_lqc_strn_n_ibd_tbl) %>% dplyr::select(starts_with('id'), fract_sites_IBD) %>%  pivot_wider(names_from = id2, values_from = fract_sites_IBD) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., 1)))%>% column_to_rownames('id1') %>% as.matrix() 
  
s_lqc_ibd_pht<- Heatmap(s_lqc_sibd_pht_mtx, rect_gp = gpar(type = "none"), name='IBD', col = colFun, show_row_names = TRUE, row_names_side = c("left"),
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = F, cluster_columns = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
              heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
              # border='black',
              cell_fun = function(j, i, x, y, width, height, fill) {
                if(i >= j) { ## diagonal(i==j), lower triangle (i>j), upper triangle (j>i)
                  grid.rect(x, y, width, height, gp = gpar(fill = fill, col = fill))
                  grid.text(sprintf("%s", s_lqc_ibdf_nsites_anno[i, j]), x, y, gp = gpar(fontsize = 13))
        }
})



s_lqc_ibd_pht

```

```{r, fig.height=5, fig.width=6}
##Merge strain numbers to IBD results
s_strn_n_ibd_tbl = ss_afm_pf7M_cln_gt_hmmibd_f[["msc14.strain_qcd"]] %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., strn_qcd_n, by = c('sample1'='strain_qcd')) %>% left_join(., strn_qcd_n, by = c('sample2'='strain_qcd'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1, '\n[', n_1, ']'), id2 = paste0(sample2, '\n[', n_2, ']'))  

##Get the IBD and the number of sites for each pairwise comparison to annotate heatmap
s_ibdf_nsites_anno <- s_strn_n_ibd_tbl %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., s_strn_n_ibd_tbl) %>% mutate(ibdf_nsites = paste0(round(fract_sites_IBD,2), '\n(',N_informative_sites,')'))%>% dplyr::select(starts_with('id'), ibdf_nsites) %>% pivot_wider(names_from = id2, values_from = ibdf_nsites) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., '1')))%>% column_to_rownames('id1')


##Create mirrored matrix and plot
sibd_pht_mtx <- s_strn_n_ibd_tbl %>% rename('id2' = id1,  'id1' = id2) %>% bind_rows(., s_strn_n_ibd_tbl) %>% dplyr::select(starts_with('id'), fract_sites_IBD) %>%  pivot_wider(names_from = id2, values_from = fract_sites_IBD) %>% arrange(id1) %>% mutate(across(starts_with('SC'), ~replace_na(., 1)))%>% column_to_rownames('id1')  
  
sibd_pht<- Heatmap(sibd_pht_mtx,rect_gp = gpar(type = "none"), name='IBD', col = colFun, show_row_names = TRUE, row_names_side = c("left"),
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = F, cluster_columns = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
              heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),
              # border='black',
              cell_fun = function(j, i, x, y, width, height, fill) {
                if(i >= j) {
                  grid.rect(x, y, width, height, gp = gpar(fill = fill, col = fill))
                  grid.text(sprintf("%s", s_ibdf_nsites_anno[i, j]), x, y, gp = gpar(fontsize = 13))
        }
})


sibd_pht
  
```

```{r, fig.height=5, fig.width=6}
##Merge strain numbers to IBD results
s_strn_n_ibd_tbl = ss_afm_pf7M_cln_gt_hmmibd_f[["msc14.strain_qcd"]] %>% dplyr::select(sample1, sample2, fract_sites_IBD,N_informative_sites) %>% left_join(., strn_qcd_n, by = c('sample1'='strain_qcd')) %>% left_join(., strn_qcd_n, by = c('sample2'='strain_qcd'), suffix = c('_1', '_2')) %>% mutate(id1 = paste0(sample1, '\n[', n_1, ']'), id2 = paste0(sample2, '\n[', n_2, ']'))  

##Get the IBD and the number of sites for each pairwise comparison to annotate heatmap
s_combi_ibdf_nsites_anno <- s_strn_n_ibd_tbl %>% 
  rename('id2' = id1,  'id1' = id2) %>% 
  bind_rows(., s_lqc_strn_n_ibd_tbl) %>% 
  mutate(ibdf_nsites = paste0(round(fract_sites_IBD,2), '\n(',N_informative_sites,')'))%>% 
  dplyr::select(starts_with('id'), ibdf_nsites) %>% 
  pivot_wider(names_from = id2, values_from = ibdf_nsites) %>% 
  arrange(id1) %>%
  mutate(across(starts_with('SC'), ~replace_na(., '1')))%>%
  column_to_rownames('id1')


##Create mirrored matrix and plot
##!!! TIP - Refer to this if you want more complex arrangements https://github.com/jokergoo/ComplexHeatmap/issues/712 - eg different colours for the different half matrices
s_combi_ibd_pht <- s_strn_n_ibd_tbl %>% 
  rename('id2' = id1,  'id1' = id2) %>% 
  bind_rows(., s_lqc_strn_n_ibd_tbl) %>% 
  dplyr::select(starts_with('id'), fract_sites_IBD) %>%  
  pivot_wider(names_from = id2, values_from = fract_sites_IBD) %>% 
  arrange(id1) %>% 
  mutate(across(starts_with('SC'), ~replace_na(., 1)))%>%
  column_to_rownames('id1') %>% 
  as.matrix()%>%
  Heatmap(., name='IBD', col = colFun, show_row_names = TRUE, row_names_side = c("left"),
                 show_column_names = TRUE, show_row_dend = T, 
                 show_column_dend = T, row_title_rot = 0, cluster_rows = F, cluster_columns = F, column_names_gp = gpar(fontsize = 15), row_names_gp = gpar(fontsize = 15), 
              heatmap_legend_param = list(title_gp = gpar(fontsize = 12, fontface = "bold"), labels_gp = gpar(fontsize = 12)),border='black',
              cell_fun = function(j, i, x, y, width, height, fill) {
        grid.text(sprintf("%s", s_combi_ibdf_nsites_anno[i, j]), x, y, gp = gpar(fontsize = 13))
})

s_combi_ibd_pht
  
```




```{r}

# k1_ibd_sites_f <- list(gt_fb_vtx_p1_bi_strn_pf7_asx_gam_ibd, gfv_p1_bi_pf7_k7_all_ibd) %>%
# k8_ibd_sites_f <- list(ss_ag_k6r_pf7M_cln_gt_hmmibd, s_k6r_pf7M_cln_gt_hmmibd) %>%
k8_ibd_sites_f <- map(ss_afm_pf7M_cln_gt_hmmibd[7:8], ~{
  tb1=.x
  list(tb1) %>%
  map(~{
    .x%>% 
      mutate(across(chr,as.character)) %>% 
  left_join(.,pf_chr_size, by ='chr') %>% 
  mutate(across(chr, ~factor(., levels = c(1:14)))) %>%
  unite(., 'comps',sample1, sample2, sep='vs', remove = F) %>%  
  group_by(chr) %>% 
  mutate(chr_s = 0, chr_e = size) %>% 
  ungroup() %>% 
  mutate(across(different, as.character))

    })

})
```


```{r}
## Function for chromosome painting

chrom_paintr_ibd <- function(gtypes_comps_tbl, ttl, expn= 100000, x_font_sz = 13) {
  gtypes_comps_tbl %>% mutate(across(start, ~(./expn)),across(end, ~(./expn)), across(size, ~(./expn))) %>%
      ggplot()+ 
      geom_segment(
    aes(x = 0-(2000/expn), xend = size+(2000/expn), y = chr, yend = chr),
    linewidth = 3.5, lineend = "round", colour = "black"
  ) + 
      geom_segment(
    aes(x = 0, xend = size, y = chr, yend = chr),
    linewidth = 2.8, lineend = "round", colour = "white"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  chr, yend =  chr,
      colour = rltd
    ),
    linewidth = 2.8
  ) +
  # scale_color_manual(values = c('Same'= '#D4AF37', 'Different'= '#2C3E50','Both'='lightgrey', 'No SNP'='white'))+ ##prev colours "black" ,"#ffd94d" 
  scale_color_manual(values = c('Same'= '#DACB7D', 'Different'= '#2C3E50','Both'='#A9B7C5', 'No SNP'='white'))+
  # facet_wrap(comps~., ncol = 5) +
      labs(title = ttl) +
      labs(colour = "Allele \ncomparison", x = 'Position (bases)', y ="Chromosome") +
  theme_classic() +
  theme(axis.text.y = element_text(size =  13),
        axis.text.x = element_text(size =  x_font_sz),
        # axis.ticks.x = element_blank(),
        # axis.title = element_text(size = 9),
        axis.title = element_blank(),
        # legend.title = element_text(size = 9, face = "bold"),
        # legend.text = element_text(size = 9),
        title = element_text(size = 12, face = "bold"),
        legend.position = "none",
        plot.background = element_rect(fill = "#bfe5fc")) 

  }


```


```{r, fig.height=14, fig.width=24}
ss_afm_k6r_pf7M_ibd <- k8_ibd_sites_f[[1]][[1]] %>% mutate(across(comps, ~str_replace(., 'vs', '_vs_')), rltd = case_when(different == '0' ~ 'Same', different == '1' ~ 'Different')) 

ss_afm_k6r_pf7M_ibd_p <- ss_afm_k6r_pf7M_ibd %>% split(., ~factor(comps)) %>% imap(., ~chrom_paintr_ibd(.x, .y))
```


```{r, fig.height=2.7, fig.width=6}
ss_afm_k6r_pf7M_ibd_p
# plot_grid(plotlist = c(ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC3_A_vs_SC4_G")], ss_ag_k6r_pf7M_ibd_p[c("SC3_A_vs_SC4_G")]), ncol = 2)
```

```{r, fig.height=14, fig.width=24}
ss_ag_k6r_pf7M_ibd <- k8_ibd_sites_f[[1]][[1]] %>% mutate(across(comps, ~str_replace(., 'vs', '_vs_')), rltd = case_when(different == '0' ~ 'Same', different == '1' ~ 'Different')) 

ss_ag_k6r_pf7M_ibd_p <- ss_ag_k6r_pf7M_ibd %>% split(., ~factor(comps)) %>% imap(., ~chrom_paintr_ibd(.x, .y))
```

```{r, fig.height=2.7, fig.width=6}

# plot_grid(plotlist = c(ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC1_A_vs_SC8A")], ss_ag_k6r_pf7M_ibd_p[c("SC1_A_vs_SC8A")]), ncol = 2)
```

```{r, fig.height=2.7, fig.width=6}
ss_ag_k6r_pf7M_ibd_p
# plot_grid(plotlist = c(ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC3_A_vs_SC4_G")], ss_ag_k6r_pf7M_ibd_p[c("SC3_A_vs_SC4_G")]), ncol = 2)
```

```{r, fig.height=2.7, fig.width=12}

plot_grid(plotlist = c(ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC2_A_vs_SC2_G")], ss_ag_k6r_pf7M_ibd_p[c("SC2_A_vs_SC2_G")],ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC6_A_vs_SC6_G")],ss_ag_k6r_pf7M_ibd_p[c("SC6_A_vs_SC6_G")]), ncol = 4)
```

```{r, fig.height=5.5, fig.width=6}
plot_grid(
plot_grid(plotlist = ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 1),
plot_grid(plotlist =ss_ag_k6r_pf7M_ibd_p[c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 1) , nrow = 1
)
```

```{r, fig.height=14, fig.width=24}
s_k6r_ibd <- k8_ibd_sites_f[[2]][[1]]  %>% mutate(across(comps, ~str_replace(., 'vs', '_vs_')), rltd = case_when(different == '0' ~ 'Same', different == '1' ~ 'Different')) 

s_k6r_ibd_comps = s_k6r_ibd %>% split(., ~factor(comps)) %>% names
s_k6r_ibd_sc7_comps <- c("SC1_vs_SC7", "SC2_vs_SC7", "SC3_vs_SC7", "SC4_vs_SC7", "SC5_vs_SC7", "SC6_vs_SC7")
s_k6r_xlab_sc7 <- ifelse(s_k6r_ibd_comps %in% s_k6r_ibd_sc7_comps, 13, 0)

s_k6r_ibd_plts <- s_k6r_ibd %>% split(., ~factor(comps)) %>% imap(., ~chrom_paintr_ibd(.x, .y))
```

```{r}
s_k6r_ibd_plts_xlab_mod = s_k6r_ibd %>% split(., ~factor(comps)) %>% list(., names(.), s_k6r_xlab_sc7) %>% pmap(., ~chrom_paintr_ibd(..1, ..2, x_font_sz = ..3))

```

```{r}
s_k6r_ibd_plts[1:3]
```


```{r, fig.height=6, fig.width=8}

plot_grid(plotlist = c(ss_afm_chromp[["msc14.strain_qcd"]][c("SC1_vs_SC2")],ss_afm_chromp[["msc14.strain_qcd"]][c("SC2_vs_SC5")], ss_afm_chromp[["msc14.strain_qcd"]][c("SC2_vs_SC6")],ss_afm_chromp[["msc14.strain_qcd"]][c("SC6_vs_SC7")],ss_afm_chromp[["msc14.strain_qcd"]][c("SC2_vs_SC7")]), ncol = 3)
```


```{r, fig.height=16, fig.width=12}
## Without position labels in all plots but bottom row
plot_grid(
  plot_grid(plotlist = s_chromp_xlab_mod[["msc14.strain_qcd"]][1:5], nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod[["msc14.strain_qcd"]][6:9], s_k6r_ibd_plts_xlab_mod[1]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod[["msc14.strain_qcd"]][10:12], s_k6r_ibd_plts_xlab_mod[c(6,2)]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod[["msc14.strain_qcd"]][13:14], s_k6r_ibd_plts_xlab_mod[c(10,7,3)]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod[["msc14.strain_qcd"]][15], s_k6r_ibd_plts_xlab_mod[c(13,11,8,4)]),nrow = 1),
  plot_grid(plotlist = s_k6r_ibd_plts_xlab_mod[c(15,14,12,9,5)], nrow = 1),
  ncol = 1
)
```

```{r, fig.width= 12, fig.height=15}
##Put together the plots for annotation


plot_grid(plt_ovrl[["msc14.strain_qcd"]] +
            theme(plot.margin = margin(t=0.7, b=0.3, unit = "cm")),
          plot_grid(grid.grabExpr(draw(ss_afm_hmap_ibs_m14_noSC8[["msc14.stage_ag_strain_qcd"]])),
                    plot_grid(
                      plot_grid(plotlist =ss_afm_chromp[["msc14.stage_ag_strain_qcd"]][c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 1),
                      plot_grid(plotlist =ss_ag_k6r_pf7M_ibd_p[c("SC2_A_vs_SC2_G", "SC6_A_vs_SC6_G")], ncol = 1) , 
                      nrow = 1
                    ),
                    ncol = 2,
                    labels = c('B', 'C'),
                    label_size = 24,
                    vjust= 0.1
          )+
              theme(plot.margin = margin(t=0.7, b=0.3, unit = "cm")),
              plot_grid(grid.grabExpr(draw(ss_afm_hmap_ibs_m14_noSC8[["msc14.strain_qcd"]])), 
              # grid.grabExpr(draw(sibd_pht)),
              grid.grabExpr(draw(s_combi_ibd_pht)),
              nrow = 1,
              rel_widths = c(.45, .55),
              # labels = c('D', 'E', 'F'),
              labels = c('D', 'E'),
              label_size = 24,
              vjust= 0.1) +
            theme(plot.margin = margin(t=0.7, b=0.3, unit = "cm")),
          nrow = 3,
          rel_heights = c(.3, .4, .3),
          labels = c('A', ''),
          label_size = 24,
          vjust= 1
)

``` 

```{r, fig.height=16, fig.width=12}
## Without position labels in all plots but bottom row
s_chromp_xlab_mod_m14_noSC8 <- s_chromp_xlab_mod[["msc14.strain_qcd"]][!str_detect(names(s_chromp_xlab_mod[["msc14.strain_qcd"]]), "_SC8")]
s_k6r_ibd_plts_xlab_mod_m14_noSC8 <- s_k6r_ibd_plts_xlab_mod[!str_detect(names(s_k6r_ibd_plts_xlab_mod), "_SC8")]

plot_grid(
  plot_grid(plotlist = s_chromp_xlab_mod_m14_noSC8[1:5], nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod_m14_noSC8[6:9], s_k6r_ibd_plts_xlab_mod_m14_noSC8[1]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod_m14_noSC8[10:12], s_k6r_ibd_plts_xlab_mod_m14_noSC8[c(6,2)]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod_m14_noSC8[13:14], s_k6r_ibd_plts_xlab_mod_m14_noSC8[c(10,7,3)]), nrow = 1),
  plot_grid(plotlist = c(s_chromp_xlab_mod_m14_noSC8[15], s_k6r_ibd_plts_xlab_mod_m14_noSC8[c(13,11,8,4)]),nrow = 1),
  plot_grid(plotlist = s_k6r_ibd_plts_xlab_mod_m14_noSC8[c(15,14,12,9,5)], nrow = 1),
  ncol = 1
)
```


```{r, fig.height=8, fig.width=20}
# k8_ibd_sites_f %>%
#   map(~{
#     .x%>%
# ggplot() +
#   geom_segment(
#     aes(x = 0, xend = chr_e, 
#         y = comps, yend = comps),
#     linewidth = 2.5, lineend = "round", colour = "grey75"
#   ) +
#   geom_segment(
#     aes(
#       x = start, xend = end,
#       y =  comps, yend =  comps,
#       colour = different
#     ),
#     linewidth = 2.5
#   ) +
#   scale_color_manual(values = c('0'='#D4AF37', '1'='#2C3E50'))+
#   facet_wrap(chr~., ncol = 5)+
#   theme_classic() +
#   theme(strip.text = element_text(size = 12, face = 'bold'),
#         axis.text = element_text(size = 12))
#     
#     
#   })

```

```{r, fig.height=14, fig.width=24}
# g8_ibd_chrom_ps <- k8_ibd_sites_f %>%
#   map(~{
#     
# a_plt<- .x %>%
# ggplot() +
#   geom_segment(
#     aes(x = 0, xend = chr_e, y = chr, yend = chr),
#     linewidth = 2, lineend = "round", colour = "grey75"
#   ) +
#   geom_segment(
#     aes(
#       x = start, xend = end,
#       y =  chr, yend =  chr,
#       colour = different
#     ),
#     linewidth = 2
#   ) +
#   scale_color_manual(values = c('0'='black', '1'='#ffd94d'))+
#   facet_wrap(comps~., ncol = 5) +
#   theme_classic() +
#   theme(strip.text = element_text(size = 12, face = 'bold'),
#         axis.text = element_text(size = 12))
# a_plt
# })
```


```{r, fig.height=6, fig.width=20}
# k8_ibd_sites_f %>%
#   map(~{
#     
# b_plt<-.x %>% 
#   ggplot(aes(x=end,y=Nsnp, color = different, fill = different)) +
#   geom_col() + 
#   ggplot2::facet_wrap(comps ~ ., ncol = 5)+theme_classic() + 
#   scale_color_manual(values = c('0'='black', '1'='#ffd94d'))+ 
#   scale_fill_manual(values = c('0'='black', '1'='#ffd94d'))+
#   theme(strip.text = element_text(size = 14, face = 'bold'),
#         axis.text = element_text(size = 13),
#         axis.text.x = element_text(angle = 45, hjust = 1))
# b_plt
# })
```

```{r, fig.height=12, fig.width=14}
# cowplot::plot_grid(a_plt+ 
#                         theme(axis.text.x = element_blank(),
#                               axis.ticks.x = element_blank(),
#                               axis.title.x = element_blank() ), 
#                    b_plt ,
#                    nrow = 2,
#                    ncol = 1,
#                    labels = "auto",
#                    align = "h"
#                    )
```


```{r, fig.height=6, fig.width=20}
k8_ibd_sites_f[[1]] %>%
  map(~{
    .x%>% 
      mutate(across(Nsnp, as.character)) %>%
ggplot() +
  geom_segment(
    aes(x = 0, xend = chr_e, 
        y = comps, yend = comps),
    linewidth =  1.5, lineend = "round", colour = "grey75"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  comps, yend =  comps,
      colour = different
    ),
    linewidth = 1.5
  ) +
  scale_color_manual(values = c('0'='black', '1'='#ffd94d'))+
  geom_label_repel(aes(label=Nsnp, x=(start + end)/2, y = comps), size=4,label.padding = unit(0.05, "lines"),alpha = 1,label.size = 0.01, color = alpha('black', .2)) +
  facet_wrap(chr~., ncol = 7)+
  theme_classic() +
  theme(strip.text = element_text(size = 14, face = 'bold'),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1)) 
  })
```

```{r, fig.height=6, fig.width=20}
k8_ibd_sites_f[[2]][[1]] %>% group_by(chr) %>% group_split() %>%
  map(~{
    .x%>% 
      mutate(across(Nsnp, as.character)) %>%
ggplot() +
  geom_segment(
    aes(x = 0, xend = chr_e, 
        y = comps, yend = comps),
    linewidth =  1.5, lineend = "round", colour = "grey75"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  comps, yend =  comps,
      colour = different
    ),
    linewidth = 1.5
  ) +
  scale_color_manual(values = c('0'='black', '1'='#ffd94d'))+
  geom_label_repel(aes(label=Nsnp, x=(start + end)/2, y = comps), size=4,label.padding = unit(0.05, "lines"),alpha = 1,label.size = 0.01, color = alpha('black', .2)) +
  facet_wrap(chr~., ncol = 7)+
  theme_classic() +
  theme(strip.text = element_text(size = 14, face = 'bold'),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1)) 
  })
```


```{r, fig.height=7, fig.width=20}
k8_ibd_sites_f[[1]] %>%
  map(~{
    .x%>% 
ggplot() +
  geom_segment(
    aes(x = 0, xend = chr_e, 
        y = chr, yend = chr),
    linewidth = 3.5, lineend = "round", colour = "grey75"
  ) +
  geom_segment(
    aes(
      x = start, xend = end,
      y =  chr, yend =  chr,
      colour = Nsnp#,
      # fill = Nsnp,
    ),
    linewidth = 3
  ) +
  # scale_color_manual(values = c('0'='black', '1'='#ffd94d'))+
  facet_wrap(comps~., ncol = 6) +
  theme_classic() +
  theme(strip.text = element_text(size = 14, face = 'bold'),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1))
  })
```

HMMIBD END!!!!!!!!!!!!!

igraph
```{r}
# #==========================================================
# # Function to create an adjacency matrix
# #=========================================================
# construct_adj_matrix = function(Result, Entry){
#   
#   sample_names = unique(c(as.character(Result$individual1), as.character(Result$individual2)))
#   sample_count = length(sample_names)
#   adj_matrix = array(data = NA, dim = c(sample_count, sample_count), 
#                      dimnames = list(sample_names, sample_names))
#   for(i in 1:nrow(Result)){
#     indi = as.character(Result$individual1[i])
#     indj = as.character(Result$individual2[i])
#     adj_matrix[indi, indj] = Result[i, Entry]
#   }
#   return(adj_matrix)
# }
# 
# 
# #==========================================================
# # Function to filter highly related samples within sites
# #==========================================================
# rm_highly_related_within = function(Result, Cities, Edge){
#   
#   if(Edge){
#     # Keep edges statistically diff. from one
#     Result_filtered = Result[Result$`r97.5.` < 1-eps,] 
#   } else {
#     # First construct an adjaceny matrix of 97.5% CI IBD
#     adj_matrix = construct_adj_matrix(Result, Entry = 'r97.5.')
#     sample_names = rownames(adj_matrix)
#     
#     # Second, modify s.t. only edges NOT statistically diff. from one
#     adj_matrix_high = adj_matrix
#     adj_matrix_high[adj_matrix < 1-eps] = 0
#     
#     # Third, create a graph and list its components
#     G = graph_from_adjacency_matrix(adjmatrix = adj_matrix_high, 
#                                     mode='upper', 
#                                     diag=F, weighted = 'w', 
#                                     add.colnames = T)
#     S = components(graph = G) # Lists components
#     
#     # Forth, draw a single sample per site within a component 
#     rep_samples = unlist(lapply(1:S$no, function(s){
#       samples_ids = which(S$membership==s)
#       if(length(samples_ids) == 1){
#         rep_samples = sample_names[samples_ids] # If only one sample per component, keep
#       } else { # Keep one per city
#         cities = Cities[sample_names[samples_ids]]
#         # Pick sample with earliest data (deterministically for comparison across plots)
#         rep_samples = sapply(unique(cities), function(city){
#           samples_to_choose_from = names(cities)[cities == city]
#           rep_sample = samples_to_choose_from[which.min(SNPData[samples_to_choose_from, 'COLLECTION.DATE'])]
#         })
#       }
#       return(rep_samples)
#     }))
#     
#     # Fifth, restructure Result s.t. only one sample per city per component
#     inds = Result$individual1 %in% rep_samples & Result$individual2 %in% rep_samples
#     Result_filtered = Result[inds,]
#   }
#   return(Result_filtered)
# }
# 
# #==========================================================
# # Function to return city subgraph
# #==========================================================
# extract_city_graph = function(x, city1, city2, col_edge = 'white', 
#                               a = 0, b = 1, c = 2){
#   inds = V(x)$site == city1|V(x)$site == city2 # indices for cities
#   z = induced_subgraph(x, vids = which(inds)) # extract subgraph
#   Jitter_sign = rep(1, length(V(z)$site ))
#   Jitter_sign[V(z)$site == city2] = -1
#   # vertical layout
#   Y = as.numeric(V(z)$site == city1) + (Jitter[V(z)$name] * Jitter_sign)
#   attr(z, 'layout') = cbind(X = V(z)$date, Y) # Full layout
#   V(z)$color = M_cols[V(z)$name] # Colour by clonal component
#   E(z)$weight[is.na(E(z)$weight)] = 0 # NAs introduced when filter edges 
#   
#   # Scale edge widths (not weights) and transparancy to range a, b, c
#   weights_rescaled = a + (b-a) * (E(z)$weight - min(E(z)$weight)) / (max(E(z)$weight) - min(E(z)$weight))
#   E(z)$width = weights_rescaled * c
#   E(z)$color = sapply(weights_rescaled, function(x) adjustcolor(col_edge, alpha.f = x)) # Make transparency a function of rhat
#   v_names = do.call(rbind, strsplit(attributes(E(z))$vnames, split = "\\|"))
#   edge_col_ind = which(M_cols[v_names[,1]] == M_cols[v_names[,2]] & M_cols[v_names[,1]] != "#FFFFFF")
#   if(any(edge_col_ind)){
#     E(z)$color[edge_col_ind] = M_cols[v_names[,1]][edge_col_ind]
#   }
#   return(z)}

```

```{r}
# #################################################################################
# #' Script to generate and plot a graph of relatedness between samples
# #################################################################################
# rm(list = ls())
# library(igraph) 
# library(RColorBrewer)
# source('../igraph_functions.R') # for construct_adj_matrix
# eps <- 0.01 # threshold below which LCI is considered zero in Taylor et al. 2019
# a = 0; b = 0.5; c = 0.25 # Scaling parameters for plotting edge width and transparancy
# PDF <- TRUE
# 
# # Load relatedness results and metadata
# freqs_used <- "Taylor2020"
# load(sprintf('../../RData/mles_CIs_extended_freqs%s_meta.RData', freqs_used)) 
# load(sprintf('../../RData/metadata_extended.RData'))
# load("../../RData/sids_to_remove_from_graphs.RData")
# 
# # Define city cols
# cities <- unique(metadata$City)
# cols_cities <- array(c(rev(brewer.pal(5, 'Spectral')), 
#                        brewer.pal(length(cities)-5, 'Dark2')), 
#                      dimnames = list(cities))
# 
# # Remove samples with one or more NA estimate
# sids_to_keep <- setdiff(metadata$SAMPLE.CODE, sids_remv) 
# inds_to_keep <- mle_CIs$individual1 %in% sids_to_keep & mle_CIs$individual2 %in% sids_to_keep
# mles_to_keep <- mle_CIs[inds_to_keep,]
# 
# A_est <- construct_adj_matrix(mles_to_keep, Entry = 'rhat')
# 
# if(PDF) pdf("../../Plots/Relatedness_graphs.pdf")
# 
# for(high_relatedness_threshold in c(0,0.25)){
#   A_est[A_est < high_relatedness_threshold] <- 0
#   fields::image.plot(A_est)
#   
#   # Create graph 
#   G_est <- graph_from_adjacency_matrix(A_est, mode='upper', diag=F, weighted = T) 
#   art_pnts <- names(articulation.points(G_est))
#   ccompnts <- components(G_est)
#   
#   # Add metadata
#   V(G_est)$marker_count <- metadata[V(G_est)$name,  "snp_count"]
#   V(G_est)$city <- metadata[V(G_est)$name, "City"]
#   V(G_est)$color <- cols_cities[V(G_est)$city] 
#   
#   weights_rescaled = a + (b-a) * (E(G_est)$weight - min(E(G_est)$weight)) / (max(E(G_est)$weight) - min(E(G_est)$weight))
#   
#   # Plot graph 
#   plot(G_est, 
#        main = sprintf("High relatedness threshold: %s", high_relatedness_threshold), 
#        cex.main = 0.75, 
#        vertex.label.color = "black", 
#        vertex.label.cex = 1, 
#        vertex.frame.color = NA,
#        vertex.label = NA, 
#        vertex.shape = c("square","circle")[metadata[V(G_est)$name, "PloSGen2020"] + 1],
#        vertex.size = 0.5*log(metadata[V(G_est)$name, "snp_count"]),
#        edge.width = weights_rescaled * c, 
#        edge.color = sapply(weights_rescaled, function(x)adjustcolor('black', alpha.f = x)))
#   
#   # Legend
#   cities_ <- unique(V(G_est)$city)
#   legend('bottomleft', pch = 16, bty = 'n', cex = 0.5, pt.cex = 1, 
#          inset = -0.1, col = cols_cities[cities_], legend = cities_)
#   
#   # R estimate range
#   range(A_est, na.rm = T)
# }
# 
# if(PDF) dev.off()
```

