---
title: "MSC DE field VS lab strain VS strain DE and visualizations"
date: '2023-08-10'
author: 'Jesse Rop'
output: 
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(SeuratData)
  library(SeuratDisk)
  library(pheatmap)
  library(plotly)
  library(rmarkdown)
  library(RColorBrewer)
  library(MAST)
  library(EnhancedVolcano)
  # library(org.Pf.plasmo.db)
  library(clusterProfiler) ##github installation newest version https://github.com/YuLab-SMU/clusterProfiler/issues/434#issuecomment-1048633162 remotes::install_github(c("YuLab-SMU/DOSE", "YuLab-SMU/clusterProfiler", "YuLab-SMU/enrichplot" ))
  library(Platypus)
  library(ggvenn)
  library(tradeSeq)
  library(MetBrewer)
  library(ComplexHeatmap)
  library(fgsea)
  library(msigdbr)
  library(slingshot)
  library(enrichplot)
  library(chromoMap)
  library(pdftools)
  library(SCpubr)
  library(GGally)
  library(circlize)
  library(readxl)
  library("ggVennDiagram")
  library(UpSetR)
  library(ComplexUpset)
  library(openxlsx)
  library(ggpubr)
  # library(ggpairs)
  library(conflicted)
  library(ggforce)
  conflict_prefer("map", "purrr")
  conflict_prefer("filter", "dplyr")
  conflict_prefer("rename", "dplyr")
  conflicts_prefer(SingleCellExperiment::colData)
})
```

# DE within MSC14 field parasites and DE on MSC14 VS MCA V3 parasites

## Variable declarations, read in input files and exploratory visualizations

```{r setup}
##Set working directory
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali')
```

```{r}
##Assign colour variables for the different plots



sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle'
)

##Field colours
fvl_cols <- c('Field' ='#aea677',
              'Lab' = "#e7f1f9") #"#ffe48d"

strain_cols_n <-c("#910000", "#ff1a8d", "#007c71", "#0ea300", "#00deca", "#9a96c7", "#5d57a4", "#a8875b", "#ffd18d", "#b4d2ff", "#cdb8c5", "#9770b3", "#C9E0F1", "#a40000" , "#00b7a7", "#ffcd12","thistle","#a40000" , "#00b7a7", "#e7f1f9", 'black', 'grey')
#, 'lavender', 'lightgray', 'grey'
names(strain_cols_n) <- c('SC2', 'SC1','SC5',  'SC4', 'SC3',  'SC6', 'SC7', 'SC8', '7G8', 'NF54','NF54_A', 'S2_b', 'S3_b', 'S1', 'S2', 'S3', 'Doublet', 'SC1_3i','Negative', 'Lab', 'Less5', NA)

#00BFFF n #E0115F -- Good blue colour to use somewhere

##Strain colours
# dset_cols <- c('Lab' ="#a9cee8",
dset_cols <- c('Lab' ="#e7f1f9",
               'Field' = "#aea677") # selected as mix of strain colours https://artincontext.org/color-mixer/

# strain_cols <- c(met.brewer(name = 'Austria')[c(1,7, 4,6)], 'light gray')
# # names(strain_cols) = c('Strain 0', 'Strain 1', 'Strain 2', 'Doublet', 'Negative')
# names(strain_cols) = c('G1', 'G2', 'G3', 'Doublet', 'Negative')

strain_dset_lvls = c( 'SC1', 'SC2','SC3',  'SC4', 'SC5',  'SC6', 'SC7', 'SC8', 'NF54', '7G8', 'NF54_A', 'Lab', 'Doublet')
m_stgs <- c('asx', 'f', 'fh', 'fl', 'mh')
pts <- c('asexual_pt', 'female_pt', 'female_pt', 'female_pt', 'male_pt')

fld_strns = paste0('SC', 1:8)
strain_dset_lvls = c( fld_strns, 'NF54', '7G8', 'NF54_A', 'Lab', 'Doublet')
sample_id_lst <-as.list(c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" ) %>% 
                          str_sort(., numeric = T))

strain_dset_lvls_m = paste(rep(str_remove(sample_id_lst[1:4], "SC"), each=length(fld_strns)), fld_strns, sep = '_')
strain_dset_lvls_don = c(strain_dset_lvls_m, strain_dset_lvls)
strain_dset_lvls_don_col = randomcoloR::distinctColorPalette(k = length(strain_dset_lvls_m)) %>% set_names(strain_dset_lvls_m) %>% c(., strain_cols_n)

strns_lab_comps <- combn(c(paste0("SC", 1:8), "Lab"), m = 2, simplify = F) %>% map(., ~paste(.[[1]], .[[2]], sep = "_vs_")) %>% unlist()

fld_strns_nf54_7g8_comps <- combn(c(paste0("SC", 1:8), "NF54", "7G8"), m = 2, simplify = F) %>% map(., ~paste(.[[1]], .[[2]], sep = "_vs_")) %>% unlist()

## Get colours randomly - do once and save output as the colors change everytime it is ran
# fld_strns_nf54_7g8_comps_col = randomcoloR::distinctColorPalette(k = length(fld_strns_nf54_7g8_comps), runTsne = F) %>% set_names(fld_strns_nf54_7g8_comps)
fld_strns_nf54_7g8_comps_col = c(SC1_vs_SC2 = "#EE643B", SC1_vs_SC3 = "#82AEE6", SC1_vs_SC4 = "#C2705C", SC1_vs_SC5 = "#B397DE", SC1_vs_SC6 = "#E6E68D", SC1_vs_SC7 = "#733BA0", SC1_vs_SC8 = "#E3BAE6", SC1_vs_NF54 = "#5AD446", SC1_vs_7G8 = "#ECAC50", SC2_vs_SC3 = "#695BE7", SC2_vs_SC4 = "#EDE054", SC2_vs_SC5 = "#4DC2E4", SC2_vs_SC6 = "#ABEBA6", SC2_vs_SC7 = "#E9BE9D", SC2_vs_SC8 = "#A49CA2", SC2_vs_NF54 = "#672EE6", SC2_vs_7G8 = "#EC45D6", SC3_vs_SC4 = "#A2B978", SC3_vs_SC5 = "#7AD8E1", SC3_vs_SC6 = "#6D8AD7", SC3_vs_SC7 = "#9E5C89", SC3_vs_SC8 = "#4D8899", SC3_vs_NF54 = "#5EEADE", SC3_vs_7G8 = "#67EE8C", SC4_vs_SC5 = "#BEE2DE", SC4_vs_SC6 = "#E1489C", SC4_vs_SC7 = "#E58DD1", SC4_vs_SC8 = "#807AE7", SC4_vs_NF54 = "#6CB69E", SC4_vs_7G8 = "#D274E6", SC5_vs_SC6 = "#61ECB7", SC5_vs_SC7 = "#C63FEB", SC5_vs_SC8 = "#9C8556", SC5_vs_NF54 = "#B1E575", SC5_vs_7G8 = "#ADD2EC", SC6_vs_SC7 = "#CAE33A", SC6_vs_SC8 = "#ABEBCD", SC6_vs_NF54 = "#B2AF48", SC6_vs_7G8 = "#E7A5AC", SC7_vs_SC8 = "#E46378", SC7_vs_NF54 = "#D6DEB0", SC7_vs_7G8 = "#E7D4E0", SC8_vs_NF54 = "#9BF43C", SC8_vs_7G8 = "#49B36F", NF54_vs_7G8 = "#E6E7D5", Field_vs_Lab = "#E9A3DC")
```



```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id)
  )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                 TRUE ~ as.character(.)))) 

```

### read in the intergrated objects for downstream MAST DE and visualizations analysis
```{r, fig.width = 32, fig.height = 7}

msc_gam_combi_noLE_de <- readRDS("data/processed/DB52e_star/msc_gam_combi_noLE_de.RDS")
msc_gcombi_noLE_sub_sce <- readRDS("data/processed/DB52e_star/msc_gcombi_noLE_sub_sce.RDS")

```  

### Tradeseq fit-gam and condition test in farm
commands to paste to farm for running tradeseq
```{r, eval=FALSE}
##!! NB - Run in farm for speed - parallelization
# AMMENDED ON 160421 TO submit jobs independently using the script /lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/multipurpose_scripts/tradeseq.R submitted by /lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/scripts/tseq_submsn.sh 
#!/usr/bin/env

##Check if all the positional variables need are provided, if not then exits and prints out the requirements

.libPaths('/lustre/scratch126/tol/teams/lawniczak/users/jr35/local_libs/R410_libs')

##load libraries
library(tradeSeq)
library(RColorBrewer)
library(SingleCellExperiment)
library(slingshot)
library(BiocParallel)
library(purrr)

##Initializes the positional variables
args = commandArgs(trailingOnly=TRUE)

i <- args[1]

msc_g_noLE_sub_sce_unbal_f10 <- readRDS('/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_inp/msc_g_noLE_sub_sce_unbal_f10.RDS')
msc_y21_sub_ptime_unbal_f10 <- readRDS('/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_inp/msc_y21_sub_ptime_unbal_f10.RDS')
msc_y21_sub_cw_unbal_f10 <- readRDS('/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_inp/msc_y21_sub_cw_unbal_f10.RDS')

# condtns <- c('dset','dset','dset','dset','dset','strain_dset','strain_dset','strain_dset','strain_dset','strain_dset' ,'strain','strain','strain','strain','strain')
condtns <- c('strain','strain','strain','strain','strain','dset','dset','dset','dset','dset')
comps <- rep(c('fh','mh'), 2)


##fitting GAMs
knots_no <- c(6,6,6,6,6,8,7,8,9,8)
# knots_no <- c(8,7,8,9,8,6,6,6,8,8)

sce_obj = c(msc_g_noLE_sub_sce_unbal_f10,msc_g_noLE_sub_sce_unbal_f10)
ptime_obj = c(msc_y21_sub_ptime_unbal_f10, msc_y21_sub_ptime_unbal_f10)
cw_obj = c(msc_y21_sub_cw_unbal_f10, msc_y21_sub_cw_unbal_f10)

i=as.numeric(i)

## Subset the relevant cells for each run
substd_sce = sce_obj[[i]][,!is.na(sce_obj[[i]]@colData[,condtns[i]])]
substd_ptime = ptime_obj[[i]][!is.na(sce_obj[[i]]@colData[,condtns[i]])]
substd_cw = cw_obj[[i]][!is.na(sce_obj[[i]]@colData[,condtns[i]])]

## Tradeseq script
msc_y21_ref_ts_sce_ub_de  <- fitGAM(counts = counts(substd_sce), 
                                  pseudotime = substd_ptime, 
                                  cellWeights = substd_cw, 
                                  nknots =knots_no[i],
                                  genes = rownames(substd_sce)[rowSums(counts(substd_sce) != 0) > 5],
                                  conditions = factor(SummarizedExperiment::colData(substd_sce)[, condtns[i]]), 
                                  verbose = T, 
                                  parallel = T)


## Save tradeseq object
saveRDS(msc_y21_ref_ts_sce_ub_de, paste0('/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_op/tseq_obj_', condtns[i], '_', comps[i], '.RDS'))

##Running condition test
msc_y21_condRes_ub_de <- conditionTest(msc_y21_ref_ts_sce_ub_de, pairwise = T)

## Save conditional test object
saveRDS(msc_y21_condRes_ub_de, paste0('/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_op/tseq_cond_obj_', condtns[i], '_', comps[i], '.RDS'))

```

transfer tradeseq results files from farm to local
```{r, eval=T}
# bst <- c(rep('_unbal', 10), rep('', 10), rep('', 5))
# bst <- c(rep('', 15))
m_stgs<-names(msc_gcombi_noLE_sub_sce)

for (ts in c('cond_obj', 'obj')){
    for (st in m_stgs) {
      for (ds in c('dset', 'lf_strain')){
      system(paste0('rsync -av --delete jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_op/tseq_',ts,'_', ds, '_', st,'.RDS /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/Mali/data/processed/DB52e_star/pseudotime_tradeseq_op/'))
    }
  }
}


```

Read in tradeseq files from local
```{r}
list.files('data/processed/DB52e_star/pseudotime_tradeseq_op', pattern = "tseq_obj.*_M.*DS$", full.names = T) %>% 
  str_sort(., numeric = T)  %>% 
  set_names(nm = (str_remove_all(.,'data/processed/DB52e_star/.*obj_|\\.RDS') %>% str_replace_all(., c('dset'='FieldVlab', 'strain'='StrainsDE', 'lf_strain'='LFStrainsDE')) ))
```


```{r}
ts_sce_ub_de_k7 <- list.files('data/processed/DB52e_star/pseudotime_tradeseq_op', pattern = "tseq_obj.*_M.*DS$", full.names = T) %>% 
  str_sort(., numeric = T)  %>% 
  set_names(nm = (str_remove_all(.,'data/processed/DB52e_star/.*obj_|\\.RDS') %>% str_replace_all(., c('dset'='FieldVlab', 'strain'='StrainsDE', 'lf_strain'='LFStrainsDE')) )) %>%
  map(., ~readRDS(.x))
```

```{r}
de_order <- c(paste0("FieldVlab_", m_stgs[1:8]), paste0("lf_StrainsDE_", m_stgs[9:28]))
# ts_sce_ub_de_k7 <- ts_sce_ub_de_k7[c(de_order, paste0(de_order[1:10], '_unbal'))]
ts_sce_ub_de_k7 <- ts_sce_ub_de_k7[c(de_order)]

```

Read in tradeseq condition DE files from local
```{r}
condRes_de_k7 <- list.files('data/processed/DB52e_star/pseudotime_tradeseq_op', pattern = "tseq_cond_obj.*_M.*DS$", full.names = T) %>% str_sort(., numeric = T)  %>% 
  set_names(nm = (str_remove_all(.,'data/processed/DB52e_star/.*obj_|\\.RDS') %>% str_replace_all(., c('dset'='FieldVlab', 'strain'='StrainsDE', 'lf_strain'='LFStrainsDE')) ))  %>%
  map(., ~readRDS(.x))
```

```{r}
condRes_de_k7 <- condRes_de_k7[c(de_order)]
condRes_de_k7 <- condRes_de_k7[!is.na(names(condRes_de_k7))]
ts_sce_ub_de_k7 <- ts_sce_ub_de_k7[names(condRes_de_k7)]

```


Read in tradeseq files from local
```{r}
# msc_y21_ref_ts_sce_ub_de <- readRDS('data/processed/DB52e_star/pseudotime_tradeseq/msc_y21_ref_ts_sce_ub_de.RDS') %>% 
#   set_names(paste0(names(.), '_', c(rep('FieldVlab', 5), rep('StrainsDE', 5))))
```

#### Visualize cells tradeseq trajectories {.tabset}

#### Read in and format tradeseq condition test results (lab VS field, field VS field) (Performed in farm)  


```{r}
##NOTE!!! replace the conditional test file with the newly generated one which applied 7 clusters
##Ajusting the global and pairwise pvalues - old p values discarded
msc_g_noLE_condRes_ub_de <- map(condRes_de_k7, ~.x %>%
                             mutate(across(starts_with('pvalue'), ~p.adjust(., "fdr"))) %>%
                             arrange(desc(waldStat)) %>%
                             rownames_to_column('gene_id'))

```



```{r}
##save condition results for GSEA
saveRDS(msc_g_noLE_condRes_ub_de, 'data/processed/DB52e_star/msc_g_noLE_condRes_ub_de.RDS')
```

```{r}
msc_g_noLE_ref_ts_sce_ub_de <- ts_sce_ub_de_k7
```


```{r}
### based on mean smoother
### All these genes are in the top 22 with waldstat > 20
cond_gns_005_ub_de <- pmap(list(msc_g_noLE_condRes_ub_de, msc_g_noLE_ref_ts_sce_ub_de),~{
  
  b = levels(..2$tradeSeq$conditions)
  
  print(b)
  
  if (dim(..1)[2] > 4) {
     ..1 %>% rename_with(., ~str_replace_all(., c('s7' = b[7],
                                               's6' = b[6],
                                               's5' = b[5],
                                               's4' = b[4],
                                               's3' = b[3],
                                               's2' = b[2],
                                               's1' = b[1],
                                               # ' ' = '',
                                               'vm'='_vs_m',
                                               'vS'='_vs_S',
                                               'vL'='_vs_L',
                                               'vN'='_vs_N',
                                               'v7'='_vs_7',
                                               #'4 \\(O'='4\\(O',
                                               'cond'=''))
  )
  } else {
  
    
    ..1 %>% mutate(across(.cols = -c("gene_id"), .names = '{.col}_2')) %>% rename_with(., ~str_replace_all(., c('2' = paste0(b[1], "_vs_", b[2]),
                                               #'4 \\(O'='4\\(O',
                                               'cond'='')))
  
  }
  
})


```

```{r}

cond_gns_005_ub_de <- c(cond_gns_005_ub_de) %>%
  map(., ~left_join(., Pf3D7_genes_plasmoDB %>% dplyr::select(gene_id, Gene, Ka_Ks, `Transcript Product Description`, TM_domains, `Computed GO Functions`), by = 'gene_id'))

```


```{r}
# msc_y21_ref_ts_sce_ub_de <- msc_y21_ref_ts_sce_ub_de[c(1:15)]
# cond_gns_005_ub_de <- cond_gns_005_ub_de[c(1:15)]
```

#### Plot smoothers function


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
tde_smooth_plt <- function(tde_obj_lst = msc_g_noLE_ref_ts_sce_ub_de[10],
                           cond_obj_lst = cond_gns_005_ub_de[10],
                           pw_comp= 'waldStat_SC2_vs_SC6', 
                           n_gns = c(1:5)
)
{
  
  list(tde_obj_lst, cond_obj_lst) %>% 
    pmap(.,possibly(~{
      
      gns = ..2 %>% filter(!!rlang::sym(pw_comp) > 20) %>% arrange(desc(!!rlang::sym(pw_comp)))  %>% pull(gene_id) %>% .[n_gns]
      
      gn_names = ..2 %>% filter(!!rlang::sym(pw_comp) > 20) %>% arrange(desc(!!rlang::sym(pw_comp))) %>% pull(Gene) %>% .[n_gns]
      print(gn_names)
      
      tde = ..1
      
      list(gns, gn_names) %>% 
        pmap(., possibly(~{
          plotSmoothers(tde, 
                        assays(tde)$counts,
                        gene = ..1,
                        alpha = 1, 
                        border = TRUE) +
            labs(title = ..2)
        }, 'missing'))
    }, 'missing'))
}

```


#### smoothers for select genes that are differentially expressed between strains {.tabset}


```{r, results='asis', warning=F, message=F}

tde_smooth_plt(tde_obj_lst = msc_g_noLE_ref_ts_sce_ub_de[str_detect(names(msc_g_noLE_ref_ts_sce_ub_de), "MSC1_")],
               cond_obj_lst = cond_gns_005_ub_de[str_detect(names(cond_gns_005_ub_de), "MSC1_")],
               pw_comp= 'waldStat_SC1_vs_SC2',
               n_gns = c(1:5)
)

```


```{r, results='asis', warning=F, message=F}

tde_smooth_plt(tde_obj_lst = msc_g_noLE_ref_ts_sce_ub_de[str_detect(names(msc_g_noLE_ref_ts_sce_ub_de), "MSC3_")],
               cond_obj_lst = cond_gns_005_ub_de[str_detect(names(cond_gns_005_ub_de), "MSC3_")],
               pw_comp= 'waldStat_SC1_vs_SC2',
               n_gns = c(1:5)
)

```


Heatmap plotting function
```{r}
### Annotation groups for the header ie column
tseq_de_hmap <- function(tseq = msc_g_noLE_ref_ts_sce_ub_de[[1]], de_tbl , pnts = 30, rnames = F, comp_grps){
  
  # gns = de_tbl %>% filter(if_any(matches("pvalue.*"), ~. < 0.05)) %>% pull(gene_id)
  gns = de_tbl %>% filter(pvalue < 0.05) %>% pull(gene_id)

  # print(str(gns))
  ##Edited from the original version to allow the labelling of rows with gene names
  msc_yhatSmooth <- predictSmooth(tseq, gene = gns, nPoints = pnts, tidy = FALSE) %>% 
        t() %>%scale() %>% t() %>% as.data.frame()%>% 
        rownames_to_column('gene_id') %>% 
        left_join(., Pf3D7_genes_plasmoDB  %>% dplyr::select("Gene", "gene_id"), by = 'gene_id') %>% 
        dplyr::select(-c(gene_id)) %>% 
        column_to_rownames('Gene')
  
  # print(str(msc_yhatSmooth))
  msc_col_annot <- predictSmooth(tseq, gene = gns, nPoints = pnts, tidy = T)  %>% 
        mutate(point = paste0("point", rep(rep(1:pnts,length(unique(tseq@colData$tradeSeq$conditions))),length(gns))), 
               lin_id = paste0('lineage', lineage,'_condition', condition, '_',point)) %>%
        rownames_to_column('No')  %>% 
        dplyr::select(lin_id,time,condition)%>% 
        distinct()%>% 
        column_to_rownames('lin_id') %>%
        rename('Pseudotime' = time, 'Condition' = condition)
  # print(str(msc_col_annot))

  hmap<- pheatmap::pheatmap(msc_yhatSmooth %>% as.data.frame() %>% as.matrix(),
                         cluster_cols = FALSE,
                         show_rownames = {{rnames}}, 
                         show_colnames = F, 
                         main = comp_grps,
                         legend = T,
                         # silent = TRUE#, 
                         annotation_row = Pf3D7_genes_plasmoDB %>%filter(!duplicated(gene_id)) %>% 
                           dplyr::select("Gene", "Ka_Ks") %>% column_to_rownames("Gene"),
                         annotation_col = (msc_col_annot %>% dplyr::select(Pseudotime, Condition))
                         # 
      )
  # str(hmap)
  return(hmap)
  # print(hmap)
  # print(pheatmap::pheatmap(msc_yhatSmooth %>% as.data.frame() %>% as.matrix()))
      cat('\n')
      cat("\n")
    
}
```



#### heatmaps for DE between field and lab in the asexual, female and males {.tabset}

```{r, results='asis', warning=F, message=F}

hmap_plt<- list(msc_g_noLE_ref_ts_sce_ub_de, 
                # map(cond_gns_005_ub_de, ~.x %>% filter(if_any(matches("pvalue"), ~. < 0.05)) %>% pull(gene_id)), 
                cond_gns_005_ub_de,
                rep(30, length(cond_gns_005_ub_de)), 
                rep(F, length(cond_gns_005_ub_de)), 
                names(msc_g_noLE_ref_ts_sce_ub_de)) %>% 
  pmap(., possibly(~{
  tseq_de_hmap(tseq = ..1, de_tbl= ..2, pnts = ..3, rnames = ..4, comp_grps = ..5)
}, 'missing genes'))
  
  # tseq_de_hmap(tseq = msc_g_noLE_ref_ts_sce_ub_de[[1]], gns= map(cond_gns_005_ub_de[1], ~.x %>% filter(if_any(matches("pvalue"), ~. < 0.05)) %>% pull(gene_id)) %>% flatten() %>% as.character(), pnts = 30, rnames = F, title = "Strain DE")
  #                      
hmap_plt

```


#### Correlation between waldStat association signal strength between comparisons

```{r, warning=F, message=F}

lowerfun <- function(data, mapping) {
  ggplot(data = data, mapping = mapping) +
    geom_point(size = 0.3, alpha = 0.3) +
    geom_abline(slope = 1, intercept = 0, color = "blue")
}

cond_gns_005_ub_de%>% 
  map(. %>% dplyr::select('gene_id', waldStat) %>% filter(waldStat<5000)) %>% 
  bind_rows(., .id = "stages") %>% 
  pivot_wider(names_from = 'stages', values_from = 'waldStat') %>% 
  column_to_rownames('gene_id') %>% 
  ggpairs(lower = list(continuous = GGally::wrap(lowerfun)),
          upper = list(continuous = wrap("cor", size = 5)))
```


## MAST DE adjusting for pseudotime

MAST DE on the entire dataset before subsampling. Subsampling is not necessary for MAST and also for tradeseq
### Prepare dataset
Normalizing and scaling the data in the subseted seurat objects - may have been done before on the larger parent object but just repeating since will not change anything as it starts from the raw counts.

Adding pseudotime metadata from the single cell experiment object 
Scaling is for visualization


```{r}
##REMEMBER TO REMOVE CLUSTERS WITH VERY FEW CELLS AND DOUBLETS SEURAT
# msc_gam_combi_noLE_de <- msc_gam_combi_noLE_de[, cells_de]
```


```{r, message=FALSE, warning=FALSE}
# Add to metadata the dataset/conditions and ptime cat
# DefaultAssay(msc_gam_combi_noLE_de[[1]]) <- "RNA"
# DefaultAssay(msc_gam_combi_noLE_de[[2]]) <- "RNA"
DefaultAssay(msc_gam_combi_noLE_de[[1]]) <- "RNA"

all_gns = unique(rownames(msc_gam_combi_noLE_de[[1]]))
# 
# msc_g_noLE_v3_combi_scld <- msc_gam_combi_noLE_de %>%
#   NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000, verbose = FALSE) %>% 
#   ScaleData(., features = all_gns,  verbose = FALSE) %>%
#   .[,rownames(fld_vs_lb_mdata)] %>% 
#   AddMetaData(., fld_vs_lb_mdata)
# 
# 
# 
# msc_g_noLE_v3_combi_afm_seu <- list(msc_g_noLE_sub_sce_unbal_f10, pts) %>%
#     pmap(~msc_g_noLE_v3_combi_scld[, colnames(..1[,!is.na(..1@colData[,'strain_dset'])])] 
#     )


##Subset original seurat object with only cells in the subsetted DE objects and add pseudotime as metadata
# msc_g_noLE_v3_combi_afm_seu <- list(msc_g_noLE_sub_sce_unbal_f10, pts) %>%
#   pmap(~msc_gam_combi_noLE_de[, colnames(..1[,!is.na(..1@colData[,'strain_dset'])])] %>%
#          AddMetaData(., SummarizedExperiment::colData(..1)[, ..2, drop=F] %>% data.frame() ) %>%
#          NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000, verbose = FALSE) %>%
#          ScaleData(., features = all_gns,  verbose = FALSE)
#   )

##Doing MAST for the lf_strain - lab_field_strain
pts <- rep(c("female_pt", "male_pt"), 14)
# msc_g_noLE_v3_combi_afm_seu <- list(msc_gcombi_noLE_sub_sce[6:15], pts_lfs, c(rep('strain_dset', 5), rep('lf_strain', 5))) %>%
msc_g_noLE_v3_combi_afm_seu <- list(msc_gcombi_noLE_sub_sce, pts, c(rep('strain_dset', 8), rep('lf_strain', 20)), rep(msc_gam_combi_noLE_de, times = 28)) %>%
  pmap(~..4[, colnames(..1[,!is.na(..1@colData[,..3])])] %>%
         AddMetaData(., SummarizedExperiment::colData(..1)[, ..2, drop=F] %>% data.frame() ) %>%
         NormalizeData(., normalization.method = "LogNormalize", scale.factor = 10000, verbose = FALSE) %>%
         ScaleData(., features = all_gns,  verbose = FALSE)
  )
```


### Actual MAST DE between pairwise strains and fieldvslab 

MAST DE adjusting for pseudotime as a latent variable

```{r, message=FALSE, warning=FALSE}
de_idents <- c(rep('dset', 8), rep('lf_strain', 20))
msc_g_noLE_v3_combi_strn_mast<- map2(msc_g_noLE_v3_combi_afm_seu, de_idents, ~SetIdent(.x, value=.y))

```

```{r, message=FALSE, warning=FALSE}
saveRDS(msc_g_noLE_v3_combi_strn_mast, 'data/processed/DB52e_star/msc_g_noLE_v3_combi_strn_mast.RDS')

```

```{r, message=FALSE, warning=FALSE, eval=F}
##!! NB Rerun only when the DE inputs change - takes looooonng
## Mast DE of dataset with male low (ml) included - shouldnt have any contribution as each stage handled differently
# strns_ub_pairwise_mkrs_pt_ml <- list(msc_g_noLE_v3_combi_strn_mast, rep(pts,2), de_idents) %>%
#   pmap( ~{
#     # ide = ..3
#     a = combn(levels(droplevels(..1@meta.data[,..3])), 2,simplify = F) 
#     obj = ..1
#     pseudos = ..2
#     map(a, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], latent.vars = pseudos, logfc.threshold = 0, test.use = 'MAST') %>% rownames_to_column('gene_id')) %>%
#       set_names(map(a, ~paste(., collapse = '_vs_') %>% str_remove_all(' '))) %>%
#       bind_rows(., .id = 'pw_comparison')
#     
#   })
# saveRDS(strns_ub_pairwise_mkrs_pt_ml,'data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml.RDS')


## Mast DE of dataset with male low (ml) included - shouldnt have any contribution as each stage handled differently
## Including lab and field strain DE
## !!!min.cells.group added for mh in the strain_dset where there's only balancing in lab VS field - we need to remove later

pts_all<- c(rep(pts,3))
pts_all<- pts

strns_ub_pairwise_mkrs_pt_ml <- list(msc_g_noLE_v3_combi_strn_mast, pts, de_idents, names(msc_g_noLE_v3_combi_strn_mast)) %>%
  pmap(., possibly( ~{
    # ide = ..3
    a = combn(levels(droplevels(..1@meta.data[,..3])), 2,simplify = F) 
    obj = ..1
    pseudos = ..2
    
    print(paste("running ...", ..4))
    
    map(a, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], latent.vars = pseudos, logfc.threshold = 0, test.use = 'MAST', min.cells.group = 1) %>% rownames_to_column('gene_id')) %>%
      set_names(map(a, ~paste(., collapse = '_vs_') %>% str_remove_all(' '))) %>%
      bind_rows(., .id = 'pw_comparison')
    
  }, 'missing genes'))

saveRDS(strns_ub_pairwise_mkrs_pt_ml,'data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml.RDS')
# c(strns_ub_pairwise_mkrs_pt_ml_lf, strns_ub_pairwise_mkrs_pt_ml)



```


```{r, message=FALSE, warning=FALSE, eval=F}
##!! NB Rerun only when the DE inputs change - takes looooonng
### Getting all genes for gsea - saved as RDS and analysis done in functional_GO_n_lit_comparison.RMD

# strns_ub_pairwise_mkrs_pt_ml_gsea_full <- list(msc_g_noLE_v3_combi_strn_mast, rep(pts,2), de_idents) %>%
#   pmap( ~{
#     # ide = ..3
#     a = combn(levels(..1@meta.data[,..3]), 2,simplify = F) 
#     obj = ..1
#     pseudos = ..2
#     map(a, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], latent.vars = pseudos, logfc.threshold = 0, test.use = 'MAST', min.pct = -Inf, min.diff.pct = -Inf) %>% rownames_to_column('gene_id')) %>%
#       set_names(map(a, ~paste(., collapse = '_vs_') %>% str_remove_all(' '))) %>%
#       bind_rows(., .id = 'pw_comparison')
#     
#   })
# saveRDS(strns_ub_pairwise_mkrs_pt_ml_gsea_full,'data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml_gsea_full.RDS')

```

```{r, message=FALSE, warning=FALSE}
strns_ub_pairwise_mkrs_pt_ml<- readRDS('data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml.RDS')
# strns_ub_pairwise_mkrs_pt_ml<- c(strns_ub_pairwise_mkrs_pt_ml[1:16], strns_ub_pairwise_mkrs_pt_ml_new)
# names(strns_ub_pairwise_mkrs_pt_ml) <- names(msc_g_noLE_v3_combi_strn_mast)

# strns_ub_pairwise_mkrs_pt_ml_all <- c( strns_ub_pairwise_mkrs_pt_ml,strns_ub_pairwise_mkrs_pt_ml_lf)
# saveRDS(strns_ub_pairwise_mkrs_pt_ml_all,'data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml_all.RDS')
# 
# strns_ub_pairwise_mkrs_pt_ml_all<- readRDS('data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml_all.RDS')
# strns_ub_pairwise_mkrs_pt_ml <- strns_ub_pairwise_mkrs_pt_ml_all
```

MAST DE not adjusting for pseudotime (no pseudotime - npt)
```{r, message=FALSE, warning=FALSE, eval=F}
##!! NB Rerun only when the DE inputs change - takes looooonng - also not necessary. Only used to assess effect of adjusting for pseudotime
strns_ub_pairwise_mkrs_npt_ml <- list(msc_g_noLE_v3_combi_strn_mast, rep(pts,2), de_idents) %>%
  pmap( ~{
    a = combn(levels(..1@meta.data[,..3]), 2,simplify = F) 
    obj = ..1
    pseudos = ..2
    map( a, ~FindMarkers(obj, ident.1 = .x[1], ident.2 = .x[2], logfc.threshold = 0, test.use = 'MAST') %>% rownames_to_column('gene_id')) %>%
      set_names(map(a, ~paste(., collapse = '_vs_') %>% str_remove_all(' '))) %>%
      bind_rows(., .id = 'pw_comparison')
    
  })

saveRDS(strns_ub_pairwise_mkrs_npt_ml,'data/processed/DB52e_star/strns_ub_pairwise_mkrs_npt_ml.RDS')

```

```{r, message=FALSE, warning=FALSE, eval=FALSE}
strns_ub_pairwise_mkrs_npt_ml<- readRDS('data/processed/DB52e_star/strns_ub_pairwise_mkrs_npt_ml.RDS')

```

#### Comparing MAST DE results when we adjust for pseudotime and when we dont

```{r, message=FALSE, warning=FALSE}
# strns_ub_pairwise_mkrs_npt_ml <- strns_ub_pairwise_mkrs_npt_ml %>% map(. %>% mutate(across(pw_comparison, ~str_replace_all(.,c('Field_vs_Lab' = 'Field_vs_Lab')))))

```

#### Comparing MAST DE results when we adjust for pseudotime and when we dont
```{r, fig.width= 8, fig.height=3, eval=F}
## !! NB - also not necessary. Only used to assess effect of adjusting for pseudotime
list(strns_ub_pairwise_mkrs_pt_ml, strns_ub_pairwise_mkrs_npt_ml) %>%
  pmap(
    ~..1 %>% 
      left_join(., ..2, by = c('gene_id', 'pw_comparison'), suffix =c(".pt", ".npt")) %>%
      ggplot(., aes(x=log10(p_val_adj.pt), y=log10(p_val_adj.npt))) + 
      geom_point() + 
      facet_grid(.~pw_comparison, scales = "free_x") +
      geom_vline( xintercept = -3)+
      geom_hline( yintercept = -3) +
      geom_abline()
  )

```


```{r}
##!!!!!NOTE - remove the empty MSC1_fle
# strns_ub_pairwise_mkrs_pt_ml_sst <- strns_ub_pairwise_mkrs_pt_ml[c(1:5, 7:20)]
# msc_g_noLE_v3_combi_afm_seu_sst <- msc_g_noLE_v3_combi_afm_seu[c(1:5,7:20)]
strns_ub_pairwise_mkrs_pt_ml_sst <- strns_ub_pairwise_mkrs_pt_ml
msc_g_noLE_v3_combi_afm_seu_sst <- msc_g_noLE_v3_combi_afm_seu
cond_gns_005_ub_de_sst <- cond_gns_005_ub_de
pts_sst <- pts

```

### MAST VS Tradeseq analsys
####  MAST gene QC
List genes expressed in less than 5 cells in each subset objects used as MAST input that were also removed when running tradeseq fitGAM

```{r}
map2(strns_ub_pairwise_mkrs_pt_ml_sst, msc_g_noLE_v3_combi_afm_seu_sst, ~.x %>% filter(gene_id %in% rownames(.y)[!(rowSums(.y@assays$RNA@counts != 0) > 5)])) %>% bind_rows(., .id = 'Stage')

```

Remove genes above ????NB NB - refer line 734
```{r}
strns_ub_pairwise_mkrs_pt_ml_sst <- map2(strns_ub_pairwise_mkrs_pt_ml_sst, msc_g_noLE_v3_combi_afm_seu_sst, ~.x %>% filter(gene_id %in% rownames(.y)[(rowSums(.y@assays$RNA@counts != 0) > 5)]))

```

#### MAST & Tradeseq merge
Merge MAST results with tde results so as to use cut-off accounting for both methods (P<0.05 MAST & waldstat >10 TDE)
```{r, fig.width= 8, fig.height=3, warning=FALSE}
mast_tde_corr_tbls_w_na <- c(
  list(cond_gns_005_ub_de_sst[1:8], strns_ub_pairwise_mkrs_pt_ml_sst[1:8]) %>% 
    # pmap(~{ left_join(..1 %>% dplyr::select(gene_id, waldStat, pvalue), ..2, by = 'gene_id') %>%
    pmap(~{ full_join(..2, ..1 %>% dplyr::select(gene_id, waldStat, pvalue),  by = 'gene_id')}),
  list(cond_gns_005_ub_de_sst[c(9:28)], strns_ub_pairwise_mkrs_pt_ml_sst[c(9:28)]) %>%
    pmap(
      ~..1 %>% 
        dplyr::select(gene_id, starts_with(c('waldStat_','pvalue_'))) %>%
        pivot_longer(.,cols = starts_with(c('waldStat_','pvalue_')), 
                     # names_pattern = "(.*)_(SC._vs.*)",
                     names_pattern = "(waldStat|pvalue)_(.*)",
                     names_to = c( ".value", "ts_pw_comp")) %>% 
        # left_join(., ..2, by = c('gene_id', 'ts_pw_comp'='pw_comparison')) 
        full_join(..2, ., by = c('gene_id', 'pw_comparison'='ts_pw_comp')) 
    )
)

```


##### MAST & Tradeseq NA genes 
Check the genes with missing values in either test - tradeseq/MAST
Thrown out probably because of not meeting the log fold threshold
```{r, fig.width= 8, fig.height=3, warning=FALSE}
##mast missing
map(mast_tde_corr_tbls_w_na, ~.x %>% dplyr::count(pw_comparison, is.na(waldStat), is.na(p_val_adj))) %>% bind_rows(.id = 'stage') %>% filter(`is.na(p_val_adj)` == T)

##tde missing
map(mast_tde_corr_tbls_w_na, ~.x %>% dplyr::count(pw_comparison, is.na(waldStat), is.na(p_val_adj))) %>% bind_rows(.id = 'stage') %>% filter(`is.na(waldStat)` == T)

```

Remove mast missing genes above
```{r}
map(mast_tde_corr_tbls_w_na,  ~.x %>% filter(is.na(waldStat)) %>% ggplot(., aes(x = pw_comparison, y =avg_log2FC)) +geom_violin())
```

```{r}
map(mast_tde_corr_tbls_w_na,  ~.x %>% filter(is.na(waldStat)) %>% ggplot(., aes(x = log10(p_val_adj), y =avg_log2FC)) +geom_point())
```



```{r}
map(mast_tde_corr_tbls_w_na,  ~.x %>% filter(is.na(p_val_adj)) %>% ggplot(., aes(x = waldStat, y =log10(pvalue))) +geom_point() + geom_hline(yintercept = 0.05) +geom_vline(xintercept = 10))
```

```{r}
map(mast_tde_corr_tbls_w_na,  ~.x %>% filter(is.na(waldStat) |is.na(p_val_adj)) %>% dim)

```

```{r}
mast_tde_corr_tbls <- map(mast_tde_corr_tbls_w_na,  ~.x %>% filter(!is.na(waldStat) & !is.na(p_val_adj)))

```

```{r}
##Number of DE genes
map(mast_tde_corr_tbls, ~.x |> filter(p_val_adj < 0.05 & waldStat>10) |> pull(gene_id) |> unique() |> length())

```

```{r}
saveRDS(mast_tde_corr_tbls, 'data/processed/DB52e_star/mast_tde_corr_tbls.RDS')

```


```{r}

##Function to save tables in excel as sheets

createWorksheet <- function(x, y){
  wb <- createWorkbook()
  for(i in seq_along(x)){
    addWorksheet(wb, names(x)[i])
    writeData(wb, sheet = names(x)[i], x[[i]])
    saveWorkbook(wb, file = y, overwrite = TRUE)
  }
}

```


```{r}
##!!!! NOTE - MANUSCRIPT MAST TRADESEQ DE table
##Preprocess the list of field_lab DE dfs with appropriate names to workbook with each stage as a sheet and add important metadata
mast_tde_de_export <- mast_tde_corr_tbls %>% map(., ~filter(., waldStat > 10 & p_val_adj <0.05 &(str_detect(pw_comparison, 'Field_vs_Lab') | str_detect(pw_comparison, 'SC._vs_SC.') | str_detect(pw_comparison, 'NF54_vs_7G8'))) %>% dplyr::rename(c('tradeseq_waldStat' = waldStat, 'tradeseq_pvalue' = pvalue)) %>% dplyr::rename_with(., ~paste0('MAST_',.), c('p_val', 'avg_log2FC', 'pct.1', 'pct.2', 'p_val_adj')) %>% left_join(.,Pf3D7_genes_plasmoDB %>% dplyr::select(gene_id, Gene), by = 'gene_id'))

names(mast_tde_de_export) <- str_replace_all(str_remove(names(mast_tde_de_export), 'FieldVlab_|lf_StrainsDE_'), c('asx'='Asexual DE', 'fh'='Female (HE) DE', 'fl'='Female (LE) DE', 'mh'='Male (HE) DE'))

##Export the list of field_lab DE dfs with appropriate names to workbook with each stage as a sheet
# createWorksheet(mast_tde_de_export[c(1,3:5)], "../Mali/v3_science_manuscript/tables/fvl_mast_tde_de_export.xlsx") 

## Write out field versus lab and powered separate field and separate lab as RDS so as to include the GO analysis
saveRDS(mast_tde_de_export[c(1:8, 25:28)], "../Mali/v3_science_manuscript/tables/fvl_n_powered_mast_tde_de_export.RDS")
  
##Get weak powered analysis
mast_tde_de_export_wk <- mast_tde_de_export[c(17:19,21:24)]

##Retain only weak SC6 comparisons in msc14 female
mast_tde_de_export_wk["MSC14_f_don"] <- map(mast_tde_de_export_wk["MSC14_f_don"],  ~filter(.x, str_detect(pw_comparison, "SC6")))

names(mast_tde_de_export_wk) <- paste0(2:(length(mast_tde_de_export_wk)+1), ".", str_replace_all(names(mast_tde_de_export_wk), c("f_don" = "female", "m_don" = "male")))

##Export the list of field strain_strain DE dfs with appropriate names to workbook with each stage as a sheet
# createWorksheet(mast_tde_de_export[c(17:24)], "../Mali/v3_science_manuscript/tables/data S13 - Field strains MAST & tradeseq DE.xlsx") 
createWorksheet(mast_tde_de_export_wk, "../Mali/v3_science_manuscript/tables/data S13 - Low power field strains MAST & tradeseq DE.xlsx") 
```

#### MAST & Tradeseq correlation
function to plot correlation between MAST and tradeseq and visualize disparities between the 2 methods for the different pairwise strain comparisons
```{r}
pw_corr_plot <- function(g_tbls = mast_tde_corr_tbls){
  for (i in 1:length(g_tbls)) {
    tryCatch({
      pt<-ggplot(g_tbls[[i]], aes(x=waldStat, y=log10(p_val_adj))) +
              # geom_point() + 
              # facet_grid(.~pw_comparison, scales = "free_x") +
              geom_point(aes(color = pw_comparison)) +
              stat_cor(method = "pearson", label.x = 50, label.y = -6)+ 
              geom_vline( xintercept = 10)+
              geom_hline( yintercept = log10(0.05))+
              theme_classic()+
              theme(axis.text=element_text(size=20), 
                    axis.title=element_text(size=18,face="bold"), 
                    legend.text = element_text(size = 16) ,
                    legend.title = element_text(size = 18, face = "bold"),
                    strip.text.x = element_text(size = 16, face = "bold"))
      return(pt)
    }, error = function(e) {
      message(as.character('ss')) # exclude this line to return NULL quietly
      return(NULL)
    })
  }
}
```

```{r}
pw_corr_plot <- function(g_tbls = mast_tde_corr_tbls){

      ggplot(g_tbls, aes(x=waldStat, y=-log10(p_val_adj))) +
              # geom_point() + 
              # facet_grid(.~pw_comparison, scales = "free_x") +
              geom_point(aes(color = pw_comparison)) +
              scale_color_manual(values = fld_strns_nf54_7g8_comps_col, name = "Comparison") +
              stat_cor(method = "pearson", label.x = 80, label.y = 5)+ 
              geom_vline( xintercept = 10)+
              geom_hline( yintercept = -log10(0.05))+
              labs(x = "tradeSeq waldStat", y = "Adjusted\nMAST P(-log10)") +
              theme_classic()+
              theme(axis.text=element_text(size=20), 
                    axis.title=element_text(size=18,face="bold"), 
                    legend.text = element_text(size = 16) ,
                    legend.title = element_text(size = 18, face = "bold"),
                    strip.text.x = element_text(size = 16, face = "bold"))
}
```


Correlation between MAST and tradeseq 
```{r}
# pw_corr_plot(g_tbls = mast_tde_corr_tbls)
de_corr_plts <- map(mast_tde_corr_tbls, ~{
  pw_corr_plot(.x %>% filter(!str_detect(pw_comparison, 'SC._vs_[L|N|7].*'))%>% filter(!str_detect(pw_comparison, 'NF54_A')))
  } )

de_corr_plts
```



##Comparing between samples
```{r, fig.width= 8, fig.height=3}
# mast_tde_corr_tbls[[8]] %>% inner_join(., mast_tde_corr_tbls[[12]], by = c('gene_id', 'pw_comparison'))  %>% ggplot(., aes(x=p_val_adj.x, y=p_val_adj.y)) +
mast_tde_corr_tbls[[2]] %>% inner_join(., mast_tde_corr_tbls[[4]], by = c('gene_id', 'pw_comparison'))  %>% 
  ggplot(., aes(x=waldStat.x, y=waldStat.y)) +
    # geom_point() + 
    # facet_grid(.~pw_comparison, scales = "free_x") +
    geom_point(aes(color = pw_comparison)) +
    stat_cor(method = "pearson", label.x = 0, label.y = 0)+ 
    # geom_vline( xintercept = 10)+
    # geom_hline( yintercept = log10(0.05))+
    theme_classic()+
    theme(axis.text=element_text(size=20), 
          axis.title=element_text(size=18,face="bold"), 
          legend.text = element_text(size = 16) ,
          legend.title = element_text(size = 18, face = "bold"),
          strip.text.x = element_text(size = 16, face = "bold"))

```


overlap in genes upregulated in second strain versus the first
Asexual
```{r, fig.width= 8, fig.height=3}

# strns_comps =  c('SC1_vs_SC2', 'SC2_vs_SC3', 'SC2_vs_SC6', 'SC1_vs_SC3')
# strns_comps =  c("SC1_vs_SC2", "SC1_vs_SC3", "SC1_vs_SC6", "SC2_vs_SC3", "SC2_vs_SC6", "SC3_vs_SC6")

list(strns_ub_pairwise_mkrs_pt_ml_sst[11]) %>%
  pmap(., ~{
    mast_de_tbl = ..1 %>% filter(p_val_adj < 0.05) %>% arrange(p_val_adj)
    strns_comps = mast_de_tbl %>% pull(pw_comparison) %>% unique()
    print(length(strns_comps))
    
    map(strns_comps, ~ mast_de_tbl %>% 
          filter(p_val_adj < 0.05  & pw_comparison== .x ) %>% 
          pull(gene_id)) %>% 
      set_names(.,strns_comps) %>% 
      ggvenn(., fill_color =c("#0073C2FF", "#EFC000FF", "#868686FF", "#9186dd"), stroke_size = 0.5, set_name_size = 4, text_size = 10, show_percentage = F )
  }
  ) 
```
UpSet plot
```{r}
## Get relevant strain comparisons and arrange appropriately
strns_comps_w_lab =  map(strns_ub_pairwise_mkrs_pt_ml_sst, ~unique(.$pw_comparison)) %>% unlist() %>% unique() 
strns_comps_w_lab =  c(sort(strns_comps_w_lab[!(strns_comps_w_lab %in% c("NF54_vs_7G8","NF54_vs_NF54_A", "7G8_vs_NF54_A","Field_vs_Lab"))]), c("NF54_vs_7G8","NF54_vs_NF54_A", "7G8_vs_NF54_A","Field_vs_Lab"))
## Remove donor strain VS field comparison
strns_comps =  strns_comps_w_lab[!(str_detect(strns_comps_w_lab, ".*SC[0-9]*_vs_[NF54_A$|NF54$|7G8$]") | str_detect(strns_comps_w_lab, ".*vs_NF54_A"))]

list(strns_ub_pairwise_mkrs_pt_ml_sst[10]) %>%
  pmap(., ~{
    mast_de_tbl = ..1
    map(strns_comps, ~ mast_de_tbl %>% 
          filter(p_val_adj < 0.05  & pw_comparison== .x ) %>% 
          pull(gene_id)) %>% 
      set_names(.,strns_comps) %>% fromList(.) %>% UpSetR::upset()
  }
  ) 

# gt_fb_vtx_p1_bi_qc_pf7 %>% dplyr::select(id, matches('G1_._._qc')) %>% filter(if_any(starts_with('G'), ~!is.na(.)))  %>% column_to_rownames('id') %>% split.default(., names(.)) %>% map(., ~.x %>% filter(if_all(everything(), ~!is.na(.))) %>% rownames()) %>% fromList(.) %>% upset()
```

```{r, fig.width= 8, fig.height=3}

# strns_comps =  c('SC2_vs_SC5', 'SC2_vs_SC7', 'SC5_vs_SC7')
# strns_comps =  c("SC2_vs_SC4", "SC2_vs_SC5", "SC2_vs_SC6", "SC2_vs_SC7", "SC4_vs_SC5", "SC4_vs_SC6", "SC4_vs_SC7", "SC5_vs_SC6", "SC5_vs_SC7", "SC6_vs_SC7" )

list(strns_ub_pairwise_mkrs_pt_ml_sst[c(5, 8:10)]) %>%
  pmap(., possibly( ~{
    mast_de_tbl = ..1 %>% filter(p_val_adj < 0.05) %>% arrange(p_val_adj)
    strns_comps = mast_de_tbl %>% pull(pw_comparison) %>% unique()
    print(strns_comps)
    print(length(strns_comps))
    
    map(strns_comps, ~ mast_de_tbl %>% 
          filter(p_val_adj < 0.05  & pw_comparison== .x ) %>% 
          pull(gene_id)) %>% 
      set_names(.,strns_comps) %>% 
      ggvenn(., fill_color =c("#0073C2FF", "#EFC000FF", "#868686FF", "#9186dd"), stroke_size = 0.5, set_name_size = 4, text_size = 12, show_percentage = F )
  }, "less than 2 groups")
  ) 
```

UpSet plot
```{r}
list(strns_ub_pairwise_mkrs_pt_ml_sst[c(9:10,14:15)]) %>%
  pmap(., ~{
    mast_de_tbl = ..1
    map(strns_comps, ~ mast_de_tbl %>% 
          filter(p_val_adj < 0.05  & pw_comparison== .x ) %>% 
          pull(gene_id)) %>% 
      set_names(.,strns_comps) %>% fromList(.) %>% UpSetR::upset()
  }
  ) 

# gt_fb_vtx_p1_bi_qc_pf7 %>% dplyr::select(id, matches('G1_._._qc')) %>% filter(if_any(starts_with('G'), ~!is.na(.)))  %>% column_to_rownames('id') %>% split.default(., names(.)) %>% map(., ~.x %>% filter(if_all(everything(), ~!is.na(.))) %>% rownames()) %>% fromList(.) %>% upset()
```

```{r}

mast_tde_corr_tbls[[6]] %>%  
  filter(p_val_adj < 0.05 & !str_detect(pw_comparison, '.*S.*_vs_Lab')) %>% 
  dplyr::select(pw_comparison, gene_id) %>%
  split(.$pw_comparison) %>%
  map(., . %>% pull(gene_id)) %>% 
  .[c(1:3, 5)] %>%
  ggvenn(., set_name_size =2.5)

```

Tradeseq smoother plots for gams SC5 VS SC7 to compare with ggplot ones
```{r, results='asis', warning=F, message=F}

tde_smooth_plt(tde_obj_lst = msc_g_noLE_ref_ts_sce_ub_de[9:12],
               cond_obj_lst = cond_gns_005_ub_de_sst[9:12],
               pw_comp= 'waldStat_SC1_vs_SC2', 
               n_gns = c(1:5)
)
```

#### MAST & Tradeseq association strength disparities
##### Tradeseq disparities
Genes significatnt in MAST (adjusted P < 0.05) but not in tradeseq (waldStat < 10)

###### Assess smoothened expression of individuals genes 
Make dataframe with genes counts per cell for plotting ggplots of gene expression over pseudotime
```{r}
#pts8<-  c(pts, pts)

gn_tble_gg_smoother <- list(msc_g_noLE_v3_combi_afm_seu_sst, pts_sst) %>%
  pmap(~..1@assays$RNA@data %>% as.matrix() %>% t() %>% as.data.frame() %>% rownames_to_column('cell_bc') %>% 
         left_join(., ..1@meta.data  %>% 
                     dplyr::select(..2, strain_dset, stageHL)%>% 
                     rownames_to_column('cell_bc'), by = 'cell_bc') 
  )

```


Grouping plots from same stage and pairwise strain comparison into separate pages

```{r}

stg_strn_mast_gns_smooth_plts <- function(gn_xprsn_tbls =  gn_tble_gg_smoother, 
                                          gns_strn = map(strns_ub_pairwise_mkrs_pt_ml_sst, 
                                                         . %>%
                                                           filter(p_val_adj < 10e-20 & (avg_log2FC > 1 | avg_log2FC < -1))  %>% 
                                                           arrange(p_val_adj) %>%
                                                           dplyr::select(gene_id, pw_comparison)), 
                                          stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                          comp_g = 'pw_comparison',
                                          top_gns = 10,
                                          ptimes = pts_sst
) {
  list(gn_xprsn_tbls, gns_strn, stage_nms, ptimes) %>%
    pmap(., ~{
      mast_de_tbl = ..1
      gns = ..2
      nms = ..3
      psdotime = ..4
      
      list(strns_comps) %>%
        pmap(possibly(~{
          pw_mast <- mast_de_tbl %>% 
            filter(str_detect(strain_dset, str_replace_all(..1, c('_vs_'='|', 'SC' = '', 'm[0-9]*_' = ''))))
          
          gns_l = gns %>% filter(!!rlang::sym(comp_g) == .x) %>% pull(gene_id) %>% head(., n = top_gns)
          # pdf(paste0('../plots/',nms,..1,'.pdf' ))
          print(gns_l)
          
          plot_list <- map(gns_l, 
                           ~ggplot(pw_mast, aes(x = !!rlang::sym(psdotime) , y = log2((!!rlang::sym(.x))+1), color = strain_dset)) + 
                             geom_point() + 
                             geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs")) +
                             theme(legend.position="none")
          )
          
          print(length(plot_list))
          
          # extract the legend from one of the plots
          lege <- cowplot::get_legend(
            # create some space to the left of the legend
            plot_list[[1]] + 
              guides(color = guide_legend(nrow = 1)) +
              theme(legend.position = "bottom")
          )
          
          
          ttl_lege <- plot_grid(ggdraw() + 
                                  draw_label(paste0(nms, '_', ..1),
                                             fontface = 'bold',
                                             x = 0,
                                             hjust = 0
                                  ),
                                lege, ncol = 2, nrow = 1)
          
          pl1 <- plot_grid(plotlist = plot_list, ncol = 4, nrow = 5) 
          print(class(pl1))
          
          pls <- plot_grid(ttl_lege, pl1, rel_heights = c(0.1, 1), ncol = 1)
          print(pls)
          
          return(list(pl1) %>% set_names(..1))
          
          f          # save_plot(paste0("../plots/", nms, '_', ..1, "top_mast_de_gns_smooth.pdf"), plot = pl1, base_height = 13, base_width = 16)
          
        }
        , 'e')
        ) 
    }
    )
}

```

Function to save plots in pdfs named by stage and pairwise strain comparison in which it is DE. The plots are generated above with the stage_strain_pw_de_comp function
```{r}
stg_strn_mast_gns_smooth_plts_saver = function(plts_cnm = 'top_sig', plt_lst = top_sig_gns_plts) {
  imap(unlist(map(plt_lst, 
                  ~unlist(.x, recursive = F, use.names = T)), 
              recursive = F), 
       possibly(~save_plot(paste0("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/plots/mali2021_all/strn_mast_de_gns_smoothers/", .y, "_",plts_cnm,".pdf"), 
                           plot = .x, 
                           base_height = 13, 
                           base_width = 16), 
                otherwise = 'error'))
}
```


Table of genes significatnt in MAST (adjusted P < 0.05) but not in tradeseq (waldStat < 10)
```{r, fig.width= 8, fig.height=3}

##MASt signal is in ok ie pvalue less than 0.0001 - TDE signal too low i.e waldstat <10

tde_disparities <- list(mast_tde_corr_tbls, strns_ub_pairwise_mkrs_pt_ml_sst) %>%
  pmap(
    ~..1 %>%
      filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>%
      filter(waldStat < 10 & p_val_adj < 0.05) 
  )

```



```{r}
# a=msc_gam_combi_noLE_de[, colnames(msc_g_noLE_v3_combi_afm_seu_sst[[1]])]@assays$RNA@counts %>% as.matrix() %>% t() %>% as.data.frame() %>% mutate(totl_gn = rowSums(.)) %>% mutate(across(starts_with('PF'), ~log1p((./totl_gn)*10000))) %>% dplyr::select(-c(totl_gn)) %>% rownames_to_column('cell_bc') %>% left_join(., msc_g_noLE_v3_combi_afm_seu_sst[[1]]@meta.data  %>% dplyr::select(asexual_pt, strain_dset, stageHL)%>% rownames_to_column('cell_bc'), by = 'cell_bc') %>% ggplot(., aes(x = asexual_pt, y = `PF3D7-1234900`, color = strain_dset, shape = stageHL)) + geom_point() + geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs"))
# print(a)
```


```{r}
# strn_0_1_3_mkrs_pt_bsu <- readRDS('data/processed/DB52e_star/strn_0_1_3_mkrs_pt_bsu.RDS')

```

Visualize disparities
```{r}
tde_disparities_gns_plts <- stg_strn_mast_gns_smooth_plts(gn_xprsn_tbls = gn_tble_gg_smoother, 
                                                          gns_strn = map(tde_disparities, 
                                                                         . %>%
                                                                           # filter(p_val_adj < 0.9) %>%
                                                                           arrange(desc(waldStat)) %>% 
                                                                           dplyr::select(gene_id, pw_comparison)), 
                                                          stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                                          ptimes = pts_sst,
                                                          comp_g = 'pw_comparison',
                                                          top_gns = 18)

```

Save tradeseq disparities
```{r}
stg_strn_mast_gns_smooth_plts_saver(plts_cnm = 'tde_disparities', plt_lst = tde_disparities_gns_plts)
```

##### MAST disparities
Genes significatnt in  tradeseq (waldStat > 10) but not in  MAST (adjusted P > 0.05)
```{r, fig.width= 8, fig.height=3}
##TDE signal is in check - MAST signal too low

mast_disparities <- list(mast_tde_corr_tbls, strns_ub_pairwise_mkrs_pt_ml_sst) %>%
  pmap(
    ~..1 %>%
      filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>%
      filter(waldStat > 10 & p_val_adj > 0.05) 
  )



```



Correlation between MAST and tradeseq 
```{r}
# pw_corr_plot(g_tbls = mast_tde_corr_tbls)
map(mast_disparities, ~{
  pw_corr_plot(.x %>% filter(!str_detect(pw_comparison, 'SC._vs_[L|N|7].*'))%>% filter(!str_detect(pw_comparison, 'NF54_A')))
  } )

```
Individual gene tradeseq normalised counts visualization smoothers

```{r, results='asis', warning=F, message=F}
tde_smooth_plt(tde_obj_lst = msc_g_noLE_ref_ts_sce_ub_de[9:10],
               cond_obj_lst = cond_gns_005_ub_de_sst[9:10],
               pw_comp= 'waldStat_SC1_vs_SC2', 
               n_gns = c(1:5)
)

```


individual genes Seurat normalized counts with MAST disparities plots
```{r}
mast_disparities_gns_plts <- stg_strn_mast_gns_smooth_plts(gn_xprsn_tbls = gn_tble_gg_smoother, 
                                                           gns_strn = map(mast_disparities, 
                                                                          . %>%
                                                                            # filter(p_val_adj < 0.9) %>%
                                                                            arrange(desc(waldStat)) %>% 
                                                                            dplyr::select(gene_id, pw_comparison)), 
                                                           stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                                           ptimes = pts_sst,
                                                           comp_g = 'pw_comparison',
                                                           top_gns = 10)



```

##### Borderline genes
Save bordeline significant genes
```{r}
stg_strn_mast_gns_smooth_plts_saver(plts_cnm = 'mast_disparities', plt_lst = mast_disparities_gns_plts)
```


Top gene tradeseq and MAST
```{r}
top_tde_mast_sig_gns_plts <-stg_strn_mast_gns_smooth_plts(gn_xprsn_tbls = gn_tble_gg_smoother,  
                                                          gns_strn = map(mast_tde_corr_tbls, 
                                                                         . %>%
                                                                           filter(p_val_adj < 0.05 & waldStat > 10) %>%
                                                                           filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>%
                                                                           arrange(desc(waldStat)) %>% 
                                                                           dplyr::select(gene_id, pw_comparison)), 
                                                          stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                                          ptimes = pts_sst,
                                                          comp_g = 'pw_comparison',
                                                          top_gns = 10)



```


Save bordeline significant genes
```{r}
stg_strn_mast_gns_smooth_plts_saver(plts_cnm = 'top_tde_mast_sig', plt_lst = top_tde_mast_sig_gns_plts)
```

#### MAST sig genes with tradeseq smoothened heatmaps 

function for getting genes to plot
```{r}
##Function for making table of genes and accompanying annotations for heatmap - table kept out of the plotting functions since we will be modifying it alot and it might be helpful to look at the results before making the heatmap
# strns_lab_comps <- c('SC2_vs_SC5', 'SC2_vs_SC7', 'SC5_vs_SC7', 'SC2_vs_Lab', 'SC5_vs_Lab', 'SC7_vs_Lab')

## Get all the strains present in the comparisons
# strns_lab_comps <- c('SC1_vs_SC2', 'SC1_vs_SC3', 'SC1_vs_SC6', 'SC2_vs_SC3', 'SC2_vs_SC5',  'SC2_vs_SC6', 'SC2_vs_SC7', 'SC3_vs_SC6',  'SC5_vs_SC7','SC1_vs_Lab', 'SC2_vs_Lab', 'SC3_vs_Lab',  'SC5_vs_Lab', 'SC6_vs_Lab', 'SC7_vs_Lab')
# strns_lab_comps <- c("SC1_vs_SC2", "SC1_vs_SC3", "SC1_vs_SC6", "SC2_vs_SC3", "SC2_vs_SC4", "SC2_vs_SC5", "SC2_vs_SC6", "SC2_vs_SC7", "SC3_vs_SC6",  "SC4_vs_SC5", "SC4_vs_SC6",  "SC4_vs_SC7", "SC5_vs_SC6", "SC5_vs_SC7", "SC6_vs_SC7","SC1_vs_Lab", "SC2_vs_Lab", "SC3_vs_Lab", "SC4_vs_Lab", "SC5_vs_Lab", "SC6_vs_Lab", "SC7_vs_Lab")

lfstrains_comps <- strns_comps_w_lab

strn_comps_gns_anno_2_plt = function(
    pval = 10e-10,
    wstat= 10,
    stg_pw_gns = mast_tde_corr_tbls[[1]],
    avglfc_top = 0.5,
    avglfc_botom = -0.5,
    comp_lvls='Field_vs_Lab',
    sig_strns_annotn_2_keep = 'Field_vs_Lab',
    strns_annotn_2_omit = 'SC._vs_Lab',
    nf54_batch_omit = "EMPTY",
    topgn = 1000000) {
  
  gns2plt <- stg_pw_gns %>% 
    filter(p_val_adj < pval &  (avg_log2FC > avglfc_top | avg_log2FC < avglfc_botom ) & waldStat > wstat) %>% 
    mutate(lfc_rank = dense_rank(dplyr::desc(abs(avg_log2FC)))) %>%
    filter(lfc_rank <= topgn) %>% ### !!! NOTE - This will work unpredictably when omitting some comparisons (pw_comparisons) downstream
    mutate(dir = case_when(avg_log2FC > avglfc_top ~ 'Upregulated', 
                           avg_log2FC < avglfc_botom ~ 'Downregulated')) %>%
    mutate(across(dir, ~factor(., levels=c('Upregulated', 'Downregulated')))) %>%
    dplyr::select(pw_comparison, gene_id, dir) %>% 
    mutate(across(pw_comparison, ~droplevels(factor(., levels = comp_lvls)))) %>%
    arrange(pw_comparison, dir) %>%
    pivot_wider(names_from = 'pw_comparison', values_from = 'dir') %>%
    filter(if_any(matches(sig_strns_annotn_2_keep), ~!is.na(.))) %>%
    filter(if_any(-c(matches(strns_annotn_2_omit), matches(nf54_batch_omit), gene_id), ~!is.na(.))) %>%
    {
      if(T %in% (str_detect(names(.),strns_annotn_2_omit))) 
        mutate(.,FieldvsLab_c = case_when(
          if_any(matches(strns_annotn_2_omit), ~ . == 'Downregulated') & if_any(matches(strns_annotn_2_omit), ~ . == 'Upregulated') ~ 'both',
          if_any(matches(strns_annotn_2_omit), ~ . == 'Downregulated') ~ 'Downregulated',
          if_any(matches(strns_annotn_2_omit), ~ . == 'Upregulated') ~ 'Upregulated')) 
      else 
        .
    } %>% 
    dplyr::select(-c(matches(strns_annotn_2_omit),matches(nf54_batch_omit))) %>%
    arrange(across(-one_of("gene_id")))%>%
    column_to_rownames('gene_id')
  
  return(gns2plt)
  
}

```


```{r}

hmap_gn_tbls <- list(mast_tde_corr_tbls, 
                     c(as.list(rep('Field_vs_Lab', 8)),rep(list(lfstrains_comps), 20)),
                     c(as.list(rep('Field_vs_Lab', 8)),rep('^[7|N|m[0-9]*_S].*_vs_[7|N|m[0-9]*_S]*', 20))) %>%
  pmap(#possibly(
    ~{
      strn_comps_gns_anno_2_plt(pval = 0.05,
                                wstat= 10,
                                stg_pw_gns = ..1,
                                avglfc_top = 0,
                                avglfc_botom = -0,
                                comp_lvls=..2,
                                sig_strns_annotn_2_keep = ..3,
                                strns_annotn_2_omit = 'm[0-9]*_SC._vs_[L|N|7].*')
      
    }
    #, 'e')
  )


```

heatmap function
```{r}
# sVs_order <- paste0(c('SC1', 'SC2', 'SC3', 'SC4', 'SC5', 'SC6', 'SC7'), ), c('7G8', 'NF54','NF54_A',  'Lab'))
sVs_order <- strain_dset_lvls_don
# sVs_order_minus_nf54O <- c('SC1', 'SC2', 'SC3', 'SC4', 'SC5', 'SC6', 'SC7','7G8', 'NF54','Lab')
sVs_order_minus_nf54O <- strain_dset_lvls_don[!strain_dset_lvls_don %in%"NF54_A"]

##Function for plotting heatmap
hmap_pw_strns_lab_tde = function(ts_dset= msc_g_noLE_ref_ts_sce_ub_de[[2]],
                                 gns_tbl = hmap_gn_tbls[[2]],
                                 cap_no = 7,
                                 split_col = 'Condition',
                                 ptime = pts[[2]],
                                 col_map = strain_cols_n,
                                 strn_row_anno_order = sVs_order,
                                 strn_conds = 'Condition',
                                 pt_labl = 'Pseudotime',
                                 hmap_mode='seurat_smooth',
                                 seu_dset = msc_g_noLE_v3_combi_afm_seu_sst[[2]],
                                 seu_comps = 'dset',
                                 chunk_size = 0.02,
                                 rnames = F,
                                 rsize = 6,
                                 lgsize = 10, ## legend titles font size (pseudotime & Scaled expression)
                                 lblsize = 10,
                                 lg_width = 4,
                                 csize = 12, ## Column/conditions header font size (strains lab etc)
                                 anno_fsize = 15,
                                 anno_trsize = 15,
                                 strn2dsample = 'Lab',
                                 shw_anno_head = T,
                                 shw_left_bar_anno=T,
                                 botom_pad = 26,
                                 strain2omit_top_bar = 'Placeholder_when_we_dont_want_to_omit_any_group') {
  ## If else option to specify the commands based on the selected  heatmap mode - tde_smooth, seurat, seurat_smooth
  ## tde_smooth - plotting the smoothened tradeseq gene expression from fitting GAMs (generalised additive models)
  if (hmap_mode == 'tde_smooth') {
    
    ##Cells groupings order, pseudotime, points calculation
    cells_anno = predictSmooth(ts_dset, gene = rownames(gns_tbl), nPoints = 20, tidy = T)  %>% 
      mutate(point = paste0("point", rep(rep(1:20,length(unique(ts_dset@colData$tradeSeq$conditions))),length(rownames(gns_tbl)))), 
             lin_id = paste0('lineage', lineage,'_condition', condition, '_',point)) %>%
      rownames_to_column('No')  %>% 
      dplyr::select(lin_id,time,condition)%>% 
      distinct()%>% 
      column_to_rownames('lin_id') %>%
      # filter(!str_detect(condition, strain2omit_top_bar)) %>%
      filter(!(condition %in% strain2omit_top_bar)) %>%
      rename('Pseudotime' = time, 'Condition' = condition) %>%
      mutate(across(Condition, ~droplevels(factor(., levels = strn_row_anno_order))))
    
    ##Get cells order
    cells_order <- rownames(cells_anno)
    
    ##Order and scale gene expression matrix appropriately
    xprsn_mtx <- predictSmooth(ts_dset, gene = rownames(gns_tbl), nPoints = 20, tidy = FALSE) %>%
      .[,cells_order] %>%
      t() %>%
      scale() %>% 
      .[,rownames(gns_tbl)] %>%
      t() %>% 
      as.matrix()
    
    ## seurat - plotting the raw seurat gene expression
  } else if (hmap_mode == 'seurat') {
    
    cells_anno = seu_dset@meta.data %>% 
      rownames_to_column('cell_bc') %>%
      # filter(!str_detect(!!rlang::sym(seu_comps), strain2omit_top_bar))%>%
      filter(!(!!rlang::sym(seu_comps) %in% strain2omit_top_bar)) %>%
      mutate(across(!!rlang::sym(seu_comps), ~droplevels(factor(., levels = strn_row_anno_order)))) %>% 
      arrange(!!rlang::sym(seu_comps), !!rlang::sym(ptime)) %>% 
      dplyr::add_count(!!rlang::sym(seu_comps), name ='grp_t') %>%  
      group_by(!!rlang::sym(seu_comps)) %>%
      mutate(samp = sample(x=n())) %>%
      ungroup() %>%
      filter(!(!!rlang::sym(seu_comps) == strn2dsample & samp > min(grp_t))) %>%
      column_to_rownames('cell_bc') %>%
      dplyr::select(!!rlang::sym(seu_comps), !!rlang::sym(ptime)) %>%
      rename('Pseudotime' = !!rlang::sym(ptime), 'Condition' = !!rlang::sym(seu_comps))
    print('lkl')
    print(cells_anno %>% dplyr::count(Condition))
    
    print('lkl')
    print(head(rownames(cells_anno)))
    
    print('lkl')
    print(head(cells_anno ))
    
    cells_order <- rownames(cells_anno)
    
    xprsn_mtx <- seu_dset@assays$RNA@scale.data %>%
      .[,cells_order] %>%
      t()  %>% 
      .[,rownames(gns_tbl)]%>% 
      t()  %>% 
      as.matrix()
    print('lkl')
    print(dim(cells_anno))
    print('lkl')
    print(dim(xprsn_mtx))
    
  }
  
  ## seurat_smooth - plotting the smoothened seurat gene expression where several cells are grouped together and the mean is calculated. The mean of this group is then used to represent all the cells therein. The group (strain/lab cluster) with the lowest is used to decided the number of chunks to make. These results in the largest cluster being equal to the smallest. - NEED TO CONFIRM IF THIS IS ACCURATE
  else if (hmap_mode == 'seurat_smooth') {
    ##Heatmaps with smoothing grouping cells within 0.1 pseudotime values
    
    chunked_cells <- seu_dset@meta.data %>% 
      mutate(ptime_cat = cut(!!rlang::sym(ptime), 
                             breaks = c(seq(min(!!rlang::sym(ptime)), max(!!rlang::sym(ptime)), chunk_size), max(!!rlang::sym(ptime))), 
                             right = F, 
                             include.lowest = T)) %>% 
      mutate(grp_id = paste0(!!rlang::sym(seu_comps), str_remove_all(ptime_cat, '\\[|\\)'))) %>%
      # filter(!str_detect(!!rlang::sym(seu_comps), strain2omit_top_bar))%>% 
      filter(!(!!rlang::sym(seu_comps) %in% strain2omit_top_bar)) %>%
      mutate(across(grp_id, ~str_replace_all(., ' |,', '_')))%>% 
      rownames_to_column('cell_bc')
    
    
    ##Prepare the Stage and ptime annotation for the columns ie cell groupings
    
    cells_anno = chunked_cells %>% 
      mutate(across(!!rlang::sym(seu_comps), ~droplevels(factor(., levels = strn_row_anno_order)))) %>% 
      arrange(!!rlang::sym(seu_comps), !!rlang::sym(ptime))  %>% 
      group_by(grp_id) %>%
      mutate(pt_mean = mean(!!rlang::sym(ptime))) %>% 
      distinct(grp_id, pt_mean, .keep_all = T) %>% 
      column_to_rownames('grp_id') %>%
      dplyr::select(!!rlang::sym(seu_comps), pt_mean) %>%
      rename('Pseudotime' = pt_mean, 'Condition' = !!rlang::sym(seu_comps))
    
    cells_order <- rownames(cells_anno)
    
    ##Get matrix for counts and convert to dataframe
    counts_t <-  seu_dset@assays$RNA@scale.data %>% as.matrix() %>% t() %>% as.data.frame() %>% rownames_to_column('cell_bc')
    
    xprsn_mtx <- chunked_cells %>% 
      dplyr::select(cell_bc, grp_id) %>% 
      left_join(., counts_t, by='cell_bc') %>% 
      group_by(grp_id) %>% 
      summarise(across(starts_with('PF3D7'), mean)) %>%
      column_to_rownames('grp_id') %>%
      .[,rownames(gns_tbl)]%>%
      t()  %>%
      .[,cells_order] %>%
      as.matrix()
    
    # Troubleshooting
    # print(dim(cells_anno))
    # print(dim(xprsn_mtx))
    
    # print(head(rownames(cells_anno)))
    # print(head(colnames(xprsn_mtx)))
    
  }
  
  
  
  ##Cells annotation
  ##Pseudotime bar color function
  ptime_color_func = colorRamp2(breaks = c(min(cells_anno[cells_order, 'Pseudotime']), 
                                           max(cells_anno[cells_order, 'Pseudotime'])), 
                                colors=c("white",  "#004225"))
  
  ##Combine cells annotation (strains/conditions and pseudotime) colors
  column_ha = HeatmapAnnotation(df = cells_anno[cells_order, , drop = FALSE] %>% 
                                  dplyr::select(strn_conds, pt_labl),
                                col = c(list(cells_anno[cells_order, , drop = FALSE] %>% 
                                               pull(!!rlang::sym(strn_conds)) %>% 
                                               unique()%>% 
                                               as.character() %>% 
                                               col_map[.]) %>% set_names(strn_conds),
                                        list(ptime_color_func) %>% set_names(pt_labl)
                                ),
                                show_legend= list(F) %>% set_names(strn_conds) %>% unlist(),
                                # annotation_legend_param = list(list(direction = "horizontal")) %>% set_names(pt_labl),
                                annotation_legend_param = list(list(direction = "horizontal", 
                                                                    title_gp=gpar(fontsize = lgsize), 
                                                                    labels_gp = gpar(fontsize = lblsize),
                                                                    legend_width = unit(4.5, "cm"),
                                                                    labels_rot = 90)) %>%
                                  set_names(pt_labl),
                                annotation_name_gp = gpar(fontsize = anno_trsize, fontface = "bold"),
                                show_annotation_name = shw_anno_head#,
                                # list(anno_block(labels = strn_row_anno_order %>% str_replace(., "train ", "") ) %>% set_names(strn_conds))
                                
  )
  
  ##Genes annotation colors
  ##Bar coloring
  row_ha = rowAnnotation(df = gns_tbl[rownames(xprsn_mtx),colnames(gns_tbl), drop = FALSE],
                         col = rep(list(c('Upregulated' = '#e7a06b', 'Downregulated' = '#9cb4cf' , 'both' = '#7068af')),  # #9bbbe0 	#e0c09b #7068af '#BF4184' '#0079C1' '#9E384D' '#E976A8' '#FF8040' '#A65100'
                                   each = length(colnames(gns_tbl)))%>% 
                           set_names( colnames(gns_tbl)),
                         # annotation_legend_param = rep(list(list(
                         #   title = "Direction",
                         #   labels = c('Upregulated', 'Downregulated', 'both'),
                         #   at = c('Upregulated', 'Downregulated', 'both')
                         # )), times=length(colnames(gns_tbl[rownames(xprsn_mtx),colnames(gns_tbl), drop = FALSE])))%>% set_names( colnames(gns_tbl[rownames(xprsn_mtx),colnames(gns_tbl), drop = FALSE]))
                         show_legend= rep(F, times = length(colnames(gns_tbl))) %>% set_names(colnames(gns_tbl)[1:length(colnames(gns_tbl))]),
                         # show_legend= rep(F, times = length(colnames(gns_tbl))-1) %>% set_names(colnames(gns_tbl)[2:length(colnames(gns_tbl))]),
                         annotation_name_gp = gpar(fontsize = anno_fsize, fontface = "bold"),
                         show_annotation_name = shw_left_bar_anno
  )
  
  xprsn_mtx %>%
    Heatmap(.,
            cluster_columns = F,
            show_column_names = F,
            cluster_rows = F,
            row_labels = Pf3D7_genes_plasmoDB %>%
              left_join(data.frame(gene_id=rownames(xprsn_mtx)), ., by=c('gene_id')) %>%  pull(Gene),
            show_row_names = {{rnames}},
            left_annotation = row_ha,
            top_annotation = column_ha,
            column_split = cells_anno[cells_order,split_col],
            use_raster = F,
            # heatmap_legend_param = list(direction = "horizontal", legend_gp = gpar(fontsize = lgsize)),
            heatmap_legend_param = list(direction = "horizontal", 
                                        title_gp = gpar(fontsize = lgsize),
                                        labels_gp = gpar(fontsize = lblsize),
                                        labels_rot = 90,
                                        # grid_width = unit(6, "mm"),
                                        legend_width = unit(lg_width, "cm")),
            row_names_gp = grid::gpar(fontsize = rsize),
            # column_labels = strn_row_anno_order %>% str_replace(., "train", "") %>% set_names(strn_row_anno_order),
            column_title_gp = gpar(fontsize = csize, fontface = "bold"),
            # column_title_rot = 90,
            name = "Scaled expression"
    ) %>%
    ##Merging legends into 1 column. Could be set to F
    draw(.,newpage = F, merge_legends = T, heatmap_legend_side = "bottom", annotation_legend_side = "bottom", padding = unit(c(botom_pad, 2, 2, 2), "mm"))
}

```


All sig genes MAST padj <0.05 and tradeseq waldstat 10
Tradeseq smooth
NB!!! - RRERUNNING TRADESEQ TO CHANGE CONDITION NAMES SHOULD SORT THE ERROR IN TRADESEQ PLOTS
```{r}


fVl_order <- c('Field', 'Lab')

hmap_gn_tbls <- list(mast_tde_corr_tbls, 
                     c(as.list(rep('Field_vs_Lab', 8)),rep(list(lfstrains_comps), 20)),
                     c(as.list(rep('Field_vs_Lab', 8)),rep('^[7|N|S].*_vs_[7|N|S]*', 20))
                     # c(as.list(rep('Field_vs_Lab', 8)),rep('^[7|N|m[0-9]*_S].*_vs_[7|N|m[0-9]*_S]*',8))
                     ) %>%
  pmap(#possibly(
    ~{
    strn_comps_gns_anno_2_plt(pval = 0.05, 
                              wstat= 10,
                              stg_pw_gns = ..1,
                              avglfc_top = 0,
                              avglfc_botom = -0,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              # strns_annotn_2_omit = 'm[0-9]*_SC._vs_[L|N|7].*',
                              # strns_annotn_2_omit = 'SC._vs_[L|N|7].*', ##Retain FieldvsLab_c which adds left bar annotation for field combined versus lab combined
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*|FieldvsLab_c', ##Remove FieldvsLab_c which adds left bar annotation for field combined versus lab combined
                              nf54_batch_omit='^.*_vs_NF54_A')
    
  }
  # , 'e')
  )

list(msc_g_noLE_ref_ts_sce_ub_de, 
     hmap_gn_tbls, 
     pts_sst, 
     c(rep(list(fVl_order), 8),rep(list(sVs_order), 20)),
     c(rep(list(fvl_cols), 8),rep(list(strain_dset_lvls_don_col), 20)),
     rep(c(F), 28), 
     rep(c(8), 28),
     rep(13, 28)) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='tde_smooth',
                            rnames = ..6,
                            rsize = ..7,
                            anno_fsize = ..8)}
     , 'e')
  )
```

Raw seurat log normalised and scaled
In the field versus lab, the field is separated into distinct strains for heatmap visualization but the analysis is for bulk field versus lab - need to confirm
```{r, fig.height=5, fig.width=7}

list(msc_g_noLE_ref_ts_sce_ub_de[9:28], 
     hmap_gn_tbls[9:28], 
     c(pts_sst[9:28]), 
     c(rep(list(sVs_order), 20)),
     c(rep(list(strain_dset_lvls_don_col), 20)),
     c(msc_g_noLE_v3_combi_afm_seu_sst[9:28]),
     rep('strain_dset', 20),
     c(rep(c(0.02,0.1), 1), rep(0.02, 18)),
     c(rep(c(F), 18), rep(c(T), 2)), 
     c(rep(c(6,10), 1), rep(c(8), 18)), 
     rep(c(15), 20),
     rep(13, 20)
)%>%
  pmap(
    possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            csize  = ..11,
                            anno_fsize = ..12,
                            strn2dsample = 'Lab')}
    , 'e')
  )

```



In the field versus lab, the field strains are grouped into field for heatmap visualization recapitulating the analysis which is for bulk field versus lab - need to confirm
- doesnt include field vs lab on left bars in strain comps
```{r, fig.width=8, fig.height=4.5}

all_labvfld_hms<- list(msc_g_noLE_ref_ts_sce_ub_de, 
     hmap_gn_tbls, 
     pts_sst, 
     c(rep(list(fVl_order), 8),rep(list(sVs_order), 20)),
     c(rep(list(fvl_cols), 8),rep(list(strain_dset_lvls_don_col), 20)),
     c(msc_g_noLE_v3_combi_afm_seu_sst),
     c(rep('dset', 8), rep('lf_strain', 20)),
     rep(c(0.02), 28),
     c(rep(F, 8), rep(F, 20)), 
     c(6,10,8,8,8,10,10,10,9,10,8,10,10,10,9,10,10,10,9,10,10,10,9,10,10,10,9,10),
     c(rep(16, 8),rep(15, 20)),
     c(rep(F, 8),rep(F, 20)),
     c(rep(F, 8),rep(T, 20)),
     c(rep(14.5, 8),rep(14, 20)),
     c(rep(14.5, 8),rep(13, 20)),
     c(rep(16, 8),rep(15, 20)),
     c(rep(0.4, 8),rep(18, 20)),
     c(rep('', 8), rep('NF54_A', 20))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            shw_anno_head = ..12,
                            shw_left_bar_anno = ..13,
                            lgsize = ..14,
                            lblsize = ..15,
                            strn2dsample = 'Lab',
                            csize = ..16,
                            botom_pad = ..17,
                            strain2omit_top_bar = ..18)}
    , 'e')
  )

all_labvfld_hms
```

Saame as above but only for the donor strains and with gene names
```{r, fig.width=8, fig.height=4.5}
don_objs <- names(msc_g_noLE_ref_ts_sce_ub_de)[str_detect(names(msc_g_noLE_ref_ts_sce_ub_de), "_don$")]

all_labvfld_hms_w_gns_labl<- list(msc_g_noLE_ref_ts_sce_ub_de[don_objs], 
     hmap_gn_tbls[don_objs], 
     pts_sst[1:8], 
     c(rep(list(sVs_order), 8)),
     c(rep(list(strain_dset_lvls_don_col), 8)),
     c(msc_g_noLE_v3_combi_afm_seu_sst[str_remove(don_objs, "lf_StrainsDE_")]),
     rep('lf_strain', 8),
     rep(c(0.02), 8),
     rep(T, 8), 
     rep(11, 8),
     c(rep(14, 8)),
     c(rep(F, 8)),
     c(rep(T, 8)),
     c(rep(11, 8)),
     c(rep(11, 8)),
     c(rep(2.8, 8)),
     c(rep(15, 8)),
     c(rep(18, 8)),
     c(rep('', 8))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            shw_anno_head = ..12,
                            shw_left_bar_anno = ..13,
                            lgsize = ..14,
                            lblsize = ..15,
                            lg_width = ..16,
                            strn2dsample = 'Lab',
                            csize = ..17,
                            botom_pad = ..18,
                            strain2omit_top_bar = ..19)}
    , 'e')
  )

all_labvfld_hms_w_gns_labl
```

```{r}
## Make plots similar to above but without nf54VS7g8
hmap_gn_tbls_no_lab <- list(mast_tde_corr_tbls[9:28], 
                     c(rep(list(lfstrains_comps), 20)),
                     c(rep('^[7|N|S].*_vs_[7|N|S]*', 20))
                     # c(as.list(rep('Field_vs_Lab', 8)),rep('^[7|N|m[0-9]*_S].*_vs_[7|N|m[0-9]*_S]*',8))
                     ) %>%
  pmap(#possibly(
    ~{
    strn_comps_gns_anno_2_plt(pval = 0.05, 
                              wstat= 10,
                              stg_pw_gns = ..1,
                              avglfc_top = 0,
                              avglfc_botom = -0,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              # strns_annotn_2_omit = 'm[0-9]*_SC._vs_[L|N|7].*',
                              # strns_annotn_2_omit = 'SC._vs_[L|N|7].*', ##Retain FieldvsLab_c which adds left bar annotation for field combined versus lab combined
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*|FieldvsLab_c|NF54_vs_7G8', ##Remove FieldvsLab_c which adds left bar annotation for field combined versus lab combined
                              nf54_batch_omit='^.*_vs_NF54_A')
    
  }
  # , 'e')
  )

all_labvfld_hms_no_lab<- list(msc_g_noLE_ref_ts_sce_ub_de[9:28], 
     hmap_gn_tbls_no_lab, 
     pts_sst[9:28], 
     c(rep(list(sVs_order), 20)),
     c(rep(list(strain_dset_lvls_don_col), 20)),
     c(msc_g_noLE_v3_combi_afm_seu_sst[9:28]),
     c(rep('lf_strain', 20)),
     rep(c(0.02), 20),
     c(rep(F, 20)), 
     c(9,10,8,10,10,10,9,10,10,10,9,10,10,10,9,10,10,10,9,10),
     c(rep(15, 20)),
     c(rep(F, 20)),
     c(rep(T, 20)),
     c(rep(14, 20)),
     c(rep(13, 20)),
     c(rep(15, 20)),
     c(rep(17.5, 20)),
     rep(list(c('NF54_A', 'NF54', '7G8')), 20)) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            shw_anno_head = ..12,
                            shw_left_bar_anno = ..13,
                            lgsize = ..14,
                            lblsize = ..15,
                            strn2dsample = 'Lab',
                            csize = ..16,
                            botom_pad = ..17,
                            strain2omit_top_bar = ..18)}
    , 'e')
  )

all_labvfld_hms_no_lab

```

### Genetic distance (IBD) and number of DE genes correlation - START
```{r, eval=T}
# wdir<- '/Users/jr35/Google Drive/My Drive/PHD_documents/sware/hmmIBD-master'
ss_afm_pf7M_cln_gt_hmmibd_f <- map("msc14.strain_qcd", ~read_delim(paste0('data/processed/ibd/', .x, '.hmm_fract.txt'))) 


```

```{r, eval=T}
## Remove SC8 from the msc14 results
# ss_afm_pf7M_cln_gt_hmmibd_f_m14 <- ss_afm_pf7M_cln_gt_hmmibd_f[[8]] %>% filter(sample1 != "SC8" & sample2 != "SC8" )
ss_afm_pf7M_cln_gt_hmmibd_f_m14 <- ss_afm_pf7M_cln_gt_hmmibd_f[[1]]

```

```{r, fig.width= 12, fig.height=4, eval=T}
##read in IBD estimates and create id for the strain comparisons
ss_afm_pf7M_cln_gt_hmmibd_f_m14 <- ss_afm_pf7M_cln_gt_hmmibd_f_m14 %>% mutate(pw_comp = paste(sample1, sample2, sep = '_vs_'))

##Get the numbers of parasites per stage between strains
strn_no <-msc_gcombi_noLE_sub_sce[c("MSC14_f", "MSC14_f_don3")] %>% map(., ~.x@colData[, c('strain')] %>% table() %>% data.frame()) %>% bind_rows(., .id = 'stage') %>% dplyr::rename('strn'='.', 'count'='Freq') %>% 
  mutate(across(strn, ~str_remove(., "M14_")),
         id = paste0(stage, '_', strn)) %>% 
  data.frame()


##get the number of genes that are DE for each pair of strains for each stage and merge in IBD results and plot
pw_comp_de_gns <- hmap_gn_tbls[c("lf_StrainsDE_MSC14_f","lf_StrainsDE_MSC14_f_don3")] %>% 
  map(., ~rownames_to_column(., 'gn')) %>% 
  bind_rows(., .id = 'stage') %>% 
  pivot_longer(cols = starts_with('SC'), names_to = 'pw_comp', values_to = 'de_dir', values_drop_na = T) %>% 
  dplyr::count(stage,  pw_comp) %>% 
  separate_wider_delim(., pw_comp, names=(c('strn1', 'strn2')), delim = '_vs_', cols_remove = F) %>% 
  mutate(across(stage, ~str_remove(., 'lf_StrainsDE_')), id1 = paste0(stage, '_', strn1), id2 = paste0(stage, '_', strn2)) %>% 
  ##Add the number of genes per minimum cell - to account for every cell number
  left_join(.,strn_no[, c('count', 'id')], by=c('id1'= 'id')) %>% 
  left_join(.,strn_no[, c('count', 'id')], by=c('id2'= 'id'), suffix = c('_1', '_2')) %>% 
  mutate(min_cnt = pmin(count_1, count_2), de_gns_pcell_min = n/min_cnt) %>% 
  rowwise() %>%
  mutate(sum_cnt = sum(count_1, count_2), de_gns_pcell_sum = n/sum_cnt)

##Plot of DE genes correlation with IBD accounting for parasites numbers
pw_comp_de_gns %>% left_join(., ss_afm_pf7M_cln_gt_hmmibd_f_m14, by='pw_comp')  %>% 
  mutate(across(stage, ~str_replace_all(., c('fh'='Female (HE)', 'fl' = 'Female (LE)'))),
         pw_comp_no = paste0(sample1,'(',count_1,')',' vs ',sample2,'(',count_2,')',' [', n,']'))%>%
  ggplot(., aes(x=fract_sites_IBD, y=de_gns_pcell_min )) + 
  geom_line()+ 
  geom_point()+ 
  geom_text_repel(aes(label = pw_comp_no, color=pw_comp), size = 6) +
  facet_wrap(.~stage,  scales = 'free_y') +
  theme_bw() +
  stat_cor(method = "pearson", label.x = 0, label.y = .25, size = 6)+
  labs(y='DE genes per least\nabundant cell in comparison', x='IBD genome proportion', color = 'Pairwise\ncomparison')+
  theme(axis.title = element_text(size =16),
        axis.text = element_text(size =15),
        strip.text =element_text(size=16,face="bold"),
        legend.title = element_text(size =13),
        legend.text = element_text(size =12),
        legend.position = 'None')


```

### Genetic distance (IBD) and number of DE genes correlation - END

Raw seurat log normalised and scaled
field vs lab MAST results visualized with field split by strains 
```{r}
## MAST Pvalue thresholds all field versus lab pval < 10e-10 and each strain versus field pval<0.05 
mast_tde_corr_tbls_fVl_fg <-list(mast_tde_corr_tbls[1:8], mast_tde_corr_tbls[9:16]) %>% pmap(~left_join(..1 %>% filter(p_val_adj < 0.05), ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison), by = 'gene_id') %>% pivot_longer(cols = c('pw_comparison.x', 'pw_comparison.y'), names_to = 'pw_class', values_to = 'pw_comparison' ) %>% distinct())

hmap_gn_sVl_tbls <- list(mast_tde_corr_tbls_fVl_fg) %>%
  pmap(~{
    ..1 %>%
      filter(p_val_adj < 0.05 &  (avg_log2FC > 0 | avg_log2FC < -0 ) & waldStat > 10) %>% 
      mutate(dir = case_when(avg_log2FC > 0 ~ 'Upregulated', 
                             avg_log2FC < -0 ~ 'Downregulated')) %>%
      mutate(across(dir, ~factor(., levels=c('Upregulated', 'Downregulated')))) %>%
      dplyr::select(pw_comparison, gene_id, dir) %>% 
      mutate(across(pw_comparison, ~factor(., levels = c('Field_vs_Lab', 'SC2_vs_Lab', 'SC5_vs_Lab', 'SC7_vs_Lab')))) %>%
      arrange(pw_comparison, dir) %>%
      filter(!is.na(pw_comparison)) %>%
      pivot_wider(names_from = 'pw_comparison', values_from = 'dir') %>%
      filter(if_any(matches('.*_vs_Lab'), ~!is.na(.))) %>%
      {
        if(T %in% (str_detect(names(.),'SC._vs_SC.'))) 
          mutate(.,FieldvsLab_c = case_when(
            if_any(contains('Lab'), ~ . == 'Downregulated') & if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'both',
            if_any(contains('Lab'), ~ . == 'Downregulated') ~ 'Downregulated',
            if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'Upregulated')) 
        else 
          .
      } %>% 
      dplyr::select(-c(matches('SC._vs_SC.'))) %>%
      arrange(across(-one_of("gene_id")))%>%
      column_to_rownames('gene_id') })

list(msc_g_noLE_ref_ts_sce_ub_de[1:8], 
     hmap_gn_sVl_tbls, 
     pts_sst[1:8], 
     rep(list(sVs_order), 8),
     rep(list(strain_dset_lvls_don_col), 8),
     msc_g_noLE_v3_combi_afm_seu_sst[1:8],
     rep('strain_dset', 8),
     c(0.02, 0.1, 0.02, 0.1, 0.02, 0.1, 0.02,0.1),
     c(F, F, F, F, F, F, F, F), 
     c(6, 10, 6, 10, 6, 10, 6, 10),
     c(rep(13, 8))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab')}
    , 'e')
  )
```

All significant? Raw seurat log normalised and scaled -include field vs lab on left bars in strain comps - IN PAPER SUPPLEMENTARY
```{r, fig.height=5, fig.width=7}
## MAST Pvalue thresholds all field versus lab pval < 10e-10 and each strain versus field pval<0.05 
# mast_tde_corr_tbls_fVl_fg <-list(mast_tde_corr_tbls[6:10], mast_tde_corr_tbls[1:5]) %>% pmap(~left_join(..1 %>% filter(p_val_adj < 0.05), ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC, p_val_adj), by = 'gene_id') %>% pivot_longer(cols = c('pw_comparison.x', 'pw_comparison.y'), names_to = 'pw_class', values_to = 'pw_comparison' ) %>% distinct())

## !!! Combining the results from field versus Lab DE to those of the field strain_vs_strain DE so that we can visualize the field versus lab DE alongside strain DE

mast_tde_corr_tbls_fVl_fg <- list(mast_tde_corr_tbls[9:16], mast_tde_corr_tbls[1:8]) %>% 
  pmap(~left_join(..1%>% filter(p_val_adj < 0.05), 
                  ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC,waldStat, p_val_adj),  
                  by = 'gene_id', 
                  suffix = c('_fVSlab', '_sVSs'))%>% 
         pivot_longer(.,cols = starts_with(c('pw_comparison', 'waldStat','p_val_adj', 'avg_log2FC')), names_pattern = "(.*).(.VS.*$)", names_to = c( ".value", "f_strains_v_lab")) 
  )

hmap_gn_sVl_tbls <- list(mast_tde_corr_tbls_fVl_fg) %>%
  pmap(~{
    ..1 %>%
      filter(p_val_adj < 0.05 &  (avg_log2FC > 0 | avg_log2FC < -0 ) & waldStat > 10) %>% 
      mutate(dir = case_when(avg_log2FC > 0 ~ 'Upregulated', 
                             avg_log2FC < -0 ~ 'Downregulated')) %>%
      mutate(across(dir, ~factor(., levels=c('Upregulated', 'Downregulated')))) %>%
      dplyr::select(pw_comparison, gene_id, dir) %>% 
      filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>% 
      mutate(across(pw_comparison, ~droplevels(factor(., levels = c(strns_lab_comps, 'Field_vs_Lab'))))) %>%
      distinct() %>%
      arrange(pw_comparison, dir) %>%
      filter(!is.na(pw_comparison)) %>%
      pivot_wider(names_from = 'pw_comparison', values_from = 'dir') %>%
      filter(if_any(matches('SC._vs_SC.'), ~!is.na(.))) %>%
      # {
      #     if(T %in% (str_detect(names(.),'SC._vs_Lab'))) 
      #         mutate(.,FieldvsLab_c = case_when(
      #             if_any(contains('Lab'), ~ . == 'Downregulated') & if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'both',
      #             if_any(contains('Lab'), ~ . == 'Downregulated') ~ 'Downregulated',
      #             if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'Upregulated')) 
      #     else 
      #         .
      # } %>% 
      dplyr::select(-c(matches('SC._vs_Lab'))) %>%
      arrange(across(-one_of("gene_id")))%>%
      column_to_rownames('gene_id') })

all_strn_fld_hms <- list(msc_g_noLE_ref_ts_sce_ub_de[1:8], 
     hmap_gn_sVl_tbls, 
     pts_sst[1:8], 
     rep(list(sVs_order), 8),
     rep(list(strain_dset_lvls_don_col), 8),
     msc_g_noLE_v3_combi_afm_seu_sst[1:8],
     rep('strain_dset', 8),
     c(0.02, 0.1, 0.02, 0.1, 0.02, 0.1, 0.02,0.1),
     c(F, F, F, F, F, F, F, F), 
     c(6, 10, 6, 10, 6, 10, 6, 10),
     c(rep(19, 8)),
     c(rep(F, 8)),
     c(rep(16, 8)),
     c(rep(16, 8)),
     c(rep(17, 8)),
     c(rep(16, 8))) %>%
  pmap(
     possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            shw_anno_head = ..12,
                            strn2dsample = 'Lab',
                            # rsize = 6,
                            lgsize = ..13,
                            lblsize = ..14,
                            csize = ..15,
                            anno_trsize = ..16,
                            botom_pad = 21)}
     , 'e')
  )

all_strn_fld_hms
```



Top MAST genes table (avglfc_top >1.5 and avglfc_top <-1.5)
tradeseq smooth
```{r}

# sVs_order <- c('Strain 0', 'Strain 1', 'Strain 2', 'Lab')
fVl_order <- c('Field', 'Lab')

hmap_gn_tbls_lfc1.5 <- list(mast_tde_corr_tbls, 
                     c(as.list(rep('Field_vs_Lab', 8)),rep(list(lfstrains_comps), 20)),
                     c(as.list(rep('Field_vs_Lab', 8)),rep('^[7|N|S].*_vs_[7|N|S]*', 20)))%>%
  pmap(possibly(~{
    strn_comps_gns_anno_2_plt(#pval = 1e-10, 
                              pval = 0.05,
                              wstat= 10,
                              stg_pw_gns = ..1,
                              avglfc_top = 1.5,
                              avglfc_botom = -1.5,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*')
    
  }
  , 'e')
  )

list(msc_g_noLE_ref_ts_sce_ub_de, 
     hmap_gn_tbls, 
     pts_sst, 
     c(rep(list(fVl_order), 8),rep(list(sVs_order), 20)),
     c(rep(list(fvl_cols), 8),rep(list(strain_dset_lvls_don_col), 20)),
     c(F,F,F,F,F,T,T,T,T,T,F,T,T,T,T,T,T,T,T,T, rep(T,8)), 
     c(6,10,8,8,8,10,10,10,9,10,8,10,10,10,9,10,10,10,9,10, rep(9,8)),
     c(rep(13, 8),rep(13, 20))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='tde_smooth',
                            rnames = ..6,
                            rsize = ..7,
                            anno_fsize = ..8)}
    , 'e')
  )
```



Top average log 2 FC (>1.5 and <1.5) genes Raw counts - no smoothing -No Field vs lab comparison on bars in the left of strain De plots
Field vs lab comparisons - Field not split by strain
Raw Seurat log normalised and scaled 

```{r}

list(c(msc_g_noLE_ref_ts_sce_ub_de[1:4], msc_g_noLE_ref_ts_sce_ub_de), 
     c(hmap_gn_tbls_lfc1.5[1:4], hmap_gn_tbls_lfc1.5), 
     c(pts_sst[1:4], pts_sst), 
     c(rep(list(fVl_order), 4),rep(list(sVs_order), 28)),
     c(rep(list(fvl_cols), 4),rep(list(strain_dset_lvls_don_col), 28)),
     c(msc_g_noLE_v3_combi_afm_seu_sst[1:4], msc_g_noLE_v3_combi_afm_seu_sst),
     c(rep('dset', 2),rep('strain_dset', 2),rep('lf_strain', 28)),
     c(0.02,0.1,0.02,0.02,0.02,0.02,0.1,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02, rep(0.02,8)),
     c(F,F,F,F,F,F,rep(T,26)), 
     c(6,10,8,8,8,10,10,10,8.2,10,10,8,10,10,10,8.2,10,10,8,8,10,10,8,8, rep(9,8)),
     c(rep(13, 4),rep(13, 28))) %>%
  pmap(
    possibly(
      ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                              gns_tbl=..2,
                              cap_no = 20,
                              split_col = 'Condition',
                              ptime = ..3,
                              strn_row_anno_order = ..4,
                              col_map = ..5,
                              strn_conds = 'Condition',
                              pt_labl = 'Pseudotime',
                              hmap_mode='seurat',
                              seu_dset = ..6,
                              seu_comps = ..7,
                              chunk_size = ..8,
                              rnames = ..9,
                              rsize = ..10,
                              anno_fsize = ..11,
                              strn2dsample = 'Lab')}
      , 'e')
  )


```

Field vs lab comparisons - Field split by strain
Same as above just field stratified by strain in field by lab comparisons
Top MAST genes table (avglfc_top >1.5 and avglfc_top <-1.5) - no gene annotations
```{r}

list(msc_g_noLE_ref_ts_sce_ub_de[1:8], 
     hmap_gn_tbls_lfc1.5[1:8], 
     c(pts_sst[1:8]), 
     c(rep(list(sVs_order), 8)),
     c(rep(list(strain_dset_lvls_don_col), 8)),
     c(msc_g_noLE_v3_combi_afm_seu_sst [1:8]),
     c(rep('strain_dset', 8)),
     c(0.02, 0.1, 0.02, 0.1, 0.02, 0.1, 0.02,0.1),
     c(F, F, F, F, F, F, F, F), 
     c(6, 10, 6, 10, 6, 10, 6, 10),
     c(rep(13, 8))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab')}
    , 'e')
  )
```

Top MAST genes table (avglfc_top >1.5 and avglfc_top <-1.5)
Heatmap with custom mean smoothing
```{r, eval=FALSE}
### NB TAKES LOOOONG
list(c(msc_g_noLE_ref_ts_sce_ub_de[1:8],msc_g_noLE_ref_ts_sce_ub_de), 
     c(hmap_gn_tbls_lfc1.5[1:8], hmap_gn_tbls_lfc1.5), 
     c(pts_sst[1:8], pts_sst), 
     c(rep(list(fVl_order), 8),rep(list(sVs_order), 20)),
     c(rep(list(fvl_cols), 8),rep(list(strain_dset_lvls_don_col), 20)),
     c(msc_g_noLE_v3_combi_afm_seu_sst[1:8], msc_g_noLE_v3_combi_afm_seu_sst),
     c(rep('dset', 8),rep('strain_dset', 4),rep('lf_strain', 20)),
     c(0.02,0.1,0.02,0.02,0.02,0.02,0.1,0.1,rep(0.02,16)),
     c(F,F,F,F,F,F,rep(T,18)), 
     c(6,10,8,8,8,10,10,10,8.2,10,10,8,10,10,10,8.2,10,10,8,8,10,10,8,8),
     c(rep(13, 8),rep(13, 20))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 20,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat_smooth',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11)}
    , 'e')
  )
```

Top MAST genes table (avglfc_top >1.5 and avglfc_top <-1.5) - includes Field vs lab aggregate - FIELD VS LAB JUST ADDED FOR genes that pass cut off for STRAIN VS STRAIN COMPARISON to see how they compare to lab. 
Field VS lab comparison is all combined field cells versus the lab done statistically- Need to check on this clearly
```{r}
## MAST Pvalue thresholds all field versus lab pval < 10e-10 and each strain versus field pval<0.05 
# mast_tde_corr_tbls_fVl_fg <-list(mast_tde_corr_tbls[6:10], mast_tde_corr_tbls[1:5]) %>% pmap(~left_join(..1 %>% filter(p_val_adj < 0.05), ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC, p_val_adj), by = 'gene_id') %>% pivot_longer(cols = c('pw_comparison.x', 'pw_comparison.y'), names_to = 'pw_class', values_to = 'pw_comparison' ) %>% distinct())

mast_tde_corr_tbls_fVl_fg <- list(mast_tde_corr_tbls[9:16], mast_tde_corr_tbls[1:8]) %>% 
  pmap(~left_join(..1%>% filter(p_val_adj < 0.05), 
                  ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC,waldStat, p_val_adj),  
                  by = 'gene_id', 
                  suffix = c('_fVSlab', '_sVSs'))%>% 
         pivot_longer(.,cols = starts_with(c('pw_comparison', 'waldStat','p_val_adj', 'avg_log2FC')), names_pattern = "(.*).(.VS.*$)", names_to = c( ".value", "f_strains_v_lab")) 
  )

hmap_gn_sVl_tbls <- list(mast_tde_corr_tbls_fVl_fg) %>%
  pmap(~{
    ..1 %>%
      filter(p_val_adj < 1e-10 &  (avg_log2FC > 1.5 | avg_log2FC < -1.5 ) & waldStat > 10) %>% 
      mutate(dir = case_when(avg_log2FC > 1.5 ~ 'Upregulated', 
                             avg_log2FC < -1.5 ~ 'Downregulated')) %>%
      mutate(across(dir, ~factor(., levels=c('Upregulated', 'Downregulated')))) %>%
      dplyr::select(pw_comparison, gene_id, dir) %>% 
      filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>% 
      mutate(across(pw_comparison, ~droplevels(factor(., levels = c(strns_lab_comps, 'Field_vs_Lab'))))) %>%
      distinct() %>%
      arrange(pw_comparison, dir) %>%
      filter(!is.na(pw_comparison)) %>%
      pivot_wider(names_from = 'pw_comparison', values_from = 'dir') %>%
      filter(if_any(matches('SC._vs_SC.'), ~!is.na(.))) %>%
      # {
      #     if(T %in% (str_detect(names(.),'SC._vs_Lab'))) 
      #         mutate(.,FieldvsLab_c = case_when(
      #             if_any(contains('Lab'), ~ . == 'Downregulated') & if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'both',
      #             if_any(contains('Lab'), ~ . == 'Downregulated') ~ 'Downregulated',
      #             if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'Upregulated')) 
      #     else 
      #         .
      # } %>% 
      dplyr::select(-c(matches('SC._vs_Lab'))) %>%
      arrange(across(-one_of("gene_id")))%>%
      column_to_rownames('gene_id') })

top_strn_hms <- list(msc_g_noLE_ref_ts_sce_ub_de[1:8], 
     hmap_gn_sVl_tbls, 
     pts_sst[1:8], 
     rep(list(sVs_order), 8),
     rep(list(strain_dset_lvls_don_col), 8),
     msc_g_noLE_v3_combi_afm_seu_sst[1:8],
     rep('strain_dset', 8),
     c(0.02,0.1, 0.02,0.1, 0.02,0.1, 0.02,0.1),
     c(T,T, T,T, T,T, T,T), 
     c(rep(10, 8)),
     c(rep(17, 8)),
     c(rep(15, 8)),
     c(rep(14, 8)),
     c(rep(17, 8)),
     c(rep(16.5, 8))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab',
                            lgsize = ..12,
                            lblsize = ..13,
                            csize = ..14,
                            botom_pad = ..15)}
    , 'e')
  )
top_strn_hms

                            # split_col = 'Condition',  )
```

```{r, fig.height= 5, fig.width=7.3}
## MAST Pvalue thresholds all field versus lab pval < 10e-10 and each strain versus field pval<0.05 
# mast_tde_corr_tbls_fVl_fg <-list(mast_tde_corr_tbls[6:10], mast_tde_corr_tbls[1:5]) %>% pmap(~left_join(..1 %>% filter(p_val_adj < 0.05), ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC, p_val_adj), by = 'gene_id') %>% pivot_longer(cols = c('pw_comparison.x', 'pw_comparison.y'), names_to = 'pw_class', values_to = 'pw_comparison' ) %>% distinct())


hmap_gn_sVl_pp_lfc1.5 <- list(mast_tde_corr_tbls[9:28], 
                     c(rep(list(lfstrains_comps), 20)),
                     c(rep('^[7|N|S].*_vs_[7|N|S]*', 20)))%>%
  pmap(possibly(~{
    strn_comps_gns_anno_2_plt(#pval = 1e-10, 
                              pval = 1e-10,
                              wstat= 10,
                              stg_pw_gns = ..1,
                              avglfc_top = 1.1,
                              avglfc_botom = -1.1,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*|.*4_A.*')
    
  }
  , 'e')
  )

top_strn_hms <- list(msc_g_noLE_ref_ts_sce_ub_de[9:28], 
     hmap_gn_sVl_pp_lfc1.5, 
     pts_sst[9:28], 
     rep(list(sVs_order), 20),
     rep(list(strain_dset_lvls_don_col), 20),
     msc_g_noLE_v3_combi_afm_seu_sst[9:28],
     rep('lf_strain', 20),
     c(0.02,0.1,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02,0.02, rep(0.02, 8)),
     c(T,T,T,T,T,T,T,T,T,T,T,T, rep(T, 8)), 
     c(10,10,14,8.2,10,14,8.2,10,10,14,8.2,10, rep(10, 8)),
     c(rep(17, 20)),
     c(rep(15, 20)),
     c(rep(14, 20)),
     c(rep(17, 20)),
     c(rep(18, 20))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab',
                            lgsize = ..12,
                            lblsize = ..13,
                            csize = ..14,
                            botom_pad = ..15,
                            strain2omit_top_bar = rep(c("NF54_A", "SC6"), 20))}
    , 'e')
  )
top_strn_hms

                            # split_col = 'Condition',  )
```

Top 10 DE genes
```{r, fig.height= 5, fig.width=7.3}
## Plot significant genes waldstat>10 and MAST P <0.05 that are top 10 based on absolute log fold change of pairwise DE comparisons for the powered analysis
hmap_gn_sVl_pp_top10 <- list(mast_tde_corr_tbls[25:28], 
                     c(rep(list(lfstrains_comps), 4)),
                     c(rep('^[7|N|S].*_vs_[7|N|S]*', 4)))%>%
  pmap(possibly(~{
    strn_comps_gns_anno_2_plt(#pval = 1e-10, 
                              pval = 0.05,
                              wstat= 10,
                              topgn = 10,
                              stg_pw_gns = ..1,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*|.*4_A.*')
    
  }
  , 'e')
  )

top10_strn_hms <- list(msc_g_noLE_ref_ts_sce_ub_de[25:28], 
     hmap_gn_sVl_pp_top10, 
     pts_sst[25:28], 
     rep(list(sVs_order), 4),
     rep(list(strain_dset_lvls_don_col), 4),
     msc_g_noLE_v3_combi_afm_seu_sst[25:28],
     rep('lf_strain', 4),
     rep(0.02, 4),
     rep(T, 4), 
     rep(11, 4),
     c(rep(14, 4)),
     c(rep(15, 4)),
     c(rep(14, 4)),
     c(rep(16, 4)),
     c(rep(19, 4))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab',
                            lgsize = ..12,
                            lblsize = ..13,
                            csize = ..14,
                            botom_pad = ..15,
                            strain2omit_top_bar = rep(c("NF54_A", "SC6"), 4))}
    , 'e')
  )
top10_strn_hms

                            # split_col = 'Condition',  )
```


Top 10 DE genes with lab
```{r, fig.height= 5, fig.width=7.3}
## Plot significant genes waldstat>10 and MAST P <0.05 that are top 10 based on absolute log fold change of pairwise DE comparisons for the powered analysis
hmap_gn_sVl_pp_top10_inc_lab <- list(mast_tde_corr_tbls[9:16], 
                     c(rep(list(lfstrains_comps), 8)),
                     c(rep('^[7|N|S].*_vs_[7|N|S]*', 8)))%>%
  pmap(possibly(~{
    strn_comps_gns_anno_2_plt(#pval = 1e-10, 
                              pval = 0.05,
                              wstat= 10,
                              topgn = 1500, ## works unpredictably as some of the groups are removed after passing the ranking based on avgLogFC
                              stg_pw_gns = ..1,
                              comp_lvls=..2,
                              sig_strns_annotn_2_keep = ..3,
                              strns_annotn_2_omit = 'SC._vs_[L|N|7].*|.*4_A.*')
    
  }
  , 'e')
  )

top10_strn_hms_inc_lab <- list(msc_g_noLE_ref_ts_sce_ub_de[9:16], 
     hmap_gn_sVl_pp_top10_inc_lab, 
     pts_sst[9:16], 
     rep(list(sVs_order), 8),
     rep(list(strain_dset_lvls_don_col), 8),
     msc_g_noLE_v3_combi_afm_seu_sst[9:16],
     rep('lf_strain', 8),
     rep(0.02, 8),
     rep(T, 8), 
     rep(11, 8),
     c(rep(14, 8)),
     c(rep(15, 8)),
     c(rep(14, 8)),
     c(rep(16, 8)),
     c(rep(19, 8))) %>%
  pmap(possibly(
    ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
                            gns_tbl=..2,
                            cap_no = 25,
                            split_col = 'Condition',
                            ptime = ..3,
                            strn_row_anno_order = ..4,
                            col_map = ..5,
                            strn_conds = 'Condition',
                            pt_labl = 'Pseudotime',
                            hmap_mode='seurat',
                            seu_dset = ..6,
                            seu_comps = ..7,
                            chunk_size = ..8,
                            rnames = ..9,
                            rsize = ..10,
                            anno_fsize = ..11,
                            strn2dsample = 'Lab',
                            lgsize = ..12,
                            lblsize = ..13,
                            csize = ..14,
                            botom_pad = ..15,
                            strain2omit_top_bar = rep(c("NF54_A"), 4))}
    , 'e')
  )
top10_strn_hms_inc_lab

                            # split_col = 'Condition',  )
```

Top MAST genes table (avglfc_top >1.5 and avglfc_top <-1.5) - includes Field vs lab aggregate - FIELD VS LAB JUST ADDED FOR genes that pass cut off for STRAIN VS STRAIN COMPARISON to see how they compare to lab. 
Field VS lab comparison is all combined field cells versus the lab done statistically- Need to check on this clearly
```{r}
## MAST Pvalue thresholds all field versus lab pval < 10e-10 and each strain versus field pval<0.05 
# mast_tde_corr_tbls_fVl_fg <-list(mast_tde_corr_tbls[6:10], mast_tde_corr_tbls[1:5]) %>% pmap(~left_join(..1 %>% filter(p_val_adj < 0.05), ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC, p_val_adj), by = 'gene_id') %>% pivot_longer(cols = c('pw_comparison.x', 'pw_comparison.y'), names_to = 'pw_class', values_to = 'pw_comparison' ) %>% distinct())
# 
# mast_tde_corr_tbls_fVl_fg <- list(mast_tde_corr_tbls[9:16], mast_tde_corr_tbls[1:8]) %>% 
#   pmap(~left_join(..1%>% filter(p_val_adj < 0.05), 
#                   ..2 %>% filter(p_val_adj < 0.05) %>% dplyr::select(gene_id, pw_comparison, avg_log2FC,waldStat, p_val_adj),  
#                   by = 'gene_id', 
#                   suffix = c('_fVSlab', '_sVSs'))%>% 
#          pivot_longer(.,cols = starts_with(c('pw_comparison', 'waldStat','p_val_adj', 'avg_log2FC')), names_pattern = "(.*).(.VS.*$)", names_to = c( ".value", "f_strains_v_lab")) 
#   )
# 
# hmap_gn_sVl_tbls <- list(mast_tde_corr_tbls_fVl_fg) %>%
#   pmap(~{
#     ..1 %>%
#       filter(p_val_adj < 1e-10 &  (avg_log2FC > 1.5 | avg_log2FC < -1.5 ) & waldStat > 10) %>% 
#       mutate(dir = case_when(avg_log2FC > 1.5 ~ 'Upregulated', 
#                              avg_log2FC < -1.5 ~ 'Downregulated')) %>%
#       mutate(across(dir, ~factor(., levels=c('Upregulated', 'Downregulated')))) %>%
#       dplyr::select(pw_comparison, gene_id, dir) %>% 
#       filter(!str_detect(pw_comparison, 'SC._vs_Lab')) %>% 
#       mutate(across(pw_comparison, ~droplevels(factor(., levels = c(strns_lab_comps, 'Field_vs_Lab'))))) %>%
#       distinct() %>%
#       arrange(pw_comparison, dir) %>%
#       filter(!is.na(pw_comparison)) %>%
#       pivot_wider(names_from = 'pw_comparison', values_from = 'dir') %>%
#       filter(if_any(matches('SC._vs_SC.'), ~!is.na(.))) %>%
#       # {
#       #     if(T %in% (str_detect(names(.),'SC._vs_Lab'))) 
#       #         mutate(.,FieldvsLab_c = case_when(
#       #             if_any(contains('Lab'), ~ . == 'Downregulated') & if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'both',
#       #             if_any(contains('Lab'), ~ . == 'Downregulated') ~ 'Downregulated',
#       #             if_any(contains('Lab'), ~ . == 'Upregulated') ~ 'Upregulated')) 
#       #     else 
#       #         .
#       # } %>% 
#       dplyr::select(-c(matches('SC._vs_Lab'))) %>%
#       arrange(across(-one_of("gene_id")))%>%
#       column_to_rownames('gene_id') })
# 
# top_strn_hms <- list(msc_g_noLE_ref_ts_sce_ub_de[9:16], 
#      hmap_gn_sVl_tbls, 
#      pts_sst[1:8], 
#      rep(list(sVs_order), 8),
#      rep(list(strain_dset_lvls_don_col), 8),
#      msc_g_noLE_v3_combi_afm_seu_sst[1:8],
#      rep('strain_dset', 8),
#      c(0.02,0.1,0.02,0.02,0.02,0.02,0.02,0.02),
#      c(T,T,T,T,T,T,T,T), 
#      c(10,10,14,8.2,10,14,8.2,10),
#      c(rep(17, 8)),
#      c(rep(15, 8)),
#      c(rep(14, 8)),
#      c(rep(17, 8)),
#      c(rep(16.5, 8))) %>%
#   pmap(possibly(
#     ~{hmap_pw_strns_lab_tde(ts_dset= ..1,
#                             gns_tbl=..2,
#                             cap_no = 25,
#                             split_col = 'Condition',
#                             ptime = ..3,
#                             strn_row_anno_order = ..4,
#                             col_map = ..5,
#                             strn_conds = 'Condition',
#                             pt_labl = 'Pseudotime',
#                             hmap_mode='seurat',
#                             seu_dset = ..6,
#                             seu_comps = ..7,
#                             chunk_size = ..8,
#                             rnames = ..9,
#                             rsize = ..10,
#                             anno_fsize = ..11,
#                             strn2dsample = 'Lab',
#                             lgsize = ..12,
#                             lblsize = ..13,
#                             csize = ..14,
#                             botom_pad = ..15)}
#     , 'e')
#   )
# top_strn_hms

                            # split_col = 'Condition',  )
```


```{r}
##Check overlap between field DE genes across donors to see which genes are likely to always be DE between strains in a donor - competition genes
sig_fde <- mast_tde_corr_tbls[str_detect(names(mast_tde_corr_tbls), "StrainsDE_MSC[0-9]*_f_don$")] %>% map(~.x %>% filter(( p_val_adj < 0.05 & waldStat > 10)) %>% distinct(gene_id, .keep_all = T) %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id")) 

sig_mde <- mast_tde_corr_tbls[str_detect(names(mast_tde_corr_tbls), "StrainsDE_MSC[0-9]*_m_don$")] %>% map(~.x %>% filter((p_val_adj < 0.05 & waldStat > 10)) %>% distinct(gene_id, .keep_all = T) %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id")) 

names(sig_fde) <- str_remove(names(sig_fde), "lf_StrainsDE_")
names(sig_mde) <- str_remove(names(sig_mde), "lf_StrainsDE_")

sig_mde[[3]]
```


```{r}
## Overlap between significant DE genes between samples - venn
## female
ggvenn(map(sig_fde, ~.x$gene_id))

## male
ggvenn(map(sig_mde, ~.x$gene_id))
```


```{r}
##Check overlap between field DE genes across donors to see which genes are likely to always be DE between strains in a donor - competition genes
sig_fde_wk <- mast_tde_corr_tbls[str_detect(names(mast_tde_corr_tbls), "StrainsDE_MSC[1|3|13]*_f_don$")] %>% map(~.x %>% filter(( p_val_adj < 0.05 & waldStat > 10)) %>% distinct(gene_id, .keep_all = T) %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id")) 

sig_mde_wk <- mast_tde_corr_tbls[c("lf_StrainsDE_MSC1_m_don", "lf_StrainsDE_MSC13_m_don", "lf_StrainsDE_MSC14_m_don")] %>% map(~.x %>% filter((p_val_adj < 0.05 & waldStat > 10)) %>% distinct(gene_id, .keep_all = T) %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id")) 

names(sig_fde_wk) <- str_remove_all(names(sig_fde_wk), "lf_StrainsDE_|_f_don")
names(sig_mde_wk) <- str_remove_all(names(sig_mde_wk), "lf_StrainsDE_|_m_don")

sig_mde_wk[[2]]
```


```{r}
## Overlap between significant DE genes between samples - venn
## female
fde_wk_ven <- ggvenn(map(sig_fde_wk, ~.x$gene_id),
       show_percentage = F,
         text_size = 6,
         set_name_size = 6,
  ) + coord_fixed(ratio = 2/4)+
    theme(#aspect.ratio = 3/4,
    # plot.background = element_rect(fill = "yellow"),
    plot.margin = margin(0,0,0,0))+
      scale_y_continuous(limits = c(-2, 2))

## male
mde_wk_ven <-ggvenn(map(sig_mde_wk, ~.x$gene_id),
       show_percentage = F,
         text_size = 6,
         set_name_size = 6,
  ) + coord_fixed(ratio = 2/4)+
    theme(#aspect.ratio = 3/4,
    # plot.background = element_rect(fill = "yellow"),
    plot.margin = margin(0,0,0,0))+
      scale_y_continuous(limits = c(-2, 2))

fde_wk_ven
mde_wk_ven
```

```{r}
## Overlap between significant DE genes between samples - list
## female
# Create a list of all the combinations
# a <- map(sig_fde, ~.x$Gene)
a <- map(sig_fde_wk, ~.x$Gene)
combs <- 
  unlist(lapply(1:length(a), 
                function(j) combn(a, j, simplify = FALSE)),
         recursive = FALSE)
names(combs) <- sapply(combs, function(i) paste0(names(i), collapse = ""))


overlaps <- map(combs, ~(Reduce(base::intersect,.x)))
overlaps_nonempty <- overlaps[unlist(map(overlaps, ~length(.x) > 0))]
overlaps_nonempty[str_detect(names(overlaps_nonempty), ".*_.*_")]

```


```{r}
## Overlap between significant DE genes between samples - list
## male
# Create a list of all the combinations
# a <- map(sig_fde, ~.x$Gene)
a <- map(sig_mde_wk, ~.x$Gene)
combs <- 
  unlist(lapply(1:length(a), 
                function(j) combn(a, j, simplify = FALSE)),
         recursive = FALSE)
names(combs) <- sapply(combs, function(i) paste0(names(i), collapse = ""))


overlaps <- map(combs, ~(Reduce(base::intersect,.x)))
overlaps_nonempty <- overlaps[unlist(map(overlaps, ~length(.x) > 0))]
overlaps_nonempty[str_detect(names(overlaps_nonempty), ".*_.*_")]

```

```{r}
##Check overlap between field DE genes across donors to see which genes are likely to always be DE between strains in a donor - competition genes
lab_de <- mast_tde_corr_tbls[c("lf_StrainsDE_M_NF54_7G8_f", "lf_StrainsDE_M_NF54_7G8_m")] %>% map(., ~.x %>% filter(p_val_adj < 0.05 & waldStat > 10) %>% distinct(gene_id, .keep_all = T)  %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id"))

fld_de <- mast_tde_corr_tbls[c("lf_StrainsDE_MSC14_f_don3", "lf_StrainsDE_MSC3_m_don")] %>% map(., ~.x %>% filter((p_val_adj < 0.05 & waldStat > 10)) %>% distinct(gene_id, .keep_all = T)  %>% left_join(., Pf3D7_genes_plasmoDB, by = "gene_id"))


names(lab_de) <- str_remove(names(lab_de), "lf_StrainsDE_")
names(fld_de) <- str_remove(names(fld_de), "lf_StrainsDE_")

fld_de[[2]]
```

```{r}
## Overlap between significant DE genes between samples - venn
## female
fld_lab_de <- map2(fld_de, lab_de, ~{
  ggvenn(list("Donor strains" = .x$gene_id, 
         "NF54 vs 7G8" = .y$gene_id),
          fill_color = c("#54d1ef", "#92b938"),
         show_percentage = T,
         text_size = 5,
         set_name_size = 5,
  ) + coord_fixed(ratio = 2/4)+
    theme(#aspect.ratio = 3/4,
    # plot.background = element_rect(fill = "yellow"),
    plot.margin = margin(0,0,0,0))+
      scale_y_continuous(limits = c(-1, 2))
})


fld_lab_de
```

```{r}
top_sig_gns_plts <- stg_strn_mast_gns_smooth_plts(gn_xprsn_tbls = c( gn_tble_gg_smoother), 
                                                  gns_strn = map(strns_ub_pairwise_mkrs_pt_ml_sst, 
                                                                 . %>%
                                                                   filter(p_val_adj < 10e-20 & (avg_log2FC > 1 | avg_log2FC < -1))  %>% 
                                                                   arrange(p_val_adj) %>%
                                                                   dplyr::select(gene_id, pw_comparison)), 
                                                  stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                                  comp_g = 'pw_comparison',
                                                  ptimes = pts_sst,
                                                  top_gns = 10)

stg_strn_mast_gns_smooth_plts_saver(plts_cnm =  'top_sig_gns_plts', plt_lst = top_sig_gns_plts)

```

Save top significant pairwise strain DE genes by MAST adjusted p value (<10e-20)
```{r, fig.width=16, fig.height=8}
                                   
p2 =EnhancedVolcano(strns_ub_pairwise_mkrs_pt_ml_sst[[15]] %>% filter(pw_comparison == 'SC2_vs_SC7'),
                    lab = left_join(strns_ub_pairwise_mkrs_pt_ml_sst[[15]], Pf3D7_genes_plasmoDB, by = 'gene_id')%>% filter(pw_comparison == 'SC2_vs_SC7') %>% pull(Gene),
                    x = 'avg_log2FC',
                    y = 'p_val_adj',
                    hline = c(10e-10),
                    pCutoff = 0.05,
                    FCcutoff = 1,
                    pointSize = 3.0,
                    labSize = 4.0,
                    title = '',
                    subtitle = "",
                    titleLabSize = 0,
                    subtitleLabSize = 0,
                    legendLabels =c(expression(Log[2] ~ FC), "p-value"))

p3 =EnhancedVolcano(strns_ub_pairwise_mkrs_pt_ml_sst[[15]] %>% filter(pw_comparison == 'SC5_vs_SC7'),
                    lab = left_join(strns_ub_pairwise_mkrs_pt_ml_sst[[15]], Pf3D7_genes_plasmoDB, by = 'gene_id')%>% filter(pw_comparison == 'SC5_vs_SC7') %>% pull(Gene),
                    x = 'avg_log2FC',
                    y = 'p_val_adj',
                    hline = c(10e-10),
                    pCutoff = 0.05,
                    FCcutoff = 1,
                    pointSize = 3.0,
                    labSize = 4.0,
                    title = '',
                    subtitle = "",
                    titleLabSize = 0,
                    subtitleLabSize = 0,
                    legendLabels =c(expression(Log[2] ~ FC), "p-value"))

plot_grid(p2,p3, ncol = 3  )

```


Strains pairwise DE overlap genes
venn diagram overlap in genes upregulated in first strain versus the second
```{r, fig.width= 8, fig.height=3}

# strns_comps =  c('SC2_vs_SC5', 'SC2_vs_SC7', 'SC5_vs_SC7')
list(strns_ub_pairwise_mkrs_pt_ml_sst[11:14]) %>%
  pmap(., ~{
    mast_de_tbl = ..1
    map(strns_comps, ~ mast_de_tbl %>% 
          filter(p_val_adj < 0.05  & pw_comparison== .x & avg_log2FC > 0.2) %>% 
          pull(gene_id)) %>% 
      set_names(.,strns_comps) %>% 
      ggvenn(., fill_color =c("#0073C2FF", "#EFC000FF", "#868686FF", "#866686FF"), stroke_size = 0.5, set_name_size = 4 )
  }
  ) 

```

overlap in genes upregulated in second strain versus the first

```{r, fig.width= 8, fig.height=3}

list(mast_tde_corr_tbls[11], names(mast_tde_corr_tbls[11])) %>%
  pmap(., ~{
    mast_de_tbl = ..1
    plt<-map(strns_comps, ~ mast_de_tbl %>% 
               filter(p_val_adj < 0.05  & waldStat>10  & pw_comparison== .x ) %>% 
               pull(gene_id)) %>% 
      set_names(.,strns_comps %>% str_replace(., '_vs_', ' vs ')) %>% 
      ggvenn()+ 
      # scale_fill_discrete(name = "Condition", values = c("#0073C2FF", "#EFC000FF", "#868686FF"))+
      scale_x_continuous(expand = expansion(mult = .15,add = .15))+ 
      scale_y_continuous(expand = expansion(mult = .15,add = .15))#+ 
    # guides(fill = guide_legend(title = "Title")) +
    # theme(legend.position = "bottom")
    
    print(plt)
    # export::graph2ppt(plt,file= paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/plots/mali2021_all/strain_de/venns/', ..2))
  }
  ) 
```


UpSet plot
```{r, fig.width= 15, fig.height=4.8}
# list(mast_tde_corr_tbls[6:10]) %>%
upst_strn_de_plts <- list(mast_tde_corr_tbls[c(9:10,12,14:16,27)], c(0,3,2,3,3,2,3), c(50,50,8,50,50,8,50)) %>%
  pmap(., ~{
    ..1 %>% filter(p_val_adj < 0.05 & waldStat>10) %>% 
      pivot_wider(id_cols = 'gene_id', names_from = pw_comparison, values_from = p_val_adj) %>% 
      # dplyr::select(matches('^[SC|7|N].*')) %>% 
      dplyr::select(-c(matches('SC._vs_[L|N|7].*'),matches('^.*_vs_NF54_A'), gene_id))%>% 
      dplyr::filter(if_any(everything(), ~!is.na(.))) %>% 
      mutate(across(everything(), ~ifelse(is.na(.), F, T))) %>% 
      ComplexUpset::upset(., colnames(.), 
                          height_ratio=1.1, 
                          width_ratio=0.3, 
                          # min_size=..2, ## !!NOTE activate for female (LE) which has very many comparisons with intersection size 1
                          n_intersections =..3,
                          set_sizes=FALSE,
                          # set_sizes=(
                          #   upset_set_size()+
                          #     theme(axis.text.x=element_text(size = 18, angle=90),
                          #           axis.title.x=element_text(size = 20))
                          # ),
                          base_annotations=list(
                            'Intersection size'=intersection_size(
                              text=list(
                                # vjust=0.5,
                                # hjust=0.5,
                                # angle=90,
                                size = 5.2#,
                                # fontface = 'bold'
                                # text_colors=c(on_background='black', on_bar='black')
                              ) 
                            ) + 
                              labs(y= 'DE genes\nintersection')+
                              scale_y_continuous(expand = expansion(mult = c(0, 0)))),
                          matrix=(
                            intersection_matrix()), 
                          name= 'Pairwise \nstrain comparison',
                          themes=upset_modify_themes(
                            list(
                              'intersections_matrix'=theme(axis.text.y=element_text(size=18),
                                                           axis.title.x=element_text(size=18, face = 'bold')),
                              #'overall_sizes'=theme(axis.text.x=element_text(size=20)),
                              'Intersection size' = theme(axis.text.y=element_text(size=18), 
                                                          axis.title.y=element_text(size=18, face = 'bold'))
                                                          #axis.title.y=element_text(size=18, face = 'bold'))
                            )
                          )
      ) 
  })

upst_strn_de_plts
```



```{r}
##Check overlap between msc14 female and msc3 male DE strain DE genes - venn diagram
f_de_gns <- mast_tde_corr_tbls[["lf_StrainsDE_MSC14_f_don3"]] %>% filter(p_val_adj < 0.05 & waldStat>10) %>% mutate(across(pw_comparison, ~droplevels(factor(.,  levels = strns_lab_comps)))) 
f_de_gns <- split(f_de_gns, f = f_de_gns$pw_comparison) %>% map(~.x %>% pull(gene_id))
comp_nms <- names(f_de_gns)

m_de_gns <- mast_tde_corr_tbls[["lf_StrainsDE_MSC3_m_don2"]] %>% filter(p_val_adj < 0.05 & waldStat>10) %>% mutate(across(pw_comparison, ~droplevels(factor(.,  levels = strns_lab_comps)))) 
m_de_gns <- split(m_de_gns, f = m_de_gns$pw_comparison) %>% map(~.x %>% pull(gene_id))
      
```


```{r}
## Overlap between significant DE genes between samples - venn
## female
names(f_de_gns) <- str_replace(names(f_de_gns), "_vs_", ".")
fde_ven <- ggvenn(f_de_gns,
       show_percentage = F,
         text_size = 6,
         set_name_size = 6,
       fill_color = as.character(fld_strns_nf54_7g8_comps_col[comp_nms]),
  ) + 
  # coord_fixed(ratio = 2/4)+
    theme(aspect.ratio = 3.6/4,
    # plot.background = element_rect(fill = "yellow"),
    plot.margin = margin(0,0,0,0))+
      scale_y_continuous(limits = c(-1.95, 1.95))

## male
mde_ven <- ggplot()+
geom_circle(aes(x0 = 10, y0 = 10, r = 6, fill = fld_strns_nf54_7g8_comps_col[names(m_de_gns)], alpha = 0.5), inherit.aes = FALSE)+
geom_text(aes(x = 10, y = 10, label = length(m_de_gns[[1]])), size = 6, face ="plain") +
  labs(title = str_replace(names(m_de_gns), "_vs_", "."))+
  theme_void()+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "plain", size = 16),
        plot.margin = margin(t = 1.2, r = 2, b = 2, l = 2, "cm"))+
  coord_fixed()

fde_ven
mde_ven
```


Put together the plots for the DE analysis
```{r}
svs_denst_lfs_plts <- readRDS("plots/mali2021/svs_denst_lfs_plts.RDS")
```      
                   
```{r, fig.width= 8.5, fig.height=10}
plts <- pmap(list(all_strn_fld_hms[c(1:2,4)], svs_denst_lfs_plts[c(1:2,4)], upst_strn_de_plts[c(1:3)]), ~{
  HM_gb = grid.grabExpr(draw(..1))
  
  cowplot::plot_grid( HM_gb,
                      plot_grid(..2+
                                  theme(axis.text=element_text(size=14),
                                        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
                                        axis.title=element_text(size=16,face="bold"), 
                                        legend.text = element_text(size = 16) ,
                                        legend.title = element_text(size = 17, face = "bold"),
                                        legend.position="bottom",
                                        plot.background = element_rect(color = "lightgrey"),
                                        plot.margin=margin(t=0.6,r=0.2,b=0.6,l=0.2,unit="cm")
                                  )+ 
                      guides(color = guide_legend(override.aes = list(size = 4), ncol = 2, title.position = "top")), 
                                ..3, 
                                ncol=2, 
                                rel_widths = c(.35,.65)
                      ) +
                        theme(plot.margin=margin(b=0.8,unit="cm")) + panel_border(),
                      ncol=1, 
                      rel_heights = c(5.6,4.4)
  )
}

)
# theme(plot.background = element_rect(color = "black")
```

```{r, fig.width= 14, fig.height=18}
plot_grid(plotlist = plts, nrow = 2, ncol = 2, labels = c('A','B','C','D'), label_size = 24, vjust = 1.1)
```                      

```{r, fig.width= 12, fig.height=6}
##Put together the plots for main fig 4 on female HE DE distribution along pseudotime and DE

##modify patchwork upset plot appropriately
upst_strn_de_plts[[7]][[1]] <- upst_strn_de_plts[[7]][[1]] + theme(axis.title.y = element_text(size = 15,face="plain"))
upst_strn_de_plts[[7]][[1]] <- upst_strn_de_plts[[7]][[1]] + theme(axis.text.y = element_text(size = 14))
# upst_strn_de_plts[[7]][[2]] <- upst_strn_de_plts[[7]][[2]]+ theme(text = element_text(size = 8))
upst_strn_de_plts[[7]][[2]] <- upst_strn_de_plts[[7]][[2]]+ theme(axis.text.y = element_text(size = 15))
upst_strn_de_plts[[7]][[2]] <- upst_strn_de_plts[[7]][[2]]+ theme(axis.title.x = element_text(size = 15,face="plain"))

strain_fhDE <- plot_grid(
  plot_grid(svs_denst_lfs_plts[[19]]+
              theme(axis.text=element_text(size=14),
                    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
                    axis.title=element_text(size=15,face="plain"), 
                    legend.text = element_text(size = 14) ,
                    legend.title = element_text(size = 15, face = "plain"),
                    # legend.position="bottom",
                    # plot.background = element_rect(color = "lightgrey"),
                    plot.margin=margin(t=0.6,r=0.6,b=0.2,l=0.6,unit="cm")
              )+ 
              guides(color = guide_legend(override.aes = list(size = 4), ncol = 1, title.position = "top")), 
            ggdraw(upst_strn_de_plts[[7]])+
              theme(plot.margin = margin(t=0.6, b=0.1, unit = "cm")), 
            nrow=2, 
            rel_heights = c(.38,.62),
            labels = c('C', 'D'),
            label_size = 24,
            vjust= 1
  ),
  
  
  plot_grid(NULL, 
            grid.grabExpr(draw(top10_strn_hms[['lf_StrainsDE_MSC14_f_don3']])),
            NULL,
            nrow=3, 
            rel_heights = c(.05,.9,.05),
            labels = c('E','', ''),
            label_size = 24,
            vjust= 1 ),
  ncol=2,
  rel_widths = c(.4,.6)
)
strain_fhDE

# theme(plot.margin = margin(t=0.7, b=0.3, unit = "cm"))
```


```{r, fig.width= 6, fig.height=8}
##Put together the plots for main fig 4 on female HE DE distribution along pseudotime and DE

strain_fhDE <- list(svs_denst_lfs_plts[19:20], list(fde_ven, mde_ven), top10_strn_hms[c('lf_StrainsDE_MSC14_f_don3', 'lf_StrainsDE_MSC3_m_don2')], list(c("E", "F", "G"), c("H", "I", "J"))) %>%
  pmap(., ~{
  plot_grid(
  plot_grid(..1+
              theme(axis.text=element_text(size=14),
                    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
                    axis.title=element_text(size=15,face="plain"), 
                    legend.text = element_text(size = 14) ,
                    legend.title = element_text(size = 15, face = "plain"),
                    legend.position="bottom",
                    # plot.background = element_rect(color = "lightgrey"),
                    plot.margin=margin(t=0.6,r=0.6,b=0.2,l=0.6,unit="cm")
              )+ 
              guides(color = guide_legend(override.aes = list(size = 4), nrow = 1, title.position = "top")), 
            ggdraw(..2)+
              theme(plot.margin = margin(t=0.6, b=0.1, unit = "cm")), 
            ncol=2, 
            rel_widths = c(.45,.55),
            labels = c(..4[1], ..4[2]),
            label_size = 24,
            vjust= 1
  ),
  
  
  plot_grid(NULL, 
            grid.grabExpr(draw(..3)),
            NULL,
            nrow=3, 
            rel_heights = c(.05,.9,.05),
            labels = c(..4[3],'', ''),
            label_size = 24,
            vjust= 0.8 ),
  nrow=2,
  rel_heights = c(.4,.6)
)
  })
strain_fhDE

# theme(plot.margin = margin(t=0.7, b=0.3, unit = "cm"))
```

```{r, fig.width= 12, fig.height=7}
##Put together the plots for main fig 4 
main_de_plt <- plot_grid(plotlist = strain_fhDE, ncol = 2)
main_de_plt
```


```{r, fig.width= 12, fig.height=7}
##save svg for sk21
save_plot('plots/mali2021_all/fig4/main_de_plt.svg',main_de_plt, base_width = 12, base_height = 7)
```

```{r} 
##read plot grid for the scmap, strain pca and strain distribution generated in a different notebook
# stg_strn_scmap_pca_tbl_fig4_grid <- readRDS('plots/mali2021_all/fig4/stg_strn_scmap_pca_tbl_fig4_grid.RDS')
# 
# ## umap with baby
# all_stg_umap_fldHL <- readRDS('plots/mali2021_all/fig4/all_stg_umap_fldHL.RDS')
```


```{r, fig.width= 12, fig.height=14, eval=F}
##!! NB for old main fig
##Put together the plots for main fig 4 on female HE DE distribution along pseudotime and DE
fig_main <- plot_grid(stg_strn_scmap_pca_tbl_fig4_grid +
            theme(plot.margin = margin(b=0.6, unit = "cm")),
          strain_fhDE,
          plot_grid(NULL, all_stg_umap_fldHL, nrow = 1, rel_widths = c(.5,.5)),
nrow=3,
rel_heights = c(.32,.40, .28)
)

fig_main

save_plot('plots/mali2021_all/fig4/final_main_pt.svg',fig_main, base_width = 12, base_height = 14)
```

      
Count genes for manuscript 
```{r}
##Field VS lab genes
map(mast_tde_corr_tbls[1:8], . %>% filter(p_val_adj < 0.05 & waldStat>10) %>% pull(gene_id) %>% unique() %>% length())

##strain DE genes
map(mast_tde_corr_tbls[6:10], . %>% filter(p_val_adj < 0.05 & waldStat>10) %>% dplyr::filter(str_detect(pw_comparison, 'SC._vs_SC.')) %>% pull(gene_id) %>% unique() %>% length())

##Number of DE genes in genes present in both field and lab parasites
map(mast_tde_corr_tbls[9:28], . %>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter((str_detect(pw_comparison, 'SC._vs_SC.') | str_detect(pw_comparison, 'NF54_vs_7G8'))) %>% pull(gene_id) %>% unique() %>% length())

##Number of DE genes in genes present in both field and lab parasites
map(mast_tde_corr_tbls[9:28], . %>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'SC5_vs_SC7') %>% pull(gene_id) %>% unique() %>% length())

map(mast_tde_corr_tbls[9:28], . %>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'SC2_vs_SC5') %>% pull(gene_id) %>% unique() %>% length())

map(mast_tde_corr_tbls[9:28], . %>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'SC2_vs_SC7') %>% pull(gene_id) %>% unique() %>% length())

##Number of DE genes in genes present in both field and lab parasites
map(mast_tde_corr_tbls[9:28], . %>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'NF54_vs_7G8') %>% pull(gene_id) %>% unique() %>% length())

##Counts per strain in balanced object to compare DE in field VS DE in lab
msc_g_noLE_v3_combi_strn_mast[[13]]@meta.data %>% dplyr::count(lf_strain)
```

```{r}
##save as RDS object to be able to add the GO analysis in the functional_GO_n_lit_comparison.Rmd notebook
saveRDS(all_labvfld_hms[1:8], 'plots/mali2021_all/DE_heatmaps/all_labvfld_hms_fvl.RDS')

```     
               
Get genes that are expressed in 25% of the both field and lab parasites to get conclude on whether we seem more genes in the 7G8 vs NF54 comparison compared to the field strain comparisons 

```{r}
##Get genes expressed in atleast 25% of subset of interest in the integrated object ##WHY 69 GENES????

##1. lab parasites - day10A dataset with 7g8 and nf54
msc_g_noLE_v3_combi_strn_mast_fh_labB <- lapply(msc_g_noLE_v3_combi_strn_mast[9:16], function(x) x[,x@meta.data$lf_strain %in% c('NF54', '7G8')])
gn_prev_fh_labB <- lapply(msc_g_noLE_v3_combi_strn_mast_fh_labB, function(x) rowSums(x@assays$RNA@counts > 0)/dim(x)[2])
gns_fh_labB_25p <- lapply(gn_prev_fh_labB, function(x) names(x[x>=0.25] ))

##1. Field parasites
msc_g_noLE_v3_combi_strn_mast_fh_fld <- lapply(msc_g_noLE_v3_combi_strn_mast[9:16], function(x) x[,grepl('SC', x@meta.data$lf_strain)])
gn_prev_fh_fld <- lapply(msc_g_noLE_v3_combi_strn_mast_fh_fld, function(x) rowSums(x@assays$RNA@counts > 0)/dim(x)[2])
gn_prev_fh_fld_25p <- lapply(gn_prev_fh_fld, function(x) names(x[x>=0.25] ))

##Gene genes present in atleast 25% of both lab and field parasite subsets 
fld_labB_overlap_gns <- map2(gns_fh_labB_25p,gn_prev_fh_fld_25p, ~.x[.x %in% .y])

## The 7f8 vs nf54 and pure strain comparisons for msc13_m and msc14_f do not need the identification of genes that are identified both in lab and field and therefore adding NULL to represent these in the list
fld_labB_overlap_gns <- c(fld_labB_overlap_gns, rep(list(NULL), 12) %>% set_names(names(msc_g_noLE_v3_combi_strn_mast[17:28])))
     
```

DE genes in fld and lab that are in both 
UpSet plot

```{r, fig.width= 15, fig.height=4.8}
# list(mast_tde_corr_tbls[6:10]) %>%
upst_strn_de_plts_fld_labB_gns <- list(mast_tde_corr_tbls[c(9:10,12,14:16)], fld_labB_overlap_gns[str_remove(names(mast_tde_corr_tbls[c(9:10,12,14:16)]), "lf_StrainsDE_")]) %>%
  pmap(., ~{
    ..1 %>% 
      filter(gene_id %in% ..2) %>%
      filter(p_val_adj < 0.05 & waldStat>10) %>% 
      pivot_wider(id_cols = 'gene_id', names_from = pw_comparison, values_from = p_val_adj) %>% 
      # dplyr::select(matches('^[SC|7|N].*')) %>% 
      dplyr::select(-c(matches('SC._vs_[L|N|7].*'),matches('^.*_vs_NF54_A'), gene_id))%>% 
      dplyr::filter(if_any(everything(), ~!is.na(.))) %>% 
      mutate(across(everything(), ~ifelse(is.na(.), F, T))) %>% 
      ComplexUpset::upset(., colnames(.), 
                          height_ratio=1.1, 
                          width_ratio=0.3, 
                          set_sizes=FALSE,
                          base_annotations=list(
                            'Intersection size'=intersection_size(
                              text=list(
                                size = 5.2
                              ) 
                            ) + 
                              labs(y= 'DE genes\nintersection')+
                              scale_y_continuous(expand = expansion(mult = c(0, 0)))),
                          matrix=(
                            intersection_matrix()), 
                          name= 'Pairwise \nstrain comparison',
                          themes=upset_modify_themes(
                            list(
                              'intersections_matrix'=theme(axis.text.y=element_text(size=18),
                                                           axis.title.x=element_text(size=18, face = 'bold')),
                              'Intersection size' = theme(axis.text.y=element_text(size=18), 
                                                          axis.title.y=element_text(size=20, face = 'bold'))
                            )
                          )
      ) 
  })

upst_strn_de_plts_fld_labB_gns
```


```{r, fig.width= 6, fig.height=4.8}
# list(mast_tde_corr_tbls[6:10]) %>%
plt_nms <- names(mast_tde_corr_tbls[c(9:10,12,14:18,23:24,27)])
upst_strn_de_plts_fld_labB_gns <- list(mast_tde_corr_tbls[plt_nms], fld_labB_overlap_gns[str_remove(plt_nms, "lf_StrainsDE_")], c('Y', 'Y', 'Y', 'Y', 'Y', 'Y', 'N', 'N', 'N', 'N', 'N')) %>%
  pmap(., ~{
    
    
    tb<- ..1 %>% filter(p_val_adj < 0.05 & waldStat>10) %>% 
    pivot_wider(id_cols = 'gene_id', names_from = pw_comparison, values_from = p_val_adj)  %>% 
    dplyr::select(-c(matches('SC._vs_[L|N|7].*'),matches('^.*_vs_NF54_A')))%>% 
    dplyr::filter(if_any(contains('vs'), ~!is.na(.))) %>% 
    mutate(across(contains('vs'), ~ifelse(is.na(.), F, T)))
    
    if(..3 == 'Y')
    
    {
    ComplexUpset::upset(tb %>% 
    mutate(lab_field = if_else(gene_id %in% ..2, 'Yes', 'No')), 
                        tb %>% dplyr::select(contains('vs')) %>% colnames(), 
                          height_ratio=1.1, 
                          width_ratio=0.3, 
                          set_sizes=FALSE,
                          min_size = 0, ## !!  REMOVING COMBINATIONS WITH SIZE OF 1
                          base_annotations=list(
                            'Intersection size'=intersection_size(
                              text=list(
                                size = 4.8,
                                vjust=0.5,
                                hjust=-0.1,
                                angle=90
                              ),
                              bar_number_threshold = 1,
                              # counts=FALSE,
                              mapping=aes(fill=lab_field) 
                            ) + 
                              labs(y= 'DE genes\nintersection')+
                              scale_y_continuous(expand = expansion(mult = c(0, 0.35)))+
                              scale_fill_manual(name = "Lab & Field", values = c('No' = '#d2a7d2', 'Yes'= '#40e0d0'))+
                              theme(legend.position = 'top',
                                    legend.margin = margin(b=-9),
                                    legend.text = element_text(size = 13),
                                    legend.title = element_text(size = 14))),
                          matrix=(
                            intersection_matrix()), 
                          name= 'Pairwise strain comparison',
                          themes=upset_modify_themes(
                            list(
                              'intersections_matrix'=theme(axis.text.y=element_text(size=14),
                                                           axis.title.x=element_text(size=15)),
                              'Intersection size' = theme(axis.text.y=element_text(size=14), 
                                                          axis.title.y=element_text(size=15, face = 'plain'))
                            )
                          )
      ) 
    }
    else
    {
    ComplexUpset::upset(tb, 
                        tb %>% dplyr::select(contains('vs')) %>% colnames(), 
                          height_ratio=1.1, 
                          width_ratio=0.3, 
                          set_sizes=FALSE,
                          base_annotations=list(
                            'Intersection size'=intersection_size(
                              text=list(
                                size = 4,
                                vjust=0.5,
                                hjust=-0.1,
                                angle=90
                              ),
                              bar_number_threshold = 1,
                              # counts=FALSE 
                            ) + 
                              labs(y= 'DE genes\nintersection')+
                              scale_y_continuous(expand = expansion(mult = c(0, 0.35)))),
                          matrix=(
                            intersection_matrix()), 
                          name= 'Pairwise strain comparison',
                          themes=upset_modify_themes(
                            list(
                              'intersections_matrix'=theme(axis.text.y=element_text(size=14),
                                                           axis.title.x=element_text(size=15)),
                              'Intersection size' = theme(axis.text.y=element_text(size=14), 
                                                          axis.title.y=element_text(size=15))
                            )
                          )
      ) 
    }
  })

upst_strn_de_plts_fld_labB_gns
```

Put together the plots for the DE analysis
                      
```{r, fig.width= 9, fig.height=10}
# pt_indx <- c(9:10,12,14:18,23:24,27)
# 
# plts <- pmap(list(all_labvfld_hms[pt_indx], svs_denst_lfs_plts[pt_indx-8], upst_strn_de_plts), ~{
# 
# plts <- pmap(list(all_labvfld_hms[c(9:10,12,14:16)], svs_denst_lfs_plts[c(1:2,4,6:8)], upst_strn_de_plts), ~{
#   HM_gb = ggdraw(grid.grabExpr(draw(..1)))
#   plot_grid(
#   cowplot::plot_grid( HM_gb+
#                         theme(plot.margin=margin(b=0.2,t=0.8,unit="cm")),
#                       ..2+
#                                   theme(axis.text=element_text(size=14),
#                                         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
#                                         axis.title=element_text(size=16,face="bold"), 
#                                         legend.text = element_text(size = 16) ,
#                                         legend.title = element_text(size = 17, face = "bold"),
#                                         legend.position="bottom",
#                                         plot.margin=margin(b=2.1,t=2.1,unit="cm"),
#                                         plot.background = element_rect(color = "lightgrey")#,
#                                         # plot.margin=margin(t=1,r=0.2,b=1,l=0.2,unit="cm")
#                                   )+ 
#                       guides(color = guide_legend(override.aes = list(size = 4), ncol = 2, title.position = "top")),
#                       ncol=2, 
#                       rel_widths = c(7.2, 2.8),
#                       labels = c('A','B'), 
#                       label_size = 24, 
#                       vjust = 1.2
#                       ) + 
#                         panel_border(),
#                                 ggdraw(..3) +
#                         theme(plot.margin=margin(b=0.2,t=0.8,unit="cm")), 
#                                 ncol=1,
#                                 rel_heights = c(.56,.44),
#                       labels = c('','C'), 
#                       label_size = 24, 
#                       vjust = 1.2
#                       )+ panel_border() 
#   #)
# }
# 
# )
# # theme(plot.background = element_rect(color = "black")
# plts
```
                     
```{r, fig.width= 6, fig.height=8}
# plts <- pmap(list(all_labvfld_hms[plt_nms], upst_strn_de_plts_fld_labB_gns[plt_nms]), ~{

# plts <- pmap(list(all_labvfld_hms[c(9:12,14:28)], unlist(list(upst_strn_de_plts_fld_labB_gns[1:2], list(NULL), upst_strn_de_plts_fld_labB_gns[4], upst_strn_de_plts_fld_labB_gns[6:10], rep(list(NULL), 4), upst_strn_de_plts_fld_labB_gns[15:16], rep(list(NULL), 2), upst_strn_de_plts_fld_labB_gns[19], list(NULL)), recursive = F)), ~{
pt_nms <- names(all_labvfld_hms[c(9:12,14:28)])

upst_plts<- map(pt_nms, ~ifelse(.x %in% names(upst_strn_de_plts_fld_labB_gns), upst_strn_de_plts_fld_labB_gns[.x], no = list()))
  
plts <- pmap(list(all_labvfld_hms[pt_nms], unlist(upst_plts, recursive = F)), ~{
 
  HM_gb = ggdraw(grid.grabExpr(draw(..1)))
  plot_grid( HM_gb+
               theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")), 
             ggdraw(..2) +
               theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")),
             ncol=1,
             rel_heights = c(.52,.480)
  )+ panel_border() 
  #)
}

)
# theme(plot.background = element_rect(color = "black")
plts
```                  

                    
```{r, fig.width= 12, fig.height=10}
hm_wk_de <- map(all_labvfld_hms_w_gns_labl[c("lf_StrainsDE_MSC1_f_don", "lf_StrainsDE_MSC3_f_don", "lf_StrainsDE_MSC13_f_don", "lf_StrainsDE_MSC1_m_don", "lf_StrainsDE_MSC13_m_don", "lf_StrainsDE_MSC14_m_don")], ~ggdraw(grid.grabExpr(draw(.x))))

```              
   
                  
```{r, fig.width= 12, fig.height=15}
plot_grid(
  plot_grid(
  plotlist = hm_wk_de,
  ncol = 2,
  labels = c("A", "D", "B", "E", "C", "F"), 
          label_size = 24, 
          vjust = 1.1,
  # rel_heights = c(.3,.3,.3),
  byrow = F),
plot_grid(fde_wk_ven, mde_wk_ven, labels = c("G", "H"), 
          label_size = 24, 
          vjust = 1.1),
nrow = 2,
rel_heights = c(.85, .15)
)
```           

```{r, fig.width= 6, fig.height=8}
pts_sigde <- pmap(list(all_labvfld_hms[c("lf_StrainsDE_MSC14_f", "lf_StrainsDE_MSC3_m")], upst_strn_de_plts_fld_labB_gns[c("lf_StrainsDE_MSC14_f", "lf_StrainsDE_MSC3_m")], fld_lab_de), ~{
  HM_gb = ggdraw(grid.grabExpr(draw(..1)))
  plot_grid( HM_gb + theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")), 
             plot_grid(
               ggdraw(..2) + theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")),
               ..3, #+ coord_flip(),
               ncol = 2,
               rel_widths = c(.7, .3)
               ),
             ncol=1,
             rel_heights = c(.52,.480)
  )+ panel_border() 
  #)
}

)
# theme(plot.background = element_rect(color = "black")
pts_sigde
```                  

```{r, results='asis', fig.height=16.5, fig.width=12}
## Save RDS for concatenating to manuscript plot
DE_go_gtpt <- readRDS("plots/DE_heatmaps/DE_go_gtpt.RDS")
DE_go_sep_ont_gtpt <- readRDS("plots/DE_heatmaps/DE_go_sep_ont_gtpt.RDS")
ont_lg <- readRDS("plots/DE_heatmaps/ont_lg.RDS")
```

```{r, fig.width= 6, fig.height=8}
de_f_pt <- plot_grid( ggdraw(grid.grabExpr(draw(all_labvfld_hms[["lf_StrainsDE_MSC14_f_don3"]]))), 
           ggdraw(upst_strn_de_plts_fld_labB_gns[["lf_StrainsDE_MSC14_f_don3"]]),
           plot_grid(NULL, DE_go_sep_ont_gtpt[[1]], NULL, ncol = 3, rel_widths = c(.1,.8,.1))+ panel_border(),
           nrow=3,
           rel_heights = c(.45,.30,.25), 
          labels = c('B','', 'D'), 
          label_size = 24, 
          vjust = 1.1
  )+ panel_border() 
  #)

de_m_pt <- plot_grid( ggdraw(grid.grabExpr(draw(all_labvfld_hms[["lf_StrainsDE_MSC3_m_don"]]))), 
           plot_grid(DE_go_sep_ont_gtpt[[2]], DE_go_sep_ont_gtpt[[3]], DE_go_sep_ont_gtpt[[4]], ncol = 3, rel_widths = c(.36,.33,.31))+ panel_border(),
           nrow=2,
           rel_heights = c(.55,.45), 
          labels = c('C','E'), 
          label_size = 24, 
          vjust = 1.1
  )+ panel_border() 


```                 


```{r, fig.width= 12, fig.height=12}
plot_grid(
plot_grid(
  de_f_pt,
  de_m_pt,
  ncol = 2),
ont_lg,
plot_grid(plotlist = map(de_corr_plts[c("lf_StrainsDE_MSC14_f_don3","lf_StrainsDE_MSC3_m_don")], ~plot_grid(.x) + panel_border()), 
          ncol = 2, 
          labels = c('F','G'), 
          # scale = 0.95, 
          label_size = 24, 
          vjust = 1.1),
nrow = 3,
rel_heights = c(.70,.04,.26)
)

```        


```{r, fig.width= 6, fig.height=10}
# de_f_pt <- plot_grid( ggdraw(grid.grabExpr(draw(all_labvfld_hms[["lf_StrainsDE_MSC14_f_don3"]]))), 
#            ggdraw(upst_strn_de_plts_fld_labB_gns[["lf_StrainsDE_MSC14_f_don3"]]),
#            plot_grid(NULL, DE_go_gtpt[[1]], NULL, ncol = 3, rel_widths = c(.15,.7,.15)),
#            de_corr_plts[["lf_StrainsDE_MSC14_f_don3"]],
#            nrow=4,
#            rel_heights = c(.42,.28,.1, .3)
#   )+ panel_border() 
#   #)
# 
# de_m_pt <- plot_grid( ggdraw(grid.grabExpr(draw(all_labvfld_hms[["lf_StrainsDE_MSC3_m_don"]]))), 
#            plot_grid(NULL, DE_go_gtpt[[2]], NULL, ncol = 3, rel_widths = c(.15,.7,.15)),
#            de_corr_plts[["lf_StrainsDE_MSC3_m_don"]],
#            nrow=3,
#            rel_heights = c(.5,.2,.3)
#   )+ panel_border() 
# 
# de_f_pt
# de_m_pt
```                 


```{r, fig.width= 6, fig.height=8}
pts_sigde_hc <- pmap(list(all_labvfld_hms[c("lf_StrainsDE_MSC14_f_don3", "lf_StrainsDE_MSC3_m_don")], c(upst_strn_de_plts_fld_labB_gns[c("lf_StrainsDE_MSC14_f_don3")],list(NULL))), ~{
  HM_gb = ggdraw(grid.grabExpr(draw(..1)))
  plot_grid( HM_gb+
               theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")), 
             ggdraw(..2) +
               theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")),
             ncol=1,
             rel_heights = c(.52,.480)
  )+ panel_border() 
  #)
}
)

pts_sigde_hc
```                 
    
   
 
```{r, fig.width= 6, fig.height=6}
pts_sigde_lab <- pmap(list(all_labvfld_hms[c("lf_StrainsDE_M_NF54_7G8_f", "lf_StrainsDE_M_NF54_7G8_m")], fld_lab_de, list(c("A", "C"), c("B", "D"))), ~{
  HM_gb = ggdraw(grid.grabExpr(draw(..1)))
  plot_grid( HM_gb+
               theme(plot.margin=margin(b=0.01,t=0.01,unit="cm")), 
             ..2,
             ncol=1,
             rel_heights = c(.68,.32),
             labels = ..3,
             label_size = 24, 
             vjust = 1.4
  )+ panel_border() 
  #)
}
)

pts_sigde_lab
```






```{r, fig.width= 12, fig.height=9}
plot_grid(
plot_grid(plotlist = pts_sigde_lab,
          ncol = 2, 
          # labels = c('A','B'), 
          # label_size = 24, 
          vjust = 1.1),
plot_grid(plotlist = map(de_corr_plts[c("lf_StrainsDE_M_NF54_7G8_f","lf_StrainsDE_M_NF54_7G8_m")], ~plot_grid(.x) + panel_border()), 
          ncol = 2, 
          labels = c('E','F'), 
          # scale = 0.95, 
          label_size = 24, 
          vjust = 1.1),
nrow = 2,
rel_heights = c(.7,.3)
)

```        

```{r, fig.width= 12, fig.height=16}
plot_grid(plts[[1]], plts[[3]], plts[[5]], plts[[7]], nrow = 2, ncol =2, labels = c('A','B', 'C', 'D'), label_size = 24, vjust = 1.1)
```      


```{r, fig.width= 12, fig.height=16}
plot_grid(plotlist = plts[c(2,4,6)], nrow = 2, ncol =2, labels = c('A','B', 'C', ''), label_size = 24, vjust = 1.1)
```


```{r, fig.width= 12, fig.height=8}
plot_grid(plotlist = plts[c(7,4)], ncol = 2, labels = c('A','B'), label_size = 24, vjust = 1.1)
```

```{r, fig.width= 12, fig.height=10}
plot_grid(
plot_grid(plotlist = pts_sigde, ncol = 2, labels = c('C','D'), label_size = 24, vjust = 1.1),
plot_grid(plotlist = de_corr_plts[c(15,12)], ncol = 2, labels = c('E','F'), scale = 0.95, label_size = 24, vjust = 1.1),
nrow = 2,
rel_heights = c(.7,.3)
)
```


```{r, fig.width= 12, fig.height=10}
plot_grid(
plot_grid(plotlist = pts_sigde, ncol = 2, labels = c('C','D'), label_size = 24, vjust = 1.1),
plot_grid(plotlist = de_corr_plts[c(15,12)], ncol = 2, labels = c('E','F'), scale = 0.95, label_size = 24, vjust = 1.1),
nrow = 2,
rel_heights = c(.7,.3)
)
```

```{r}
##Number of DE genes in genes present in both field and lab parasites
mast_tde_corr_tbls[['lf_StrainsDE_MSC14_f']] %>% filter(gene_id %in% fld_labB_overlap_gns)%>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'SC5_vs_SC7') %>% dim

##Number of DE genes in genes present in both field and lab parasites
mast_tde_corr_tbls[['lf_StrainsDE_MSC14_f']] %>% filter(gene_id %in% fld_labB_overlap_gns)%>% filter(p_val_adj < 0.05 & waldStat>10)%>% dplyr::filter(pw_comparison == 'NF54_vs_7G8') %>% dim

```


Borderline significant MAST genes (p_val_adj < 0.05 & p_val_adj > 10e-4)
Visualize
```{r, eval=FALSE}
## !! NB takes looooong
border_sig_gns_plts <- stg_strn_mast_gns_smooth_plts(gn_xprsn_tbls = c(gn_tble_gg_smoother), 
                                                     gns_strn = map(strns_ub_pairwise_mkrs_pt_ml_sst, 
                                                                    . %>%
                                                                      filter(p_val_adj < 0.05 & p_val_adj > 10e-4)  %>% 
                                                                      arrange(p_val_adj) %>%
                                                                      dplyr::select(gene_id, pw_comparison)), 
                                                     stage_nms = names(strns_ub_pairwise_mkrs_pt_ml_sst), 
                                                     comp_g = 'pw_comparison',
                                                     ptimes = pts_sst,
                                                     top_gns = 10)

```

Save bordeline significant genes
```{r, eval=FALSE}
## !! NB takes looooong
stg_strn_mast_gns_smooth_plts_saver(plts_cnm = 'borderline_sig', plt_lst = border_sig_gns_plts)
```
