---
title: "Functional GO enrichment analysis and comparison with findings in literature"
date: '2022-09-10'
author: 'Jesse Rop'
output:
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---

```{r}
suppressPackageStartupMessages({
    library(Seurat)
    library(dplyr) 
    library(stringr)
    library(purrr)
    library(readr)
    library(tibble)
    library(ggplot2)
    library(tidyr)
    library(cowplot)
    library(SeuratData)
    library(SeuratDisk)
    library(pheatmap)
    library(plotly)
    library(rmarkdown)
    library(RColorBrewer)
    library(MAST)
    library(EnhancedVolcano)
    # library(org.Pf.plasmo.db)
    library(clusterProfiler) ##github installation newest version https://github.com/YuLab-SMU/clusterProfiler/issues/434#issuecomment-1048633162 remotes::install_github(c("YuLab-SMU/DOSE", "YuLab-SMU/clusterProfiler", "YuLab-SMU/enrichplot" ))
    library(Platypus)
    library(ggvenn)
    library(tradeSeq)
    library(MetBrewer)
    library(ComplexHeatmap)
    library(fgsea)
    library(msigdbr)
    library(slingshot)
  library(enrichplot)
  library(readxl)
  library(DT)
  library(knitr)
  library(conflicted)
  library(openxlsx)
  library(gt)
  library(ggforce)
  conflicts_prefer(purrr::map)
  conflicts_prefer(dplyr::filter)
  conflicts_prefer(dplyr::rename)
  # library(knitr)
})
```

# DE within MSC14 field parasites and DE on MSC14 VS MCA V3 parasites

## Variable declarations, read in input files and exploratory visualizations

```{r setup}
##Set working directory
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/')
##Get common resources
```

```{r}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id)
  )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                 TRUE ~ as.character(.)))) 
```

```{r}
##Get common resources
# source('scripts/msc_common_functions.R')
```

```{r}
##Assign colour variables for the different plots
sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   # 'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   'Female'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   ## 'Male (HE)'= 'mediumpurple',#'#039911',
                   'Male'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle'
)

monet_col <- met.brewer('Monet',9,override.order = T)
col_branches1 = c("commited"= monet_col[4],
                 "developing"=monet_col[5],
                 "branching"=monet_col[6],"early female"=monet_col[2],
                 "early male"=monet_col[8],"late female"=monet_col[1],"late male"=monet_col[7])

```


```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB_go <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_190322.csv") %>%
    dplyr::rename("gene_id" = `Gene ID` ) %>%
    group_by(gene_id) %>%
    add_count(name='gn_iso') %>%
    ungroup() %>%
    mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                   TRUE ~ as.character(.)))) %>%
    mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
    dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>%
    mutate(across(where(is.character), ~na_if(., "N/A")),
           Gene = coalesce(`Gene_name`, gene_id)
    )  %>%
    group_by(Gene) %>%
    add_count() %>%
  ungroup() %>%
  mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                   TRUE ~ as.character(.))))

```

```{r}
##Field datasets (msc1,3,13,14,1272C)
sample_id_lst <-as.list(c("MSC1", "MSC3", "MSC1272", "MSC13", "MSC14" ) %>% 
    str_sort(., numeric = T))

# sample_id_lst <-sample_id_lst[4]

msc_qcd_anotd_stgstrqc <- readRDS('data/processed/DB52e_star/msc_qcd_anotd.RDS') #%>% 
  # set_names(sample_id_lst)

msc_qcd_anotd_stgstrqc <- msc_qcd_anotd_stgstrqc[1:4]

```

```{r}
##Field datasets (msc1,3,13,14,1272C)
# mast_tde_corr_tbls <- readRDS('data/processed/msc14/5736STDY11771546_DB52e_star/mast_tde_corr_tbls.RDS')
mast_tde_corr_tbls <- readRDS('data/processed/DB52e_star/mast_tde_corr_tbls.RDS')
```

```{r}
##Field datasets (msc1,3,13,14,1272C)
msc_g_noLE_condRes_ub_de <- readRDS('data/processed/DB52e_star/msc_g_noLE_condRes_ub_de.RDS')
```

Read in tradeseq files from local
```{r}
# msc14_ref_ts_sce <- readRDS('data/processed/msc14/5736STDY11771546_DB52e_star/pseudotime_tradeseq/msc14_ref_ts_sce.RDS')
msc_g_noLE_v3_combi_strn_mast <- readRDS('data/processed/DB52e_star/msc_g_noLE_v3_combi_strn_mast.RDS')
```

### Exploratory Visualization
#### Field parasites UMAP split by strain {.tabset}
```{r, results='asis', fig.width=12, fig.height=3.5}
# cat('## Visualize field parasites {.tabset}   \n')
  msc_qcd_anotd_stgstrqc  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('##### Tab',.y,'   \n')
        p <- DimPlot(.x, reduction = "umap",  split.by = 'Strain_DN', group.by = 'stageHL', 
        cols = sp_fld_colors, pt.size = 1.5)  + 
          theme(axis.text=element_text(size=18), 
      axis.title=element_text(size=20,face="bold"), 
      # legend.text = element_blank(), 
      plot.title = element_blank(),
      strip.text.x = element_text(size = 24, face = "bold") )
        print(p)
        cat('\n  ')
        cat("\n")
      })

```


## MSC DE

### female (high count VS low count) and male (high count VS low count)

#### Female high vs female low DE list with annotation {.tabset}
```{r}
##female
# fh_fl_marker_list_acc_strn_3.13.14 <- readRDS( "data/processed/fh_fl_marker_list_acc_strn_3.13.14.RDS") %>% 
  # set_names(sample_id_lst[2:4])
fh_fl_marker_list_acc_strn_3.13.14 <- readRDS( "data/processed/DB52e_star/fh_fl_marker_list_acc_strn_3.13.14.RDS") 

fh_fl_marker_list_acc_strn_3.13.14 <- fh_fl_marker_list_acc_strn_3.13.14 %>%
    map(. %>% rownames_to_column('gene_id') %>% 
            left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') %>%
          arrange(desc(avg_log2FC)) %>%
          dplyr::select(Gene,gene_id, p_val,  avg_log2FC, pct.1, pct.2, p_val_adj, `Product Description`, Ka_Ks,TM_domains, `Computed GO Components`, `Computed GO Functions`, `Computed GO Processes`,`Curated GO Components`, `Curated GO Functions`, `Curated GO Processes`, `# Exons in Gene`,`Annotated 3' UTR length`,`Annotated 5' UTR length`,`Gene Strand`,`Gene Type`, `CDS Length`))

```

```{r, results='asis', eval=F}
##This code only runs when knitting. when running the chunks we get the error, 'Error in cat(knitr::knit_print(.x %>% head(., n = 50))) : argument 1 (type 'list') cannot be handled by 'cat'
data.frame() %>%
  DT::datatable() %>%
  knitr::knit_print() %>%
  attr('knit_meta') %>%
  knitr::knit_meta_add() %>%
  invisible()


imap(fh_fl_marker_list_acc_strn_3.13.14, ~{
      cat('##### ',.y,'\n')
      cat('Upregulated in Female \n')
      cat('\n')
      cat(knitr::knit_print(.x%>% head(., n=50) ))
      
      cat('Upregulated in Female (LE) \n')
      cat(knitr::knit_print(.x%>% tail(., n=50)))
      cat('\n')})
```


#### male (high count VS low count) DE with annotation {.tabset}
```{r}
##male
mh_ml_marker_list_acc_strn_13.14 <- readRDS( "data/processed/DB52e_star/mh_ml_marker_list_acc_strn_13.14.RDS") 

mh_ml_marker_list_acc_strn_13.14 <- mh_ml_marker_list_acc_strn_13.14 %>%
    map(. %>% rownames_to_column('gene_id') %>% 
            left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') %>%
          arrange(desc(avg_log2FC)) %>%
          dplyr::select(Gene,gene_id, p_val,  avg_log2FC, pct.1, pct.2, p_val_adj,gene_id, `Product Description`, Ka_Ks,TM_domains, `Computed GO Components`, `Computed GO Functions`, `Computed GO Processes`,`Curated GO Components`, `Curated GO Functions`, `Curated GO Processes`, `# Exons in Gene`,`Annotated 3' UTR length`,`Annotated 5' UTR length`,`Gene Strand`,`Gene Type`, `CDS Length`))
```

```{r, results='asis', eval=FALSE}
##This code only runs when knitting. when running the chunks we get the error, 'Error in cat(knitr::knit_print(.x %>% head(., n = 50))) : argument 1 (type 'list') cannot be handled by 'cat'
data.frame() %>%
  DT::datatable() %>%
  knitr::knit_print() %>%
  attr('knit_meta') %>%
  knitr::knit_meta_add() %>%
  invisible()


imap(mh_ml_marker_list_acc_strn_13.14, ~{
      cat('##### ',.y,'\n')
      cat('Upregulated in Female \n')
      cat('\n')
      cat(knitr::knit_print(.x%>% head(., n=50) ))
      
      cat('Upregulated in Female (LE) \n')
      cat(knitr::knit_print(.x%>% tail(., n=50)))
      cat('\n')})
```


### Average expression of gene sets

Function 
```{r}
makeSetScore <- function(seur, gene.set,assay='RNA', slt = 'counts') {
  # Get mean expression of genes of interest per cell
  mean.exp <- colMeans(x = as.matrix(GetAssayData(seur, slot = slt, assay = assay))[gene.set, ], na.rm = TRUE)
  
  # Add mean expression values in 'object@meta.data$gene.set.score'
  if (all(names(x = mean.exp) == rownames(x = seur@meta.data))) {
    return(mean.exp)
  }
}

```


Function avg expression across all cell per gene
```{r}
# per_gn_all_cells_makeSetScore <- function(seur, gene.set,assay='RNA') {
#   # Get mean expression of genes of interest across all cells
#   mean.exp <- rowMeans(x = as.matrix(GetAssayData(seur, slot = 'data', assay = assay))[gene.set, ], na.rm = TRUE)
#   
#   # Add mean expression values in 'object@meta.data$gene.set.score'
#   if (all(names(x = mean.exp) == rownames(x = seur@meta.data))) {
#     return(mean.exp)
#   }
# }

```

#### Average expression of 511 Lasonder translationally repressed genes in MSC {.tabset}
```{r, message=FALSE}
lasonder_tr <- read_xlsx('data/raw/published_lit/lasonder/supp_512_TR_genes.xlsx', skip = 4)
tr_gns <- lasonder_tr %>% dplyr::rename('gene_id' = `Gene ID...1`) %>% pull(gene_id) %>% str_replace(.,'_', '-')
tr_gns <- tr_gns[str_detect(tr_gns, 'PF3D7-')]
```


```{r}

lasonder_avg_exp <- msc_qcd_anotd_stgstrqc %>%
  map(~makeSetScore(., tr_gns, assay = 'RNA') %>% 
        data.frame %>% 
        setNames('lasonder_avg') 
      )

msc_qcd_anotd_stgstrqc <- list(msc_qcd_anotd_stgstrqc, lasonder_avg_exp) %>%
  pmap(~AddMetaData(..1, ..2))
```


```{r, results='asis', message=FALSE}
lasonder_avg_umap <-  msc_qcd_anotd_stgstrqc  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        p <- FeaturePlot(., features = "lasonder_avg", dims = c(1,2), label = T, label.size = 4, repel = T) + scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral")))+ 
            theme(axis.text=element_text(size=12), 
                  axis.title=element_text(size=12,face="bold"), 
                  plot.title = element_blank(),
                  legend.text = element_text(size=12),
                  legend.position="bottom",
                  legend.key.width = unit(1, "cm")
            )+
            coord_fixed() 
        print(p)
        cat('\n')
        cat("\n")
      })

lasonder_avg_umap
```


```{r, results='asis', message=FALSE}
trunc_axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(2.2, "cm")
)

lasonder_avg_umap <-  msc_qcd_anotd_stgstrqc  %>% 
      purrr::imap(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        #p<-do_FeaturePlot(., features = "lasonder_avg", dims = c(1,2),  label.size = 4,label = T,repel = T,label.color = "black") + 
        p<-FeaturePlot(., features = "lasonder_avg", dims = c(1,2),label = T,repel = T, label.size = 5) +
          scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral")))+
              theme_classic() +
              theme(#axis.text=element_text(size=18), 
                    #axis.title=element_text(size=18,face="bold"), 
                    legend.text = element_text(size=20),
                    legend.position="bottom",
                    legend.title =element_blank(),
                    plot.title = element_blank(),
                    axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
        axis.title = element_text(hjust = 0))+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)+
              # coord_fixed()+
        guides(color = guide_colorbar(barwidth = 10, barheight =0.6),
         x= trunc_axis,
         y= trunc_axis)
        return(p)
        # print(p)
        cat('\n')
        cat("\n")
      })


lasonder_avg_umap
```


##### Female LE, HE, Male LE, HE absence/present genes (detected in 25% of the population) overrepresentation test {.tabset}

```{r}
## Copy the gene lists to make do GO overrepresentation test
mfet_gns_3.13.14_don_colps<- readRDS("data/processed/DB52e_star/msc14/mf_le_he/mfet_gns_3.13.14_don_colps.RDS")
mfet_gns_3.13.14_lst<- readRDS("data/processed/DB52e_star/msc14/mf_le_he/mfet_gns_3.13.14_lst.RDS")
```


Avg expression of FLE genes in all clusters


```{r}
mfet_gns_3.13.14_don_colps %>% filter(fh_fl_only == 'y') %>% dim
fhe_fle_gns <- mfet_gns_3.13.14_don_colps %>% filter(fh_fl_only == 'y') %>% pull(gene_id)

fhe_fle_avg_exp <- msc_qcd_anotd_stgstrqc[2:4] %>%
  map(~makeSetScore(., fhe_fle_gns, assay = 'RNA') %>% 
        data.frame %>% 
        setNames('fhe_fle_avg') 
      )

msc_qcd_anotd_stgstrqc[2:4] <- list(msc_qcd_anotd_stgstrqc[2:4], fhe_fle_avg_exp) %>%
  pmap(~AddMetaData(..1, ..2))
```

```{r, results='asis', message=FALSE}

fhe_fle_umap <-  msc_qcd_anotd_stgstrqc[2:4]  %>% 
      purrr::imap(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        #p<-do_FeaturePlot(., features = "lasonder_avg", dims = c(1,2),  label.size = 4,label = T,repel = T,label.color = "black") + 
        p<-FeaturePlot(., features = "fhe_fle_avg", dims = c(1,2),label = T,repel = T, label.size = 5) +
          scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral")))+
              theme_classic() +
              theme(#axis.text=element_text(size=18), 
                    #axis.title=element_text(size=18,face="bold"), 
                    legend.text = element_text(size=20),
                    legend.position="bottom",
                    legend.title =element_blank(),
                    plot.title = element_blank(),
                    axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
        axis.title = element_text(hjust = 0))+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)+
              # coord_fixed()+
        guides(color = guide_colorbar(barwidth = 10, barheight =0.6),
         x= trunc_axis,
         y= trunc_axis)
        return(p)
        # print(p)
        cat('\n')
        cat("\n")
      })


fhe_fle_umap
```


```{r, results='asis', message=FALSE}
# msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% ggplot(aes(x = stageHL, y = fhe_fle_avg)) + geom_boxplot()
```


```{r}
mfet_gns_3.13.14_don_colps %>% filter(mh_ml_only == 'y') %>% dim
mhe_mle_gns <- mfet_gns_3.13.14_don_colps %>% filter(mh_ml_only == 'y') %>% pull(gene_id)

mhe_mle_avg_exp <- msc_qcd_anotd_stgstrqc[3:4] %>%
  map(~makeSetScore(., mhe_mle_gns, assay = 'RNA') %>% 
        data.frame %>% 
        setNames('mhe_mle_avg') 
      )

msc_qcd_anotd_stgstrqc[3:4] <- list(msc_qcd_anotd_stgstrqc[3:4], mhe_mle_avg_exp) %>%
  pmap(~AddMetaData(..1, ..2))
```

```{r, results='asis', message=FALSE}

mhe_mle_umap <-  msc_qcd_anotd_stgstrqc[3:4]  %>% 
      purrr::imap(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        #p<-do_FeaturePlot(., features = "lasonder_avg", dims = c(1,2),  label.size = 4,label = T,repel = T,label.color = "black") + 
        p<-FeaturePlot(., features = "mhe_mle_avg", dims = c(1,2),label = T,repel = T, label.size = 5) +
          scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral")))+
              theme_classic() +
              theme(#axis.text=element_text(size=18), 
                    #axis.title=element_text(size=18,face="bold"), 
                    legend.text = element_text(size=20),
                    legend.position="bottom",
                    legend.title =element_blank(),
                    plot.title = element_blank(),
                    axis.line = element_line(arrow = arrow(angle = 20, length = unit(0.15, "inches"), ends = "last", type = "closed")),
        axis.title = element_text(hjust = 0))+
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL)+
              # coord_fixed()+
        guides(color = guide_colorbar(barwidth = 10, barheight =0.6),
         x= trunc_axis,
         y= trunc_axis)
        return(p)
        # print(p)
        cat('\n')
        cat("\n")
      })


mhe_mle_umap
```

```{r, results='asis', message=FALSE}
barp_theme <- theme(axis.text.x = element_blank(),
        axis.title=element_text(size=16,face="plain"), 
        axis.text = element_text(size=15), 
        axis.ticks.x = element_blank(),
        strip.text =element_text(size=16,face="plain", colour = 'black'),
        panel.border=element_rect(colour="black",fill='transparent'),
        legend.title = element_text(size=16,face="bold"),
        legend.text = element_text(size=15),
        legend.position = 'bottom')

ffl_core_gns_barp <- msc_qcd_anotd_stgstrqc[2:4]  %>% 
  map(., ~.x@meta.data)%>% 
  bind_rows(. , .id = "donor") %>%
  mutate(across(donor, ~droplevels(factor(., levels = sample_id_lst)))) %>%
  ggplot(aes(x=stageHL, y = fhe_fle_avg)) +
  geom_boxplot()+
  geom_sina(size = 0.2) +
  facet_wrap(. ~ donor, scales = "free_x")+
  geom_tile(aes(x = stageHL, y = -1, fill = stageHL, height = 0.9)) +
  scale_fill_manual(values = sp_fld_colors)+
  labs(x = "Stage", y = "F+FL core \ngenes avg. \nexpression")+
  theme_classic() +
  barp_theme +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
      
mml_core_gns_barp <- msc_qcd_anotd_stgstrqc[3:4]  %>% 
  map(., ~.x@meta.data)%>% 
  bind_rows(. , .id = "donor") %>%
  mutate(across(donor, ~droplevels(factor(., levels = sample_id_lst)))) %>%
  ggplot(aes(x=stageHL, y = mhe_mle_avg)) +
  geom_boxplot()+
  geom_sina(size = 0.2) +
  facet_wrap(. ~ donor, scales = "free_x") +
  geom_tile(aes(x = stageHL, y = -0.5, fill = stageHL, height = 0.4)) +
  scale_fill_manual(values = sp_fld_colors)+
  labs(x = "Stage", y = "M+ML core \ngenes avg. \nexpression")+
  theme_classic()+
  barp_theme +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
  
      
ffl_core_gns_barp
mml_core_gns_barp
```

```{r, results='asis', message=FALSE}
msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(str_detect(stageHL, 'Female')) %>% rownames_to_column('gene_id') %>% mutate(across(gene_id, factor)) %>% ggplot(aes(x = stageHL, y = fhe_fle_avg)) + geom_boxplot() + geom_line(aes( group = gene_id))+ geom_point(aes(group=gene_id))
```


```{r}
fhe_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'data', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female') %>% rownames()], na.rm = TRUE)
fle_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'data', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames()], na.rm = TRUE)

list(fhe_avg, fle_avg) %>% set_names(str_remove(c('fhe_avg', 'fle_avg'), '_avg')) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id') %>% ggplot(aes(x=fle, y=fhe)) + geom_point() + labs(x = 'Female (LE)', y = 'Female') + geom_abline() + theme_classic() + theme(axis.text = element_text(size = 15), axis.title = element_text(size = 16)) 

```

```{r}
fhe_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'counts', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female') %>% rownames()], na.rm = TRUE)
fle_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'counts', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames()], na.rm = TRUE)

list(fhe_avg, fle_avg) %>% set_names(str_remove(c('fhe_avg', 'fle_avg'), '_avg')) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id') %>% ggplot(aes(x=fle, y=fhe)) + geom_point() + labs(x = 'Female (LE)', y = 'Female') + geom_abline() + theme_classic() + theme(axis.text = element_text(size = 15), axis.title = element_text(size = 16)) 

```



```{r}
fhe_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'data', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female') %>% rownames()], na.rm = TRUE)
fle_avg <- rowMeans(as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'data', assay = 'RNA'))[fhe_fle_gns, msc_qcd_anotd_stgstrqc[[4]]@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames()], na.rm = TRUE)

list(fhe_avg, fle_avg) %>% set_names(str_remove(c('fhe_avg', 'fle_avg'), '_avg')) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id') %>% ggplot(aes(x=fle, y=fhe)) + geom_point() + labs(x = 'Female (LE)', y = 'Female') + geom_abline() + theme_classic() + theme(axis.text = element_text(size = 15), axis.title = element_text(size = 16)) 

# list(fhe_avg, fle_avg) bind_cols(fhe_avg, fle_avg) 

# rowMeans(x = as.matrix(GetAssayData(msc_qcd_anotd_stgstrqc[[4]], slot = 'data', assay = assay))[gene.set, ], na.rm = TRUE)
```



```{r}
##Average expression of a gene in each stage
##stage variables
# stgs <- c('Late ring','Early trophozoite' ,'Female', 'Female (LE)', 'Male', 'Male (LE)')
stg_key <- c('Late ring' = 'LT', 'Early trophozoite' = 'ET' , 'Female' = 'F', 'Female (LE)' = 'FLE', 'Male' = 'M', 'Male (LE)' = 'MLE', 'Gametocyte (female)' = 'GF', 'Gametocyte (male)' = 'GM')

##get avg expression
stg_expr_cellProp_lst <- pmap(list(msc_qcd_anotd_stgstrqc), ~{
  dset <- ..1
  stg <- as.character(unique(dset@meta.data$stageHL_ref))
  print(stg)
  stg_nms <- stg_key[stg]
  print(stg_nms)
    
    ex <- map(stg, ~rowMeans(as.matrix(GetAssayData(dset, slot = 'data', assay = 'RNA'))[, dset@meta.data %>% filter(stageHL_ref == .x) %>% rownames()], na.rm = TRUE)) %>% 
    set_names(paste0(stg_nms, "_exp")) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id')
    
    cp <- map(stg, ~rowSums(as.matrix(GetAssayData(dset, slot = 'data', assay = 'RNA'))[, dset@meta.data %>% filter(stageHL_ref == .x) %>% rownames()] > 0)/ length(dset@meta.data %>% filter(stageHL_ref == .x) %>% rownames())) %>% 
    set_names(paste0(stg_nms, "_celProp")) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id')
    
    ex_cp_df <- left_join(ex, cp, by = 'gene_id') %>% mutate(lasonder = ifelse(gene_id %in% tr_gns, 'y', 'n'))
    
  }) 

```



```{r}
## Average expression of a gene in each strain stage (AFM) combination
ss_stg_expr_cellProp_lst <- pmap(list(msc_qcd_anotd_stgstrqc), ~{
    dset <- ..1
    strn_stg <- dset@meta.data %>% dplyr::count(Strain_DN, stage_afm) %>% dplyr::select(-c(n)) %>% as.list()
    strn_stg_nms <- pmap(strn_stg, ~paste(..1, ..2, sep = '_'))
  
    ##get cells of interest
    ss_bcodes <- pmap(strn_stg, ~{dset@meta.data %>% filter(Strain_DN == ..1 & stage_afm == ..2) %>% rownames()})

    # print(strn_stg_nms)
    
    ##get avg expression
    ss_exp <- map(ss_bcodes, ~{GetAssayData(dset, slot = 'data', assay = 'RNA')[, .x] %>% as.matrix() %>% rowMeans(., na.rm = TRUE)}) %>% 
      set_names(paste0(strn_stg_nms, '_exp')) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id')
    
    ##Proportion of cells in each stage expressing a gene
    ss_prop_w_gn <-map(ss_bcodes, ~{rowSums(as.matrix(GetAssayData(dset, slot = 'data', assay = 'RNA')[, .x])> 0)/length(.x)}) %>% 
      set_names(paste0(strn_stg_nms, '_celProp')) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id')
    
    ##Number instead of proportion
    # ss_num_w_gn <-map(ss_bcodes, ~{rowSums(as.matrix(GetAssayData(dset, slot = 'data', assay = 'RNA')[, .x])> 0)}) %>% 
    #   set_names(strn_stg_nms) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols  %>% rownames_to_column('gene_id')
    
    
    ##Merge
    ss_stg_expr_cellProp <- left_join(ss_exp, ss_prop_w_gn, by = 'gene_id') %>% mutate(lasonder = ifelse(gene_id %in% tr_gns, 'y', 'n'))

  }) 

```


```{r}
##gametocyte commiited rings genes paper https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7710746/
Pf3D7_genes_plasmoDB %>% filter(Gene %in% str_to_upper(c('ap2-g', 'hda', 'arom', 'lsd2', 'rga', 'PF3D7-0406500', 'ep1', 'PF3D7-0113400', 'PF3D7-1252500', 'surf13.1', 'surf1.2', 'surf4.2', 'Rfge7', 'gexp5', 'vap1', 'tra3', 'msrp1', 'cpp', 'nat', 'sbp1', 'kahrp'))) %>% pull(Gene)
```

## GO and GSEA analysis of DE genes

### GO enrichment function for gene table from plasmoDb
Term2gene and term2name files generated manually in dplyr. Allows the computation of GO enrichment for 'function', 'component' and 'processes' separately generating results for each

GO table preparation from plasmodb gene table 


```{r}

go_tbl_prep <- function(go_term = 'Function', sufx = 's'){
  go_tbl <- Pf3D7_genes_plasmoDB_go[c('gene_id', paste0(c(rep('Curated ', 2), rep('Computed ',2)), rep('GO ', 4),go_term, c(sufx, ' '), c('','IDs')))] %>% 
    pivot_longer(-c('gene_id'),
                   names_to = c( "method_cc",".value"),
                   names_pattern = "(C.*d) (G.*)")  %>% 
    rename_all(., ~str_remove(., 'GO ')) %>% 
    separate_rows(., c(contains('IDs'), contains(paste0(go_term, sufx))), sep=';') %>%
    distinct(., .keep_all = T) %>% 
    filter(!if_all(contains(paste0(go_term, sufx)), ~is.na(.))) %>% 
    dplyr::rename("Gene"= gene_id, "TERM" = paste0(go_term, sufx), "GO" = paste0(go_term, ' IDs'))
  
  go_gaf <- list('TERM2GENE' = go_tbl %>% dplyr::select(GO,Gene) %>% distinct(., .keep_all = T), 'TERM2NAME'= go_tbl %>% dplyr::select(GO,TERM) %>% dplyr::rename("GOID"= GO) %>% distinct(., .keep_all = T)) 

}


```


GO enrichment function - calculates GO overrepresentation for upregulated and downregulated genes and also GSEA using lfc
```{r}

go_e_fun  <- function(dset_lst, dset_lst_full, t2gn_lst, lst_names, unvrs){
  
  field_overGO <- dset_lst %>% 
    filter(avg_log2FC>0 ) %>%
    pull(gene_id) %>%  
    enricher(., TERM2GENE = t2gn_lst$TERM2GENE, TERM2NAME=t2gn_lst$TERM2NAME, universe = unvrs)
  
  lab_overGO <- dset_lst %>% 
    filter(avg_log2FC<0 ) %>%
    pull(gene_id) %>% 
    enricher(., TERM2GENE = t2gn_lst$TERM2GENE, TERM2NAME=t2gn_lst$TERM2NAME, universe = unvrs)
  
  f_vs_l <- dset_lst_full %>% 
    dplyr::select(gene_id,avg_log2FC) %>% 
    filter(!is.na(avg_log2FC)) %>%
    arrange(desc(avg_log2FC)) %>%
    # rownames_to_column() %>% 
    # filter(rowname %in% go_tbl$gene_id) %>% 
    deframe()
  
  print(str(f_vs_l))
  
  ## Filtering out genes based on Pvalue may reduce power of the analysis https://www.biostars.org/p/247647/ & https://resources.qiagenbioinformatics.com/manuals/clcgenomicsworkbench/650/Gene_set_enrichment_analysis.html
  gsea = GSEA(f_vs_l , TERM2GENE = t2gn_lst$TERM2GENE, TERM2NAME=t2gn_lst$TERM2NAME, pvalueCutoff = 0.05)
  
  go_e_lst <- list(field_overGO, lab_overGO, gsea) %>% set_names(lst_names)
  return(go_e_lst)

}

```

GO enrichment function for set of genes and universe - same as above but only for over-representation 
```{r}
go_e_fun_overrep_only  <- function(gns, unvrs, t2gn_lst){
  
  overGO <- gns %>%  
    enricher(., TERM2GENE = t2gn_lst$TERM2GENE, TERM2NAME=t2gn_lst$TERM2NAME, universe = unvrs)
  
    return(overGO)

}

```

```{r}
go_e_fun_overrep_only_fcp  <- function(gns, unvrs, t2gn_lst){
 
  map(list(Pf3D7_66_gaf_F, Pf3D7_66_gaf_C, Pf3D7_66_gaf_P), ~enricher(gns, TERM2GENE = .x$TERM2GENE, TERM2NAME=.x$TERM2NAME, universe = unvrs) %>% .@result %>% filter(p.adjust<0.05)) %>%
            set_names(c('Molecular function', 'Cellular component', 'Biological process')) %>%
            bind_rows(., .id='Ontology')
  

}



```


#### Function GO enrichment {.tabset}

```{r}
field_lab_exprsn <- map(msc_g_noLE_v3_combi_strn_mast[1:8], ~rownames(.)[(rowSums(.@assays$RNA@counts) > 0)])

```

```{r, message=FALSE, warning=FALSE}
## !!!!! NB WRONG - NEED TO RERUN FULL MAST DE IN THE 'msc_FvL_strain_DE ....' notbook
# strns_ub_pairwise_mkrs_pt_ml_gsea_full<- readRDS('data/processed/msc14/5736STDY11771546_DB52e_star/strns_ub_pairwise_mkrs_pt_ml_gsea_full.RDS')
strns_ub_pairwise_mkrs_pt_ml <- readRDS('data/processed/DB52e_star/strns_ub_pairwise_mkrs_pt_ml.RDS')
strns_ub_pairwise_mkrs_pt_ml_gsea_full <- strns_ub_pairwise_mkrs_pt_ml
```

```{r, message=FALSE}
go_tbl_functions <- go_tbl_prep(go_term = 'Function', sufx = 's')

go_function = list(mast_tde_corr_tbls[1:8]  %>% map(.  %>% filter(!is.na(avg_log2FC) & p_val_adj<0.05 & waldStat >10)),
                   strns_ub_pairwise_mkrs_pt_ml_gsea_full[1:8],
                     field_lab_exprsn) %>%
  pmap(., ~go_e_fun(dset_lst = ..1, dset_lst_full = ..2, t2gn_lst= go_tbl_functions, lst_names = c('field_overGO', 'lab_overGO', 'fieldVSlab_gsea'), unvrs = ..3))

```



```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(go_function)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        tryCatch(
          print(enrichplot::dotplot(.x)), 
          error = function(e){function() 'No overepresented GO terms'})
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```


#### Components GO enrichment {.tabset}
```{r, message=FALSE}
go_tbl_components <- go_tbl_prep(go_term = 'Component', sufx = 's')

go_components = list(mast_tde_corr_tbls[1:8]  %>% map(.  %>% filter(!is.na(avg_log2FC)& p_val_adj<0.05 & waldStat >10)),
                     strns_ub_pairwise_mkrs_pt_ml_gsea_full[1:8],
                     field_lab_exprsn) %>%
  pmap(., ~go_e_fun(dset_lst = ..1,dset_lst_full = ..2, t2gn_lst= go_tbl_components, lst_names = c('field_overGO', 'lab_overGO', 'fieldVSlab_gsea'), unvrs = ..3))


```

```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(go_components)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('##### ',.y,'   \n')
        tryCatch(print(enrichplot::dotplot(.x)), error = function(e){function() 'No overepresented GO terms'})
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```


### GO enrichment with GO annotation file (gaf) from plasmoDB. 
Groups the computation of GO enrichment for 'function', 'component' and 'processes' together generating only one result dataframe for all the terms.

#### P.falciparum annotations
Read and preprocess .gaf file
```{r, message=FALSE, eval=FALSE}

##!!!NOTE - Download .gaf file from plasmoDB (https://plasmodb.org/common/downloads/Current_Release/Pfalciparum3D7/gaf/ ) rename to end with .gz and then use gunzip to decompress it

## Instrucions on using clusterprofiler with gaf file provided on plasmoDB https://support.bioconductor.org/p/p134523/#p134554 
## parse_gff function https://github.com/YuLab-SMU/clusterProfiler/issues/421#issuecomment-1019300572

# Pf3D7_63_gaf <- clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-63_Pfalciparum3D7_GO.gaf")
# Pf3D7_63_gaf[1] = Pf3D7_63_gaf[1] %>% map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))
# 
# saveRDS(Pf3D7_63_gaf, "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_63_gaf.rds")

Pf3D7_66_gaf <- clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-66_Pfalciparum3D7_GO.gaf")
Pf3D7_66_gaf[1] = Pf3D7_66_gaf[1] %>% map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))

saveRDS(Pf3D7_66_gaf, "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf.rds")
```

```{r, message=FALSE}
## Instrucions on using clusterprofiler with gaf file provided on plasmoDB https://support.bioconductor.org/p/p134523/#p134554 
## parse_gff function https://github.com/YuLab-SMU/clusterProfiler/issues/421#issuecomment-1019300572

Pf3D7_63_gaf <- readRDS("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf.rds")
```

```{r, message=FALSE, eval=FALSE}
## Instrucions on using clusterprofiler with gaf file provided on plasmoDB https://support.bioconductor.org/p/p134523/#p134554 
## parse_gff function https://github.com/YuLab-SMU/clusterProfiler/issues/421#issuecomment-1019300572
## GAF column names can be found here
## GAF file split into the different aspects using awk before parsing since parse_gff doesn't separate aspects - awk -F"\t" 'NR == 1 || $9 == "F" { print $0 }' PlasmoDB-63_Pfalciparum3D7_GO.gaf > PlasmoDB-63_Pfalciparum3D7_GO_F.gaf

## !! NB - Not much information on converting .gaf file to format needed for GSEA analysis: best function seems to be clusterProfiler::parse_gff https://github.com/YuLab-SMU/clusterProfiler/issues/421 - It adds more GOIDs which may be child GOs

Pf3D7_66_gaf_F <-  clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-66_Pfalciparum3D7_GO_F.gaf")
Pf3D7_66_gaf_F[1] = Pf3D7_66_gaf_F[1] %>% map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))

saveRDS(Pf3D7_66_gaf_F, "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_F.rds")

Pf3D7_66_gaf_P <-  clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-66_Pfalciparum3D7_GO_P.gaf")
Pf3D7_66_gaf_P[1] = Pf3D7_66_gaf_P[1] %>% map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))

saveRDS(Pf3D7_66_gaf_P, "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_P.rds")

Pf3D7_66_gaf_C <-  clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-66_Pfalciparum3D7_GO_C.gaf")
Pf3D7_66_gaf_C[1] = Pf3D7_66_gaf_C[1] %>% map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))

saveRDS(Pf3D7_66_gaf_C, "/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_C.rds")
```

```{r, message=FALSE}
## Instrucions on using clusterprofiler with gaf file provided on plasmoDB https://support.bioconductor.org/p/p134523/#p134554 
## parse_gff function https://github.com/YuLab-SMU/clusterProfiler/issues/421#issuecomment-1019300572

Pf3D7_66_gaf_F <- readRDS("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_F.rds")
Pf3D7_66_gaf_P <- readRDS("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_P.rds")
Pf3D7_66_gaf_C <- readRDS("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3D7_66_gaf_C.rds")

Pf3D7_63_gaf_txt <- read_delim("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-66_Pfalciparum3D7_GO.gaf", skip = 1, delim = '\t', col_types = paste(rep('c', 17), collapse = ''), col_names = paste0('X', (1:17)))

Pf3D7_63_gaf_terms_cat <- Pf3D7_63_gaf_txt %>% filter(!is.na(X9)) %>% distinct(pick(X5, X9)) %>% rename('go_id' = X5, 'fpc_class' = X9) 
```


##### Field VS Lab GO overrepresentation test & GSEA {.tabset}
Consider this for GSEA https://github.com/satijalab/seurat/issues/532 
```{r, message=FALSE}
go_std_pdb = list(mast_tde_corr_tbls[1:8]  %>% map(.  %>% filter(!is.na(avg_log2FC) & p_val_adj<0.05 & waldStat >10)),
                  strns_ub_pairwise_mkrs_pt_ml_gsea_full[1:8],
                     field_lab_exprsn) %>%
  pmap(., ~go_e_fun(dset_lst = ..1,dset_lst_full = ..2, t2gn_lst= Pf3D7_63_gaf, lst_names = c('field_overGO', 'lab_overGO', 'fieldVSlab_gsea'), unvrs = ..3))
# go_processes = go_e_fun(go_term = 'Process', sufx = 'es')
```


```{r, results='asis', fig.width=9, fig.height=6.5}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(go_std_pdb)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        tryCatch(
          print(enrichplot::dotplot(.x, font.size=20) + 
                  theme_classic() +
                  # https://stackoverflow.com/questions/9690648/avoid-clipping-of-points-along-axis-in-ggplot
                  coord_cartesian(clip = 'off') +
                  labs(title = .y)+
                  theme(legend.title=element_text(size=20,face="bold"), 
                        legend.text=element_text(size=20),
                        axis.text.x = element_text(size=20, angle = 90, vjust = 0.5, hjust=1),
                        axis.text.y = element_text(size=20, face = "bold"),
                        axis.title = element_text(size=20, face="bold"),
                        legend.key = element_rect(colour = "transparent", fill = "transparent")
                        )), 
          error = function(e){function() 'No overepresented GO terms'})
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```

```{r, fig.width=6, fig.height=3}
# str_wrap(df$category, width = 30)

fvl_gsea_plts <- unlist(go_std_pdb)[grepl('gsea', names(unlist(go_std_pdb)))]  %>% 
      purrr::imap(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        plt <- tryCatch(
          .x@result  %>% filter(p.adjust < 0.05) %>% arrange(p.adjust) %>% group_by(NES<0) %>% 
            slice_head(., n = 5) %>%
              mutate(go_labs = str_wrap(Description, width = 30)) %>%
            ggplot(aes(reorder(go_labs, NES), NES)) +
            geom_col(aes(fill = setSize/100), width = 0.8, position = position_dodge(width = 0.2), ) +
            coord_flip() +
            labs(x="GO terms", y="NES",title = .y) +
              theme_classic()+ 
                  theme(legend.title=element_text(size=12,face="bold"), 
                        legend.text=element_text(size=11),
                        axis.text.x = element_text(size=14, angle = 90, vjust = 0.5, hjust=1),
                        axis.text.y = element_text(size=14, face = "bold", lineheight = 0.8),
                        axis.title = element_text(size=15, face="bold"),
                        legend.key = element_rect(colour = "transparent", fill = "transparent")
                        ), 
          error = function(e){function() 'No overepresented GO terms'})
        return(plt)
        cat('\n \n \n ')
        cat("\n\n\n")
      })

fvl_gsea_plts
```

```{r}
createWorksheet <- function(x, y){
  wb <- createWorkbook()
  for(i in seq_along(x)){
    addWorksheet(wb, names(x)[i])
    writeData(wb, sheet = names(x)[i], x[[i]])
    saveWorkbook(wb, file = y, overwrite = TRUE)
  }
}
```

Put together GO tables together for manuscript
```{r}
##!!!! NOTE - MANUSCRIPT field vs lab DE and GSEA table

## Extract the asx, fh, fl and m GSEA relevant files to export
fvl_gsea_std_pdb <- unlist(go_std_pdb)[grepl('gsea', names(unlist(go_std_pdb)))][c(1,3:5)]

##Prepare names for worksheet
names(fvl_gsea_std_pdb) <- str_replace_all(str_remove_all(names(fvl_gsea_std_pdb), 'FieldVlab_|.fieldVSlab'), c('asx_'='Asexual ', 'fh_'='Female ', 'fl_'='Female (LE) ', 'mh_'='Male ')) 

##Read the DE results
fvl_n_powered_mast_tde_de_export <- readRDS("../Mali/v3_science_manuscript/tables/fvl_n_powered_mast_tde_de_export.RDS")

##Write both the DE and GO results
createWorksheet(c(fvl_n_powered_mast_tde_de_export[str_detect(names(fvl_n_powered_mast_tde_de_export), "_FvL")],fvl_gsea_std_pdb), "../Mali/v3_science_manuscript/tables/data S12 - Field versus lab MAST & tradeseq DE & GSEA.xlsx") 

```

Read the objects for manuscript image
```{r, fig.width= 12, fig.height=17}
##read in RDS object with heatmaps and density plot generated in the msc_field_lab_strains_DE.Rmd to combine with the gsea barplot generated here
all_labvfld_hms_fvl <- readRDS('plots/DE_heatmaps/all_labvfld_hms_fvl.RDS')
fvl_denst_plt <- readRDS('plots/DE_heatmaps/fvl_denst_plt.RDS')
```

Read the objects for manuscript image
```{r, fig.width= 8.5, fig.height=13}
# fvl_de <- pmap(list(all_labvfld_hms_fvl[c(1,3:5)], fvl_denst_plt[c(1,3:5)], fvl_gsea_plts[c(1,3:5)]), ~{
#   HM_gb = grid.grabExpr(draw(..1))
#   
#   cowplot::plot_grid( HM_gb,
#                       plot_grid(..2+
#                                   theme(axis.text=element_text(size=14),
#                                         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
#                                         # axis.title.y = element_blank(),
#                                         axis.title.y=element_text(size=14,face="bold"),
#                                         axis.title.x=element_text(size=14,face="bold"), 
#                                         legend.text = element_text(size = 13.5) ,
#                                         legend.title = element_text(size = 14, face = "bold"),
#                                         legend.position="bottom",
#                                         plot.background = element_rect(color = "lightgrey"),
#                                         plot.margin=margin(t=2,r=0.1,b=2.8,l=0.1,unit="cm"),
#                                         legend.box.margin = margin(t = 0, r = 0.8, b = 0, l = 0, unit = "cm")
#                                   )+ 
#                       guides(color = guide_legend(override.aes = list(size = 4), ncol = 2, title.position = "top")), 
#                                 ..3+
#                          # scale_x_discrete(labels = function(l) str_wrap(l, width = 15)) +
#                           scale_fill_continuous(name = 'Size') +
#                                   theme(axis.text.x = element_text(size=12, angle = 90, vjust = 0.5, hjust=1),
#                                         axis.text.y = element_text(size=15.5, face='plain', lineheight = 0.8),
#                                         axis.title=element_text(size=13,face="bold"), 
#                                         # axis.title.y=element_blank(), 
#                                         legend.text = element_text(size = 13) ,
#                                         legend.title = element_text(size = 13, face = "bold"),
#                                         legend.position="bottom",
#                                         legend.key.height = unit(0.2, "cm"),
#                                         legend.margin = margin(0),
#                                         plot.background = element_rect(color = "lightgrey"),
#                                         plot.title = element_blank(),
#                                         plot.margin=margin(t=0.1,r=0.1,b=0.8,l=0.1,unit="cm"),
#                                         legend.box.margin = margin(t = 0, r = 3.0, b = 0, l = 0, unit = "cm")
#                                   ) +
#                         guides(fill = guide_legend()),
#                                 ncol=2, 
#                                 rel_widths = c(.3,.7)
#                       )  + 
#                         panel_border(),
#                       ncol=1, 
#                       rel_heights = c(.42,.58)
#   )+ panel_border()
# }
# 
# )
```


```{r, fig.width= 12, fig.height=16}
# plot_grid(plotlist = fvl_de, nrow = 2, ncol = 2, labels = c('A','B','C','D'), label_size = 24, vjust = 1.1)
```           



##### Female low vs female high GO overrepresentation test & GSEA {.tabset}

```{r}
##Get the universe - all genes with average expression greater than 0 across all 'Female' & 'Female (LE)' cells
fh_FLE_expresn <- msc_qcd_anotd_stgstrqc[2:4] %>% 
  map(., . %>% subset(., subset =stageHL %in% c('Female', 'Female (LE)')) %>% 
  AverageExpression(., assays = 'RNA', group.by = 'stageHL') %>% 
  data.frame() %>% 
  filter(if_any(everything(), ~. >0)) %>% 
  rownames())



```

```{r, message=FALSE}
fh_fl_go_std_pdb <- list(fh_fl_marker_list_acc_strn_3.13.14 %>% map(.  %>% filter(!is.na(avg_log2FC) & p_val_adj<0.05)),
                         fh_fl_marker_list_acc_strn_3.13.14,
                     fh_FLE_expresn) %>%
  pmap(., ~go_e_fun(dset_lst = ..1,dset_lst_full = ..2, t2gn_lst= Pf3D7_66_gaf_F, lst_names = c('fh_overGO', 'fl_overGO', 'fhVSfl_gsea'), unvrs = ..3))


```

```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(fh_fl_go_std_pdb)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        tryCatch(print(enrichplot::dotplot(.x, font.size=18)), error = function(e){function() 'No overepresented/gsea GO terms'})
        tryCatch(print(cnetplot(.x, showCategory=10)), error = function(e){function() 'No overepresented/gsea GO terms'})
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```

##### Female low vs female high pairwise {.tabset}

```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(fh_fl_go_std_pdb)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        fh_fl_pwt = tryCatch(pairwise_termsim(.x), error = function(e){function() 'No overepresented/gsea GO terms'})
        tryCatch(print(emapplot( fh_fl_pwt, node_scale=1.5,layout="kk")), error = function(e){function() 'No overepresented/gsea GO terms'})
        # print(upsetplot(fh_fl_pwt))
        # p1 <- treeplot(fh_fl_pwt)
        # p2 <- treeplot(fh_fl_pwt, hclust_method = "average")
        # print(aplot::plot_list(p1, p2, tag_levels='A'))
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```

##### Female LE, HE, Male LE, HE absence/present genes (detected in 25% of the population) overrepresentation test {.tabset}

```{r}
## Copy the gene lists to make do GO overrepresentation test
# mfet_gns_3.13.14_don_colps<- readRDS("data/processed/msc14/5736STDY11771546_DB52e_star/mf_le_he/mfet_gns_3.13.14_don_colps.RDS")

```



```{r}
##Get the universe - all genes with average expression greater than 0 across all msc14 cells
m_f_expresn <- msc_qcd_anotd_stgstrqc[2:4] %>% 
  map(., . %>% AverageExpression(., assays = 'RNA', group.by = 'stageHL') %>% 
  data.frame() %>% 
  filter(if_any(everything(), ~. >0)) %>% 
  rownames())

m_f_expresn <- list(unique(unlist(m_f_expresn)))



```

```{r, message=FALSE}
##Different classes of genes
f_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(is.na(MLE_genes) & is.na(M_genes) & FLE_genes == 'y' & F_genes == 'y' & is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

m_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(is.na(FLE_genes) & is.na(F_genes) & MLE_genes == 'y' & M_genes == 'y' & is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1])%>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})


##Present in fh or mh and absent in all the others
##fh
fh_only_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(is.na(MLE_genes) & is.na(M_genes) & is.na(FLE_genes) & F_genes == 'y' & is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##mh
mh_only_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(is.na(FLE_genes) & is.na(F_genes) & is.na(MLE_genes)  & M_genes == 'y' & is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##Present in fl, ml and either present or absent the others - This is because these cells express very few genes
##fl
fl_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(FLE_genes == 'y') %>% pull(gene_id)),
                          m_f_expresn[1])  %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##ml
ml_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(MLE_genes == 'y') %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##Present in fl, ml and absent in asexuals but either present or absent in the others - This is because these cells express very few genes
##fl
fl_noasex_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter( FLE_genes == 'y' & is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##ml
ml_noasex_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter( MLE_genes == 'y'& is.na(LR_genes) & is.na(ET_genes)) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})


##Present in all stages
##fl
all_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(if_all(contains('_genes'), ~!is.na(.) )) %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##Present in all gam stages and absent in asexual stages
gam_noasex_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(if_all(matches('[F|M|FLE|MLE]_genes'), ~!is.na(.) )) %>% filter(if_all(matches('.[R|T]_genes'), ~is.na(.) ))  %>% pull(gene_id)),
                          m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

##Present in all asexual stages and absent in gam stages

asex_nogam_abspres_go_std_pdb <- list(list(mfet_gns_3.13.14_don_colps %>% filter(if_all(matches('.[R|T]_genes'), ~!is.na(.) )) %>% filter(if_all(matches('[F|M|FLE|MLE]_genes'), ~is.na(.) ))  %>% pull(gene_id)),
                                      m_f_expresn[1]) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

```


```{r, results='asis', fig.height=4, fig.width=12}
go_bar_fn <-function(go_res  ) {
  
  tbl <- go_res %>% separate_wider_delim(., GeneRatio, '/', names = c('nume', 'denom')) %>% mutate('Gene ratio' =as.numeric(nume)/as.numeric(denom) )%>% mutate(Description = str_wrap(Description, width = 30))  %>% arrange(`Gene ratio`)
  
  tbl %>%
    mutate(across(Description, ~factor(., levels = tbl$Description))) %>% 
    ggplot(aes(x = `Gene ratio`, y=Description, size=Count, color = p.adjust)) + 
    geom_point() + 
    theme_bw()+
    facet_wrap( .~Ontology, scales =  "free_y") +
  theme(legend.box = "horizontal") #+
    # guide(color = guide_legend(override.aes = list(size = 3.5), nrow = 2,title.position = "top"))
  
}
```


```{r, results='asis', fig.height=8, fig.width=18}
# plt1<- list(all_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology)),
#             names(all_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology))),
#             c(40, 22, 10),
#             c(6, 5, 2)) %>%
#   pmap(., ~go_bar_fn(..1, ..2, ..3, ..4)) %>% 
#   plot_grid(plotlist = ., nrow = 1, rel_widths = c(.50,.40,.10))
# plt1
```


```{r, results='asis', fig.height=4, fig.width=12}
go_bar_fn <-function(go_res , ttl, strw_int = 20, b_length = 4, x_digits = 1) {
  
  # tbl <- go_res %>% separate_wider_delim(., GeneRatio, '/', names = c('nume', 'denom')) %>% mutate('Gene ratio' =as.numeric(nume)/as.numeric(denom) )%>% mutate(Description = str_wrap(Description, width = 30))  %>% arrange(`Gene ratio`)
  tbl <- go_res %>% separate_wider_delim(., GeneRatio, '/', names = c('nume', 'denom')) %>% mutate('Gene ratio' =as.numeric(nume)/as.numeric(denom) ) %>% arrange(`Gene ratio`)
  
  tbl %>% 
    # group_by(Ontology) %>%
    # add_count(Ontology, name = 'ont_size') %>%
    # mutate(strw_int = case_when(ont_size<3~60,ont_size>3~30)) %>%
    mutate(across(Description, ~str_replace_all(. , c('-'=' ','oxidoreductase'='oxido reductase')) ))%>%
    mutate(go_labs = str_wrap(Description, width = strw_int)) %>%
    mutate(p=-log10(p.adjust))%>%
    ggplot(aes(y = reorder(go_labs, `Gene ratio`), x=`Gene ratio`, size=Count, color = p.adjust)) +
    geom_point() + 
    scale_x_continuous(labels = function(x) format(x, digits = x_digits), breaks=scales::breaks_pretty(n=3)) +
    scale_color_continuous(labels = function(x) format(x, digits = 2), breaks=scales::breaks_pretty(n=3)) +
    scale_size_continuous(labels = function(x) format(x, digits = 1), breaks=scales::breaks_pretty(n=2)) +
    theme_bw()+
    coord_cartesian(clip = 'off') +
    # facet_wrap( .~Ontology, scales =  "free", space = "free_x") + 
     # ggforce::facet_row(vars(Ontology), scales = 'free', space = 'free') +
    labs(y='Description', title = str_replace(ttl, ' ', '\n'))+
  theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.x = element_text(size = 16),
        plot.title = element_text(size = 16, face = 'bold'),
        legend.position = "bottom",
        legend.box = 'horizontal',
        legend.spacing.x = unit(0.01, "cm"),
        legend.box.margin = margin(t = 0, r = 3.1, b = 0, l = 0, unit = "cm"),
        legend.title = element_text(size = 12, face = 'bold'),
        legend.text = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')#,
        # panel.spacing = unit(0, "lines"),  # Remove panel spacing
        # plot.margin = margin(0, 0, 0, 0, "lines"),  # Remove plot margin
        # aspect.ratio = c(.4,.4,.2)
        )+  # Adjust facet widths) +
    guides(color = guide_colorbar(title.position = "left", order = 2, barwidth=b_length, barheight = 0.3,title = 'P: '),
           size = guide_legend(title.position = "left", order = 1,title = 'Count:'))
  
}
```


```{r}
##Save the above image for plot grid in functional_GO_n_lit_comparison.RMD for manuscript
stg_strn_distb_bar = readRDS("plots/figS26/stg_strn_distb_bar.RDS")
```


Combine mfet_gns_3.13.14_don_colps with expression information for genes

```{r}
## Remove MSC3 Male information from the avg expression and cell proportion table since we are only interested in clusters where we have both LE and canonical
stg_expr_cellProp_lst["MSC3"] <- map(stg_expr_cellProp_lst["MSC3"], ~.x %>% select(-c(contains("M_"))))

##write genes to manuscript
stg_gns_ms <- map2(stg_expr_cellProp_lst[2:4], mfet_gns_3.13.14_lst[names(stg_expr_cellProp_lst[2:4])], ~{.y %>% full_join(., .x, by = 'gene_id')%>%
  left_join(., Pf3D7_genes_plasmoDB[,c("gene_id", "Gene", "Product Description")], by = 'gene_id') %>% dplyr::select(gene_id, ends_with('_genes'), ends_with('_exp'), ends_with('_celProp'),  Gene, `Product Description`)})# %>% 

names(stg_gns_ms) <- paste0(2:4, ".", names(stg_gns_ms), " Core genes")
```

```{r}
##write collapsed core genes that feed into the GO 
comb_gns <- list(mfet_gns_3.13.14_don_colps %>% dplyr::select(gene_id, ends_with('_genes'))) %>% set_names("5.Combined core genes for GO")

```

```{r}
##write genes to manuscript
# stg_gns_ms <- mfet_gns_3.13.14_don_colps %>% dplyr::select(gene_id, F_genes, FLE_genes, M_genes, MLE_genes, LR_genes, ET_genes, Gene, `Product Description`) %>% list() %>% set_names(c('2.Stage genes'))
```

Write the GO terms from above into file
```{r}
##!!!! NOTE - GO stage genes and GO table
##Combine stage genes table and GO results

stg_gns_go <- map(list( fh_only_abspres_go_std_pdb[[1]],f_abspres_go_std_pdb[[1]], mh_only_abspres_go_std_pdb[[1]], m_abspres_go_std_pdb[[1]], asex_nogam_abspres_go_std_pdb[[1]]),
                  ~.x %>% 
                    group_by(Ontology) %>%  
                    arrange(`p.adjust`, .by_group = T)  %>% 
                    ungroup()  %>%
                    rename(c('Gene\nRatio'= GeneRatio))) %>% 
  set_names(c('6.Female core GO', '7.Female (canonical+LE) core GO', '8.Male core GO', '9.Male core (canonical+LE) GO', '10.Asexual core GO'))

##Export the list of field strain_strain DE dfs with appropriate names to workbook with each stage as a sheet
# createWorksheet(c(stg_gns_ms, stg_gns_go, ussr_avg_exprsn_dta), "../Mali/v3_science_manuscript/tables/data S11 - Donor parasite stage genes, GO, FHE vs FLE unspliced_spliced ratios and average expression.xlsx")
createWorksheet(c(stg_gns_ms, comb_gns, stg_gns_go), "../Mali/v3_science_manuscript/tables/data S11 - Donor parasite stage core genes, average expression and GO of core gene sets.xlsx")

```

```{r}
##Print the GO tables using and arrange in the order fh ony, fl+fh only, mh ony, asex only and ml+mh only
go_gt_lst <- map(list( fh_only_abspres_go_std_pdb[[1]],f_abspres_go_std_pdb[[1]], mh_only_abspres_go_std_pdb[[1]], asex_nogam_abspres_go_std_pdb[[1]],m_abspres_go_std_pdb[[1]]),
                 ~.x %>% 
                   group_by(Ontology) %>% 
                   arrange(`p.adjust`) %>% 
                   slice_head(., n=2) %>% 
                   ungroup()  %>% 
                   dplyr::select(c('Ontology', 'Description', 'GeneRatio')) %>% rename(c('Gene\nRatio'= GeneRatio)) %>%  
                   mutate(Description = str_wrap(Description, width = 20)) %>%  
                   gt() %>% 
                   data_color(method = 'factor', columns = 'Ontology', target_columns = c('Description', 'Gene\nRatio'), palette = 'Pastel2') %>% cols_hide('Ontology') %>% 
                   tab_options(table.width = pct(20), column_labels.font.weight = 'bold'))

```


```{r}
f_go_tbl<- f_abspres_go_std_pdb[[1]]  %>% dplyr::select(c('Ontology', 'Description', 'GeneRatio', 'p.adjust')) %>% mutate(across(p.adjust, ~round(., digits = 2))) %>% gt() %>% data_color(method = 'factor', columns = 'Ontology', target_columns = c('Description', 'GeneRatio', 'p.adjust')) %>% cols_hide('Ontology')

m_go_tbl<- m_abspres_go_std_pdb[[1]]  %>% dplyr::select(c('Ontology', 'Description', 'GeneRatio', 'p.adjust')) %>% mutate(across(p.adjust, ~round(., digits = 2))) %>% gt() %>% data_color(method = 'factor', columns = 'Ontology', target_columns = c('Description', 'GeneRatio', 'p.adjust')) %>% cols_hide('Ontology')

```

```{r, results='asis', fig.height=4, fig.width=12}
# plt1<- list(f_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology)),
#             names(f_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology))),
#             c(40, 22, 10),
#             c(6, 5, 2)) %>%
#   pmap(., ~go_bar_fn(..1, ..2, ..3, ..4)) %>% 
#   plot_grid(plotlist = ., nrow = 1, rel_widths = c(.50,.31,.19))

plt1<- list(f_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology)),
            names(f_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology))),
            c(40, 22),
            c(6, 5)) %>%
  pmap(., ~go_bar_fn(..1, ..2, ..3, ..4)) %>% 
  plot_grid(plotlist = ., nrow = 1, rel_widths = c(.50,.31,.19))
plt1
```


```{r, results='asis', fig.height=4, fig.width=10}
plt2<- list(m_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology)),
            names(m_abspres_go_std_pdb[[1]] %>% split(factor(.$Ontology))),
            c(10, 30),
            c(2, 6),
            c(2, 2)
            # c(10, 30, 13),
            # c(2, 6, 3),
            # c(2, 2, 2)
            ) %>% 
  pmap(., ~go_bar_fn(go_res=..1 , ttl=..2, strw_int = ..3, b_length = ..4, x_digits = ..5)) %>% 
  plot_grid(plotlist = ., nrow = 1, rel_widths = c(.25,.47,.28))
plt2
```

```{r}
##Save the above image for plot grid in functional_GO_n_lit_comparison.RMD for manuscript
stg_strn_distb_bar = readRDS("plots/figS26/stg_strn_distb_bar.RDS")
```

```{r, results='asis', fig.height=10, fig.width=10}
fld_gn_expression_specificity_upset_vn_don3 = readRDS("plots/figS26/fld_gn_expression_specificity_upset_vn_don3.RDS")
```


```{r, results='asis', fig.height=10, fig.width=10}
fh_fl_vlcano = readRDS("plots/figS26/fh_fl_vlcano.RDS")
```

```{r, results='asis', fig.height=10, fig.width=10}
fh_fl_gsea_bplot = readRDS("plots/figS26/fh_fl_gsea_bplot.RDS")
```

```{r, results='asis', fig.height=10, fig.width=10}
fhe_fle_ussr = readRDS("plots/figS26/fhe_fle_ussr.RDS")
```

```{r, results='asis', fig.height=10, fig.width=10}
fhe_fle_ussr_wilcx = readRDS("plots/figS26/fhe_fle_ussr_wilcx.RDS")
```



```{r, results='asis', fig.height=4, fig.width=12}
vlcno_gsea_lasonder <- plot_grid(fh_fl_vlcano+
  theme(
    axis.text.y  = element_text(size = 15),
    axis.text.x = element_text(size = 15),#, angle = 45
    axis.title.y  = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 16, face = "bold"),
    # axis.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position="none",
    text = element_text(size = 5),
    plot.caption = element_blank(),
    legend.margin = margin(t=-15),
    # plot.margin = unit(c(1,2,1,1),"cm")
    # legend.margin=margin()
  ), 
                         fh_fl_gsea_bplot +
  theme(
    axis.text.y  = element_text(size = 15),
    axis.text.x = element_text(size = 15),#, angle = 45
    axis.title.y  = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 16, face = "plain"),
    # axis.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 12, face = "bold"),
    legend.position="bottom",
    legend.margin = margin(t=-0.1),
    plot.margin = margin(t=1.5, l=0.5, r=0.5, unit = "cm")
    # legend.margin=margin()
  ),
                         lasonder_avg_umap[[1]]+
  theme(legend.text = element_text(size = 11),
        plot.margin = margin(t=1, unit = "cm")),
                         ncol = 3,
                         rel_widths = c(.33,.41,.26),
              labels = c('D', 'E', 'F'),
              label_size = 20)

vlcno_gsea_lasonder
```



```{r, results='asis', fig.height=16.5, fig.width=12}
##modify patchwork upset plot appropriately

plt= plot_grid(stg_strn_distb_bar[[1]]+
                 theme(axis.title.y = element_text(size = 16, face = 'plain'),
                       axis.text.y = element_text(size = 15),
                       legend.title = element_text(size = 16),
                       legend.text = element_text(size = 15),
                       strip.text =element_text(size=17,face="bold", colour = 'black'),
                       legend.margin = margin(t=-5)),
               ggdraw(fld_gn_expression_specificity_upset_vn_don3)+
                theme(plot.margin = margin(t=0.5, unit = "cm")),
               plt1, 
               plt2,
               vlcno_gsea_lasonder,
               nrow = 5, 
               rel_heights = c(.17,.26,.195,.175,.20),
              labels = c('A', 'B', 'C', 'D', ''),
              label_size = 20
               )
plt
```



```{r, results='asis', fig.height=16.5, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961

gt_grob <- function(gt_object, wth = NULL, hgt = NULL, ...){
  
  out_name <- file.path(
    tempfile(pattern = "file", tmpdir = tempdir(), fileext = ".png")
  )
  gtsave(gt_object, out_name, ...)
  gtsave(gt_object, filename = "obj2.png", path = "plots/")
  in_png <- png::readPNG(out_name)
  on.exit(file.remove(out_name), add=TRUE)
  grid::rasterGrob(in_png, width = wth, height = hgt)
  
}

# f_gt_plt <- gt_grob(f_go_tbl)
# m_gt_plt <- gt_grob(m_go_tbl)
## !!! in case of error "s$close()" check https://stackoverflow.com/questions/73964325/error-in-sclose-attempt-to-apply-non-function-when-calling-gtsave 
all_stg_gtpt <- map(go_gt_lst, ~gt_grob(.x))
all_stg_gtpt
```


```{r, results='asis', fig.height=16, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961

##Get legend for the GO ontologies
onts = c('Biological process', 'Cellular component', 'Molecular function')
ont_lg <- cowplot::get_legend(data.frame(ont = onts) %>% ggplot(aes( y=1, fill=ont)) + scale_fill_manual(values = brewer.pal(n = 3, name = 'Pastel2') %>% set_names(onts)) + geom_bar() + theme(legend.direction = 'horizontal', legend.title = element_text(size = 14, face = 'bold'), legend.text = element_text(size = 14))+ guides(fill = guide_legend(title = 'Ontology:', nrow = 1)))

##PUT plots together for the manuscript
plt= plot_grid(stg_strn_distb_bar[[1]]+
                 theme(axis.title.y = element_text(size = 16, face = 'plain'),
                       axis.text.y = element_text(size = 15),
                       legend.title = element_text(size = 16),
                       legend.text = element_text(size = 15),
                       strip.text =element_text(size=17,face="bold", colour = 'black'),
                       legend.margin = margin(t=-5)),
               NULL,
               ggdraw(fld_gn_expression_specificity_upset_vn_don3)+
                theme(plot.margin = margin(t=0.5, unit = "cm")),
               plot_grid(plotlist = all_stg_gtpt[1:8], ncol = 5),
               ont_lg,
               NULL,
               # us_s_lasonder,
               nrow = 7, 
               rel_heights = c(.17,0.01,.32,.27,0.02,0.01,.20),
              labels = c('A','', 'B', 'C', '', '', ''),
              label_size = 20
               )
plt
save_plot('plots/figS26/final_pt.svg',plt, base_width = 12, base_height = 16.5)
```







```{r, results='asis', fig.height=4, fig.width=12}
## combine plots for LE exploration
core_g_lg <- cowplot::get_legend(mml_core_gns_barp + 
  theme(legend.position = 'bottom') +
    guides(fill=guide_legend(title.position = "left", nrow = 1, title = "Stage")))


fl_ml_mosq_pt <- plot_grid(
  plot_grid(plotlist = map(list(ffl_core_gns_barp, mml_core_gns_barp), ~.x +
                               theme(legend.position = 'none',
                                     # plot.margin = margin(t=1, unit = "cm")
                                     )),
              rel_widths = c(.6,.4)),
  core_g_lg,
    nrow = 2,
    rel_heights = c(.8,.2))

fl_ml_mosq_pt
```

```{r, results='asis', fig.height=5.5, fig.width=12}
## combine plots for LE exploration
go_mosq_fl_pt <- plot_grid(
  plot_grid(plotlist = all_stg_gtpt[1:5], ncol = 5, rel_widths = c(.24,.18,.19,.20,.19)),
  ont_lg,
  NULL,
  fl_ml_mosq_pt,
  nrow = 4,
  rel_heights = c(.55,.02,.03,.40),
  labels = c("D","","","E"),
  label_size = 20
               
)
go_mosq_fl_pt
```

genes in lasonder among gene sets for MANUSCRIPT
```{r}
a= mfet_gns_3.13.14_don_colps %>% filter(fh_fl_only == 'y') %>% pull(gene_id) 
table(a %in% tr_gns)
prop.table(table(a %in% tr_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(fh_only == 'y') %>% pull(gene_id) 
table(a %in% tr_gns)
prop.table(table(a %in% tr_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(mh_ml_only == 'y') %>% pull(gene_id) 
table(a %in% tr_gns)
prop.table(table(a %in% tr_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(mh_only == 'y') %>% pull(gene_id) 
table(a %in% tr_gns)
prop.table(table(a %in% tr_gns))

```


#### P.berghei annotations 
```{r, message=FALSE}
##ortholog file to translate to falciparum IDs
ortho <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/ortho4.csv")

```

genes in kevin hart pycrr4
```{r}
hart_male_tr_gns <- read_xlsx('data/raw/published_lit/Hart_deadenylases_male_gam/DE_due_pyccr4_deltn.xlsx', sheet = 'P adjusted Significant') %>% rename('gene_id'=GeneID...1) %>% left_join(., ortho, by=c('gene_id'='yoelii') ) 

hart_gns <- hart_male_tr_gns%>% filter(!is.na(falciparum)) %>% dplyr::pull(falciparum) %>% str_replace(., '_', '-')

a= mfet_gns_3.13.14_don_colps %>% filter(fh_fl_only == 'y') %>% pull(gene_id) 
table(a %in% hart_gns)
prop.table(table(a %in% hart_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(fh_only == 'y') %>% pull(gene_id) 
table(a %in% hart_gns)
prop.table(table(a %in% hart_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(mh_ml_only == 'y') %>% pull(gene_id) 
table(a %in% hart_gns)
prop.table(table(a %in% hart_gns))

a= mfet_gns_3.13.14_don_colps %>% filter(mh_only == 'y') %>% pull(gene_id) 
table(a %in% hart_gns)
prop.table(table(a %in% hart_gns))

```

```{r, results='asis', fig.height=4, fig.width=12}
tbl <- c(f_abspres_go_std_pdb,m_abspres_go_std_pdb)  %>% set_names('Female', 'Male') %>%  bind_rows(., .id = 'Sex') %>% separate_wider_delim(., GeneRatio, '/', names = c('nume', 'denom')) %>% mutate('Gene ratio' =as.numeric(nume)/as.numeric(denom) ) %>% arrange(`Gene ratio`)
  
  tbl %>% mutate(across(Description, ~factor(., levels = tbl$Description))) %>% 
    ggplot(aes(x = `Gene ratio`, y=Description, size=Count, color = p.adjust)) + 
    geom_point() + 
    theme_bw()+
    facet_grid( Sex~Ontology, scales =  "free", space =  "free")
  
```


## GO for the DE genes

### NF54 VS 7G8

```{r, fig.width = 32, fig.height = 7}

# msc_gam_combi_noLE_de <- readRDS("data/processed/DB52e_star/msc_gam_combi_noLE_de.RDS")
msc_gcombi_noLE_sub_sce <- readRDS("data/processed/DB52e_star/msc_gcombi_noLE_sub_sce.RDS")

```  


```{r}
## get the universe for the GO overrepresentation test ie all the genes detected (average expression greater than 0) in the cell subset that go into the DE
DE_unvrs_gns <- map(msc_gcombi_noLE_sub_sce[c("MSC14_f_don3","MSC3_m_don2","M_NF54_7G8_f", "M_NF54_7G8_m")], ~{
  
  CreateSeuratObject(counts = counts(.x), data = logcounts(.x)) %>% 
  AverageExpression(., assays = 'RNA') %>% 
  data.frame() %>% 
  filter(all > 0) %>%
  rownames()
})


```

```{r}
## Get the field and lab DE genes
DEgns <- map(mast_tde_corr_tbls[c("lf_StrainsDE_MSC14_f_don3","lf_StrainsDE_MSC3_m_don2","lf_StrainsDE_M_NF54_7G8_f", "lf_StrainsDE_M_NF54_7G8_m")], ~{
  .x %>% filter(p_val_adj < 0.05 & waldStat>10) %>% pull(gene_id)
})

```


```{r, message=FALSE}
## GO
DE_go_std_pdb <- list(DEgns, DE_unvrs_gns) %>%
    pmap(., ~{go_e_fun_overrep_only_fcp(gns = ..1, unvrs = ..2)})

```


Write the GO terms from above into file
```{r}
##!!!! NOTE - GO stage genes and GO table
##Combine stage genes table and GO results

stg_gns_go <- map(DE_go_std_pdb[1:2],
                  ~.x %>% 
                    group_by(Ontology) %>%  
                    arrange(`p.adjust`, .by_group = T)  %>% 
                    ungroup()  %>%
                    rename(c('Gene\nRatio'= GeneRatio))) %>% 
  set_names(c('4.MSC14_female DE GO', '5.MSC3_male DE GO'))

```


```{r}
## Rename the lists for friendly tab names
fvl_n_powered_mast_tde_de_export_fld <- fvl_n_powered_mast_tde_de_export[c("MSC14_f_don3", "MSC3_m_don2")]
names(fvl_n_powered_mast_tde_de_export_fld) <- paste0(2:(length(fvl_n_powered_mast_tde_de_export_fld)+1), ".", str_replace_all(names(fvl_n_powered_mast_tde_de_export_fld), c("f_don3" = "female DE", "m_don2" = "male DE")))

fvl_n_powered_mast_tde_de_export_lab <- fvl_n_powered_mast_tde_de_export[c("M_NF54_7G8_f", "M_NF54_7G8_m")]
names(fvl_n_powered_mast_tde_de_export_lab) <- paste0(6:(length(fvl_n_powered_mast_tde_de_export_lab)+5), ".", str_remove(str_replace_all(names(fvl_n_powered_mast_tde_de_export_lab), c("_f" = "_female DE", "_m" = "_male DE")), "M_"))

```


```{r}
##Export the list of field strain_strain DE dfs with appropriate names to workbook with each stage as a sheet
createWorksheet(c(fvl_n_powered_mast_tde_de_export_fld, stg_gns_go, fvl_n_powered_mast_tde_de_export_lab), "../Mali/v3_science_manuscript/tables/data S12 - High power strains MAST & tradeseq DE & GO.xlsx")

```

```{r}
##Print the DE GO tables using and arrange in the order msc14 female and then msc3 male
DE_go_gt_lst <- map(DE_go_std_pdb[1:2],
                 ~.x %>% 
                   group_by(Ontology) %>% 
                   arrange(`p.adjust`) %>% 
                   slice_head(., n=4) %>% 
                   ungroup()  %>% 
                   dplyr::select(c('Ontology', 'Description', 'GeneRatio')) %>% rename(c('Gene\nRatio'= GeneRatio)) %>%  
                   mutate(Description = str_wrap(Description, width = 40)) %>%  
                   gt() %>% 
                   data_color(method = 'factor', columns = 'Ontology', target_columns = c('Description', 'Gene\nRatio'), palette = 'Pastel2') %>% cols_hide('Ontology') %>% 
                   tab_options(table.width = pct(40), column_labels.font.weight = 'bold')
                 )

```


```{r}
##Print the DE GO tables using and arrange in the order msc14 female and then msc3 male
stg_gns_go
DE_go_gt_lst 
```


```{r, results='asis', fig.height=16.5, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961
DE_go_gtpt <- map(DE_go_gt_lst, ~gt_grob(.x))
DE_go_gtpt
```


```{r, results='asis', fig.height=16.5, fig.width=12}
## Save RDS for concatenating to manuscript plot
saveRDS(DE_go_gtpt, "plots/DE_heatmaps/DE_go_gtpt.RDS")
```

```{r}
##Print the DE GO tables separately for each ontology
DE_go_gt_lst_sep_ont <- map(DE_go_std_pdb[1:2],
    ~.x %>% 
        mutate(across(Ontology, ~factor(., levels = c('Biological process', 'Cellular component', 'Molecular function')))) %>%
        group_by(Ontology) %>%
        group_split() %>%
        setNames(sort(unique(.x$Ontology)))) %>% unlist(., recursive =F) 

##!!! NOTE - manually set the NA description to 'obsolete pathogenesis' - this need to be updated either in the plasmoDB gaf or in the clusterprofiler database. Seems to be an obsolete ID as shown here https://www.ebi.ac.uk/QuickGO/term/GO:0009405 
DE_go_gt_lst_sep_ont["lf_StrainsDE_MSC3_m_don2.Biological process"] <- DE_go_gt_lst_sep_ont["lf_StrainsDE_MSC3_m_don2.Biological process"] %>% map(~mutate(.x, across(Description, ~if_else(is.na(.), "obsolete pathogenesis", .))))
      
DE_go_gt_lst_sep_ont_tbl <- map2(DE_go_gt_lst_sep_ont, c(40,10,10,10),
                 ~.x %>% 
                   # group_by(Ontology) %>% 
                   arrange(`p.adjust`) %>% 
                   slice_head(., n=5) %>% 
                   ungroup()  %>% 
                   dplyr::select(c('Ontology', 'Description', 'GeneRatio')) %>% rename(c('Gene\nRatio'= GeneRatio)) %>%  
                   mutate(Description = str_wrap(Description, width = 40)) %>%  
                   gt() %>% 
                   data_color(method = 'factor', columns = 'Ontology', target_columns = c('Description', 'Gene\nRatio'), palette = 'Pastel2') %>% cols_hide('Ontology') %>% 
                   tab_options(table.width = pct(.y), column_labels.font.weight = 'bold')
                 )

DE_go_gt_lst_sep_ont_tbl
```

```{r, results='asis', fig.height=16.5, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961
DE_go_sep_ont_gtpt <- map(DE_go_gt_lst_sep_ont_tbl, ~gt_grob(.x))
saveRDS(DE_go_sep_ont_gtpt, "plots/DE_heatmaps/DE_go_sep_ont_gtpt.RDS")

DE_go_sep_ont_gtpt
```

```{r, results='asis', fig.height=4, fig.width=11}
go_bar_fn(m_abspres_go_std_pdb[[1]], ttl = names(m_abspres_go_std_pdb[1]))
```


```{r, results='asis', fig.height=4, fig.width=11}
c(fh_only_abspres_go_std_pdb, fl_abspres_go_std_pdb, ml_abspres_go_std_pdb, fl_noasex_abspres_go_std_pdb, ml_noasex_abspres_go_std_pdb, all_abspres_go_std_pdb, gam_noasex_abspres_go_std_pdb, asex_nogam_abspres_go_std_pdb) %>% imap(., ~.x %>% filter(p.adjust < 0.05) %>% arrange(p.adjust) %>% slice_head(n = 10) %>% go_bar_fn(., ttl=.y))
# go_bar_fn(fh_only_abspres_go_std_pdb[[1]])
```


```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(fh_fl_go_std_pdb)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        tryCatch(print(enrichplot::dotplot(.x, font.size=18)), error = function(e){function() 'No overepresented/gsea GO terms'})
        tryCatch(print(cnetplot(.x, showCategory=10)), error = function(e){function() 'No overepresented/gsea GO terms'})
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```

##### Female low vs female high pairwise {.tabset}

```{r, results='asis'}
# cat('## Visualize field parasites {.tabset}   \n')
unlist(fh_fl_go_std_pdb)  %>% 
      purrr::iwalk(.,~{
        # create tabset for each group 
        cat('###### ',.y,'   \n')
        fh_fl_pwt = tryCatch(pairwise_termsim(.x), error = function(e){function() 'No overepresented/gsea GO terms'})
        tryCatch(print(emapplot( fh_fl_pwt, node_scale=1.5,layout="kk")), error = function(e){function() 'No overepresented/gsea GO terms'})
        # print(upsetplot(fh_fl_pwt))
        # p1 <- treeplot(fh_fl_pwt)
        # p2 <- treeplot(fh_fl_pwt, hclust_method = "average")
        # print(aplot::plot_list(p1, p2, tag_levels='A'))
        cat('\n \n \n ')
        cat("\n\n\n")
      })

```


```{r, results='asis', fig.height=16, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961

##Get legend for the GO ontologies
onts = c('Biological process', 'Cellular component', 'Molecular function')
ont_lg <- cowplot::get_legend(data.frame(ont = onts) %>% ggplot(aes( y=1, fill=ont)) + scale_fill_manual(values = brewer.pal(n = 3, name = 'Pastel2') %>% set_names(onts)) + geom_bar() + theme(legend.direction = 'horizontal', legend.title = element_text(size = 14, face = 'bold'), legend.text = element_text(size = 14))+ guides(fill = guide_legend(title = 'Ontology:', nrow = 1)))

##Save ontology legend above for use in other plots
saveRDS(ont_lg, "plots/DE_heatmaps/ont_lg.RDS")

##PUT plots together for the manuscript
plt= plot_grid(stg_strn_distb_bar[[1]]+
                 theme(axis.title.y = element_text(size = 16, face = 'plain'),
                       axis.text.y = element_text(size = 15),
                       legend.title = element_text(size = 16),
                       legend.text = element_text(size = 15),
                       strip.text =element_text(size=17,face="bold", colour = 'black'),
                       legend.margin = margin(t=-5)),
               NULL,
               ggdraw(fld_gn_expression_specificity_upset_vn_don3)+
                theme(plot.margin = margin(t=0.5, unit = "cm")),
               plot_grid(plotlist = all_stg_gtpt[1:8], ncol = 5),
               ont_lg,
               NULL,
               # us_s_lasonder,
               nrow = 7, 
               rel_heights = c(.17,0.01,.32,.27,0.02,0.01,.20),
              labels = c('A','', 'B', 'C', '', '', ''),
              label_size = 20
               )
plt
save_plot('plots/figS26/final_pt.svg',plt, base_width = 12, base_height = 16.5)
```

## Phenoplasm database exploration

```{r}
## Read phenoplasm data http://phenoplasm.org/about.php
phenopDB <- read_tsv('data/raw/phenoplasm_output/PhenoPlasmData.tsv')

phenopDB %>% filter(Species == 'P. falciparum 3D7') %>% filter(Stage %in% c('Oocyst', 'Ookinete', 'Sporozoite') & !Phenotype %in% c('No difference'))

```


```{r}
phenopDB_pf_pb_yoel_ortho <- phenopDB  %>% dplyr::select(Species,Gene ,Stage,Phenotype) %>% mutate(across(Phenotype, ~str_replace(., 'Difference from wild-type', 'Different (WT)'))) %>% distinct() %>% group_by(Gene, Stage) %>% mutate(across(c(Phenotype), ~paste(., collapse = ',')))  %>% ungroup() %>% distinct(pick(Gene, Stage, Phenotype), .keep_all = TRUE) %>%  pivot_wider(id_cols = c(Species, Gene), names_from = 'Stage', values_from = 'Phenotype')  %>% left_join(., ortho[,c('falciparum', 'berghei')], by= c('Gene'='berghei')) %>% left_join(., ortho[,c('falciparum', 'yoelii')], by= c('Gene'='yoelii')) %>% mutate(gene_id = coalesce(falciparum.x, falciparum.y, Gene)) %>% filter(str_detect(gene_id, 'PF3D7_')) %>%  mutate(across(gene_id, ~str_replace(., '_', '-')))

# phenopDB_pf_pb_ortho_mosq <- phenopDB  %>% dplyr::select(Species,Gene ,Stage,Phenotype) %>% mutate(across(Phenotype, ~str_replace(., 'Difference from wild-type', 'Different (WT)'))) %>% distinct() %>% group_by(Gene, Stage) %>% mutate(across(c(Phenotype), ~paste(., collapse = ',')))  %>% ungroup() %>% distinct(pick(Gene, Stage, Phenotype), .keep_all = TRUE) %>%  pivot_wider(id_cols = c(Species, Gene), names_from = 'Stage', values_from = 'Phenotype')  %>% left_join(., ortho[,c('falciparum', 'berghei')], by= c('Gene'='berghei')) %>% mutate(gene_id = coalesce(falciparum,  Gene)) %>% filter(str_detect(gene_id, 'PF3D7_')) %>%  mutate(across(gene_id, ~str_replace(., '_', '-'))) %>% dplyr::select(-c(Asexual, Gametocyte, Liver, falciparum))

```


```{r}
## Avg expression + mosquito needed genes + lasonder
phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_lst <- map(stg_expr_cellProp_lst[2:4], ~{phenopDB_pf_pb_yoel_ortho %>% dplyr::select(-c(Asexual, Gametocyte, Liver, falciparum.x,  falciparum.y)) %>% filter(if_any(c('Oocyst', 'Ookinete', 'Sporozoite'), ~!is.na(.) )) %>% left_join(., .x, by=c('gene_id'))  %>% mutate(across(lasonder, ~replace_na(., 'n'))) %>% mutate('Lasonder' = case_when(lasonder == 'y' ~ 'Yes', lasonder == 'n' ~ 'No'))})

  
```

```{r}
phenop_mosq_plts_facet <- map(phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_lst, ~.x %>% pivot_longer(cols = c(Ookinete,Oocyst,Sporozoite), names_to = 'Stage', values_to = 'Phenotype') %>% filter(!is.na(Phenotype) & !Phenotype %in% c('No difference', 'Not tested')) %>% filter(!str_detect(Phenotype, ','))) %>% 
  bind_rows(., .id= 'donor') %>%
  mutate(across(donor, ~droplevels(factor(., levels = sample_id_lst)))) %>%
  ggplot(aes(x= FLE_exp, y = F_exp, color = Lasonder, shape = Phenotype)) + 
  geom_point() + 
  geom_abline() + 
  facet_wrap(. ~ donor)+
  labs(x='Female (LE)', y='Female')  + 
  theme_bw()+
  barp_theme + 
  theme(plot.title = element_text(size=16, hjust = .5, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_line(),
        axis.text.x = element_text(size=15),
        aspect.ratio = 1) +
  guides(shape = guide_legend(title.position = "top"), color=guide_legend(title.position = "top"))



phenop_mosq_plts_facet
```

```{r}
phenop_mosq_plts_lst <- imap(phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_lst, ~.x %>% pivot_longer(cols = c(Ookinete,Oocyst,Sporozoite), names_to = 'Stage', values_to = 'Phenotype') %>% filter(!is.na(Phenotype) & !Phenotype %in% c('No difference', 'Not tested')) %>% filter(!str_detect(Phenotype, ',')) %>% 
                               ggplot(aes(x= FLE_exp, y = F_exp, color = Lasonder, shape = Phenotype)) + 
                               geom_point() + 
                               geom_abline() + 
                               coord_fixed(ratio = 1, xlim = c(0, 4.5), ylim = c(0, 4.5)) +
                              labs(x='Female (LE)', y='Female', title = .y)  + 
                              theme_classic() + 
                              theme(axis.title = element_text(size=15), 
                              axis.text = element_text(size=15), 
                              legend.title = element_text(size=15, face = 'bold'), 
                              legend.text  = element_text(size=15),
                              plot.title = element_text(size=16, hjust = .5, face = "bold"))
)
                             


phenop_mosq_plts_lst
```


```{r}

phenopDB_pf_pb_yoel_ortho_gam_gnexp_lasonder_lst <- map(stg_expr_cellProp_lst[2:4], ~{phenopDB_pf_pb_yoel_ortho %>% dplyr::select(-c(Asexual, Oocyst, Ookinete, Sporozoite, Liver, falciparum.x,  falciparum.y)) %>% filter(if_any(c('Gametocyte'), ~!is.na(.) )) %>% left_join(., .x, by=c('gene_id'))  %>% mutate(across(lasonder, ~replace_na(., 'n'))) %>% mutate('Lasonder' = case_when(lasonder == 'y' ~ 'Yes', lasonder == 'n' ~ 'No'))})
```

```{r}
map(phenopDB_pf_pb_yoel_ortho_gam_gnexp_lasonder_lst, ~.x %>% pivot_longer(cols = c(Gametocyte), names_to = 'Stage', values_to = 'Phenotype') %>% filter(!is.na(Phenotype) & !Phenotype %in% c('No difference', 'Not tested')) %>% filter(!str_detect(Phenotype, ',')) %>% ggplot(aes(x= FLE_exp, y = F_exp, color = Lasonder, shape = Phenotype)) + geom_point() + geom_abline() + coord_fixed(ratio = 1, xlim = c(0, 4.5), ylim = c(0, 4.5)) + labs(x='Female (LE)', y='Female')  + theme_classic() + 
  theme(axis.title = element_text(size=15), 
        axis.text = element_text(size=15), 
        legend.title = element_text(size=15, face = 'bold'), 
        legend.text  = element_text(size=15)))



```


```{r}

pheno_pf_mosq_gns <- map(stg_expr_cellProp_lst[2:4], ~{phenopDB %>% filter(Species == 'P. falciparum 3D7') %>% filter(Stage %in% c('Oocyst', 'Ookinete', 'Sporozoite') & !Phenotype %in% c('No difference', 'Not tested')) %>% mutate(across(Gene, ~str_replace(., '_', '-')), across(Phenotype, ~str_replace(., 'Difference from wild-type', 'Different (WT)'))) %>% left_join(., .x, by=c('Gene'= 'gene_id'))}) 

pheno_pf_pb_mosq_gns <-  map(stg_expr_cellProp_lst[2:4], ~{phenopDB %>% filter(Species %in% c('P. falciparum 3D7', 'P. berghei ANKA')) %>% filter(Stage %in% c('Oocyst', 'Ookinete', 'Sporozoite') & !Phenotype %in% c('No difference','Not tested')) %>% left_join(., ortho %>% select(falciparum, berghei), by=c('Gene' = 'berghei')) %>% mutate(gene_id = coalesce(falciparum, Gene)) %>% mutate(across(gene_id, ~str_replace(., '_', '-')), across(Phenotype, ~str_replace(., 'Difference from wild-type', 'Different (WT)'))) %>% filter(str_detect(gene_id, 'PF3D7')) %>% left_join(., .x, by=c('gene_id'))  %>% mutate(across(lasonder, ~replace_na(., 'n'))) %>% mutate('Lasonder' = case_when(lasonder == 'y' ~ 'Yes', lasonder == 'n' ~ 'No'))}) 

# phenopDB_ortho_mosq %>% left_join(., stg_expr_cellProp, by=c('gene_id'))  %>% mutate(across(lasonder, ~replace_na(., 'n'))) %>% mutate('Lasonder' = case_when(lasonder == 'y' ~ 'Yes', lasonder == 'n' ~ 'No')) %>% pivot_longer(cols = c(Ookinete,Oocyst,Sporozoite), names_to = 'Stage', values_to = 'Phenotype') %>% filter(!is.na(Phenotype))


```

```{r}
# phenop_mosq_plts <- pheno_pf_pb_mosq_gns %>% ggplot(aes(x= FLE_exp, y = F_exp, color = Lasonder, shape = Phenotype)) + geom_point() + geom_abline() + coord_fixed(ratio = 1, xlim = c(0, 4.5), ylim = c(0, 4.5)) + labs(x='Female (LE)', y='Female')  + theme_classic() + 
#   theme(axis.title = element_text(size=15), 
#         axis.text = element_text(size=15), 
#         legend.title = element_text(size=15, face = 'bold'), 
#         legend.text  = element_text(size=15))
# 
# phenop_mosq_plts

# pheno_pf_pb_mosq_gns  %>% ggplot(aes(x= FLE_celProp, y = F_celProp,  color= Lasonder,  shape= Phenotype)) + geom_point() + geom_abline()
```




```{r}

map(pheno_pf_pb_mosq_gns, ~.x %>% mutate(across(lasonder, ~replace_na(., 'y')))   %>% mutate(across(lasonder, ~case_when(is.na(.) ~ 'n', T ~ .))) %>% filter(FLE_exp<0.2 & F_exp<0.25) %>% ggplot(aes(x= FLE_exp, y = F_exp, color = lasonder, shape = Phenotype)) + geom_point() + geom_abline())

map(pheno_pf_pb_mosq_gns, ~.x  %>% mutate(across(lasonder, ~replace_na(., 'y')))   %>% ggplot(aes(x= FLE_celProp, y = F_celProp,  color= lasonder,  shape= Phenotype)) + geom_point() + geom_abline())
```


```{r}

pheno_pf_ag_gns <- map(stg_expr_cellProp_lst[2:4], ~{phenopDB %>% filter(Species == 'P. falciparum 3D7') %>% filter(Stage %in% c('Asexual', 'Gametocyte') & !Phenotype %in% c('No difference')) %>% mutate(across(Gene, ~str_replace(., '_', '-'))) %>% left_join(., .x, by=c('Gene'= 'gene_id')) %>% slice_sample(., n=150)})

pheno_pf_pb_ag_gns <- map(stg_expr_cellProp_lst[2:4], ~{phenopDB %>% filter(Species %in% c('P. falciparum 3D7', 'P. berghei ANKA')) %>% filter(Stage %in% c('Asexual', 'Gametocyte') & !Phenotype %in% c('No difference')) %>% left_join(., ortho, by=c('Gene' = 'berghei')) %>% mutate(gene_id = coalesce(falciparum, Gene)) %>% mutate(across(gene_id, ~str_replace(., '_', '-'))) %>% filter(str_detect(gene_id, 'PF3D7')) %>% left_join(., .x, by=c('gene_id'))  %>% slice_sample(., n=150)})

map(pheno_pf_pb_ag_gns, ~.x  %>% ggplot(aes(x= FLE_exp, y = F_exp, shape = Species,color = Phenotype)) + geom_point() + geom_abline())

map(pheno_pf_pb_ag_gns, ~.x  %>% ggplot(aes(x= FLE_celProp, y = F_celProp, shape = Species, color = Phenotype)) + geom_point() + geom_abline())
```


```{r}
##read table with fhe fle unspliced spliced counts for manuscript
fhe_fle_ussr_dta = readRDS("v3_science_manuscript/tables/fhe_fle_ussr_dta.RDS")  %>% mutate(across(everything(), ~ifelse(is.nan(.) | is.infinite(.), NA, .)))

phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_export = phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_lst[["MSC14"]] %>% dplyr::select(c("Species", "Gene", "Disruption", "Ookinete", "Oocyst", "Sporozoite",  "gene_id", "F_exp", "FLE_exp", "F_celProp", "FLE_celProp", "Lasonder"))

##Put the unspliced and average expression tables in one list
ussr_avg_exprsn_dta <- list(fhe_fle_ussr_dta, phenopDB_pf_pb_yoel_ortho_mosq_gnexp_lasonder_export) %>% set_names(c('8.Unspliced-spliced ratios', '9.Msq phenoplasm gns & exp'))
```



```{r, results='asis', fig.height=4, fig.width=12}
us_s_lasonder <- plot_grid(fhe_fle_ussr, 
                         phenop_mosq_plts_lst[[3]] + 
  theme(legend.position = 'bottom',
        legend.direction = 'vertical',
        legend.spacing.x = unit(-0.04, units = 'cm'),
        legend.box.spacing = unit(-0.08, units = 'cm')),
                         fhe_fle_umap[[1]]+
  theme(legend.text = element_text(size = 11),
        plot.margin = margin(t=1, unit = "cm")),
  mhe_mle_umap[[1]]+
  theme(legend.text = element_text(size = 11),
        plot.margin = margin(t=1, unit = "cm")),
                         ncol = 4,
                         rel_widths = c(.22, .28,.25,.25),
              labels = c('D', 'E', 'F', 'G'),
              label_size = 20)

us_s_lasonder
```


```{r, results='asis', fig.height=4, fig.width=12}
## combine plots for LE exploration
core_g_lg <- cowplot::get_legend(mml_core_gns_barp + 
  theme(legend.position = 'bottom') +
    guides(fill=guide_legend(title.position = "top", nrow = 2, title = "Stage")))


fl_ml_mosq_pt <- plot_grid(
  phenop_mosq_plts_facet,
  plot_grid(
    plot_grid(plotlist = map(list(ffl_core_gns_barp, mml_core_gns_barp), ~.x +
                               theme(legend.position = 'none',
                                     # plot.margin = margin(t=1, unit = "cm")
                                     )),
              rel_widths = c(.6,.4)),
  core_g_lg,
    nrow = 2,
    rel_heights = c(.8,.2)),
  ncol = 2,
                         rel_widths = c(.5,.5),
              labels = c('D', 'E'),
              label_size = 20)

fl_ml_mosq_pt
```


```{r, results='asis', fig.height=16, fig.width=12}
##get gt tables as grob for merging using cowplot from gt github https://github.com/rstudio/gt/issues/961

##Get legend for the GO ontologies
onts = c('Biological process', 'Cellular component', 'Molecular function')
ont_lg <- cowplot::get_legend(data.frame(ont = onts) %>% ggplot(aes( y=1, fill=ont)) + scale_fill_manual(values = brewer.pal(n = 3, name = 'Pastel2') %>% set_names(onts)) + geom_bar() + theme(legend.direction = 'horizontal', legend.title = element_text(size = 14, face = 'bold'), legend.text = element_text(size = 14))+ guides(fill = guide_legend(title = 'Ontology:', nrow = 1)))

##Save ontology legend above for use in other plots
saveRDS(ont_lg, "plots/DE_heatmaps/ont_lg.RDS")

##PUT plots together for the manuscript
plt= plot_grid(stg_strn_distb_bar[[1]]+
                 theme(axis.title.y = element_text(size = 16, face = 'plain'),
                       axis.text.y = element_text(size = 15),
                       legend.title = element_text(size = 16),
                       legend.text = element_text(size = 15),
                       strip.text =element_text(size=17,face="bold", colour = 'black'),
                       legend.margin = margin(t=-5)),
               NULL,
               ggdraw(fld_gn_expression_specificity_upset_vn_don3)+
                theme(plot.margin = margin(t=0.5, unit = "cm")),
               plot_grid(plotlist = all_stg_gtpt[1:8], ncol = 5),
               ont_lg,
               NULL,
               us_s_lasonder,
               nrow = 7, 
               rel_heights = c(.17,0.01,.32,.27,0.02,0.01,.20),
              labels = c('A','', 'B', 'C', '', '', ''),
              label_size = 20
               )
plt
save_plot('plots/figS26/final_pt.svg',plt, base_width = 12, base_height = 16.5)
```