---
title: "MSC DE (fe/male low-expression VS fe/male hig_count) & GSEA"
date: '2023-May-27'
author: 'Jesse Rop'
output: 
  html_document:
    toc: true
    theme: united
    code_folding: hide
    df_print: paged
---
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr) 
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  library(ggplot2)
  library(tidyr)
  library(cowplot)
  library(SeuratData)
  library(SeuratDisk)
  library(pheatmap)
  library(plotly)
  library(rmarkdown)
  library(RColorBrewer)
  library(MAST)
  library(EnhancedVolcano)
  # library(org.Pf.plasmo.db)
  library(clusterProfiler) ##github installation newest version https://github.com/YuLab-SMU/clusterProfiler/issues/434#issuecomment-1048633162 remotes::install_github(c("YuLab-SMU/DOSE", "YuLab-SMU/clusterProfiler", "YuLab-SMU/enrichplot" ))
  library(Platypus)
  library(ggvenn)
  library(tradeSeq)
  library(MetBrewer)
  library(ComplexHeatmap)
  library(fgsea)
  library(msigdbr)
  library(slingshot)
  library(enrichplot)
  library(chromoMap)
  library(pdftools)
  library(SCpubr)
  library(GGally)
  library(circlize)
  library(readxl)
  library("ggVennDiagram")
  library(UpSetR)
  library(ComplexUpset)
})
```

# DE within MSC14 field parasites and DE on MSC14 VS MCA V3 parasites

## Variable declarations, read in input files and exploratory visualizations

```{r setup}
##Set working directory
knitr::opts_knit$set(root.dir = '/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali')
```

```{r}
##Assign colour variables for the different plots

sp_colors <- c('Early ring'='#D1EC9F',
                     'Late ring'='#78C679',
                     'Early trophozoite'='#FEEEAA',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                     'Gametocyte (developing)'='thistle',
                     'Gametocyte (female)'='#551A8B',
                     'Gametocyte (male)'='mediumpurple',
               'Unassigned'='red', 
               'Not assigned'='#D6CFC7',
               'Female (LE)'='#208cf7',
               # 'Female (HE)'='#551A8B',
               'Female'='#551A8B',
               'Male (LE)'='#a894d1',
               'Male'='mediumpurple',
               # 'Male (HE)'='mediumpurple',
               'Failed QC' = 'grey',
               'Gametocyte (committed)' = '#C399A2', 
               'Gametocyte (developing)' = '#66444C', 
               'Gametocyte (branching)' = '#66444C', 
               'Gametocyte (early female)' = '#749E89', 
               'Gametocyte (early male)' = '#7D87B2', 
               'Gametocyte (late female)' = '#4E6D58', 
               'Gametocyte (late male)' = '#41507B'
)

sp_fld_colors <- c('Late ring'= '#78C679',#met.brewer('Java')[2],
                   'Early trophozoite'='#FEEEAA',#met.brewer('Java')[3],
                   'Female (LE)'= met.brewer('Ingres')[5],
                   # 'Female (HE)'= '#551A8B', #met.brewer('Ingres')[7],
                   'Female'= '#551A8B', #met.brewer('Ingres')[7],
                   # 'Male (LE)'=met.brewer('Nizami')[6],
                   # 'Male (HE)'=met.brewer('Nizami')[8],
                   # 'Gametocyte (male)'=met.brewer('Austria')[5],
                   'Male (LE)'=met.brewer('Ingres')[7],#'#a2e19e',
                   'Male'= 'mediumpurple',#'#039911',
                   ## 'Male (HE)'= 'mediumpurple',#'#039911',
                   'Gametocyte (male)'='#3DED97',
                   'Gametocyte (female)'=met.brewer('Ingres')[6],
                   'Lab' = '#D3D3D3',
                   'Unassigned'='#DCDCDC',
                   'Failed QC' = 'grey',
                   'Not assigned' = 'grey',
                   'Early ring'='#D1EC9F',
                     'Late trophozoite'='#FEB24C',
                     'Early schizont'='#C9E8F1',
                     'Late schizont'='#85B1D3',
                   'Gametocyte (developing)'='thistle'
)

# strain_dset_lvls = c( "G1", "G2", "G3", "Lab")
# strain_dset_lvls = c( 'SC1', 'SC2','SC3',  'SC4', 'SC5',  'SC6', 'S7', 'Lab' )
strain_dset_lvls = c( 'SC1', 'SC2','SC3',  'SC4', 'SC5',  'SC6', 'SC7', 'SC8', 'Lab', 'Doublet')

```


```{r}
##Copy souporcell files for the optimum clusters

##Sample names
# sample_name_nms <- c("msc33", "msc48","msc49", "msc55", "msc60", "msc66") %>% sort()
sample_name_nms <- c("MSC1", "MSC3","MSC13") %>% str_sort(numeric = T)
sample_name_nms_all <- c("MSC1", "MSC3","MSC13","MSC14") %>% str_sort(numeric = T)

##Sample irods IDs
# source_irods_nms <- list("5736STDY13782211","5736STDY13782212","5736STDY13782213","5736STDY13782214","5736STDY13782215", "5736STDY13437393") %>% set_names(sample_name_nms)

source_irods_nms <- list("5736STDY11771535","5736STDY11771536","5736STDY11771544") %>% set_names(sample_name_nms)


# sample_ids <- paste0(sample_name_nms, '_', source_irods_nms) %>% set_names(sample_name_nms)
sample_ids <- paste0(sample_name_nms, '_', source_irods_nms) %>% set_names(sample_name_nms)

trip = 'Mali'


```



```{r, message=FALSE}
##Metadata from all p.falciparum genes downloaded from plasmoDB
Pf3D7_genes_plasmoDB <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/Pf3d7_Genes_Summary_270922.csv") %>% 
  dplyr::rename("gene_id" = `Gene ID` ) %>% 
  group_by(gene_id) %>% 
  add_count(name='gn_iso') %>% 
  ungroup() %>%
  mutate(across(gene_id, ~case_when(gn_iso>1 ~ source_id,
                                    TRUE ~ as.character(.)))) %>% 
  mutate(across(gene_id, ~str_replace(.,"_", "-")))%>%
  dplyr::rename('Gene_name' = `Gene Name or Symbol`, 'Ka_Ks' = `NonSyn/Syn SNP Ratio All Strains`, 'TM_domains'=`# TM Domains`) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A")),
         Gene = coalesce(`Gene_name`, gene_id)
  )  %>%
  group_by(Gene) %>% 
  add_count() %>% 
  ungroup() %>%
  mutate(across(Gene, ~case_when(n>1 ~ paste0(.,'_', gene_id),
                                 TRUE ~ as.character(.)))) 

```


```{r}
##Field datasets (MSC1,3,13,14,1272C)

msc_qcd_anotd <- readRDS("data/processed/DB52e_star/msc_qcd_anotd.RDS") 

```


```{r}
## Copy the gene lists to make do GO overrepresentation test
mfet_gns_3.13.14_don_colps <- readRDS("data/processed/msc14/5736STDY11771546_DB52e_star/mf_le_he/mfet_gns_3.13.14_don_colps.RDS")

```

```{r, eval=T}
##Copy souporcell files for the optimum clusters
trip = "Mali"

# for (k in c(6,8)) {
# for (smpl in c('stage_afm_strain_qcd', 'stage_ag_strain_qcd', 'strain_qcd')) {
for (i in 1:length(sample_name_nms_all)) {

  system(paste0('mkdir -p /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(sample_name_nms_all[i]),'/vcyto_results')
          )
        system(paste0('rsync -av jr35@farm5:/lustre/scratch126/tol/teams/lawniczak/users/jr35/phd/',trip,'/data/processed/DB52e_star/', str_to_lower(sample_name_nms_all[i]),'/vcyto_results/*.loom /Users/jr35/Google\\ Drive/My\\ Drive/PHD_documents/malaria_phd/',trip,'/data/processed/DB52e_star/', str_to_lower(sample_name_nms_all[i]),'/vcyto_results/')
          )
      
}
```



```{r}
## Read loom
# m14_s_us_loom <- SeuratWrappers::ReadVelocity("data/processed/msc14/5736STDY11771546_DB52e_star/vcyto_results/possorted_genome_bam_JNORH.loom")

m_s_us_loom <- list.files(paste0("data/processed/DB52e_star"), pattern = "possorted_genome_bam.*loom", full.names = T, recursive = T, include.dirs = T) %>%
set_names(nm = (str_remove_all(.,'data/processed/DB52e_star/|vcyto_results/') %>% str_replace_all(., '/', '_'))) %>%
  map(., ~SeuratWrappers::ReadVelocity(.))
  
m_s_us_loom <- m_s_us_loom[as.numeric(map(str_to_lower(sample_name_nms_all), ~which(str_detect(.x, c(str_replace(names(m_s_us_loom), "_possorted.*", "\\\\b"))))))]
```


```{r}
## fhe fle
m_stgs = c('Female', 'Female (LE)', 'Male', 'Male (LE)', 'Late ring', 'Early trophozoite')

m_s_us_ratio <- pmap(list(m_s_us_loom, map(names(m_s_us_loom), ~str_remove_all(., "msc[0-9]*_|.loom")), msc_qcd_anotd[1:4]), ~{
  loom = ..1
  l_id = ..2
  seu = ..3
  map(m_stgs, ~{
    tryCatch({
  rowSums( loom$unspliced[,seu@meta.data %>% 
                         filter(stageHL == .x) %>% 
                         rownames() %>% 
                         str_replace(., '-1', 'x') %>% 
                         paste0(l_id, ':',.)])/rowSums(loom$spliced[,seu@meta.data %>% 
                            filter(stageHL == .x) %>% 
                            rownames() %>% 
                            str_replace(., '-1', 'x') %>% paste0(l_id, ':',.)])
    }, error = function(e) {
      message(as.character('no unassigned cells')) # exclude this line to return NULL quietly
      return(NULL)
    })
}) %>% set_names(m_stgs) %>% discard(is.null) %>% imap(., ~data.frame(.x) %>% set_names(.y)) %>% bind_cols %>% filter(if_any(everything(), ~ .>0 & .<Inf)) %>% rownames_to_column('gene_id') %>% mutate(across(gene_id, ~str_replace(., '_', '-')))
  })

```


```{r}
pf_intronic_gns <- Pf3D7_genes_plasmoDB %>% filter(`# Exons in Gene`>1) %>% pull(gene_id)

pf_intronic_gns <- Pf3D7_genes_plasmoDB %>% filter(`# Exons in Gene`>1) %>% pull(gene_id)


map(m_s_us_ratio, ~.x %>% filter(!gene_id %in% pf_intronic_gns ))

```


```{r}
## fhe fle

map(m_s_us_ratio, ~..1 %>% pivot_longer(cols = -c('gene_id'), names_to = 'stage', values_to = 's_us_ratio') %>% filter(!is.na(s_us_ratio) & s_us_ratio!=Inf) %>% ggplot(aes(x=stage, y=log10(s_us_ratio))) + geom_boxplot())
```


#### Average expression of 511 Lasonder translationally repressed genes in MSC {.tabset}
```{r, message=FALSE}
lasonder_tr <- read_xlsx('data/raw/published_lit/lasonder/supp_512_TR_genes.xlsx', skip = 4)
tr_gns <- lasonder_tr %>% dplyr::rename('gene_id' = `Gene ID...1`) %>% pull(gene_id) %>% str_replace(.,'_', '-')
tr_gns <- tr_gns[str_detect(tr_gns, 'PF3D7-')]
```

```{r}
## fhe fle

map(m_s_us_ratio, ~..1 %>% filter(gene_id %in% tr_gns) %>% pivot_longer(cols = -c('gene_id'), names_to = 'stage', values_to = 's_us_ratio') %>% filter(!is.na(s_us_ratio) & s_us_ratio!=Inf) %>% ggplot(aes(x=stage, y=log10(s_us_ratio))) + geom_boxplot())

map(m_s_us_ratio[2:4], ~..1 %>% filter(gene_id %in% tr_gns) %>% filter(if_all(contains("Female"), ~.>0 & .<Inf)) %>% ggplot(aes(x= log10(`Female (LE)`), y=log10(`Female`))) + geom_point() + geom_abline() + geom_smooth())
```


```{r}
## genes in lasonder and in fh and fl

map2(m_s_us_ratio[2:4], msc_3_13_14_fle_p[1:3], ~.x %>% filter(`Female (LE)` != Inf & gene_id %in% .y & gene_id %in% tr_gns) %>% arrange(desc(`Female (LE)`)))
```


##gene sets in phenoplasm output
```{r}
## Copy the gene lists to make do GO overrepresentation test
phenop_lehe <- read_delim('data/raw/phenoplasm_output/phenotype_both_LE_HE.txt', delim = '\t', skip = 2)
phenop_nle <- read_delim('data/raw/phenoplasm_output/phenotype_not_in_LE.txt', delim = '\t', skip = 2)

```

```{r}
## Read Ngwa 2013 paper on the assesment of changes in transcriptome after 30 minutes activation using (SSH) technique method which amplifies induced genes while simultaneously suppressing house-keeping genes, and is thus particularly suited to discover new or unexpected genes
ngwa_activ_gams <- read_csv("data/raw/published_lit/ngwa_2013_activated_gametocytes_t_ome/initial_phase_of_mosquito_transcriptome.csv") %>% rename('gene_id' = `Plasmodb ID`) %>% mutate(across(gene_id, ~str_replace(.,'_', '-')))

```

```{r}
## Check the prevalence of the different subsets in the Ngwa genes

# mfet_gns_3.13.14_don_colps %>% filter(FLE_genes == 'y') %>% filter(gene_id %in% ngwa_activ_gams$gene_id)
# mfet_gns_3.13.14_don_colps %>% filter(MLE_genes == 'y') %>% filter(gene_id %in% ngwa_activ_gams$gene_id)

prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$MLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$MLE_genes),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$FLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$FLE_genes),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$F_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$F_genes),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$M_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$M_genes),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$fh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$fh_only),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$mh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$mh_only),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$gam_noasex == 'y' & !is.na(mfet_gns_3.13.14_don_colps$gam_noasex),]$gene_id %in% ngwa_activ_gams$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$asex_nogam == 'y' & !is.na(mfet_gns_3.13.14_don_colps$asex_nogam),]$gene_id %in% ngwa_activ_gams$gene_id))


```

```{r}
##  talman male gamete proteome - gametocytes were allowed to activate in exflagellation medium - supernatant with male gametes that had swum from the pellet
talman_male_gamete_pome <- read_xlsx("data/raw/published_lit/talman2014_male_gamete_proteins/male_gamete_proteome.xlsx", skip = 3)

##ortholog file to translate to falciparum IDs
ortho <- read_csv("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/ortho4.csv")

talman_male_gamete_pome <- talman_male_gamete_pome %>% inner_join(., ortho %>% dplyr::select(berghei, falciparum) %>% mutate(across(berghei, ~str_remove(., '0$'))), by = c('Accession Number'='berghei')) %>% mutate(across(falciparum, ~str_replace(.,'_', '-'))) %>% rename('gene_id' = falciparum)
```



```{r}
## Check the prevalence of the different subsets in the talman male gamete proteome

prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$MLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$MLE_genes),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$FLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$FLE_genes),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$F_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$F_genes),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$M_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$M_genes),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$fh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$fh_only),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$mh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$mh_only),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$gam_noasex == 'y' & !is.na(mfet_gns_3.13.14_don_colps$gam_noasex),]$gene_id %in% talman_male_gamete_pome$gene_id))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$asex_nogam == 'y' & !is.na(mfet_gns_3.13.14_don_colps$asex_nogam),]$gene_id %in% talman_male_gamete_pome$gene_id))


```


```{r}
##  talman male gamete proteome - gametocytes were allowed to activate in exflagellation medium - supernatant with male gametes that had swum from the pellet
zeeshan_kinesin13ptdKO_WTDE <- read_xls("data/raw/published_lit/zeeshan2022_kinesins/DE_kinesinKO_WTactivgam.xls", sheet = 'S1C', skip = 2)

##ortholog  translate to falciparum IDs
zeeshan_kinesin13ptdKO_WTDE <- zeeshan_kinesin13ptdKO_WTDE %>% inner_join(., ortho %>% dplyr::select(berghei, falciparum), by = c('Gene ID'='berghei')) %>% mutate(across(falciparum, ~str_replace(.,'_', '-'))) %>% rename('gene_id' = falciparum)
```


```{r}
## Check the prevalence of the different subsets in the talman male gamete proteome

prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$MLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$MLE_genes),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$FLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$FLE_genes),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$F_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$F_genes),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$M_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$M_genes),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$fh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$fh_only),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$mh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$mh_only),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$gam_noasex == 'y' & !is.na(mfet_gns_3.13.14_don_colps$gam_noasex),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$asex_nogam == 'y' & !is.na(mfet_gns_3.13.14_don_colps$asex_nogam),]$gene_id %in% (zeeshan_kinesin13ptdKO_WTDE %>% filter(`Log2Fold Change` <0) %>% pull(gene_id))))


```


```{r}
##  Hitz pfmap2 on male gametogenesis and gene expression
##  ON pfmap2 switched on; Off pfmap2 switched off; act. - activated for 15 min by XA 
## Cy5/Cy3 microarray
map2KO_WT_male_exflagellateDE <- read_xlsx("data/raw/published_lit/hitz2020_pfmap2_malegametogenesis/map2KO_WT_male_exflagellateDE.xlsx") %>% mutate(across(GeneID, ~str_replace(.,'_', '-'))) %>% rename('gene_id' = 'GeneID') 

read_xlsx("data/raw/published_lit/hitz2020_pfmap2_malegametogenesis/map2KO_WT_male_exflagellateDE.xlsx")
```


```{r}
##Correlation of 15 min XA(xanthurenic acid)-activated VS non-activated stage V gametocytes transcriptome from hitz etal 2020. They report no DE same as here. 
map2KO_WT_male_exflagellateDE %>% 
  mutate(across(-c('gene_id', 'DESCRIPTION', 'gametocyte genes (Young, 2005)'), as.numeric)) %>% 
  rowwise() %>%
    mutate(avg_XA_act_gam = mean(c_across(`stage V ON act. (repl. 1)`:`stage V ON act. (repl. 2)`)),
           avg_gam = mean(c_across( `stage V ON (repl. 1)`:`stage V ON (repl. 2)`))) %>%
  ggplot(aes(x= avg_XA_act_gam, y= avg_gam)) + 
  geom_point()
```

```{r}
## Check the prevalence of the different subsets in the talman male gamete proteome

prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$MLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$MLE_genes),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$FLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$FLE_genes),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$F_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$F_genes),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$M_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$M_genes),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$fh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$fh_only),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$mh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$mh_only),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$gam_noasex == 'y' & !is.na(mfet_gns_3.13.14_don_colps$gam_noasex),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$asex_nogam == 'y' & !is.na(mfet_gns_3.13.14_don_colps$asex_nogam),]$gene_id %in% (map2KO_WT_male_exflagellateDE %>% filter(`stage V ON act. (repl. 1)` > 3 & `stage V ON act. (repl. 2)` > 3)  %>% pull(gene_id))))


```

C3 indicated to be the earliest developmental stage, followed by C1, C0, and C2 and ending with C4 as the terminal cluster
```{r}
##  scRNAseq of mosquito midgut smartseq2 a.gambiae fed nf54 gams and microdisection used to harvest viable parasites from timepoint 2hours to 20 hours . 
## From unfertilized female gametes to zygote / ookinete
## C3 indicated to be the earliest developmental stage, followed by C1, C0, and C2 and ending with C4 as the terminal cluster
## clusters C3 and C1 align with pt0 and represent early zygote development and C0 aligns with pt1 and represents the maturing zygote to early ookinete development, whereas C2 and C4 align with pt2 and represent mid- to late ookinete development

moha_DE_cluster_markers <- read_xlsx("data/raw/published_lit/mohammed2023_scrnaseq_mosq_gut/DE_cluster_markers.xlsx", skip=1) %>% mutate(across(`Gene ID`, ~str_replace(.,'_', '-'))) %>% rename('gene_id' = 'Gene ID') 

##Very few genes for meaningful comparison
moha_DE_cluster_markers %>% filter(cluster %in% c('C_3', 'C_1') & `p_val_adj (Bonferroni correction)` < 0.05)
```

C3 indicated to be the earliest developmental stage, followed by C1, C0, and C2 and ending with C4 as the terminal cluster
```{r}
##  Kaneko gsnf2 k vs wt

kaneko_gnsf2_ko_gns <- read_xlsx("data/raw/published_lit/kaneko2023_gsnf2_male_game/kaneko2023_gsnf2_chromatin.xlsx") 

##ortholog  translate to falciparum IDs
kaneko_gnsf2_ko_gns <- kaneko_gnsf2_ko_gns %>% inner_join(., ortho %>% dplyr::select(berghei, falciparum), by = c('ID'='berghei')) %>% mutate(across(falciparum, ~str_replace(.,'_', '-'))) %>% rename('gene_id' = falciparum)
```


```{r}
## Check the prevalence of the different subsets in the talman male gamete proteome

prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$MLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$MLE_genes),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$FLE_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$FLE_genes),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$F_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$F_genes),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$M_genes == 'y' & !is.na(mfet_gns_3.13.14_don_colps$M_genes),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$fh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$fh_only),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$mh_only == 'y' & !is.na(mfet_gns_3.13.14_don_colps$mh_only),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$gam_noasex == 'y' & !is.na(mfet_gns_3.13.14_don_colps$gam_noasex),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))
prop.table(table(mfet_gns_3.13.14_don_colps[mfet_gns_3.13.14_don_colps$asex_nogam == 'y' & !is.na(mfet_gns_3.13.14_don_colps$asex_nogam),]$gene_id %in% (kaneko_gnsf2_ko_gns %>% filter(!is.na(`Target of gSNF2`) & !is.na(`male-enriched`)) %>% pull(gene_id))))


```

## MSC DE

### female high vs female low and male low vs male high 

Process takes long so saving analysis to read from file.
```{r, eval=F}
#DO NOT DELETE

# set ident
msc_qcd_anotd[1:4] <- purrr::map(msc_qcd_anotd[1:4], ~SetIdent(., value='stageHL'))

fh_fl_marker_list_3.13.14 <- msc_qcd_anotd[2:4] %>%
  purrr::map(.  %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Female') %>% rownames(),
                         ident.2= .@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         logfc.threshold = -Inf,
                         min.pct=-Inf
  ))

saveRDS(fh_fl_marker_list_3.13.14, "data/processed/DB52e_star/fh_fl_marker_list_3.13.14.RDS")

# mh_ml_marker_list_13.14 <- msc_qcd_anotd[1:4] %>%
mh_ml_marker_list_13.14 <- msc_qcd_anotd[3:4] %>%
  purrr::map(.  %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Male') %>% rownames(),
                         ident.2=.@meta.data %>% filter(stageHL == 'Male (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         logfc.threshold = -Inf,
                         min.pct=-Inf
  ))

saveRDS(mh_ml_marker_list_13.14, "data/processed/DB52e_star/mh_ml_marker_list_13.14.RDS")

```

```{r, eval=F}
#DO NOT DELETE
fh_fl_marker_list_acc_strn_3.13.14 <- msc_qcd_anotd[2:4] %>%
  purrr::map(., ~.x[,!is.na(.x@meta.data$Strain_DN)]  %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Female') %>% rownames(),
                         ident.2= .@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         logfc.threshold = -Inf,
                         min.pct=-Inf,
                         latent.vars = 'Strain_DN'
  ))

saveRDS(fh_fl_marker_list_acc_strn_3.13.14, "data/processed/DB52e_star/fh_fl_marker_list_acc_strn_3.13.14.RDS")

mh_ml_marker_list_acc_strn_13.14 <- msc_qcd_anotd[3:4] %>%
  purrr::map(., ~.x[,!is.na(.x@meta.data$Strain_DN)] %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Male') %>% rownames(),
                         ident.2=.@meta.data %>% filter(stageHL == 'Male (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         logfc.threshold = -Inf,
                         min.pct=-Inf,
                         latent.vars = 'Strain_DN'
  ))

saveRDS(mh_ml_marker_list_acc_strn_13.14, "data/processed/DB52e_star/mh_ml_marker_list_acc_strn_13.14.RDS")

```


```{r, eval=T}
#DO NOT DELETE
fh_fl_marker_list_acc_strn_3.13.14 <- readRDS("data/processed/DB52e_star/fh_fl_marker_list_acc_strn_3.13.14.RDS")  
mh_ml_marker_list_acc_strn_13.14 <- readRDS("data/processed/DB52e_star/mh_ml_marker_list_acc_strn_13.14.RDS")

```

```{r, eval=F}
#DO NOT DELETE
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 <- msc_qcd_anotd[2:4] %>%
  purrr::map(., ~.x[,!is.na(.x@meta.data$Strain_DN)] %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Female') %>% rownames(),
                         ident.2= .@meta.data %>% filter(stageHL == 'Female (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         # logfc.threshold = -Inf,
                         # min.pct=-Inf,
                         latent.vars = 'Strain_DN'
  ))

####!! MANUSCRIPT TABLE F VS FL DE
saveRDS(fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14, "data/processed/DB52e_star/fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14.RDS")

mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14 <- msc_qcd_anotd[3:4] %>%
  purrr::map(., ~.x[,!is.na(.x@meta.data$Strain_DN)] %>% FindMarkers(.,
                         ident.1 =.@meta.data %>% filter(stageHL == 'Male') %>% rownames(),
                         ident.2=.@meta.data %>% filter(stageHL == 'Male (LE)') %>% rownames(),
                         slot = "data",
                         test.use = 'MAST',
                         latent.vars = 'Strain_DN'
  ))

saveRDS(mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14, "data/processed/DB52e_star/mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14.RDS")

```


Merge DE list with annotation
```{r}
##female - !!Need to confirm up to date status - fh_fl_marker_list_3.13.14.RDS generated in above so and the chunk turned off to save time then file is stored
fh_fl_marker_list_3.13.14 <- readRDS( "data/processed/DB52e_star/fh_fl_marker_list_3.13.14.RDS")
fh_fl_marker_list_3.13.14 <- fh_fl_marker_list_3.13.14 %>%
  purrr::map(. %>% rownames_to_column('gene_id') %>% 
        left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') 
  )

##Write out fh_fl_list for msc14 publication
write.csv(fh_fl_marker_list_3.13.14[[1]], "v3_science_manuscript/tables/msc14_fh_vs_fl.csv")
```


```{r, eval=F}
fh_fl_marker_list_3.13.14 %>%
  purrr::map(. %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj))
```

```{r}
##female -!!Need to confirm up to date status - fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14.RDS generated in above so and the chunk turned off to save time then file is stored

fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 <- readRDS( "data/processed/DB52e_star/fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14.RDS")
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 <- fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 %>%
  purrr::map(. %>% rownames_to_column('gene_id') %>% 
        left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') %>%
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, p_val_adj, `Product Description` ) %>%
        arrange(desc(avg_log2FC))
  )

##Write out fh_fl_list for msc14 publication
write.csv(fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]], "v3_science_manuscript/tables/msc14_fh_vs_fl_acc_strn_lfc.25_mp.1.csv")

##male -!!Need to confirm up to date status - mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14.RDS generated in above so and the chunk turned off to save time then file is stored

mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14 <- readRDS( "data/processed/DB52e_star/mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14.RDS")
mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14 <- mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14 %>%
  purrr::map(. %>% rownames_to_column('gene_id') %>% 
        left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') %>%
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, p_val_adj, `Product Description` ) %>%
        arrange(desc(avg_log2FC))
  )

##Write out mh_ml_list for msc14 publication
write.csv(mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]], "v3_science_manuscript/tables/msc14_mh_vs_ml_acc_strn_lfc.25_mp.1.csv")
```

```{r}
##male -!!Need to confirm up to date status - mh_ml_marker_list_13.14.RDS generated in above so and the chunk turned off to save time then file is stored
mh_ml_marker_list_13.14 <- readRDS( "data/processed/DB52e_star/mh_ml_marker_list_13.14.RDS")
mh_ml_marker_list_13.14 <- mh_ml_marker_list_13.14 %>%
  purrr::map(. %>% rownames_to_column('gene_id') %>% 
        left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') 
  )


```

```{r, eval=F}
##male
mh_ml_marker_list_13.14 %>%
  purrr::map(. %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj))
```

Tabulate selec set of genes
```{r, eval=F}

fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 %>%
  purrr::map(. %>% dplyr::filter(pct.1 >0.9 &  pct.2 <0.1 ) %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2,  p_val_adj))
```

```{r, eval=F}

fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14 %>%
  purrr::map(. %>% dplyr::filter(avg_log2FC < 0 & pct.1 <0.5 &  pct.2 >0.3) %>% arrange(avg_log2FC) %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2,  p_val_adj))
```

#### female high vs female low DE visualization
```{r, fig.height=7, fig.width=7}

fh_fl_vlcano <- EnhancedVolcano(fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]],
                lab = fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]]$Gene,
                selectLab =fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% 
                  filter(p_val_adj < 0.05 & (avg_log2FC > 2 | avg_log2FC < -2) & !str_detect(Gene, 'PF3D7')) %>%
                  pull(Gene),
                x = 'avg_log2FC',
                y = 'p_val_adj',
                pCutoff = 0.05,
                FCcutoff = 2,
                pointSize = 1.5,
                labSize = 5.0,
                title = '',
                subtitle = "",
                titleLabSize = 0,
                subtitleLabSize = 0,
                axisLabSize = 24,
                drawConnectors = TRUE,
                legendLabels =c(expression(Log[2] ~ FC), "p-value"),
                gridlines.minor = FALSE)

fh_fl_vlcano

##Save the above image for plot grid in functional_GO_n_lit_comparison.RMD for manuscript
saveRDS(fh_fl_vlcano, "plots/figS26/fh_fl_vlcano.RDS")
```

```{r}
##Number of genes for manuscript
#Significant P <0.05
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05) %>% dim
#upregulated in F
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05 & avg_log2FC > 0) %>% dim
#upregulated in FLE
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05 & avg_log2FC < 0) %>% dim

#Highly DE lfc 2
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05 & (avg_log2FC > 2 | avg_log2FC < -2)) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05 & avg_log2FC > 2) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<0.05 & avg_log2FC < -2) %>% dim


##Number of genes for manuscript
#Highly significant P= 10e-70
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70 & avg_log2FC > 0) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70 & avg_log2FC < 0) %>% dim

#Highly DE lfc 2
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70 & (avg_log2FC > 2 | avg_log2FC < -2)) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70 & avg_log2FC > 2) %>% dim
fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>% filter(p_val_adj<10e-70 & avg_log2FC < -2) %>% dim
```



#### female high vs female low DE visualization
```{r, fig.height=7, fig.width=7}

EnhancedVolcano(mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]],
                lab = mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]]$Gene,
                selectLab =mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]] %>% 
                  filter(p_val_adj < 10e-15 & (avg_log2FC > 1 | avg_log2FC < -1) & !str_detect(Gene, 'PF3D7')) %>%
                  pull(Gene),
                x = 'avg_log2FC',
                y = 'p_val_adj',
                pCutoff = 10e-15,
                FCcutoff = 1,
                pointSize = 3.0,
                labSize = 7.0,
                title = '',
                subtitle = "",
                titleLabSize = 0,
                subtitleLabSize = 0,
                axisLabSize = 24,
                drawConnectors = TRUE,
                legendLabels =c(expression(Log[2] ~ FC), "p-value"),
                gridlines.minor = FALSE)
```


#### tabulate female high vs female low ovlp
```{r}

##up_in_fl_ovlp
m14_topfl_overlapping <- fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>%
  filter(avg_log2FC < 0 & p_val_adj < 1e-10) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)

##up_in_fh_ovlp
m14_topfh_overlapping <- fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14[[1]] %>%
        filter(avg_log2FC > 0 & p_val_adj < 1e-10) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)

```


#### tabulate male high vs male low ovlp
```{r}
##up_in_ml_ovlp top500
m14_topml_overlapping <-  mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]] %>%
        filter(avg_log2FC < 0 & p_val_adj < 1e-10) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)

##up_in_mh_ovlp  top500
m14_topmh_overlapping <- mh_ml_marker_list_acc_strn_lfc.25_mp1_13.14[[1]] %>%
        filter(avg_log2FC > 0 & p_val_adj < 1e-10) %>%
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)


```

#### overlap between top 500 mh,ml,fh,fl above
```{r}
ggvenn(
  list(topmh = m14_topmh_overlapping,
       topml= m14_topml_overlapping,
       topfh= m14_topfh_overlapping,
       topfl= m14_topfl_overlapping),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```

```{r, eval=T}
##FGSEA
library(fgsea)

Pf3D7_61_gaf_F <-  clusterProfiler::parse_gff("/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Pf3D7_genomes_gtfs_indexs/PlasmoDB-61_Pfalciparum3D7_GO_F.gaf")
Pf3D7_61_gaf_F[1] = Pf3D7_61_gaf_F[1] %>% purrr::map(. %>% mutate(across(Gene, ~str_replace(., '_', '-'))))

gene_ids_pf<- data.frame(Pf3D7_61_gaf_F[[1]]) %>% 
  left_join(., Pf3D7_61_gaf_F[[2]], by= c("GO" = "GOID")) %>% 
  filter(!is.na(TERM))  %>% 
  dplyr::select(TERM,Gene) %>% 
  group_by(TERM) %>% 
  summarise(genes = paste0(Gene, collapse = ",")) %>% 
  filter(str_count(genes, 'PF3D7') >10) %>% 
  deframe() %>% 
  as.list() %>% 
  purrr::map(., ~str_split_1(.x, ','))



ranks<- fh_fl_marker_list_acc_strn_3.13.14[[1]] %>% 
  rownames_to_column('gene_id')  %>%
  arrange(desc(avg_log2FC)) %>% dplyr::select(gene_id, avg_log2FC) %>% deframe(.)

fgseaRes<- fgsea(gene_ids_pf, stats = ranks, nPermSimple = 10000)

fgseaResTidy <- fgseaRes %>%
  as_tibble() %>%
  arrange(desc(NES)) 

fgseaResTidy %>% filter(padj < 0.05) %>% write_csv("v3_science_manuscript/tables/msc14_fh_fl_gsea.csv")


##clusterprofiler
gsea = clusterProfiler::GSEA(ranks , TERM2GENE = Pf3D7_61_gaf_F$TERM2GENE, TERM2NAME=Pf3D7_61_gaf_F$TERM2NAME, pvalueCutoff = 0.05)
print(enrichplot::dotplot(gsea))

```


```{r, fig.height=7, eval=T}
##FGSEA
# library(fgsea)

fh_fl_gsea_bplot<- fgseaResTidy %>% 
  filter(padj < 0.05) %>% 
  arrange(padj) %>% 
  group_by(NES<0) %>% 
  slice_head(., n = 5) %>%
  mutate(go_labs = str_wrap(pathway, width = 31)) %>%
  ggplot(aes(reorder(go_labs, NES), NES)) +
  geom_col(aes(fill = size/100)) +
  coord_flip() +
  labs(x="GO term: Function", y="NES",
       # title="Hallmark pathways NES from GSEA"
  ) + 
  theme_classic()+
  theme(
    axis.text.y  = element_text(size = 26),
    axis.text.x = element_text(size = 26),#, angle = 45
    axis.title.y  = element_text(size = 24, face = "bold"),
    axis.title.x = element_text(size = 24, face = "bold"),
    # axis.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 22, face = "bold"),
    legend.position="bottom"#,
    # plot.margin = unit(c(1,2,1,1),"cm")
    # legend.margin=margin()
  )#+ 
# guides(fill=guide_legend(title="Size\n(X100)"))

fh_fl_gsea_bplot
```


```{r}
##Save the above image for plot grid in functional_GO_n_lit_comparison.RMD for manuscript
saveRDS(fh_fl_gsea_bplot, "plots/figS26/fh_fl_gsea_bplot.RDS")
```

```{r, eval=T, fig.width= 8, fig.height=5}
##FGSEA
m_ranks<- mh_ml_marker_list_acc_strn_13.14[[1]] %>% 
  rownames_to_column('gene_id')  %>%
  arrange(desc(avg_log2FC)) %>% dplyr::select(gene_id, avg_log2FC) %>% deframe(.)

m_fgseaRes <- fgsea(gene_ids_pf, stats = m_ranks, nPermSimple = 10000)

m_fgseaResTidy <- m_fgseaRes %>%
  as_tibble() %>%
  arrange(desc(NES)) 

m_fgseaResTidy %>% filter(padj < 0.05) %>% write_csv("v3_science_manuscript/tables/msc14_mh_ml_gsea.csv")


##clusterprofiler
gsea = clusterProfiler::GSEA(m_ranks , TERM2GENE = Pf3D7_61_gaf_F$TERM2GENE, TERM2NAME=Pf3D7_61_gaf_F$TERM2NAME, pvalueCutoff = 0.05)
print(enrichplot::dotplot(gsea))


```


```{r, eval=T, fig.width= 16, fig.height=5}
##FGSEA

mh_ml_gsea_bplot <- fgseaResTidy %>% filter(padj < 0.05) %>% arrange(padj) %>% group_by(NES<0) %>% 
  slice_head(., n = 5) %>%
  ggplot(aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill = size/100)) +
  coord_flip() +
  labs(x="GO term: Function", y="Normalized\nEnrichment\nScore",
       # title="Hallmark pathways NES from GSEA"
  ) + 
  theme_classic()+
  theme(
    axis.text.y  = element_text(size = 26),
    axis.text.x = element_text(size = 26),#, angle = 45
    axis.title.y  = element_text(size = 24, face = "bold"),
    axis.title.x = element_text(size = 24, face = "bold"),
    # axis.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 22, face = "bold"),
    legend.position="bottom"#,
    # plot.margin = unit(c(1,2,1,1),"cm")
    # legend.margin=margin()
  )#+ 
# guides(fill=guide_legend(title="Size\n(X100)"))

mh_ml_gsea_bplot
```

```{r}

##Function to save tables in excel as sheets

createWorksheet <- function(x, y){
  wb <- createWorkbook()
  for(i in seq_along(x)){
    addWorksheet(wb, names(x)[i])
    writeData(wb, sheet = names(x)[i], x[[i]])
    saveWorkbook(wb, file = y, overwrite = TRUE)
  }
}

```

```{r}
##!!!! NOTE - MANUSCRIPT female vs female (LE) table

# createWorksheet(c(map(fh_fl_marker_list_acc_strn_lfc.25_mp1_3.13.14, ~.x %>% filter(p_val_adj<0.05)) %>% set_names('female(HE) VS female(LE) DE'),
#                 list(fgseaResTidy %>% filter(padj < 0.05)) %>% set_names('female(HE) VS female(LE) GSEA')),
# "../Mali/v3_science_manuscript/tables/data S11 - female vs female (LE) DE & GSEA.xlsx")

```


#### female high vs female low DE genes overlapping between samples visualization  {.tabset}

##### up_in_fh_ovlp
```{r, eval=FALSE}
ggvenn(
  list(msc3_fh = fh_fl_marker_list_3.13.14[[1]] %>% filter(p_val_adj < 1e-50  & avg_log2FC > 1) %>% pull(gene_id),
       msc13_fh= fh_fl_marker_list_3.13.14[[2]] %>% filter(p_val_adj < 1e-50  & avg_log2FC > 1) %>% pull(gene_id) ,
       msc14_fh = fh_fl_marker_list_3.13.14[[3]] %>% filter(p_val_adj < 1e-50  & avg_log2FC > 1) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```

##### up_in_fl_ovlp
```{r, eval=F}
ggvenn(
  list(msc3_fl = fh_fl_marker_list_3.13.14[[1]] %>% filter(p_val_adj < 1e-50  & avg_log2FC < -1) %>% pull(gene_id),
       msc13_fl= fh_fl_marker_list_3.13.14[[2]] %>% filter(p_val_adj < 1e-50  & avg_log2FC < -1) %>% pull(gene_id) ,
       msc14_fl = fh_fl_marker_list_3.13.14[[3]] %>% filter(p_val_adj < 1e-50  & avg_log2FC < -1) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```



##### up_in_mh_ovlp
```{r, eval=F}
ggvenn(
  list(msc13_mh = mh_ml_marker_list_13.14[[1]] %>% filter(p_val_adj < 1e-10  & avg_log2FC > 0.5) %>% pull(gene_id),
       msc14_mh= mh_ml_marker_list_13.14[[2]] %>% filter(p_val_adj < 1e-10  & avg_log2FC > 0.5) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```

##### up_in_ml_ovlp
Only one sample with male low-expression. Need to return the deleted male low-expression for MSC14
```{r, eval=F}
ggvenn(
  list(msc13_ml = mh_ml_marker_list_13.14[[1]] %>% filter(p_val_adj < 1e-10  & avg_log2FC < -0.5) %>% pull(gene_id),
       msc14_ml= mh_ml_marker_list_13.14[[2]] %>% filter(p_val_adj < 1e-10  & avg_log2FC < -0.5) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```


#### tabulate female high vs female low ovlp
```{r}

##up_in_fl_ovlp
topfl_overlapping <- fh_fl_marker_list_3.13.14[1] %>%
  purrr::map(. %>% 
        filter(avg_log2FC < 0 & p_val_adj < 1e-10) %>% 
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)) %>% 
  Reduce(base::intersect,.)

##up_in_fh_ovlp
topfh_overlapping <- fh_fl_marker_list_3.13.14[1] %>%
  purrr::map(. %>% 
        filter(avg_log2FC > 0 & p_val_adj < 1e-10) %>% 
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)) %>% 
  Reduce(base::intersect,.)

```

```{r, eval=F}

fh_fl_marker_list_3.13.14 %>%
  purrr::map(. %>% filter(avg_log2FC > 0 & p_val_adj < 1e-10) %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% arrange(avg_log2FC))

```

#### tabulate male high vs male low ovlp
```{r}
##up_in_ml_ovlp top500
topml_overlapping <- mh_ml_marker_list_13.14 %>%
  purrr::map(. %>% 
        filter(avg_log2FC < 0 & p_val_adj < 1e-10) %>% 
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)) %>% 
  Reduce(base::intersect,.)

##up_in_mh_ovlp  top500
topmh_overlapping <- mh_ml_marker_list_13.14 %>%
  purrr::map(. %>% 
        filter(avg_log2FC > 0 & p_val_adj < 1e-10) %>% 
        dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% 
        arrange(avg_log2FC) %>% 
        pull(gene_id) %>% 
        str_replace(., '-', '_') %>% 
        head(.,n=500)) %>% 
  Reduce(base::intersect,.)


```

```{r, eval=F}
##up_in_ml_ovlp top500
mh_ml_marker_list_13.14 %>%
  purrr::map(. %>% filter(avg_log2FC < 0 & p_val_adj < 1e-10) %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj) %>% arrange(avg_log2FC))

```

#### overlap between top 500 mh,ml,fh,fl above
```{r}
ggvenn(
  list(topmh = topmh_overlapping,
       topml= topml_overlapping,
       topfh= topfh_overlapping,
       topfl= topfl_overlapping),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```


#### overlap between all sets for msc3 (up in mh, low in mh, up in fh, low in fh)
```{r, eval=FALSE}
ggvenn(
  list(topmh =  mh_ml_marker_list_13.14[[1]] %>% filter(avg_log2FC > 0.5 & p_val_adj < 1e-100) %>% pull(gene_id),
       topml =  mh_ml_marker_list_13.14[[1]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-100) %>% pull(gene_id),
       topfh =  fh_fl_marker_list_3.13.14[[2]] %>% filter(avg_log2FC > 0.5 & p_val_adj < 1e-100) %>% pull(gene_id),
       topfl =  fh_fl_marker_list_3.13.14[[2]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-100) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```

#### overlap between all sets for msc14 (up in mh, low in mh, up in fh, low in fh)
```{r, eval=FALSE}
ggvenn(
  list(topmh =  mh_ml_marker_list_13.14[[2]] %>% filter(avg_log2FC > 0.5 & p_val_adj < 1e-15) %>% pull(gene_id),
       topml =  mh_ml_marker_list_13.14[[2]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-15) %>% pull(gene_id),
       topfh =  fh_fl_marker_list_3.13.14[[3]] %>% filter(avg_log2FC > 0.5 & p_val_adj < 1e-15) %>% pull(gene_id),
       topfl =  fh_fl_marker_list_3.13.14[[3]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-15) %>% pull(gene_id)),
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#A46868"),
  stroke_size = 0.5, set_name_size = 4
)
```

#### overlap between all sets for msc14 (up in mh, low in mh, up in fh, low in fh)
```{r, eval=FALSE}
fl_ml_overlap <- base::intersect(fh_fl_marker_list_3.13.14[[3]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-15) %>% pull(gene_id),
                                 mh_ml_marker_list_13.14[[2]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-15) %>% pull(gene_id))
```

#### Visualize  fh_vs_fl mh_vs_ml  the markers identified above


##### Average expression visualization {.tabset}

Function 
```{r}
makeSetScore <- function(seur, gene.set,assay='RNA', slt = 'data') {
  # Get mean expression of genes of interest per cell
  gene.set <- gene.set[gene.set %in% rownames(seur) ]
  mean.exp <- colMeans(x = as.matrix(GetAssayData(seur, slot = slt, assay = assay))[gene.set, ], na.rm = TRUE)
  
  # Add mean expression values in 'object@meta.data$gene.set.score'
  if (all(names(x = mean.exp) == rownames(x = seur@meta.data))) {
    return(mean.exp)
  }
}

```

fl_ml overlap genes
```{r}

# fl_ml_ovlp_avg_exp <- msc_qcd_anotd[1:4] %>%
#   purrr::map(~makeSetScore(., fl_ml_overlap, assay = 'RNA') %>% 
#         data.frame %>% 
#         setNames('fl_ml_overlap') 
#   )

fl_ml_ovlp_avg_exp <- msc_qcd_anotd[1:4] %>%
  purrr::map(~makeSetScore(., fh_fl_marker_list_3.13.14[[1]] %>% filter(avg_log2FC < -0.5 & p_val_adj < 1e-15) %>% pull(gene_id), assay = 'RNA') %>% 
        data.frame %>% 
        setNames('fl_ml_overlap') 
  )

# msc_qcd_anotd[1:4] <- list(msc_qcd_anotd[1:4], fl_ml_ovlp_avg_exp) %>%
#   pmap(~AddMetaData(..1, ..2)) 

msc_qcd_anotd[1:4] <- list(msc_qcd_anotd[1:4], fl_ml_ovlp_avg_exp) %>%
  pmap(~AddMetaData(..1, ..2)) 
```


```{r, results='asis', message=FALSE}
msc_qcd_anotd[1:4] <- purrr::map(msc_qcd_anotd[1:4], ~SetIdent(., value='stageHL'))
msc_qcd_anotd[1:4]  %>% 
  purrr::iwalk(.,~{
    # create tabset for each group 
    cat('###### ',.y,'   \n')
    cat('Average Female (LE) Male (LE) overlap genes expression \n')
    cat('\n')
    print(FeaturePlot(., features = "fl_ml_overlap", dims = c(1,2), pt.size = 1.5, label = T, label.size = 5, repel = T) + 
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "Spectral")))+ 
            theme(axis.text=element_text(size=12), 
                  axis.title=element_text(size=12,face="bold"), 
                  legend.text = element_text(size=12),
                  legend.position="bottom",
                  legend.key.width = unit(1, "cm")
            )+
            coord_fixed()
          #guides(colour = guide_colorbar(barwidth = 17))
    )
    # export::graph2ppt(file= paste0('/Users/jr35/Google Drive/My Drive/PHD_documents/malaria_phd/Mali/plots/markers/kes_asx_gam_gold/gam_', .y))
    
  })

```

```{r, fig.width = 20, fig.height = 5}
Idents(msc_qcd_anotd[[1]]) <- 'stageHL'
msc_qcd_anotd[[1]] %>% VlnPlot(., features = c("PF3D7-0416100", "PF3D7-1105000"), group.by = 'seurat_clusters') 
msc_qcd_anotd[[1]] %>%FeaturePlot(., reduction = "umap", features =c("PF3D7-0216000","PF3D7-0416100", "PF3D7-1232800", "PF3D7-0219400"), pt.size = 1.2,  label = T, repel = T, label.size = 5,  cols =  c("lightgrey", "red"),ncol=4) 
```


```{r, fig.width = 20, fig.height = 6}
Idents(msc_qcd_anotd[[1]]) <- 'stageHL'

msc_qcd_anotd[[1]] %>% VlnPlot(., features = c("PF3D7-1234700","PF3D7-0904200","PF3D7-0715400",  "PF3D7-1430100"), group.by = 'stageHL', ncol = 5, cols = sp_colors[c('Gametocyte (developing)', 'Early trophozoite', 'Late ring','Female','Female (LE)', 'Male',  'Male (LE)', 'Unassigned')]) 

vp <- msc_qcd_anotd[[1]] %>% VlnPlot(., features = c( "PF3D7-1234700","PF3D7-0904200","PF3D7-0715400", "PF3D7-1430100"), group.by = 'stageHL', ncol = 5, combine = F, cols = sp_colors[c('Gametocyte (developing)', 'Early trophozoite', 'Late ring','Female','Female (LE)', 'Male',  'Male (LE)', 'Unassigned')]) 


##Using match to ensure correct order
p1 <-list(vp, fh_fl_marker_list_3.13.14[[1]][match(c("PF3D7-1234700","PF3D7-0904200","PF3D7-0715400", "PF3D7-1430100"), fh_fl_marker_list_3.13.14[[1]]$gene_id), 'Gene']) %>%
  pmap(~ ..1 +labs(title = ..2) +
         theme(legend.position="none")+
         xlab('Stage') +
         theme(axis.text=element_text(size=18), 
               # axis.title=element_text(size=24,face="bold"), 
               axis.title.y  =element_text(size=18,face="bold"),
               legend.text = element_text(size=34), 
               title = element_text(size=24,face="bold"),
               axis.text.x = element_text(angle = 90),
               # axis.title.x = element_text('stageHL'),
               strip.text.x = element_text(size = 24, face = "bold")))  %>% 
  plot_grid(plotlist = .,  ncol = 4)

p1


fp <- msc_qcd_anotd[[1]] %>%FeaturePlot(., reduction = "umap", features =c("PF3D7-0715400",  "PF3D7-1234700","PF3D7-0904200","PF3D7-1430100"), pt.size = 1.2,  label = T, repel = T, label.size = 5,ncol=4, combine =F) 

list(fp, fh_fl_marker_list_3.13.14[[1]][fh_fl_marker_list_3.13.14[[1]]$gene_id %in% c("PF3D7-0715400",  "PF3D7-1234700","PF3D7-0904200","PF3D7-1430100"), 'Gene']) %>%
  pmap(~ ..1 +labs(title = ..2))  %>% plot_grid(plotlist = .,  ncol = 4)


```

## MAST DE between strains accounting for stage

```{r}
##REMEMBER TO REMOVE CLUSTERS WITH VERY FEW CELLS AND DOUBLETS SEURAT
# msc14_v3_combi <- msc14_v3_combi[, cells_de]
```


```{r, message=FALSE, warning=FALSE, eval=F}
# For performing differential expression after integration, we switch back to the original data

msc_qcd_anotd[1:4] <- purrr::map(msc_qcd_anotd[1:4], ~SetIdent(., value='Strain_DN'))

## NB!! Seurat documentation issue use ident.1 instead of cells.1 https://github.com/satijalab/seurat/issues/3001
strains_marker_list <- msc_qcd_anotd[1:4] %>% purrr::map(., ~.x[,!is.na(.x@meta.data$Strain_DN)] %>% 
                                                  FindAllMarkers(.,
                                           assay = 'RNA',
                                           slot = "data",
                                           test.use = 'MAST',
                                           latent.vars = 'stageHL'))

strains_marker_list <- strains_marker_list %>%
  purrr::map(. %>% rownames_to_column('gene_id') %>% 
        left_join(.,Pf3D7_genes_plasmoDB, by = 'gene_id') 
  )

saveRDS(strains_marker_list, "data/processed/DB52e_star/strains_marker_list.RDS")
# msc_qcd_anotd[2:5] <- purrr::map(msc_qcd_anotd[2:5], ~SetIdent(., value='seurat_clusters'))


```

```{r}
##!!Need to confirm up to date status - strains_marker_list.RDS generated above and the chunk turned off to not redundantly rerun for speed. The file is then stored
strains_marker_list <- readRDS("data/processed/DB52e_star/strains_marker_list.RDS")
```

```{r, eval=F}
strains_marker_list %>%
  purrr::map(. %>% dplyr::select(gene_id, Gene, cluster, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj))
```


Tabulate selec set of genes
```{r, eval=F}

strains_marker_list %>%
  purrr::map(. %>% dplyr::filter(avg_log2FC >2 ) %>% dplyr::select(gene_id, Gene, cluster, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj))
```

```{r, eval=F}

fh_fl_marker_list_3.13.14 %>%
  purrr::map(. %>% dplyr::filter(avg_log2FC < 0 & pct.1 <0.5 &  pct.2 >0.3) %>% arrange(avg_log2FC) %>% dplyr::select(gene_id, Gene, avg_log2FC, pct.1, pct.2, Ka_Ks, p_val_adj))
```

#### female high vs female low DE visualization
```{r, fig.height=6, fig.width=6}
EnhancedVolcano(fh_fl_marker_list_3.13.14[[1]],
                lab = fh_fl_marker_list_3.13.14[[1]]$Gene,
                x = 'avg_log2FC',
                y = 'p_val_adj',
                pCutoff = 10e-32,
                FCcutoff = 2,
                pointSize = 3.0,
                labSize = 4.0,
                title = '',
                subtitle = "",
                titleLabSize = 0,
                subtitleLabSize = 0,
                legendLabels =c(expression(Log[2] ~ FC), "p-value"))
```

Filter  genes expressed in very few cells per stage

```{r}
asx_det_gns <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Late ring', 'Early trophozoite')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

asx_det_gns <- asx_det_gns[asx_det_gns > 100] %>% names()

m_det_gn <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Male')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

m_det_gn <- m_det_gn[m_det_gn > 100] %>% names()

fl_det_gn  <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Female (LE)')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

fl_det_gn <- fl_det_gn[fl_det_gn > 100] %>% names()

ggvenn(list('as'=asx_det_gns, 'm' = m_det_gn, 'fl'= fl_det_gn))

```


```{r}
asx_det_gns <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Late ring', 'Early trophozoite')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

asx_det_gns <- asx_det_gns[asx_det_gns > 50] %>% names()

m_det_gn <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Male')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

m_det_gn <- m_det_gn[m_det_gn > 50] %>% names()

fl_det_gn  <- msc_qcd_anotd[[4]][,msc_qcd_anotd[[4]]@meta.data$stageHL %in% c('Female (LE)')]@assays$RNA@counts %>% as.matrix() %>% rowSums(.)

fl_det_gn <- fl_det_gn[fl_det_gn > 50] %>% names()

ggvenn(list('as'=asx_det_gns, 'm' = m_det_gn, 'fl'= fl_det_gn))

```
